### 2021-04-15 AI&ML/Project proposal/Project_proposal_V1.tex
Project proposal

\documentclass{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\renewcommand{\baselinestretch}{1.2} 
\title{\textbf{A decomposition based on constrained clustering algorithms for job shop scheduling problems} \\}

\author{\\[1in] Student: \\ \textit{\textbf{Mohammed El-Kholany, MSc.}}\\Alpen-Adria-Universität Klagenfurt, Austria \\[1in] \and Supervisor:\\ \textit{\textbf{Assoc. Prof. Konstantin Schekotihin}} \\  \\[1in]}
\date{April 2021}

\begin{document}

\clearpage\maketitle
\thispagestyle{empty}
\clearpage
\setcounter{page}{1}
\newpage
\section{Introduction}
Nowadays, the massive progress and development of the internet and online technologies, data generated by machines and devices, product development, quality and inventory management systems, or production planning systems has become huge and is expected to increase in the coming years. Hence, to capture long-term revenues and sustainable competitive advantages, companies must manage the knowledge and have the valuable information to make the right decision at the right time in \cite{benabdellah2019survey}. \\

It can be assumed that knowledge management is a crucial issue in the industry. To extract implicit, unknown, and potentially useful information from data, we use data mining techniques that have been responsible for many of artificial intelligence’s recent successes. Clustering is one of these techniques; clustering has always been an exploratory but critical task in the knowledge discovery process, with applications ranging from image processing, production systems, and information retrieval in \cite{benabdellah2019survey}. Clustering is a technique that aims to group datasets (objects) into clusters by defining similar objects and aggregate them in the same cluster while the dissimilar objects should be belonging to different clusters. \\

Furthermore, from an optimization perspective, the main objective is to minimize the distance between objects falling in the same cluster and maximize the distance between the others that belong to different clusters. Since clustering does not use a subset of the dataset as labels to learn a classification model, clustering differs from classification in \cite{wagstaff2001constrained}. In other words, with the terminology of machine learning, clustering is a form of an unsupervised task, which calculates the similarity between data objects without knowing the proper distribution in \cite{li2018geometric}. Due to these unsupervised characteristics, clustering is known as one of the most challenging tasks in \cite{benabdellah2019survey}. \\

Clustering has been widely applied in several disciplines; one is production-planning systems, more specifically scheduling. Scheduling is one of the most complex problems in the industry. Scheduling of operations involves resource allocation over a period of time to perform a series of tasks, where one of the most critical problems is the Job-shop Scheduling Problem (JSP). The JSP has been known as a complex and combinatorial optimization problem, and it has been proven to be NP-hard in \cite{baker1974introduction,lenstra1979computational}. Therefore, it is very difficult to reach optimal solution in a reasonable time, even for small instances. However, from a practical perspective, the JSP scale is large, where the number of operations can be up to 10000 in some workshops in \cite{zhang2010hybrid}. \\

\section{Project Objective}
The main objective is to find the most appropriate sequence of operations to optimize the typical performance indicator (makespan). Several approaches have been studied to solve JSP, and one of the most effective approaches used is decomposition in \cite{zhang2010hybrid,zhai2014decomposition}. The main idea of the decomposition is to split the problem into a series of sub-problems (windows) and solve each of them separately in a sequence, and then obtain the solution of the original problem. \\

This project intends to apply the clustering technique to decompose the problem into sub-problems, where the number of clusters corresponds to the number of sub-problems (windows), and the operations correspond to data objects. For the JSP, An instance consists of a set of jobs to be processed on machines. The number of jobs is $n$, and the number of machines is $m$. Each job contains $m$ operations with the operation precedence constraints. Each operation should be executed on machine $m$ for a certain amount of time. Since the problem at hand has precedence constraints, it’s required to apply a constrained clustering approach for the decomposition in \cite{wagstaff2001constrained}. \\

In the context of partitioning algorithms, instance-level constraints are helpful to express a priori knowledge about which operations should or should not be grouped. Consequently, two types of pairwise constraints will be considered:

\begin{enumerate}
    \item Must link constraints specify that two operations should be in the same window.
    \item Cannot link constraints specify that two operations must not be on the same window.
\end{enumerate}

It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations. The lack of these attributes guides me to extract more features that could specify which operations are similar and should be in the same window and which not that must not be in the same window. The main contribution of this project is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan. More specifically, I need to extract some features to define which operations are similar to be scheduled on the same window and get a near-optimal solution in a reasonable time.

\subsection{Feature extraction}
Some features could be extracted:
\begin{enumerate}
    \item The earliest starting time (EST) of an operation, the smaller the EST difference between the two operations, the more similarity.
    \item Work Remaining (WR) of a job, an operation that belongs to a higher WR job, will have a higher probability of being in an earlier window.
    \item Rank of operations sharing the same machine; the smaller difference the rank, the more similarity.
    \item Distance between operations sharing the same machine (using EST); the smaller distance, the more similarity.
\end{enumerate}

\section{Dataset Selection}
I select a set of benchmark instances generated in \cite{taillard1993benchmarks} as a dataset to work on. Many researchers have studied these instances and presented different approaches to solving them. I also developed different decomposition strategies to solve those. Therefore, it will be possible to measure the success/failure of my proposed model by comparing the results of the strategies developed in the past with those obtained from the proposed one.

\newpage
\bibliographystyle{ieeetr.bst}
\bibliography{references.bib}
\end{document}

### 2021-04-19 Artificial Intelligence and Machine Learning/Project proposal/Project_proposal_V1.tex
modify project proposal
worked on formulations

--- 2021-04-15 AI&ML/Project proposal/Project_proposal_V1.tex
@@
 
-\author{\\[1in] Student: \\ \textit{\textbf{Mohammed El-Kholany, MSc.}}\\Alpen-Adria-Universität Klagenfurt, Austria \\[1in] \and Supervisor:\\ \textit{\textbf{Assoc.
-Prof.
-Konstantin Schekotihin}} \\  \\[1in]}
+\author{\\[1in] Student: \\ \textit{\textbf{Mohammed El-Kholany, MSc.}}\\Alpen-Adria-Universität Klagenfurt, Austria \\[3in] }
 \date{April 2021}
@@
 Nowadays, the massive progress and development of the internet and online technologies, data generated by machines and devices, product development, quality and inventory management systems, or production planning systems has become huge and is expected to increase in the coming years.
-Hence, to capture long-term revenues and sustainable competitive advantages, companies must manage the knowledge and have the valuable information to make the right decision at the right time in \cite{benabdellah2019survey}.
+Hence, to capture long-term revenues and sustainable competitive advantages, companies must manage the knowledge and have the valuable information to make the right decision at the right time \cite{benabdellah2019survey}.
 \\
@@
 To extract implicit, unknown, and potentially useful information from data, we use data mining techniques that have been responsible for many of artificial intelligence’s recent successes.
-Clustering is one of these techniques; clustering has always been an exploratory but critical task in the knowledge discovery process, with applications ranging from image processing, production systems, and information retrieval in \cite{benabdellah2019survey}.
-Clustering is a technique that aims to group datasets (objects) into clusters by defining similar objects and aggregate them in the same cluster while the dissimilar objects should be belonging to different clusters.
+Clustering is one of these techniques; clustering has always been an exploratory but critical task in the knowledge discovery process, with applications ranging from image processing, production systems, and information retrieval \cite{benabdellah2019survey}.
+Clustering is a technique that aims to group datasets (objects) into clusters by defining similar objects and aggregate them in the same cluster while the dissimilar objects should belong to different clusters.
 \\
@@
 Furthermore, from an optimization perspective, the main objective is to minimize the distance between objects falling in the same cluster and maximize the distance between the others that belong to different clusters.
-Since clustering does not use a subset of the dataset as labels to learn a classification model, clustering differs from classification in \cite{wagstaff2001constrained}.
-In other words, with the terminology of machine learning, clustering is a form of an unsupervised task, which calculates the similarity between data objects without knowing the proper distribution in \cite{li2018geometric}.
-Due to these unsupervised characteristics, clustering is known as one of the most challenging tasks in \cite{benabdellah2019survey}.
+Since clustering does not use a subset of the dataset as labels to learn a classification model, clustering differs from classification \cite{wagstaff2001constrained}.
+In other words, with the terminology of machine learning, clustering is a form of an unsupervised task, which calculates the similarity between data objects without knowing the proper distribution~\cite{li2018geometric}.
+Due to these unsupervised characteristics, clustering is known as one of the most challenging tasks \cite{benabdellah2019survey}.
 \\
@@
 Scheduling of operations involves resource allocation over a period of time to perform a series of tasks, where one of the most critical problems is the Job-shop Scheduling Problem (JSP).
-The JSP has been known as a complex and combinatorial optimization problem, and it has been proven to be NP-hard in \cite{baker1974introduction,lenstra1979computational}.
-Therefore, it is very difficult to reach optimal solution in a reasonable time, even for small instances.
-However, from a practical perspective, the JSP scale is large, where the number of operations can be up to 10000 in some workshops in \cite{zhang2010hybrid}.
-\\
+The JSP has been known as a complex and combinatorial optimization problem, and it has been proven to be NP-hard \cite{baker1974introduction,lenstra1979computational}.
+Therefore, it is very difficult to reach an optimal solution in a reasonable time, even for small instances.
+However, from a practical perspective, the JSP scale is large, where the number of operations can be up to 10,000s in some workshops \cite{zhang2010hybrid}.
+% \\
 
@@
 The main objective is to find the most appropriate sequence of operations to optimize the typical performance indicator (makespan).
-Several approaches have been studied to solve JSP, and one of the most effective approaches used is decomposition in \cite{zhang2010hybrid,zhai2014decomposition}.
+Several approaches have been studied to solve JSP, and one of the most effective approaches used is decomposition \cite{zhang2010hybrid,zhai2014decomposition}.
 The main idea of the decomposition is to split the problem into a series of sub-problems (windows) and solve each of them separately in a sequence, and then obtain the solution of the original problem.
@@
 This project intends to apply the clustering technique to decompose the problem into sub-problems, where the number of clusters corresponds to the number of sub-problems (windows), and the operations correspond to data objects.
-For the JSP, An instance consists of a set of jobs to be processed on machines.
+For the JSP, an instance consists of a set of jobs to be processed on machines.
 The number of jobs is $n$, and the number of machines is $m$.
@@
 Each operation should be executed on machine $m$ for a certain amount of time.
-Since the problem at hand has precedence constraints, it’s required to apply a constrained clustering approach for the decomposition in \cite{wagstaff2001constrained}.
+Since the problem at hand has precedence constraints, it is required to apply a constrained clustering approach for the decomposition \cite{wagstaff2001constrained}.
 \\
@@
     \item Must link constraints specify that two operations should be in the same window.
-    \item Cannot link constraints specify that two operations must not be on the same window.
+    \item Cannot link constraints specify that two operations must not be in the same window.
 \end{enumerate}
@@
 It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations.
-The lack of these attributes guides me to extract more features that could specify which operations are similar and should be in the same window and which not that must not be in the same window.
+The lack of these attributes guides me to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window.
 The main contribution of this project is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan.
-More specifically, I need to extract some features to define which operations are similar to be scheduled on the same window and get a near-optimal solution in a reasonable time.
+More specifically, I need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time.
 
@@
     \item Work Remaining (WR) of a job, an operation that belongs to a higher WR job, will have a higher probability of being in an earlier window.
-    \item Rank of operations sharing the same machine; the smaller difference the rank, the more similarity.
+    \item Rank of operations sharing the same machine; the smaller difference in the rank, the more similarity.
     \item Distance between operations sharing the same machine (using EST); the smaller distance, the more similarity.
@@
 \section{Dataset Selection}
-I select a set of benchmark instances generated in \cite{taillard1993benchmarks} as a dataset to work on.
+I will use the benchmark instances presented in \cite{taillard1993benchmarks} as a dataset to work on.
 Many researchers have studied these instances and presented different approaches to solving them.
 I also developed different decomposition strategies to solve those.
-Therefore, it will be possible to measure the success/failure of my proposed model by comparing the results of the strategies developed in the past with those obtained from the proposed one.
+Therefore, it will be possible to measure the success/failure of my proposed model by comparing the results of the strategies developed in the past with those obtained from the proposed project.
 

### 2021-04-20 Artificial Intelligence and Machine Learning/Project proposal/Project_proposal_V1.tex
remove autogen

@@
 \usepackage[utf8]{inputenc}
+\usepackage{todonotes}
 \renewcommand{\baselinestretch}{1.2}
+
+\newcommand{\comment}[1]{\todo[inline]{#1}}
+
 \title{\textbf{A decomposition based on constrained clustering algorithms for job shop scheduling problems} \\}
@@
 Hence, to capture long-term revenues and sustainable competitive advantages, companies must manage the knowledge and have the valuable information to make the right decision at the right time \cite{benabdellah2019survey}.
-\\
 
@@
 To extract implicit, unknown, and potentially useful information from data, we use data mining techniques that have been responsible for many of artificial intelligence’s recent successes.
-Clustering is one of these techniques; clustering has always been an exploratory but critical task in the knowledge discovery process, with applications ranging from image processing, production systems, and information retrieval \cite{benabdellah2019survey}.
-Clustering is a technique that aims to group datasets (objects) into clusters by defining similar objects and aggregate them in the same cluster while the dissimilar objects should belong to different clusters.
-\\
+Clustering is one of these techniques, which is applied whenever the extracted data is not labeled, i.e., the semantics of this data is with respect to the application is unclear.
+Therefore, clustering has always been an exploratory but critical task in the knowledge discovery process, with applications ranging from image processing, production systems, and information retrieval \cite{benabdellah2019survey}.
+Clustering is a technique that aims to partition a dataset (objects) into subsets by identifying similar objects and aggregating them in the same cluster while the dissimilar objects should belong to different clusters.
+
+\comment{The next paragraph jumps to a distance, which is one of the similarity measures in the Euclidean space. One should explain this transition.}
 
@@
 Since clustering does not use a subset of the dataset as labels to learn a classification model, clustering differs from classification \cite{wagstaff2001constrained}.
-In other words, with the terminology of machine learning, clustering is a form of an unsupervised task, which calculates the similarity between data objects without knowing the proper distribution~\cite{li2018geometric}.
-Due to these unsupervised characteristics, clustering is known as one of the most challenging tasks \cite{benabdellah2019survey}.
-\\
+In other words, with the terminology of machine learning, clustering is a form of an unsupervised task, which calculates the similarity between data objects without knowing the proper attribution~\cite{li2018geometric}.
+Due to these unsupervised characteristics, clustering is known as one of the most challenging machine learning tasks \cite{benabdellah2019survey}.
 
 Clustering has been widely applied in several disciplines; one is production-planning systems, more specifically scheduling. 
+\comment{Are there any citations for the claim above?}
 Scheduling is one of the most complex problems in the industry.
@@
 However, from a practical perspective, the JSP scale is large, where the number of operations can be up to 10,000s in some workshops \cite{zhang2010hybrid}.
-% \\
 
@@
 The main idea of the decomposition is to split the problem into a series of sub-problems (windows) and solve each of them separately in a sequence, and then obtain the solution of the original problem.
-\\
 
@@
 Since the problem at hand has precedence constraints, it is required to apply a constrained clustering approach for the decomposition \cite{wagstaff2001constrained}.
-\\
 
@@
 The main contribution of this project is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan.
+\comment{Both I and we are used in the proposal, please select only one.}
 More specifically, I need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time.
@@
 
+\comment{Maybe a good idea would be to consider also some real-world instances just to test the selected similarity function and clustering algorithms.}
+
 \newpage

### 2021-04-21 Artificial Intelligence and Machine Learning/Project proposal/Project_proposal_V2.tex
updated proposal

--- 2021-04-20 Artificial Intelligence and Machine Learning/Project proposal/Project_proposal_V1.tex
@@
 
+The data sets could be composed entirely of either numeric features or symbolic features.
+For the numeric feature, we will use the Euclidean distance metric.
+However, for symbolic features, the Hamming distance is computed.
+Since our problem contains only numeric features, we will use Euclidean distance to determine the distance between the objects \cite{aloise2009np, wagstaff2001constrained,macqueen1967some}.
 \comment{The next paragraph jumps to a distance, which is one of the similarity measures in the Euclidean space. One should explain this transition.}
@@
 
-Clustering has been widely applied in several disciplines; one is production-planning systems, more specifically scheduling. 
+Clustering has been widely applied in several disciplines; one is production-planning systems~\cite{nananukul2013clustering,koskosidis1992clustering}, more specifically scheduling \cite{yashar2013multi,tong2016research}.
 \comment{Are there any citations for the claim above?}
@@
 It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations.
-The lack of these attributes guides me to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window.
+The lack of these attributes guides us to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window.
 The main contribution of this project is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan.
 \comment{Both I and we are used in the proposal, please select only one.}
-More specifically, I need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time.
+More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time.
 
@@
 \section{Dataset Selection}
-I will use the benchmark instances presented in \cite{taillard1993benchmarks} as a dataset to work on.
+We will use the benchmark instances presented in \cite{taillard1993benchmarks} as a dataset to work on.
 Many researchers have studied these instances and presented different approaches to solving them.
-I also developed different decomposition strategies to solve those.
-Therefore, it will be possible to measure the success/failure of my proposed model by comparing the results of the strategies developed in the past with those obtained from the proposed project.
+We also developed different decomposition strategies to solve those.
+Therefore, it will be possible to measure the success/failure of our proposed model by comparing the results of the strategies developed in the past with those obtained from the proposed project.
+Since the benchmark instances are randomly generated, they do not represent a real-life application.
+Hence, We will consider some real-world instances to test our proposed model concerning the selected similarity function and clustering algorithms.
 

### 2021-08-09 Artificial Intelligence and Machine Learning/Project report/Project_report_V2.tex
Project report draft

\documentclass{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{todonotes}
\usepackage{textcomp}
\usepackage{color,soul}
%\usepackage{array}
%\newcolumntype{P}[1]{>{\raggedleft\arraybackslash}p{#1}}

\usepackage{geometry}
\usepackage{changepage}
%\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
%\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
\renewcommand{\baselinestretch}{1.2}

\newcommand{\comment}[1]{\todo[inline]{#1}}

\title{\textbf{A decomposition based on constrained clustering algorithms for job shop scheduling problems} \\}

\author{\\[1in] Student: \\ \textit{\textbf{Mohammed El-Kholany, MSc.}}\\Alpen-Adria-Universität Klagenfurt, Austria \\[3in] }
\date{August 2021}

\begin{document}

\clearpage\maketitle
\thispagestyle{empty}
\clearpage
\setcounter{page}{1}
\newpage
\section{Introduction}
Nowadays, the massive progress and development of the internet and online technologies, data generated by machines and devices, product development, quality and inventory management systems, or production planning systems has become huge and is expected to increase in the coming years. Hence, to capture long-term revenues and sustainable competitive advantages, companies must manage the knowledge and have the valuable information to make the right decision at the right time \cite{benabdellah2019survey}. \\

It can be assumed that knowledge management is a crucial issue in the industry. To extract implicit, unknown, and potentially useful information from data, we use data mining techniques that have been responsible for many of artificial intelligence’s recent successes. 
Clustering is one of these techniques, which is applied whenever the extracted data is not labeled, i.e., the semantics of this data is with respect to the application is unclear. Therefore, clustering has always been an exploratory but critical task in the knowledge discovery process, with applications ranging from image processing, production systems, and information retrieval \cite{benabdellah2019survey}. Clustering is a technique that aims to partition a dataset (objects) into subsets by identifying similar objects and aggregating them in the same cluster while the dissimilar objects should belong to different clusters.\\

The data sets could be composed entirely of either numeric features or symbolic features. For the numeric feature, we will use the Euclidean distance metric. However, for symbolic features, the Hamming distance is computed. Since our problem contains only numeric features, we will use Euclidean distance to determine the distance between the objects \cite{aloise2009np, wagstaff2001constrained,macqueen1967some}.\\
%\comment{The next paragraph jumps to a distance, which is one of the similarity measures in the Euclidean space. One should explain this transition.}

Furthermore, from an optimization perspective, the main objective is to minimize the distance between objects falling in the same cluster and maximize the distance between the others that belong to different clusters. Since clustering does not use a subset of the dataset as labels to learn a classification model, clustering differs from classification \cite{wagstaff2001constrained}. In other words, with the terminology of machine learning, clustering is a form of an unsupervised task, which calculates the similarity between data objects without knowing the proper attribution~\cite{li2018geometric}. Due to these unsupervised characteristics, clustering is known as one of the most challenging machine learning tasks \cite{benabdellah2019survey}.

Clustering has been widely applied in several disciplines; one is production-planning systems~\cite{nananukul2013clustering,koskosidis1992clustering}, more specifically scheduling \cite{yashar2013multi,tong2016research}. 
%\comment{Are there any citations for the claim above?}
Scheduling is one of the most complex problems in the industry. Scheduling of operations involves resource allocation over a period of time to perform a series of tasks, where one of the most critical problems is the Job-shop Scheduling Problem (JSP). The JSP has been known as a complex and combinatorial optimization problem, and it has been proven to be NP-hard \cite{baker1974introduction,lenstra1979computational}. Therefore, it is very difficult to reach an optimal solution in a reasonable time, even for small instances. However, from a practical perspective, the JSP scale is large, where the number of operations can be up to 10,000s in some workshops \cite{zhang2010hybrid}.

\section{Literature Review}
In this section, we will review some of the most related works about data mining approaches employed to solve scheduling problems. Over the last decades, solving the scheduling problem has attracted the interest of many researchers, and a wide range of approaches has been developed for solving the scheduling problem. One of the approaches that have been widely used to solve the problem is data mining. The main idea of data mining is to explore the patterns from the obtained solutions by some heuristics and then develop some decision rules that approximate the order of the operations in each machine \cite{ismail2012production}. More precisely, it aims to get an order of the operations according to some characteristics.\\

Data mining methodologies are applied to explore the patterns in data generated by Genetic Algorithm (GA) \cite{koonce2000using}. This work aimed to generate rules to determine the priority for each job with weight. They solved the scheduling problem using GA to create some rules. Some characteristics (rank of operations, processing time, machine load, and remaining processing time) have been considered. In order to evaluate their heuristic, they tested the generated rules on some other instances that were randomly generated. The results showed that the generated rules were able to outperform the shortest processing time heuristic. One of the drawbacks is that the problem instances considered contain only 36 operations.\\

Other researchers also implemented GA to get some information of the solution but considered other features to generate decision rules \cite{harrath2002genetic}. They considered processing time, job length, remaining processing time, the rank of the operation, and machine load. To build a generic model for scheduling problems, they discretized the data by converting the quantitative data into qualitative. They sorted the operations share the same machine according to processing time. If two or more operations have the same processing time, they look at the time length of the job. If the job length is the same, sort them out based on the remaining processing time.\\

The work mentioned above had considered static JSP. However, dynamic JSP became one of the most crucial problems in real-life applications. New dispatching rules for JSP are presented using a data mining approach to obtain an efficient schedule in a dynamic Job-shop Scheduling environment \cite{shahzad2010discovering}. This study aimed to minimize makespan by creating new dispatching rules. Firstly, they solved the scheduling problems using Tabu Search and used the obtained solution as a training set to get a rule-set capable of approximating efficient solutions. They considered some other features like (Earliest possible Starting Time EST, the difference in remaining processing time for a job, and the difference in the processing time of two operations) to create a rule-set for dispatching the operations. The experiments showed that the results of mined rule-set are superior or comparable to the heuristics in the literature.\\

The variable neighborhood search (VNS) is combined with the k-means algorithm to address Dynamic JSP \cite{adibi2014clustering}. The proposed algorithm depends on extracting information from the problem's input data when an event like (machine breakdown or adding new job(s)) occurred. At any rescheduling point, cluster analysis is performed to replace jobs according to their distances. In other words, a job that belongs to a farther cluster has a greater probability to be selected in making a new adjacent than the other belonging to a closer cluster. The proposed method was compared to VNS and some common heuristics using a simulated job shop. the results indicated that the performance of the proposed model is better than the others, including some heuristics (FIFO, LIFO, and SPT).\\

Another study has presented a data mining approach to generate an initial population for population-based heuristics/metaheuristics to solve JSP \cite{nasiri2019data}. They extracted information from optimal or near-optimal solutions obtained by a heuristic procedure called "Assignment Procedure" to build decision rules to determine the priority of the operations. Once the rules are extracted, the initial solution is generated using these rules. To evaluate the performance of the proposed method, they performed a series of computational experiments on several test instances. The results showed that the proposed method with data mining generated initial population outperforms the methods with randomly generated initial population.\\

From the literature, we have found that most of the studies which used data mining techniques aimed to extract some features either from the problem instances or solutions obtained by some heuristics to build a decision-rules to order the operations assigned to a machine. Our study aims to  find the most appropriate sequence of operations to optimize the typical performance indicator (makespan). Several approaches have been studied to solve JSP, and one of the most effective approaches used is decomposition \cite{zhang2010hybrid,zhai2014decomposition}. The main idea of the decomposition is to split the problem into a series of sub-problems (windows) and solve each of them separately in a sequence, and then obtain the solution of the original problem.\\

This work intends to apply the clustering techniques to decompose the problem into sub-problems, where the number of clusters corresponds to the number of sub-problems (windows), and the operations correspond to data objects. For the JSP, an instance consists of a set of jobs to be processed on machines. The number of jobs is $n$, and the number of machines is $m$. Each job contains $m$ operations with the operation precedence constraints. Each operation should be executed on machine $m$ for a certain amount of time. Since the problem at hand has precedence constraints, it is required to apply a constrained clustering approach for the decomposition \cite{wagstaff2001constrained}.\\

It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations. The lack of these attributes guides us to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window. The main contribution of this work is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan. 
More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time.

This work consists of two main phases:

\begin{enumerate}
    \item The first phase is to apply data mining to predict the order of each operation on its machine according to some features. In this phase, we will try to get all possible features extracted from the problem itself, which significantly affect the quality of the solution. In addition, we will use some heuristics such as (FIFO, MTWR, EST) to learn the pattern of the solutions obtained by those.

    \item The second phase aims to develop, implement and test a proposed clustering algorithm that will split the problem into sub-problems where each operation is a data object, and each cluster will represent a time window. It will consider the precedence constraints between the operations and balancing between the clusters according to the number of operations per cluster.
\end{enumerate}

\section{Features extraction}
In this section, we will show which features have been extracted and how did we extract them. Based on the previous work, \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data}, the following features have been identified: priority, processing time, remaining processing time, machine load, and route position. Most of the researchers converted the attributes into qualitative data. As mentioned in these studies, the main reason for this conversion is to make the decision rules more generic, which can be applied to several scheduling problems. However, in our case, we will not convert the data into classes because we will use these features to measure the similarities/dissimilarities between the operations in the clustering process, which requires having numerical values.
We will show the features that we considered:
\subsection{\textit{Process}}
\textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are attributes dependent on the problem domain. The \textit{PT} represents the execution time of a particular operation in a machine. However, \textit{RPT} is the aggregate processing time for all subsequent operations of that job.

\subsection{\textit{Operation}}
The \textit{Operation} attribute is an ordinal value to represent the rank of each operation in its job.

\subsection{\textit{Time Length of a Job}(TLJ)}
\textit{TLJ} is an attribute that describes the total processing time needed to complete a particular job. It is the summation of processing time for all operations of that job.

\subsection{\textit{Earliest Starting Time} (EST)}
The \textit{EST} attribute represents the earliest possible time for a particular operation to be executed. It is calculated by accumulating the processing time of the predecessor(s) of that operation, and it is $Zero$, if the operation has no predecessor.

\subsection{\textit{Machine Load} (ML)}
\textit{ML} is a property of a machine that shows how much time a particular machine has to execute all the operations assigned to it. In the initial state, the \textit{ML} is the aggregate of processing time of all operations assigned to that machine. For each step, the load of a machine is decreased by the processing time of an operation assigned to the machine and has the smallest \textit{EST}.

\subsection{\textit{Starting Time} (ST)}
\textit{ST} attribute represents the starting execution time of each operation by solving the problem using heuristics FIFO, MTWR and EST developed by \cite{tassel2021reinforcement,el2020job}.

\subsection{\textit{Waiting Time} (WT)}
Since we used three different heuristics to mine information from the obtained solution, we will have three waiting times. \textit{WT} attribute is the difference between the starting time of a particular operation and the maximum value between the end time of its predecessor and when the machine became free.

\section{Proposed Algorithm}
In this section, we will show the implementation of our proposed model and how it works. The main idea of the proposed model is to split the problem into sub-problems, solve them sequentially and integrate the solutions to have the solution of the main problem. We aim to decompose the problem into time windows because this technique proved its effectiveness for solving the scheduling problems. Several ways have been presented during the last decades to decompose the problem  \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.\\

The proposed model aims to apply clustering methods where each cluster will represent a time window, and each operation is a data object. The clustering algorithms are used to group a set of data objects into one cluster. It depends on measuring the similarity/dissimilarity between the data objects according to the features of these data objects. The similar objects will be in one cluster, and the dissimilar are in different clusters. The difference between the basic clustering and our proposed is the ordering of the clusters. In general, the cluster number does not have meaning; it just means that the data objects grouped in a particular cluster are similar. However, in our case, the cluster number is meaningful because the data objects grouped in cluster $i$ should be scheduled before other data objects in cluster $i+1$. More specifically, it is not allowed to insert a data object (\textit{an operation}) in a particular cluster, and the predecessor of that operation is not assigned yet to the same cluster or the preceding cluster(s). Since we will solve each window separately, we should consider balancing the clusters concerning the number of operations assigned to each cluster. So that, we will have two types of constraints:
\begin{enumerate}
\item The predecessor(s) of an operation must be inserted into the same or previous cluster(s).
\item The number of operations assigned to each cluster should be close to each other.
\end{enumerate}

Figure 1 shows a flowchart of the proposed algorithm. As shown in the flowchart, the number of clusters is given. We calculate the cluster capacity by dividing the total number of operations by the number of clusters. In the next step, the clustering process started by generating the centroids, where the number of centroids is the number of clusters. In general, the centroids are generated randomly as well as in our model. However, we selected some operations as centroids. Although the operations are selected randomly, we followed a rule to select the centroids. For example, let us assume that a job consists of 15 operations, and three centroids should be generated. The first centroid will be an operation of rank 1 - 5, the second centroid will be 6 - 10, and the third will be from the last five operations. We followed this strategy to make them far from each other because the operations grouped in the first clusters should be at the beginning of their jobs. After generating the centroids, we perform a double-check for exceeding the number of clusters and the cluster capacity, respectively. After that, we calculate the distance between each operation and the centroid of the current cluster using Euclidean distance and assign the nearest operation to that cluster. In order to satisfy the precedence constraint, we check if the predecessor is not assigned to a cluster, it will be assigned to the current cluster, and the centroid is updated. If the maximum number of operations is reached, we will go to the next cluster. After assigning all operations into clusters, the algorithm terminates.

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{Flow_chart_9.png}
  \caption{Flowchart of the proposed Clustering Algorithm.}
  \label{fig:Flowchart of the proposed Clustering Algorithm}
\end{figure}

\section{Results}
We use the benchmark instances presented in \cite{taillard1993benchmarks} as a dataset to work on. Many researchers have studied these instances and presented different approaches to solving them. We also developed different decomposition strategies to solve those. Therefore, it will be possible to measure the success/failure of our proposed model by comparing the results of the strategies developed in the past with those obtained from the proposed algorithm. There are 12 features, as we described in the previous section. In order to know which features have a significant effect on the results, we did some combinations of the features to perform the decomposition process clustering-based and then solved the problem using the Answer Set Programming model developed in \cite{el2020job} with a timeout of 10 minutes. The combinations of the selected features are listed as follows:
\begin{enumerate}
\item \textbf{F1} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST \}}
\item \textbf{F2} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT \}} 
\item \textbf{F3} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, EST \}} 
\item \textbf{F4} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, ML \}} 
\item \textbf{F5} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT, EST \}}
\item \textbf{F6} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT, ML \}} 
\item \textbf{F7} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, EST, ML \}} 
\item \textbf{F8} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST \}} 
\item \textbf{F9} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT \}} 
\item \textbf{F10} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, EST \}} 
\item \textbf{F11} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, ML \}}
\item \textbf{F12} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, EST \}}
\item \textbf{F13} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, ML \}}
\item \textbf{F14} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, EST, ML \}}
\item \textbf{F15} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, EST, ML \}}
\end{enumerate}
We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments. The following table shows the selected features, and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}. The following table shows the results of our experiments. The first column represents the instances we tested, and the next three columns show the makespan values obtained by FIFO, MTWR, and EST heuristics. From the fifth column, we show the makespan values obtained by combining the features mentioned above. Our experiments found that the first five combinations do not provide good solutions compared to the other heuristics. However, features (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}, \textbf{F14} and \textbf{F15}) have a good impact in the cost function in the most cases. Especially, combinations \textbf{F6} and \textbf{F15} provided good results. It means that the most important features extracted from the problem itself and have a positive impact on the performance are \textit{Remaining Processing Time, Machine Load and Earliest Starting Time}. Also, The information obtained from the solutions of the other heuristics like (\textit{Starting Time} and \textit{Waiting Time}) led us to get better decomposition for the problem at hand. More specifically, combining \textit{Remaining Processing Time} with \textit{Machine Load} provides a good decomposition with/without \textit{Waiting Time} attributes obtained by the other heuristics.
\begin{table}[h!]
\setlength{\tabcolsep}{4.0pt}
  \begin{center}
    \caption{The performance of selected features.}
    \label{tab:table1}
 \begin{adjustwidth}{-3cm}{}
    \begin{tabular}{c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c}
      \textbf{Instances} & \textbf{FIFO} & \textbf{MTWR} & \textbf{EST} & \textbf{F1} & \textbf{F2} & \textbf{F3} & \textbf{F4} & \textbf{F5} & \textbf{F6} & \textbf{F7} & \textbf{F8} & \textbf{F9} & \textbf{F10} & \textbf{F11} & \textbf{F12} & \textbf{F13} & \textbf{F14} & \textbf{F15}\\
      \hline
      TA51 &3549  &3364  &3833  &3506  &3362  &3324  &3360  &3346  &\textbf{3294}  &3328  &3588  &3588  &3568  &3514  &3682  &3624  &3554  &3509 \\
      TA52 &3339  &3304  &3594  &3277  &3318  &3330  &3286  &3243  &3275  &3226  &3397  &3373  &3533  &3256  &3494  &3505  &3373  &\textbf{3184} \\
      TA53 &3160  &3168  &3443  &3382  &3585  &3347  &3543  &3517  &3250  &3338  &3251  &3522  &3373  &3585  &\textbf{3143}  &3427  &3524  &3380 \\
      TA54 &3218  &3494  &3439  &3414  &3425  &3425  &3503  &3397  &3265  &3353  &3563  &3443  &3386  &3352  &3270  &3236  &3232  &3289 \\
      TA55 &3291  &3237  &3784  &3308  &3441  &3424  &3319  &3355  &3386  &3301  &3498  &3385  &3530  &3365  &3309  &3370  &3385  &\textbf{3218} \\
      TA56 &3325  &3287  &3552  &3353  &3380  &3548  &3270  &3383  &3366  &3564  &\textbf{3229}  &3306  &3362  &3344  &3332  &3338  &3483  &3389 \\
      TA57 &3654  &3633  &3681  &3605  &3573  &3601  &3583  &3650  &\textbf{3480}  &3592  &3621  &3543  &3608  &3670  &3632  &3749  &3573  &\textbf{3480} \\
      TA58 &3299  &3591  &3764  &3352  &3412  &3370  &3509  &3707  &\textbf{3324}  &3496  &3517 &3716 &3644  &3611  &3471  &3437  &3365   &3481 \\
      TA59 &3344  &3394  &3661  &3453  &3416  &3301  &3339  &3508  &3430  &3352  &3258 &3381  &3402  &\textbf{3145}  &3250  &3253  &3284  &3290 \\
      TA60 &3129  &3257  &3399  &3483  &\textbf{3315}  &3617  &3563  &3591  &3327  &3399  &3374  &3526  &3637  &3425  &3676  &3659  &3767  &3569 \\
    \end{tabular}
\end{adjustwidth}
  \end{center}

\end{table}

\section{Conclusion}
Data mining and machine learning is an area of interest in many manufacturing companies and other fields. In order to increase productivity and decrease the operation cost, there should be high-quality schedules to be used. This study focused on data mining and clustering approaches for solving the Job-shop Scheduling Problem. The first issue in this work is to explore the knowledge in good schedules and try to use this knowledge to get more efficient schedules. We extracted some features from the problem at hand and other features from schedules obtained by heuristics. The second issue is finding a powerful way to decompose the problem into windows where each cluster is a window and solve them sequentially. In order to split the problem, we proposed a constrained clustering algorithm considering the precedence constraints of the JSP. We tested our model on benchmark instances, and the results showed that the performance of the proposed model outperforms other heuristics FIFO, MTWR, and EST; and some attributes like (\textit{Starting Time, Waiting Time, Remaining Processing Time, and Machine Load}) played an important role to split the operations in a good way. There are several ways to extend this study. Firstly, solve bigger instances and others in a real-life application in which more features could be extracted. Secondly, develop other clustering algorithms to split the problem more efficiently and get better results.

\newpage
\bibliographystyle{ieeetr.bst}
\bibliography{references.bib}
\end{document}

### 2021-08-18 JSP Encoding/Multi-shot/Statis_assignment_compression/dlO_timeout_whole_window_comp.py
adding a paper draft of  PADL

@@ -5,7 +5,7 @@ import theory
 import time
 from clingo import Function
 
-NUM_OF_TIME_WINDOWS = 2
+NUM_OF_TIME_WINDOWS = 4
 MAX_TIMEOUT = 600
 
 class Application:

### 2021-08-18 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
adding a paper draft

--- 2021-08-09 Artificial Intelligence and Machine Learning/Project report/Project_report_V2.tex
@@
-\documentclass{article}
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
+%
+\documentclass[runningheads]{llncs}
+%
 \usepackage{graphicx}
-\usepackage[utf8]{inputenc}
-\usepackage{todonotes}
-\usepackage{textcomp}
-\usepackage{color,soul}
-%\usepackage{array}
-%\newcolumntype{P}[1]{>{\raggedleft\arraybackslash}p{#1}}
-
-\usepackage{geometry}
 \usepackage{changepage}
-%\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-%\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
-\renewcommand{\baselinestretch}{1.2}
-
-\newcommand{\comment}[1]{\todo[inline]{#1}}
-
-\title{\textbf{A decomposition based on constrained clustering algorithms for job shop scheduling problems} \\}
-
-\author{\\[1in] Student: \\ \textit{\textbf{Mohammed El-Kholany, MSc.}}\\Alpen-Adria-Universität Klagenfurt, Austria \\[3in] }
-\date{August 2021}
+\usepackage{cite}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
 \begin{document}
+%
+\title{A decomposition based on constrained clustering algorithms for job shop scheduling problems\thanks{Supported by organization x.}}
+%
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-1111-2222-3333} \and
+Second Author\inst{1}\orcidID{1111-2222-3333-4444} \and
+Third Author\inst{1,2}\orcidID{2222--3333-4444-5555}}
+%
+\authorrunning{M. El-Kholany et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
+\email{\{mohammed.el-kholany, konstantin.schekotihin and martin.gebser\}@aau.at}\\
+%\url{http://www.springer.com/gp/computer-science/lncs} \and
+Technische Universität Graz, Graz, Austria\\
+\email{mgebser@ist.tugraz.at}}
+%
+\maketitle              % typeset the header of the contribution
+%
+\begin{abstract}
+Scheduling decision in production planning plays a significant role in the manufacturing industry.
+A good schedule may lead to solve production problems and improve productivity.
+Job-shop Scheduling Problem (JSP) is one of the most crucial problems in production systems.
+JSP deals with allocating tasks on resources and determining the sequence of the operations on each machine.
+The main objective of JSP is to complete the processing of all operations in a minimal time.
+One of the most effective approaches studied by many researchers for solving the JSP is decomposition.
+This work aims to apply the decomposition strategy to split the problem into a series of sub-problems using data mining methodologies.
+We will propose a clustering algorithm to group the data points (operations) into clusters, where each cluster is a Time Window.
+The decomposition process depends on two main phases; the first phase is to extract features to predict the sequence of the operations on each machine.
+These features are extracted either from the problem itself or from solutions obtained by other heuristics.
+The second phase is to develop a constrained clustering algorithm to assign each data object (an operation) into a cluster.
+We solved the problem using the Answer Set Programming model, and we tested our proposed algorithm on a set of benchmark instances, and the results showed that our proposed outperformed other heuristics in most cases.
+The results also showed that some features like \textit{Remaining Processing Time}, \textit{Machine Load} and \textit{Earliest starting Time} have a significant impact on the solution quality.
 
-\clearpage\maketitle
-\thispagestyle{empty}
-\clearpage
-\setcounter{page}{1}
-\newpage
+\keywords{Job-shop Scheduling  Problem \and Constrained Clustering \and Time Windows \and Answer Set Programming.}
+\end{abstract}
+%
+%
+%
 \section{Introduction}
@@
 Due to these unsupervised characteristics, clustering is known as one of the most challenging machine learning tasks \cite{benabdellah2019survey}.
+\\
 
@@
 \section{Literature Review}
+
 In this section, we will review some of the most related works about data mining approaches employed to solve scheduling problems.
@@
 The main contribution of this work is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan.
-More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time.
+More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time. \\
 
@@
 \section{Proposed Algorithm}
+
 In this section, we will show the implementation of our proposed model and how it works.
@@
 
-Figure 1 shows a flowchart of the proposed algorithm.
+Fig.~\ref{fig1} shows a flowchart of the proposed algorithm.
 As shown in the flowchart, the number of clusters is given.
@@
   \caption{Flowchart of the proposed Clustering Algorithm.}
-  \label{fig:Flowchart of the proposed Clustering Algorithm}
+  \label{fig1}
 \end{figure}
@@
 We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments.
-The following table shows the selected features, and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
+The following Table~\ref{tab1} shows the selected features, and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
 The following table shows the results of our experiments.
@@
     \caption{The performance of selected features.}
-    \label{tab:table1}
- \begin{adjustwidth}{-3cm}{}
+    \label{tab1}
+ \begin{adjustwidth}{-3.7cm}{}
     \begin{tabular}{c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c}
@@
 
-\newpage
-\bibliographystyle{ieeetr.bst}
-\bibliography{references.bib}
+\bibliographystyle{splncs04}
+\bibliography{references}
+
 \end{document}

### 2021-08-20 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
updated version

@@
 Second Author\inst{1}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{1,2}\orcidID{2222--3333-4444-5555}}
+Third Author\inst{1}\orcidID{2222--3333-4444-5555}}
 %
@@
 This work aims to apply the decomposition strategy to split the problem into a series of sub-problems using data mining methodologies.
-We will propose a clustering algorithm to group the data points (operations) into clusters, where each cluster is a Time Window.
+We will propose a clustering algorithm to group the data objects (operations) into clusters, where each cluster is a Time Window.
 The decomposition process depends on two main phases; the first phase is to extract features to predict the sequence of the operations on each machine.
@@
 However, for symbolic features, the Hamming distance is computed.
-Since our problem contains only numeric features, we will use Euclidean distance to determine the distance between the objects \cite{aloise2009np, wagstaff2001constrained,macqueen1967some}.\\
+Since our problem contains only numeric features, we will use Euclidean distance to measure the distance between the objects \cite{aloise2009np, wagstaff2001constrained,macqueen1967some}.\\
 %\comment{The next paragraph jumps to a distance, which is one of the similarity measures in the Euclidean space. One should explain this transition.}

### 2021-09-22 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
Updating abstract

--- 2021-08-18 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 Job-shop Scheduling Problem (JSP) is one of the most crucial problems in production systems.
-JSP deals with allocating tasks on resources and determining the sequence of the operations on each machine.
-The main objective of JSP is to complete the processing of all operations in a minimal time.
+JSP deals with allocating tasks on resources and determining the sequence of the operations on each machine while completing all tasks as early as possible.
 One of the most effective approaches studied by many researchers for solving the JSP is decomposition.
-This work aims to apply the decomposition strategy to split the problem into a series of sub-problems using data mining methodologies.
-We will propose a clustering algorithm to group the data points (operations) into clusters, where each cluster is a Time Window.
+This work investigates applying the decomposition strategy to split the problem into a series of sub-problems using data mining methodologies.
+This study proposes a constrained clustering algorithm to group the data objects (operations) into clusters, where each cluster is a Time Window.
 The decomposition process depends on two main phases; the first phase is to extract features to predict the sequence of the operations on each machine.

### 2021-09-27 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
Modifying main issues

@@
 %
+\usepackage{helvet,times,courier}
 \usepackage{graphicx}
@@
 \usepackage{cite}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage{mathtools}
+\usepackage{rotating}
+\usepackage{caption} 
+\captionsetup[table]{skip=10pt}
+\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
 % Used for displaying a sample figure. If possible, figure files should
@@
 %
-\title{A decomposition based on constrained clustering algorithms for job shop scheduling problems\thanks{Supported by organization x.}}
+\title{A decomposition based on constrained clustering algorithms for JSP\thanks{Supported by organization x.}}
 %
@@
 %
-\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{1}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{1}\orcidID{2222--3333-4444-5555}}
+\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-0002-1088-2081} \and
+Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958} \and
+Martin Gebser\inst{1,2}\orcidID{0000--0000-0000-0000}}
 %
@@
 This work investigates applying the decomposition strategy to split the problem into a series of sub-problems using data mining methodologies.
-This study proposes a constrained clustering algorithm to group the data objects (operations) into clusters, where each cluster is a Time Window.
+This study proposes a constrained clustering algorithm to group the data objects (operations) into clusters, where each cluster is a Time Window (TW).
 The decomposition process depends on two main phases; the first phase is to extract features to predict the sequence of the operations on each machine.
@@
 The second phase is to develop a constrained clustering algorithm to assign each data object (an operation) into a cluster.
-We solved the problem using the Answer Set Programming model, and we tested our proposed algorithm on a set of benchmark instances, and the results showed that our proposed outperformed other heuristics in most cases.
-The results also showed that some features like \textit{Remaining Processing Time}, \textit{Machine Load} and \textit{Earliest starting Time} have a significant impact on the solution quality.
+We solved the problem using the Answer Set Programming model, and tested our proposed algorithm on a set of benchmark instances, and the results showed that our proposed outperformed other heuristics in most cases.
+The results also showed that some features like \textit{Remaining Processing Time}, \textit{Machine Load} and \textit{Earliest Starting Time} have a significant impact on the solution quality.
 
@@
 However, from a practical perspective, the JSP scale is large, where the number of operations can be up to 10,000s in some workshops \cite{zhang2010hybrid}.
+\\
+
+This work intends to apply the clustering techniques to decompose the problem into sub-problems, where the number of clusters corresponds to the number of sub-problems (windows), and the operations correspond to data objects.
+For the JSP, an instance consists of a set of jobs to be processed on machines.
+The number of jobs is $n$, and the number of machines is $m$.
+Each job contains $m$ operations with the operation precedence constraints.
+Each operation should be executed on machine $m$ for a certain amount of time.
+Since the problem at hand has precedence constraints, it is required to apply a constrained clustering approach for the decomposition \cite{wagstaff2001constrained}.
+Although many studies proposed different constrained clustering methods in \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}.
+However, most of those checked whether two data objects must be assigned into the same cluster.
+For the JSP, it is essential to consider the sequencing between the clusters to handle the precedence constraints.
+More specifically, our proposed method will not permit to assign an operation into a particular cluster if the predecessor of that operation is not yet assigned to the same or previous cluster(s).
+\\
+
+It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations.
+The lack of these attributes guides us to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window.
+The main contribution of this work is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan.
+More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time. \\
+
+This work consists of two main phases:
+
+\begin{enumerate}
+    \item The first phase is to apply data mining to predict the order of each operation on its machine according to some features.
+In this phase, we will try to get all possible features extracted from the problem itself, which significantly affect the quality of the solution.
+In addition, we will use some heuristics such as (FIFO, MTWR, EST) to learn the pattern of the solutions obtained by those.
+
+    \item The second phase aims to develop, implement and test a proposed clustering algorithm that will split the problem into sub-problems where each operation is a data object, and each cluster will represent a Time Window (TW).
+It will consider the precedence constraints between the operations and balancing between the clusters according to the number of operations per cluster.
+\end{enumerate}
+
+The paper is organized as follows.
+The following Section describes the Job-shop Scheduling Problem in detail and shows an illustrative example of input data.
+Section $3$ shows which features we extracted and how we extracted them, followed by Section $4$, which presents the proposed algorithm.
+Section $5$ presents an extensive computational study on benchmark instances, comparing our results with the others obtained by some heuristics.
+A relevant literature review on the subject is given in Section $6$ followed by the conclusion and future work.
+
+\section{Problem Formulation}
+The Job-shop Scheduling Problem (JSP) is one of the well-known hardest combinatorial optimization problems \cite{baker1974introduction,lenstra1979computational,garey1976complexity}.
+The JSP is given a set $J = \{ J_1, J_2,...,J_n \}$  of jobs and a set $M = \{ M_1, M_2,...,M_m \}$ of Machines.
+Each job $J_i$ has a series $  O_{i,1}, O_{i,2},...,O_{i,m} $ of operations that must be processed in a predefined order.
+Each machine $M_i$ executes an operation of each job with a known and fixed processing time.
+The main objective of the JSP is to minimize the total completion time for all jobs (makespan).
+\\ The JSP assumptions are listed as follows:
+
+\begin{enumerate}
+	\item Once an operation is started, it must be completed without interruption.
+	\item Each job visits a machine once.
+	\item A machine cannot perform more than one operation at a time.
+	\item A machine is not broken.
+	\item All jobs and machines are available at time 0.
+\end{enumerate}
+It is hard to reach a solution close to the optimal in a reasonable time, even if the number of operations to be scheduled is low.
+In real-life applications, the number of operations that are to be scheduled is up to $10,000$ \cite{zhang2010hybrid}, which makes the problem more complicated.
+The JSP has attracted the interest of many researchers in the last decades to find effective methods to solve this problem.
+One of these methods is decomposition.
+The decomposition idea is to split the problem into sub-problems, solve them separately, and then integrate those solutions to solve the whole problem.
+Many decomposition strategies have been proposed for solving the JSP  \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.
+However, there is no guarantee of which method is the best because the performance varies according to the problem.
+\\
+
+A problem data can be organized in a table, where rows correspond to operations and columns refer to the name of the operation, the machine assigned to an operation, and the processing time of a particular operation.
+Table~\ref{tab1} shows the information of a simple Job-shop Scheduling Problem which consists of $3$ jobs and $3$ machines.
+For instance, in the third row of the table, $O_{1,3}$ is the third operation of the first job that will be executed by machine $1$ with processing time $5$.
+
+\begin{table}[h!]
+\fontsize{12}{9}
+\setlength{\tabcolsep}{10.0pt}
+\centering
+  \begin{center}
+    \caption{An example of a Job-shop Scheduling Problem.}
+    \label{tab1}
+ %\begin{adjustwidth}{0.0cm}{}
+    \begin{tabular}{c c c}
+      \textbf{Operation} & \textbf{Machine} & \textbf{Procession Time} \\
+      \hline
+								\\
+      $O_{1,1}$  & 2  &  9  	\\
+      $O_{1,2}$  & 3  &  3  	\\
+      $O_{1,3}$  & 1  &  5  	\\
+								\\
+      $O_{2,1}$  & 3  & 4  	\\
+      $O_{2,2}$  & 2  & 6  	\\
+      $O_{2,3}$  & 1  & 2  	\\
+								\\
+      $O_{3,1}$  & 1  & 4  	\\
+      $O_{3,2}$  & 3  & 3  	\\
+      $O_{3,3}$  & 2  & 5  	\\
+    \end{tabular}
+
+%\end{adjustwidth}
+  \end{center}
+
+\end{table}
+In our example in Table~\ref{tab1}, there are $9$ operations and $3$ machines.
+We aim to split them efficiently into two parts and schedule them separately without violating the precedence constraint.
+More specifically, we should not assign an operation in the first TW, and its predecessor is assigned to the following TW(s).
+Table~\ref{tab2} shows how the output of the proposed method should look.
+As shown in the Table~\ref{tab2}, operations $\{ O_{1,1}, O_{1,2}, O_{2,1}, O_{2,2}, O_{3,1} \}$ are assigned to the first TW and the rest to the second TW.
+It means that the first TW will be optimized and then the upcoming TW(s).
+In this study, we investigate applying a constrained clustering algorithm to decompose the JSP, where we extract some features from the problem itself and the solutions of other heuristics.
+These features have been used for the splitting process, and we present them in detail in the following section.
+
+\begin{table}[h!]
+\fontsize{12}{9}
+\setlength{\tabcolsep}{10.0pt}
+\centering
+  \begin{center}
+    \caption{Time Window Assignment.}
+    \label{tab2}
+ %\begin{adjustwidth}{0.0cm}{}
+    \begin{tabular}{c  c }
+      \textbf{Operation} & \textbf{Time Window}  \\
+      \hline
+							\\
+      $O_{1,1}$  & 1    	\\
+      $O_{1,2}$  & 1    	\\
+      $O_{2,1}$  & 1    	\\
+      $O_{2,2}$  & 1   	\\
+      $O_{3,1}$  & 1		\\
+					    		\\
+      $O_{1,3}$  & 2  	\\
+      $O_{2,3}$  & 2		\\
+      $O_{3,2}$  & 2    	\\
+      $O_{3,3}$  & 2    	\\
+    \end{tabular}
+%\end{adjustwidth}
+  \end{center}
+
+\end{table}
+
+\section{Features extraction}
+In this section, we show which features have been extracted and how did we extract them.
+Based on the previous work, \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data}, the following features have been identified: priority, processing time, remaining processing time, machine load, and route position.
+Most of the researchers converted the attributes into qualitative data.
+As mentioned in these studies, the main reason for this conversion is to make the decision rules more generic, which can be applied to several scheduling problems.
+However, in our case, we will not convert the data into classes because we will use these features to measure the similarities/dissimilarities between the operations in the clustering process, which requires having numerical values.
+We present the features that we considered:
+\subsection{\textit{Process}}
+\textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are attributes dependent on the problem domain.
+The \textit{PT} represents the execution time of a particular operation in a machine.
+However, \textit{RPT} is the aggregate processing time for all subsequent operations of that job.
+
+\subsection{\textit{Operation}}
+The \textit{Operation} attribute is an ordinal value to represent the rank of each operation in its job.
+
+\subsection{\textit{Time Length of a Job} (TLJ)}
+\textit{TLJ} is an attribute that describes the total processing time needed to complete a particular job. It is the summation of processing time for all operations of that job.
+
+\subsection{\textit{Earliest Starting Time} (EST)}
+The \textit{EST} attribute represents the earliest possible time for a particular operation to be executed.
+It is calculated by accumulating the processing time of the predecessor(s) of that operation, and it is $Zero$, if the operation has no predecessor.
+
+\subsection{\textit{Machine Load} (ML)}
+\textit{ML} is a property of a machine that shows how much time a particular machine has to execute all the operations assigned to it.
+In the initial state, the \textit{ML} is the aggregate of processing time of all operations assigned to that machine.
+For each step, the load of a machine is decreased by the processing time of an operation assigned to the machine and has the smallest \textit{EST}.
+
+\subsection{\textit{Starting Time} (ST)}
+\textit{ST} attribute represents the starting execution time of each operation by solving the problem using heuristics FIFO, MTWR and EST developed by \cite{tassel2021reinforcement,el2020job}.
+
+\subsection{\textit{Waiting Time} (WT)}
+Since we used three different heuristics to mine information from the obtained solution, we will have three waiting times.
+\textit{WT} attribute is the difference between the starting time of a particular operation and the maximum value between the end time of its predecessor and when the machine became free.
+\\
+
+The following Table~\ref{tab3} shows a sample of the features that we extracted for the decomposition process.
+Each row in the table represents an operation, and it contains its value according to each feature.
+The second column shows the \textit{Remain Processing Time (RPT)} of a job.
+For instance, the first job $J_1$ consists of $3$ operations with a total processing time of $17$.
+In the beginning, the \textit{Remain Processing Time} of $O_{1,1}$ is $17$ because no operations are executed.
+After finishing the first operation of job $1$, the \textit{RPT} became $8$ because the processing time of the completed operation $O_{1,1}$ is excluded.
+When the second operation of job $1$ is completed, the \textit{RPT} of that job is $5$.
+\\
+
+The third column represents the \textit{Earliest Starting Time} which we described earlier.
+The last column is the \textit{Machine Load} which is the summation of the processing time for all operations assigned to a machine.
+For example, Machine $3$ has to process operations $\{ O_{1,2}, O_{2,1}, O_{3,2} \}$, the load of machine $3$ at time $0$ is 10.
+We assumed that the machine would start processing the operation $O_{2,1}$ that has the smallest \textit{EST}.
+After completing the operation $O_{2,1}$, the machine will target the next operation, which has the smallest value of \textit{EST} and the load of the machine is decreased by the processing time of the completed operation $O_{2,1}$, and so on till completing all the operations assigned to a particular machine.
+
+\begin{table}[h!]
+\fontsize{12}{9}
+\setlength{\tabcolsep}{10.0pt}
+\centering
+  \begin{center}
+  \caption{Time Window Assignment.}
+  \label{tab3}
+ %\begin{adjustwidth}{0.0cm}{}
+    \begin{tabular}{c  c  c  c}
+      \textbf{Operation} & \textbf{RPT} & \textbf{EST} & \textbf{ML}\\
+      \hline
+										\\
+      $O_{1,1}$  & 17 & 0   & 20	\\
+      $O_{1,2}$  & 8  & 9   & 3		\\
+      $O_{1,3}$  & 5  & 12 & 5		\\
+										\\
+      $O_{2,1}$  & 12 & 0   & 10	\\
+      $O_{2,2}$  & 8  & 4   & 11	\\
+      $O_{2,3}$  & 2  & 10 & 7		\\
+					    					\\
+      $O_{3,1}$  & 12 & 0  & 11	\\
+      $O_{3,2}$  & 8  & 4  & 6		\\
+      $O_{3,3}$  & 5  & 7  & 5		\\
+    \end{tabular}
+%\end{adjustwidth}
+
+  \end{center} 
+\end{table}
+
+\section{Proposed Algorithm}
+
+In this section, we will show the implementation of our proposed model and how it works.
+The main idea of the proposed model is to split the problem into sub-problems, solve them sequentially and integrate the solutions to have the solution of the main problem.
+We aim to decompose the problem into TWs because this technique proved its effectiveness for solving the scheduling problems \cite{zhang2010hybrid,zhai2014decomposition}.
+Several ways have been presented during the last decades to decompose the problem \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.\\
+
+The proposed model aims to apply clustering methods where each cluster will represent a TW, and each operation is a data object.
+The clustering algorithms are used to group a set of data objects into one cluster.
+It depends on measuring the similarity/dissimilarity between the data objects according to the features of these data objects.
+The similar objects will be in one cluster, and the dissimilar are in different clusters.
+The difference between the basic clustering and our proposed is the ordering of the clusters.
+In general, the cluster number does not have meaning; it just means that the data objects grouped in a particular cluster are similar.
+However, in our case, the cluster number is meaningful because the data objects grouped in a particular cluster $i$ should be scheduled before other data objects in the successor cluster(s).
+More specifically, it is not allowed to insert a data object (\textit{an operation}) in a particular cluster, and the predecessor of that operation is not assigned yet to the same cluster or the preceding cluster(s).
+Since we will solve each window separately, we should consider balancing the clusters concerning the number of operations assigned to each cluster.
+So that, we will have two types of constraints:
+\begin{enumerate}
+\item The predecessor(s) of an operation must be inserted into the same or previous cluster(s).
+\item The number of operations assigned to each cluster should be close to each other.
+\end{enumerate}
+
+Algorithm.~\ref{alg} shows the pseudocode of the proposed algorithm.
+As shown in the pseudocode, the number of clusters is given.
+We calculate the cluster capacity by dividing the total number of operations by the number of clusters.
+In the next step, the clustering process starts by generating the centroids, where the number of centroids is the number of clusters.
+In general, the centroids are generated randomly as well as in our model.
+However, we selected some operations as centroids.
+Although the operations are selected randomly, we followed a rule to select the centroids.
+For example, let us assume that a job consists of 15 operations, and the number of clusters is $3$.
+The first centroid will be an operation of rank 1 - 5, the second centroid will be 6 - 10, and the third will be from the last five operations.
+We followed this strategy to make them far from each other because the operations grouped in the first clusters should be at the beginning of their jobs.
+After generating the centroids, we check for exceeding the number of clusters and the cluster capacity, respectively.
+After that, we calculate the distance between each operation and the centroid of the current cluster using Euclidean distance and assign the nearest operation to that cluster.
+In order to satisfy the precedence constraint, we check if the predecessor is not assigned to a cluster, it will be assigned to the current cluster, and the centroid is updated.
+If the maximum number of operations is reached, we will go to the next cluster.
+After assigning all operations into clusters, the algorithm terminates.
+
+\begin{algorithm}
+\caption{Proposed Clustering Algorithm}\label{alg}
+\begin{algorithmic}
+\Require $num\_clus$
+%\Require $n \geq 0$
+%\Ensure $y = x^n$
+\State $count\_clus \gets 0$
+
+\State $clus\_capacity \gets \ceil*{\frac{num\_oper}{num\_clus}}$
+\State Generate centroids
+\While{$count\_clus < num\_clus$}
+    \State $count\_cap \gets 0$
+    \State $curr\_clus \gets count\_clus + 1$
+	\While{$count\_cap < clus\_capacity$}
+          \State Calculate distance between data objects and a centroid   \Comment{Using Euclidean distance}
+          \State $selected\_data\_obj \gets nearest\_data\_obj$
+          %\State $selected\_data\_obj \in curr\_clus$
+          \State $curr\_clus \gets selected\_data\_obj$    \Comment{Assigning an operation to a TW}
+
+          \If{$pred\_selected\_data\_obj \notin clus$}     \Comment{Satisfying the precedence constraint}
+              \State $curr\_clus \gets pred\_selected\_data\_obj$
+          \EndIf
+          \State $update\_centroid$
+          \State $count\_cap \gets count\_cap + 1 $
+	\EndWhile
+    \State $count\_clus \gets count\_clus + 1 $
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+
+\iffalse
+\begin{figure}[h!]
+  \includegraphics[width=\linewidth]{Flow_chart_9.png}
+  \caption{Flowchart of the proposed Clustering Algorithm.}
+  \label{fig1}
+\end{figure}
+\fi
+
+%\newpage
+
+\section{Results}
+We use the benchmark instances presented in \cite{taillard1993benchmarks} as a dataset to work on.
+Many researchers have studied these instances and presented different approaches to solving them.
+We also developed different decomposition strategies to solve those.
+Therefore, it will be possible to measure the success/failure of our proposed model by comparing the results of the strategies developed in the past with those obtained from the proposed algorithm.
+There are 12 features, as we described in the previous section.
+In order to know which features have a significant effect on the results, we did some combinations of the features to perform the decomposition process clustering-based and then solved the problem using the Answer Set Programming model developed in \cite{el2020job} with a timeout of 10 minutes.
+The combinations of the selected features are listed as follows:
+\begin{enumerate}
+\fontsize{9pt}{10pt}
+\item \textbf{F1} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST \}}
+\item \textbf{F2} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT \}} 
+\item \textbf{F3} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, EST \}} 
+\item \textbf{F4} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, ML \}} 
+\item \textbf{F5} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT, EST \}}
+\item \textbf{F6} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT, ML \}} 
+\item \textbf{F7} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, EST, ML \}} 
+\item \textbf{F8} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST \}} 
+\item \textbf{F9} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT \}} 
+\item \textbf{F10} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, EST \}} 
+\item \textbf{F11} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, ML \}}
+\item \textbf{F12} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, EST \}}
+\item \textbf{F13} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, ML \}}
+\item \textbf{F14} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, EST, ML \}}
+\item \textbf{F15} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, EST, ML \}}
+\end{enumerate}
+We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments.
+The following Table~\ref{tab4} shows the selected features and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
+The following table shows the results of our experiments.
+The first column represents the instances we tested, and the next four columns show the makespan values obtained by FIFO, MTWR, and (EST with single and multi-shot) heuristics.
+From the sixth column, we show the makespan values obtained by combining the features mentioned above.
+Our experiments found that the first five combinations do not provide good solutions compared to the other heuristics.
+However, features (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}, and \textbf{F15}) have a good impact in the cost function in the most cases.
+Especially, combinations \textbf{F6} and \textbf{F15} provided good results.
+It means that the most important features extracted from the problem itself and have a positive impact on the performance are \textit{Remaining Processing Time, Machine Load and Earliest Starting Time}.
+Also, The information obtained from the solutions of the other heuristics like (\textit{Starting Time} and \textit{Waiting Time}) led us to get better decomposition for the problem at hand.
+More specifically, combining \textit{Remaining Processing Time} with \textit{Machine Load} provides a good decomposition with/without \textit{Waiting Time} attributes obtained by the other heuristics.
+In a few instances, our proposed method did not outperform other heuristics, and the average makespan of \textbf{F6} was better than the average makespan obtained by the other heuristics and close to the FIFO makespan.
+The following section illustrates the difference between our proposed algorithm and the methods implemented by other researchers.
+
+%\begin{table}
+\begin{sidewaystable}
+\setlength{\tabcolsep}{5.5pt}
+  \begin{center}
+    \caption{The performance of selected features.}
+    \label{tab4}
+%\rotatebox{90}{
+ \begin{adjustwidth}{-3.5cm}{}
+    \begin{tabular}{c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c}
+
+\hline
+      \textbf{Instances} & \textbf{FIFO} & \textbf{MTWR} & \textbf{Single-shot} & \textbf{EST} & \textbf{F1} & \textbf{F2} & \textbf{F3} & \textbf{F4} & \textbf{F5} & \textbf{F6} & \textbf{F7} & \textbf{F8} & \textbf{F9} & \textbf{F10} & \textbf{F11} & \textbf{F12} & \textbf{F13} & \textbf{F14} & \textbf{F15}\\
+      \hline
+      TA51 &3549  &3364  &3632  &3506  &3506  &3362  &3324  &3360  &3346  &\textbf{3294}  &3328  &3588  &3588  &3568  &3514  &3682  &3624  &3554  &3509 \\
+      TA52 &3339  &3304  &3615  &3773  &3277  &3318  &3330  &3286  &3243  &3275  &3226  &3397  &3373  &3533  &3256  &3494  &3505  &3373  &\textbf{3184} \\
+      TA53 &3160  &3168  &3481  &3478  &3382  &3585  &3347  &3543  &3517  &3250  &3338  &3251  &3522  &3373  &3585  &\textbf{3143}  &3427  &3524  &3380 \\
+      TA54 &3218  &3494  &3462  &3497  &3414  &3425  &3425  &3503  &3397  &3265  &3353  &3563  &3443  &3386  &3352  &3270  &3236  &3232  &3289 \\
+      TA55 &3291  &3237  &3552  &3482  &3308  &3441  &3424  &3319  &3355  &3386  &3301  &3498  &3385  &3530  &3365  &3309  &3370  &3385  &\textbf{3218} \\
+      TA56 &3325  &3287  &3610  &3605  &3353  &3380  &3548  &3270  &3383  &3366  &3564  &\textbf{3229}  &3306  &3362  &3344  &3332  &3338  &3483  &3389 \\
+      TA57 &3654  &3633  &3778  &3753  &3605  &3573  &3601  &3583  &3650  &\textbf{3480}  &3592  &3621  &3543  &3608  &3670  &3632  &3749  &3573  &\textbf{3480} \\
+      TA58 &3299  &3591  &3718  &3731  &3352  &3412  &3370  &3509  &3707  &3324  &3496  &3517 &3716 &3644  &3611  &3471  &3437  &3365   &3481 \\
+      TA59 &3344  &3394  &3613  &3398  &3453  &3416  &3301  &3339  &3508  &3430  &3352  &3258 &3381  &3402  &\textbf{3145}  &3250  &3253  &3284  &3290 \\
+      TA60 &3129  &3257  &3550  &3247  &3483  &3315  &3617  &3563  &3591  &3327  &3399  &3374  &3526  &3637  &3425  &3676  &3659  &3767  &3569 \\
+\hline
+ Average  &3331  &3373  &3601  &3547  &3413  &3423  &3429  &3428  &3470  &3340  &3395  &3430  &3475  &3504  &3427  &3426  &3460  &3454  &3379 \\
+\hline
+    \end{tabular}
+%}
+\end{adjustwidth}
+  \end{center}
+\end{sidewaystable}
+%\end{table}
 
@@
 
-From the literature, we have found that most of the studies which used data mining techniques aimed to extract some features either from the problem instances or solutions obtained by some heuristics to build a decision-rules to order the operations assigned to a machine.
-Our study aims to  find the most appropriate sequence of operations to optimize the typical performance indicator (makespan).
-Several approaches have been studied to solve JSP, and one of the most effective approaches used is decomposition \cite{zhang2010hybrid,zhai2014decomposition}.
-The main idea of the decomposition is to split the problem into a series of sub-problems (windows) and solve each of them separately in a sequence, and then obtain the solution of the original problem.\\
-
-This work intends to apply the clustering techniques to decompose the problem into sub-problems, where the number of clusters corresponds to the number of sub-problems (windows), and the operations correspond to data objects.
-For the JSP, an instance consists of a set of jobs to be processed on machines.
-The number of jobs is $n$, and the number of machines is $m$.
-Each job contains $m$ operations with the operation precedence constraints.
-Each operation should be executed on machine $m$ for a certain amount of time.
-Since the problem at hand has precedence constraints, it is required to apply a constrained clustering approach for the decomposition \cite{wagstaff2001constrained}.\\
-
-It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations.
-The lack of these attributes guides us to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window.
-The main contribution of this work is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan.
-More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time. \\
-
-This work consists of two main phases:
-
-\begin{enumerate}
-    \item The first phase is to apply data mining to predict the order of each operation on its machine according to some features.
-In this phase, we will try to get all possible features extracted from the problem itself, which significantly affect the quality of the solution.
-In addition, we will use some heuristics such as (FIFO, MTWR, EST) to learn the pattern of the solutions obtained by those.
-
-    \item The second phase aims to develop, implement and test a proposed clustering algorithm that will split the problem into sub-problems where each operation is a data object, and each cluster will represent a time window.
-It will consider the precedence constraints between the operations and balancing between the clusters according to the number of operations per cluster.
-\end{enumerate}
-
-\section{Features extraction}
-In this section, we will show which features have been extracted and how did we extract them.
-Based on the previous work, \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data}, the following features have been identified: priority, processing time, remaining processing time, machine load, and route position.
-Most of the researchers converted the attributes into qualitative data.
-As mentioned in these studies, the main reason for this conversion is to make the decision rules more generic, which can be applied to several scheduling problems.
-However, in our case, we will not convert the data into classes because we will use these features to measure the similarities/dissimilarities between the operations in the clustering process, which requires having numerical values.
-We will show the features that we considered:
-\subsection{\textit{Process}}
-\textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are attributes dependent on the problem domain.
-The \textit{PT} represents the execution time of a particular operation in a machine.
-However, \textit{RPT} is the aggregate processing time for all subsequent operations of that job.
-
-\subsection{\textit{Operation}}
-The \textit{Operation} attribute is an ordinal value to represent the rank of each operation in its job.
-
-\subsection{\textit{Time Length of a Job}(TLJ)}
-\textit{TLJ} is an attribute that describes the total processing time needed to complete a particular job. It is the summation of processing time for all operations of that job.
-
-\subsection{\textit{Earliest Starting Time} (EST)}
-The \textit{EST} attribute represents the earliest possible time for a particular operation to be executed.
-It is calculated by accumulating the processing time of the predecessor(s) of that operation, and it is $Zero$, if the operation has no predecessor.
-
-\subsection{\textit{Machine Load} (ML)}
-\textit{ML} is a property of a machine that shows how much time a particular machine has to execute all the operations assigned to it.
-In the initial state, the \textit{ML} is the aggregate of processing time of all operations assigned to that machine.
-For each step, the load of a machine is decreased by the processing time of an operation assigned to the machine and has the smallest \textit{EST}.
-
-\subsection{\textit{Starting Time} (ST)}
-\textit{ST} attribute represents the starting execution time of each operation by solving the problem using heuristics FIFO, MTWR and EST developed by \cite{tassel2021reinforcement,el2020job}.
-
-\subsection{\textit{Waiting Time} (WT)}
-Since we used three different heuristics to mine information from the obtained solution, we will have three waiting times.
-\textit{WT} attribute is the difference between the starting time of a particular operation and the maximum value between the end time of its predecessor and when the machine became free.
-
-\section{Proposed Algorithm}
-
-In this section, we will show the implementation of our proposed model and how it works.
-The main idea of the proposed model is to split the problem into sub-problems, solve them sequentially and integrate the solutions to have the solution of the main problem.
-We aim to decompose the problem into time windows because this technique proved its effectiveness for solving the scheduling problems.
-Several ways have been presented during the last decades to decompose the problem  \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.\\
-
-The proposed model aims to apply clustering methods where each cluster will represent a time window, and each operation is a data object.
-The clustering algorithms are used to group a set of data objects into one cluster.
-It depends on measuring the similarity/dissimilarity between the data objects according to the features of these data objects.
-The similar objects will be in one cluster, and the dissimilar are in different clusters.
-The difference between the basic clustering and our proposed is the ordering of the clusters.
-In general, the cluster number does not have meaning; it just means that the data objects grouped in a particular cluster are similar.
-However, in our case, the cluster number is meaningful because the data objects grouped in cluster $i$ should be scheduled before other data objects in cluster $i+1$.
-More specifically, it is not allowed to insert a data object (\textit{an operation}) in a particular cluster, and the predecessor of that operation is not assigned yet to the same cluster or the preceding cluster(s).
-Since we will solve each window separately, we should consider balancing the clusters concerning the number of operations assigned to each cluster.
-So that, we will have two types of constraints:
-\begin{enumerate}
-\item The predecessor(s) of an operation must be inserted into the same or previous cluster(s).
-\item The number of operations assigned to each cluster should be close to each other.
-\end{enumerate}
-
-Fig.~\ref{fig1} shows a flowchart of the proposed algorithm.
-As shown in the flowchart, the number of clusters is given.
-We calculate the cluster capacity by dividing the total number of operations by the number of clusters.
-In the next step, the clustering process started by generating the centroids, where the number of centroids is the number of clusters.
-In general, the centroids are generated randomly as well as in our model.
-However, we selected some operations as centroids.
-Although the operations are selected randomly, we followed a rule to select the centroids.
-For example, let us assume that a job consists of 15 operations, and three centroids should be generated.
-The first centroid will be an operation of rank 1 - 5, the second centroid will be 6 - 10, and the third will be from the last five operations.
-We followed this strategy to make them far from each other because the operations grouped in the first clusters should be at the beginning of their jobs.
-After generating the centroids, we perform a double-check for exceeding the number of clusters and the cluster capacity, respectively.
-After that, we calculate the distance between each operation and the centroid of the current cluster using Euclidean distance and assign the nearest operation to that cluster.
-In order to satisfy the precedence constraint, we check if the predecessor is not assigned to a cluster, it will be assigned to the current cluster, and the centroid is updated.
-If the maximum number of operations is reached, we will go to the next cluster.
-After assigning all operations into clusters, the algorithm terminates.
-
-\begin{figure}[h!]
-  \includegraphics[width=\linewidth]{Flow_chart_9.png}
-  \caption{Flowchart of the proposed Clustering Algorithm.}
-  \label{fig1}
-\end{figure}
-
-\section{Results}
-We use the benchmark instances presented in \cite{taillard1993benchmarks} as a dataset to work on.
-Many researchers have studied these instances and presented different approaches to solving them.
-We also developed different decomposition strategies to solve those.
-Therefore, it will be possible to measure the success/failure of our proposed model by comparing the results of the strategies developed in the past with those obtained from the proposed algorithm.
-There are 12 features, as we described in the previous section.
-In order to know which features have a significant effect on the results, we did some combinations of the features to perform the decomposition process clustering-based and then solved the problem using the Answer Set Programming model developed in \cite{el2020job} with a timeout of 10 minutes.
-The combinations of the selected features are listed as follows:
-\begin{enumerate}
-\item \textbf{F1} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST \}}
-\item \textbf{F2} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT \}} 
-\item \textbf{F3} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, EST \}} 
-\item \textbf{F4} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, ML \}} 
-\item \textbf{F5} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT, EST \}}
-\item \textbf{F6} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT, ML \}} 
-\item \textbf{F7} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, EST, ML \}} 
-\item \textbf{F8} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST \}} 
-\item \textbf{F9} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT \}} 
-\item \textbf{F10} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, EST \}} 
-\item \textbf{F11} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, ML \}}
-\item \textbf{F12} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, EST \}}
-\item \textbf{F13} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, ML \}}
-\item \textbf{F14} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, EST, ML \}}
-\item \textbf{F15} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, EST, ML \}}
-\end{enumerate}
-We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments.
-The following Table~\ref{tab1} shows the selected features, and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
-The following table shows the results of our experiments.
-The first column represents the instances we tested, and the next three columns show the makespan values obtained by FIFO, MTWR, and EST heuristics.
-From the fifth column, we show the makespan values obtained by combining the features mentioned above.
-Our experiments found that the first five combinations do not provide good solutions compared to the other heuristics.
-However, features (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}, \textbf{F14} and \textbf{F15}) have a good impact in the cost function in the most cases.
-Especially, combinations \textbf{F6} and \textbf{F15} provided good results.
-It means that the most important features extracted from the problem itself and have a positive impact on the performance are \textit{Remaining Processing Time, Machine Load and Earliest Starting Time}.
-Also, The information obtained from the solutions of the other heuristics like (\textit{Starting Time} and \textit{Waiting Time}) led us to get better decomposition for the problem at hand.
-More specifically, combining \textit{Remaining Processing Time} with \textit{Machine Load} provides a good decomposition with/without \textit{Waiting Time} attributes obtained by the other heuristics.
-\begin{table}[h!]
-\setlength{\tabcolsep}{4.0pt}
-  \begin{center}
-    \caption{The performance of selected features.}
-    \label{tab1}
- \begin{adjustwidth}{-3.7cm}{}
-    \begin{tabular}{c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c}
-      \textbf{Instances} & \textbf{FIFO} & \textbf{MTWR} & \textbf{EST} & \textbf{F1} & \textbf{F2} & \textbf{F3} & \textbf{F4} & \textbf{F5} & \textbf{F6} & \textbf{F7} & \textbf{F8} & \textbf{F9} & \textbf{F10} & \textbf{F11} & \textbf{F12} & \textbf{F13} & \textbf{F14} & \textbf{F15}\\
-      \hline
-      TA51 &3549  &3364  &3833  &3506  &3362  &3324  &3360  &3346  &\textbf{3294}  &3328  &3588  &3588  &3568  &3514  &3682  &3624  &3554  &3509 \\
-      TA52 &3339  &3304  &3594  &3277  &3318  &3330  &3286  &3243  &3275  &3226  &3397  &3373  &3533  &3256  &3494  &3505  &3373  &\textbf{3184} \\
-      TA53 &3160  &3168  &3443  &3382  &3585  &3347  &3543  &3517  &3250  &3338  &3251  &3522  &3373  &3585  &\textbf{3143}  &3427  &3524  &3380 \\
-      TA54 &3218  &3494  &3439  &3414  &3425  &3425  &3503  &3397  &3265  &3353  &3563  &3443  &3386  &3352  &3270  &3236  &3232  &3289 \\
-      TA55 &3291  &3237  &3784  &3308  &3441  &3424  &3319  &3355  &3386  &3301  &3498  &3385  &3530  &3365  &3309  &3370  &3385  &\textbf{3218} \\
-      TA56 &3325  &3287  &3552  &3353  &3380  &3548  &3270  &3383  &3366  &3564  &\textbf{3229}  &3306  &3362  &3344  &3332  &3338  &3483  &3389 \\
-      TA57 &3654  &3633  &3681  &3605  &3573  &3601  &3583  &3650  &\textbf{3480}  &3592  &3621  &3543  &3608  &3670  &3632  &3749  &3573  &\textbf{3480} \\
-      TA58 &3299  &3591  &3764  &3352  &3412  &3370  &3509  &3707  &\textbf{3324}  &3496  &3517 &3716 &3644  &3611  &3471  &3437  &3365   &3481 \\
-      TA59 &3344  &3394  &3661  &3453  &3416  &3301  &3339  &3508  &3430  &3352  &3258 &3381  &3402  &\textbf{3145}  &3250  &3253  &3284  &3290 \\
-      TA60 &3129  &3257  &3399  &3483  &\textbf{3315}  &3617  &3563  &3591  &3327  &3399  &3374  &3526  &3637  &3425  &3676  &3659  &3767  &3569 \\
-    \end{tabular}
-\end{adjustwidth}
-  \end{center}
-
-\end{table}
+From the literature, we have found that most of the studies which used data mining techniques aimed to extract some features either from the problem instances or solutions obtained by some heuristics to build a decision-rules to order the operations assigned to a machine.\\
 
@@
 In order to split the problem, we proposed a constrained clustering algorithm considering the precedence constraints of the JSP.
-We tested our model on benchmark instances, and the results showed that the performance of the proposed model outperforms other heuristics FIFO, MTWR, and EST; and some attributes like (\textit{Starting Time, Waiting Time, Remaining Processing Time, and Machine Load}) played an important role to split the operations in a good way.
+We tested our model on benchmark instances, and the results showed that the performance of the proposed model outperforms other heuristics FIFO, MTWR, and EST in the most cases; and some attributes like (\textit{Starting Time, Waiting Time, Remaining Processing Time, and Machine Load}) played an important role to split the operations in a good way.
 There are several ways to extend this study.

### 2021-10-01 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
abstract

@@
 \begin{abstract}
-Scheduling decision in production planning plays a significant role in the manufacturing industry.
-A good schedule may lead to solve production problems and improve productivity.
-Job-shop Scheduling Problem (JSP) is one of the most crucial problems in production systems.
-JSP deals with allocating tasks on resources and determining the sequence of the operations on each machine while completing all tasks as early as possible.
-One of the most effective approaches studied by many researchers for solving the JSP is decomposition.
-This work investigates applying the decomposition strategy to split the problem into a series of sub-problems using data mining methodologies.
-This study proposes a constrained clustering algorithm to group the data objects (operations) into clusters, where each cluster is a Time Window (TW).
-The decomposition process depends on two main phases; the first phase is to extract features to predict the sequence of the operations on each machine.
+Scheduling is a crucial problem appearing in various domains, such as manufacturing, transportation, or healthcare.
+In most problem definitions, the goal is to schedule a given set of operations on available resources to complete the operations as early as possible.
+Unfortunately, most scheduling problems cannot be solved efficiently.
+Therefore, the research of suitable approximation methods is of primary importance.
+
+This work suggests a novel approximation approach based on problem decomposition with data mining methodologies.
+This study proposes a constrained clustering algorithm to group the operations into clusters corresponding to time windows in which these operations must be scheduled.
+The decomposition process depends on two main phases.
+The first phase is to extract features to predict the sequence of the operations on each resource.
 These features are extracted either from the problem itself or from solutions obtained by other heuristics.
-The second phase is to develop a constrained clustering algorithm to assign each data object (an operation) into a cluster.
-We solved the problem using the Answer Set Programming model, and tested our proposed algorithm on a set of benchmark instances, and the results showed that our proposed outperformed other heuristics in most cases.
-The results also showed that some features like \textit{Remaining Processing Time}, \textit{Machine Load} and \textit{Earliest Starting Time} have a significant impact on the solution quality.
+The second phase is to develop a constrained clustering algorithm to assign each operation into a time window.
+We solved the problem using the Answer Set Programming.
+Evaluation results show that our proposed outperformed other heuristic schedulers in most cases, where features, like \textit{Remaining Processing Time}, \textit{Machine Load}, and \textit{Earliest Starting Time}, contributed significantly to the solution quality.
 
@@
 \subsection{\textit{Operation}}
-The \textit{Operation} attribute is an ordinal value to represent the rank of each operation in its job.
+The \textit{operation} attribute is an ordinal value to represent the rank of each operation in its job.
 

### 2021-10-01 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
transpose the table

@@
 
-%\begin{table}
-\begin{sidewaystable}
-\setlength{\tabcolsep}{5.5pt}
+\begin{table}
   \begin{center}
@@
     \label{tab4}
-%\rotatebox{90}{
- \begin{adjustwidth}{-3.5cm}{}
-    \begin{tabular}{c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c}
 
-\hline
-      \textbf{Instances} & \textbf{FIFO} & \textbf{MTWR} & \textbf{Single-shot} & \textbf{EST} & \textbf{F1} & \textbf{F2} & \textbf{F3} & \textbf{F4} & \textbf{F5} & \textbf{F6} & \textbf{F7} & \textbf{F8} & \textbf{F9} & \textbf{F10} & \textbf{F11} & \textbf{F12} & \textbf{F13} & \textbf{F14} & \textbf{F15}\\
-      \hline
-      TA51 &3549  &3364  &3632  &3506  &3506  &3362  &3324  &3360  &3346  &\textbf{3294}  &3328  &3588  &3588  &3568  &3514  &3682  &3624  &3554  &3509 \\
-      TA52 &3339  &3304  &3615  &3773  &3277  &3318  &3330  &3286  &3243  &3275  &3226  &3397  &3373  &3533  &3256  &3494  &3505  &3373  &\textbf{3184} \\
-      TA53 &3160  &3168  &3481  &3478  &3382  &3585  &3347  &3543  &3517  &3250  &3338  &3251  &3522  &3373  &3585  &\textbf{3143}  &3427  &3524  &3380 \\
-      TA54 &3218  &3494  &3462  &3497  &3414  &3425  &3425  &3503  &3397  &3265  &3353  &3563  &3443  &3386  &3352  &3270  &3236  &3232  &3289 \\
-      TA55 &3291  &3237  &3552  &3482  &3308  &3441  &3424  &3319  &3355  &3386  &3301  &3498  &3385  &3530  &3365  &3309  &3370  &3385  &\textbf{3218} \\
-      TA56 &3325  &3287  &3610  &3605  &3353  &3380  &3548  &3270  &3383  &3366  &3564  &\textbf{3229}  &3306  &3362  &3344  &3332  &3338  &3483  &3389 \\
-      TA57 &3654  &3633  &3778  &3753  &3605  &3573  &3601  &3583  &3650  &\textbf{3480}  &3592  &3621  &3543  &3608  &3670  &3632  &3749  &3573  &\textbf{3480} \\
-      TA58 &3299  &3591  &3718  &3731  &3352  &3412  &3370  &3509  &3707  &3324  &3496  &3517 &3716 &3644  &3611  &3471  &3437  &3365   &3481 \\
-      TA59 &3344  &3394  &3613  &3398  &3453  &3416  &3301  &3339  &3508  &3430  &3352  &3258 &3381  &3402  &\textbf{3145}  &3250  &3253  &3284  &3290 \\
-      TA60 &3129  &3257  &3550  &3247  &3483  &3315  &3617  &3563  &3591  &3327  &3399  &3374  &3526  &3637  &3425  &3676  &3659  &3767  &3569 \\
-\hline
- Average  &3331  &3373  &3601  &3547  &3413  &3423  &3429  &3428  &3470  &3340  &3395  &3430  &3475  &3504  &3427  &3426  &3460  &3454  &3379 \\
-\hline
+    \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|} \hline
+      & \multicolumn{11}{|c|}{\textbf{Instances}} \\ \cline{2-12}
+      \textbf{Approach}   & TA51                           & TA52                           & TA53                           & TA54 & TA55                           & TA56                           & TA57                           & TA58 & TA59                           & TA60 & AVG  \\ \hline
+      \textbf{FIFO}        & 3549                           & 3339                           & 3160                           & 3218 & 3291                           & 3325                           & 3654                           & 3299 & 3344                           & 3129 & 3331 \\
+      \textbf{MTWR}        & 3364                           & 3304                           & 3168                           & 3494 & 3237                           & 3287                           & 3633                           & 3591 & 3394                           & 3257 & 3373 \\
+      \textbf{Single-shot} & 3632                           & 3615                           & 3481                           & 3462 & 3552                           & 3610                           & 3778                           & 3718 & 3613                           & 3550 & 3601 \\
+      \textbf{EST}         & 3506                           & 3773                           & 3478                           & 3497 & 3482                           & 3605                           & 3753                           & 3731 & 3398                           & 3247 & 3547 \\ \hline
+      \textbf{F1}          & 3506                           & 3277                           & 3382                           & 3414 & 3308                           & 3353                           & 3605                           & 3352 & 3453                           & 3483 & 3413 \\
+      \textbf{F2}          & 3362                           & 3318                           & 3585                           & 3425 & 3441                           & 3380                           & 3573                           & 3412 & 3416                           & 3315 & 3423 \\
+      \textbf{F3}          & 3324                           & 3330                           & 3347                           & 3425 & 3424                           & 3548                           & 3601                           & 3370 & 3301                           & 3617 & 3429 \\
+      \textbf{F4}          & 3360                           & 3286                           & 3543                           & 3503 & 3319                           & 3270                           & 3583                           & 3509 & 3339                           & 3563 & 3428 \\
+      \textbf{F5}          & 3346                           & 3243                           & 3517                           & 3397 & 3355                           & 3383                           & 3650                           & 3707 & 3508                           & 3591 & 3470 \\
+      \textbf{F6}          & \textbf{3294} & 3275                           & 3250                           & 3265 & 3386                           & 3366                           & \textbf{3480} & 3324 & 3430                           & 3327 & 3340 \\
+      \textbf{F7}          & 3328                           & 3226                           & 3338                           & 3353 & 3301                           & 3564                           & 3592                           & 3496 & 3352                           & 3399 & 3395 \\
+      \textbf{F8}          & 3588                           & 3397                           & 3251                           & 3563 & 3498                           & \textbf{3229} & 3621                           & 3517 & 3258                           & 3374 & 3430 \\
+      \textbf{F9}          & 3588                           & 3373                           & 3522                           & 3443 & 3385                           & 3306                           & 3543                           & 3716 & 3381                           & 3526 & 3475 \\
+      \textbf{F10}         & 3568                           & 3533                           & 3373                           & 3386 & 3530                           & 3362                           & 3608                           & 3644 & 3402                           & 3637 & 3504 \\
+      \textbf{F11}         & 3514                           & 3256                           & 3585                           & 3352 & 3365                           & 3344                           & 3670                           & 3611 & \textbf{3145} & 3425 & 3427 \\
+      \textbf{F12}         & 3682                           & 3494                           & \textbf{3143} & 3270 & 3309                           & 3332                           & 3632                           & 3471 & 3250                           & 3676 & 3426 \\
+      \textbf{F13}         & 3624                           & 3505                           & 3427                           & 3236 & 3370                           & 3338                           & 3749                           & 3437 & 3253                           & 3659 & 3460 \\
+      \textbf{F14}         & 3554                           & 3373                           & 3524                           & 3232 & 3385                           & 3483                           & 3573                           & 3365 & 3284                           & 3767 & 3454 \\
+      \textbf{F15}         & 3509                           & \textbf{3184} & 3380                           & 3289 & \textbf{3218} & 3389                           & \textbf{3480} & 3481 & 3290                           & 3569 & 3379 \\ \hline
       \end{tabular}
-%}
-\end{adjustwidth}
   \end{center}
-\end{sidewaystable}
-%\end{table}
+\end{table}
 

### 2021-10-03 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
intro

@@
 \usepackage{caption}
+\usepackage{todonotes}
 \captionsetup[table]{skip=10pt}
@@
 %
-\title{A decomposition based on constrained clustering algorithms for JSP\thanks{Supported by organization x.}}
+\title{Decomposition-based Job-shop Scheduling with Constrained Clustering}%\thanks{Supported by organization x.}}
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Decomposition-based JSP with Constrained Clustering}
 % If the paper title is too long for the running head, you can set
@@
 \section{Introduction}
-Nowadays, the massive progress and development of the internet and online technologies, data generated by machines and devices, product development, quality and inventory management systems, or production planning systems has become huge and is expected to increase in the coming years.
+Scheduling is one of the most crucial problems in various industrial, transportation, or healthcare applications. 
+\todo{add 1-2 citations for each application} 
+These applications resulted in different definitions of scheduling problems, with Job-shop Scheduling Problem (JSP) being among the most popular ones.
+In JSP, operations of given jobs must be scheduled on available machines in an optimal way wrt.\ some pre-defined criteria.
+The latter include, for instance, minimization of makespan, i.e., the completion time of the last job, or tardiness, i.e., the sum of delays for all jobs completed after their deadlines.
+
+However, JSP is a NP-complete combinatorial optimization problem \cite{DBLP:journals/mor/GareyJS76,SOTSKOV1995237} for which no efficient algorithms are known.
+Therefore, searching for an optimal solution with modern approaches like Answer Set Programming (ASP), Mixed Integer Programming, or Constraint Programming \todo{cite at least 1 paper solving JSP with these methods} often takes too much computation time, even for small instances.
+Practical applications require solving JSP instances of vast scale with thousands of operations \cite{zhang2010hybrid}.
+As a result, many researchers focused on the development of efficient algorithms for finding high-quality approximations of the optimal schedule, including dispatching rules and other heuristic approaches~\cite{blackstone1982state}, or stochastic optimization techniques \cite{DBLP:journals/informs/VaessensAL96,DBLP:journals/jim/CalisB15}.
+The main issue of these approaches is that they require manual parametrization for a particular scheduling problem, which is tedious and error-prone.
+For instance, heuristic approaches might require the development of new heuristics or their combinations.
+Similarly, default parameters of stochastic methods might result in a finding of mediocre schedules.
+As a result, much focus of the recent research lies on the automation of the parametrization using machine learning techniques~\cite{bengio2020machine}.
+
+%Nowadays, the massive progress and development of the internet and online technologies, data generated by machines and devices, product development, quality and inventory management systems, or production planning systems has become huge and is expected to increase in the coming years.
 Hence, to capture long-term revenues and sustainable competitive advantages, companies must manage the knowledge and have the valuable information to make the right decision at the right time \cite{benabdellah2019survey}.
@@
 
-It can be assumed that knowledge management is a crucial issue in the industry.
+%It can be assumed that knowledge management is a crucial issue in the industry.
 To extract implicit, unknown, and potentially useful information from data, we use data mining techniques that have been responsible for many of artificial intelligence’s recent successes.
-Clustering is one of these techniques, which is applied whenever the extracted data is not labeled, i.e., the semantics of this data is with respect to the application is unclear.
+%Clustering is one of these techniques, which is applied whenever the extracted data is not labeled, i.e., the semantics of this data is with respect to the application is unclear.
 Therefore, clustering has always been an exploratory but critical task in the knowledge discovery process, with applications ranging from image processing, production systems, and information retrieval \cite{benabdellah2019survey}.
@@
 
-The data sets could be composed entirely of either numeric features or symbolic features.
+%The data sets could be composed entirely of either numeric features or symbolic features.
 For the numeric feature, we will use the Euclidean distance metric.
@@
 
-Furthermore, from an optimization perspective, the main objective is to minimize the distance between objects falling in the same cluster and maximize the distance between the others that belong to different clusters.
+%Furthermore, from an optimization perspective, the main objective is to minimize the distance between objects falling in the same cluster and maximize the distance between the others that belong to different clusters.
 Since clustering does not use a subset of the dataset as labels to learn a classification model, clustering differs from classification \cite{wagstaff2001constrained}.
@@
 
-Clustering has been widely applied in several disciplines; one is production-planning systems~\cite{nananukul2013clustering,koskosidis1992clustering}, more specifically scheduling \cite{yashar2013multi,tong2016research}.
+%Clustering has been widely applied in several disciplines; one is production-planning systems~\cite{nananukul2013clustering,koskosidis1992clustering}, more specifically scheduling \cite{yashar2013multi,tong2016research}.
 %\comment{Are there any citations for the claim above?}
-Scheduling is one of the most complex problems in the industry.
-Scheduling of operations involves resource allocation over a period of time to perform a series of tasks, where one of the most critical problems is the Job-shop Scheduling Problem (JSP).
-The JSP has been known as a complex and combinatorial optimization problem, and it has been proven to be NP-hard \cite{baker1974introduction,lenstra1979computational}.
-Therefore, it is very difficult to reach an optimal solution in a reasonable time, even for small instances.
-However, from a practical perspective, the JSP scale is large, where the number of operations can be up to 10,000s in some workshops \cite{zhang2010hybrid}.
-\\
 
-This work intends to apply the clustering techniques to decompose the problem into sub-problems, where the number of clusters corresponds to the number of sub-problems (windows), and the operations correspond to data objects.
+In this work, we suggest a clustering-based method that automatically decomposes a JSP instance into sub-instances which can be tackled by 
+optimization algorithms of an ASP solver. 
+General clustering algorithms are unsupervised learning methods that partition a given set of objects into disjoint clusters, where each cluster comprises closest objects wrt.\ some distance function.
+In scheduling settings, however, clusters must satisfy additional constraints, which define precedence relations between operations of a job.
+Standard constrained clustering algorithms are not applicable in scheduling scenarios since they usually allow for the definition of disjointness constraints, which specify objects that cannot appear in one cluster \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}.
+%\cite{wagstaff2001constrained}
+%where the number of clusters corresponds to the number of sub-problems (windows), and the operations correspond to data objects.
 For the JSP, an instance consists of a set of jobs to be processed on machines.
-The number of jobs is $n$, and the number of machines is $m$.
+%The number of jobs is $n$, and the number of machines is $m$.
 Each job contains $m$ operations with the operation precedence constraints.
@@
 However, most of those checked whether two data objects must be assigned into the same cluster.
-For the JSP, it is essential to consider the sequencing between the clusters to handle the precedence constraints.
-More specifically, our proposed method will not permit to assign an operation into a particular cluster if the predecessor of that operation is not yet assigned to the same or previous cluster(s).
-\\
+Therefore, our clustering method implements a novel type of constraints, which prevent the assignment of an operation to a cluster if its preceding operation is not yet assigned to the same or previous cluster.
 
@@
 \begin{table}[h!]
-\fontsize{12}{9}
+%\fontsize{12}{9}
 \setlength{\tabcolsep}{10.0pt}
@@
     \begin{tabular}{c c c}
-      \textbf{Operation} & \textbf{Machine} & \textbf{Procession Time} \\
+      \textbf{Operation} & \textbf{Machine} & \textbf{Procesing Time} \\
       \hline

### 2021-10-04 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
adding citation

@@
 \section{Introduction}
-Scheduling is one of the most crucial problems in various industrial, transportation, or healthcare applications. 
-\todo{add 1-2 citations for each application} 
+Scheduling is one of the most crucial problems in various industrial, transportation, or healthcare applications
+%\todo{add 1-2 citations for each application}
+ \cite{pezzella2008genetic,chaudhry2016research,nouiri2018effective,demirbilek2019dynamically,schoenfelder2020nurse,wang2019routing,janakbhai2021blockchain}.
+
 These applications resulted in different definitions of scheduling problems, with Job-shop Scheduling Problem (JSP) being among the most popular ones.
@@
 However, JSP is a NP-complete combinatorial optimization problem \cite{DBLP:journals/mor/GareyJS76,SOTSKOV1995237} for which no efficient algorithms are known.
-Therefore, searching for an optimal solution with modern approaches like Answer Set Programming (ASP), Mixed Integer Programming, or Constraint Programming \todo{cite at least 1 paper solving JSP with these methods} often takes too much computation time, even for small instances.
-Practical applications require solving JSP instances of vast scale with thousands of operations \cite{zhang2010hybrid}.
+Therefore, searching for an optimal solution with modern approaches like Answer Set Programming (ASP), Mixed Integer Programming, or Constraint Programming \cite{meng2020mixed,el2020job,al2017job}
+%\todo{cite at least 1 paper solving JSP with these methods} 
+often takes too much computation time, even for small instances. Practical applications require solving JSP instances of vast scale with thousands of operations \cite{zhang2010hybrid}.
 As a result, many researchers focused on the development of efficient algorithms for finding high-quality approximations of the optimal schedule, including dispatching rules and other heuristic approaches~\cite{blackstone1982state}, or stochastic optimization techniques \cite{DBLP:journals/informs/VaessensAL96,DBLP:journals/jim/CalisB15}.
@@
 \begin{enumerate}
-	\item Once an operation is started, it must be completed without interruption.
+	\item Once an operation starts, it must be completed without interruption.
 	\item Each job visits a machine once.

### 2021-10-04 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
final intro

--- 2021-10-03 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 \usepackage{todonotes}
+\usepackage[inline]{enumitem}
 \captionsetup[table]{skip=10pt}
@@
 Standard constrained clustering algorithms are not applicable in scheduling scenarios since they usually allow for the definition of disjointness constraints, which specify objects that cannot appear in one cluster \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}.
-%\cite{wagstaff2001constrained}
-%where the number of clusters corresponds to the number of sub-problems (windows), and the operations correspond to data objects.
-For the JSP, an instance consists of a set of jobs to be processed on machines.
-%The number of jobs is $n$, and the number of machines is $m$.
-Each job contains $m$ operations with the operation precedence constraints.
-Each operation should be executed on machine $m$ for a certain amount of time.
-Since the problem at hand has precedence constraints, it is required to apply a constrained clustering approach for the decomposition \cite{wagstaff2001constrained}.
-Although many studies proposed different constrained clustering methods in \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}.
-However, most of those checked whether two data objects must be assigned into the same cluster.
-Therefore, our clustering method implements a novel type of constraints, which prevent the assignment of an operation to a cluster if its preceding operation is not yet assigned to the same or previous cluster.
-
-It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations.
+Therefore, our clustering method implements a novel type of constraints that
+\begin{enumerate*}[label=\emph{(\roman*)}]
+  \item prevent the assignment of an operation to a cluster if its preceding operation is not yet assigned to the same or previous cluster, and
+  \item ensure balancing between the clusters according to the number of operations per cluster.
+\end{enumerate*}
+Further contributions of this paper can be summarized as follows:
+\begin{itemize}
+  \item A typical JSP instance describing jobs, their operations, and available machines, does not provide enough information for a clustering algorithm to find a good decomposition.
+Therefore, we suggest various heuristic feature extractors, such as First-In-First-Out or machine load.
+  \item We implement the suggested constrained clustering method and combine it with the forward selection of features.
+The latter is an automatic procedure aiming to find a subset of features allowing the clustering algorithm to compute a decomposition resulting in the best quality schedule.
+  \item Evaluation of our method on the Taillard benchmark instances \cite{taillard1993benchmarks} showed that it significantly outperforms baseline heuristic methods and pure ASP optimization algorithms wrt.\ the makespan optimization criteria within the timeout of 600 seconds.
+\end{itemize}
+%It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations.
 The lack of these attributes guides us to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window.
 The main contribution of this work is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan.
-More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time. \\
+%More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time. \\
 
-This work consists of two main phases:
+%This work consists of two main phases:
 
-\begin{enumerate}
-    \item The first phase is to apply data mining to predict the order of each operation on its machine according to some features.
+%\begin{enumerate}
+%    \item The first phase is to apply data mining to predict the order of each operation on its machine according to some features.
 In this phase, we will try to get all possible features extracted from the problem itself, which significantly affect the quality of the solution.
@@
 
-    \item The second phase aims to develop, implement and test a proposed clustering algorithm that will split the problem into sub-problems where each operation is a data object, and each cluster will represent a Time Window (TW).
+%    \item The second phase aims to develop, implement and test a proposed clustering algorithm that will split the problem into sub-problems where each operation is a data object, and each cluster will represent a Time Window (TW).
 It will consider the precedence constraints between the operations and balancing between the clusters according to the number of operations per cluster.
-\end{enumerate}
+%\end{enumerate}
 
-The paper is organized as follows.
-The following Section describes the Job-shop Scheduling Problem in detail and shows an illustrative example of input data.
-Section $3$ shows which features we extracted and how we extracted them, followed by Section $4$, which presents the proposed algorithm.
-Section $5$ presents an extensive computational study on benchmark instances, comparing our results with the others obtained by some heuristics.
-A relevant literature review on the subject is given in Section $6$ followed by the conclusion and future work.
+The paper is organized as follows: the following section provides the definition of JSP and gives an illustrative example of input data.
+Section \ref{sec:features} discusses heuristic feature extractors, followed by Section \ref{sec:method}, which presents the proposed constrained clustering algorithm.
+Section \ref{sec:eval} presents an extensive computational study on benchmark instances, comparing our results with the baseline solvers.
+A literature review is given in Section \ref{sec:literature}, followed by the conclusions and future work.
 
@@
 \section{Features extraction}
+\label{sec:features}
 In this section, we show which features have been extracted and how did we extract them.
@@
 \section{Proposed Algorithm}
-
+\label{sec:method}
 In this section, we will show the implementation of our proposed model and how it works.
@@
 \section{Results}
+\label{sec:eval}
 We use the benchmark instances presented in \cite{taillard1993benchmarks} as a dataset to work on.
@@
 \section{Literature Review}
-
+\label{sec:literature}
 In this section, we will review some of the most related works about data mining approaches employed to solve scheduling problems.
@@
 
-\section{Conclusion}
+\section{Conclusions}
+\label{sec:conclusions}
 Data mining and machine learning is an area of interest in many manufacturing companies and other fields.

### 2021-10-04 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
Section 2 and comments in 3

@@
 \begin{itemize}
-  \item A typical JSP instance describing jobs, their operations, and available machines, does not provide enough information for a clustering algorithm to find a good decomposition.
-Therefore, we suggest various heuristic feature extractors, such as First-In-First-Out or machine load.
+  \item Since a typical JSP instance describing jobs, their operations, and available machines, does not provide enough information for a clustering algorithm to find a good decomposition, we suggest various heuristic feature extractors, such as First-In-First-Out or machine load.
   \item We implement the suggested constrained clustering method and combine it with the forward selection of features.
@@
 
-\section{Problem Formulation}
-The Job-shop Scheduling Problem (JSP) is one of the well-known hardest combinatorial optimization problems \cite{baker1974introduction,lenstra1979computational,garey1976complexity}.
-The JSP is given a set $J = \{ J_1, J_2,...,J_n \}$  of jobs and a set $M = \{ M_1, M_2,...,M_m \}$ of Machines.
-Each job $J_i$ has a series $  O_{i,1}, O_{i,2},...,O_{i,m} $ of operations that must be processed in a predefined order.
+\section{Problem Definition}
+The Job-shop Scheduling Problem (JSP) is one of the well-known scheduling problems \cite{baker1974introduction,lenstra1979computational}.
+An instance of a JSP comprises a set $J = \{ J_1, J_2, \dots ,J_n \}$  of jobs and a set $M = \{ M_1, M_2, \dots ,M_m \}$ of Machines.
+Each job $J_i$ has a sequence $(O_{i,1}, O_{i,2}, \dots ,O_{i,m})$ of operations that must be processed in a predefined order.
 Each machine $M_i$ executes an operation of each job with a known and fixed processing time.
-The main objective of the JSP is to minimize the total completion time for all jobs (makespan).
-\\ The JSP assumptions are listed as follows:
+The main objective of a JSP solver is to find a schedule that minimizes some criteria like makespan, i.e., the total completion time for all jobs.
+In addition, JSP makes the following assumptions:
+\begin{itemize}
+	\item Once an operation starts, it must be completed without interruptions.
+	\item Each job has no duplicated operations, i.e., visits a machine only once.
+	\item A machine cannot perform more than one operation at a time.
+	\item All machines are available at all times.
+	\item Processing the the jobs starts at time 0.
+\end{itemize}
 
-\begin{enumerate}
-	\item Once an operation starts, it must be completed without interruption.
-	\item Each job visits a machine once.
-	\item A machine cannot perform more than one operation at a time.
-	\item A machine is not broken.
-	\item All jobs and machines are available at time 0.
-\end{enumerate}
-It is hard to reach a solution close to the optimal in a reasonable time, even if the number of operations to be scheduled is low.
-In real-life applications, the number of operations that are to be scheduled is up to $10,000$ \cite{zhang2010hybrid}, which makes the problem more complicated.
-The JSP has attracted the interest of many researchers in the last decades to find effective methods to solve this problem.
-One of these methods is decomposition.
-The decomposition idea is to split the problem into sub-problems, solve them separately, and then integrate those solutions to solve the whole problem.
-Many decomposition strategies have been proposed for solving the JSP  \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.
-However, there is no guarantee of which method is the best because the performance varies according to the problem.
-\\
-
-A problem data can be organized in a table, where rows correspond to operations and columns refer to the name of the operation, the machine assigned to an operation, and the processing time of a particular operation.
-Table~\ref{tab1} shows the information of a simple Job-shop Scheduling Problem which consists of $3$ jobs and $3$ machines.
+Let us illustrate the problem on a small example presented in Table~\ref{tab1}, which shows the information of a simple JSP comprising $3$ jobs and $3$ machines.
+In this table, rows correspond to operations, whereas columns refer to the name of the operation, the machine assigned to an operation, and the processing time of a particular operation.
 For instance, in the third row of the table, $O_{1,3}$ is the third operation of the first job that will be executed by machine $1$ with processing time $5$.
 
-\begin{table}[h!]
+\begin{table}
+\begin{minipage}{.6\textwidth}
 %\fontsize{12}{9}
-\setlength{\tabcolsep}{10.0pt}
+\setlength{\tabcolsep}{5.0pt}
 \centering
-  \begin{center}
-    \caption{An example of a Job-shop Scheduling Problem.}
+ % \begin{center}
+    \caption{A sample JSP instance.}
     \label{tab1}
@@
     \end{tabular}
-
 %\end{adjustwidth}
-  \end{center}
-
-\end{table}
-In our example in Table~\ref{tab1}, there are $9$ operations and $3$ machines.
-We aim to split them efficiently into two parts and schedule them separately without violating the precedence constraint.
-More specifically, we should not assign an operation in the first TW, and its predecessor is assigned to the following TW(s).
-Table~\ref{tab2} shows how the output of the proposed method should look.
-As shown in the Table~\ref{tab2}, operations $\{ O_{1,1}, O_{1,2}, O_{2,1}, O_{2,2}, O_{3,1} \}$ are assigned to the first TW and the rest to the second TW.
-It means that the first TW will be optimized and then the upcoming TW(s).
-In this study, we investigate applying a constrained clustering algorithm to decompose the JSP, where we extract some features from the problem itself and the solutions of other heuristics.
-These features have been used for the splitting process, and we present them in detail in the following section.
-
-\begin{table}[h!]
-\fontsize{12}{9}
-\setlength{\tabcolsep}{10.0pt}
+%  \end{center}
+\end{minipage}
+\begin{minipage}{.4\textwidth}
+  \setlength{\tabcolsep}{5.0pt}
   \centering
-  \begin{center}
-    \caption{Time Window Assignment.}
+ % \begin{center}
+    \caption{Possible decomposition.}
     \label{tab2}
@@
       $O_{3,3}$  & 2    	\\
+      \\
     \end{tabular}
 %\end{adjustwidth}
-  \end{center}
+  %\end{center}
+\end{minipage}
+\end{table}
 
-\end{table}
+Since no efficient algorithm is known for JSP \cite{garey1976complexity}, 
+modern optimization algorithms often cannot reach a solution close to the optimal one in a reasonable time, even if the number of operations to be scheduled is low.
+In real-life applications, the number of operations in an instance might be up to tens of thousands \cite{zhang2010hybrid}.
+The JSP has attracted the interest of many researchers in the last decades to find effective methods to solve this problem.
+One of these methods is decomposition.
+The decomposition idea is to split the problem into sub-problems, solve them separately, and then integrate those solutions to solve the whole problem.
+Many decomposition strategies have been proposed for solving the JSP  \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.
+However, there is no guarantee of which method is the best because the performance varies according to the problem.
+
+In our example in Table~\ref{tab1}, there are $9$ operations and $3$ machines.
+We aim to split the operations efficiently into two partitions and schedule them separately without violating the precedence constraint.
+More specifically, we should not assign an operation in the Time Window (TW) which precedes TW(s) of at least one of its predecessors.
+Table~\ref{tab2} shows one of the possible decompositions that we would like our method to return.
+Namely, operations $\{ O_{1,1}, O_{1,2}, O_{2,1}, O_{2,2}, O_{3,1} \}$ are assigned to the first TW and the rest to the second TW.
+Given this decomposition, a multi-shot optimization method, like ASP, will first find an optimal schedule of the operations in the first TW and provide it as a set of facts to the search of the optimal solution for the second TW.
+In this study, we investigate applying a constrained clustering algorithm to decompose the JSP, where we extract some features from the problem itself and the solutions of other heuristics.
+These features have been used for the splitting process, and we present them in detail in the following section.
 
@@
 We present the features that we considered:
-\subsection{\textit{Process}}
-\textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are attributes dependent on the problem domain.
+\paragraph{Process.}
+The two attributes \textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are dependent on the problem domain.
 The \textit{PT} represents the execution time of a particular operation in a machine.
 However, \textit{RPT} is the aggregate processing time for all subsequent operations of that job.
+\todo[inline]{we need to provide rationale behind the features, why do we think they are useful for the clusering algorithm.
+Example of values extracted for the sample instance in Table 1 would also improve the readability of the paper.}
+\todo[inline]{Why in all other paragraphs we are talking only about one feature at a time and here about two.
+Can we split them?
+Like \\paragraph\{Processing Time (PT)\} and \\paragraph\{Remain Processing time (RPT)\}}
 
-\subsection{\textit{Operation}}
+\paragraph{Operation.}
 The \textit{operation} attribute is an ordinal value to represent the rank of each operation in its job.
 
-\subsection{\textit{Time Length of a Job} (TLJ)}
+\paragraph{Time Length of a Job (TLJ).}
 \textit{TLJ} is an attribute that describes the total processing time needed to complete a particular job. It is the summation of processing time for all operations of that job.
 
-\subsection{\textit{Earliest Starting Time} (EST)}
+\paragraph{Earliest Starting Time (EST).}
 The \textit{EST} attribute represents the earliest possible time for a particular operation to be executed.
@@
 
-\subsection{\textit{Machine Load} (ML)}
+\paragraph{{Machine Load} (ML)}
 \textit{ML} is a property of a machine that shows how much time a particular machine has to execute all the operations assigned to it.
@@
 
-\subsection{\textit{Starting Time} (ST)}
+\paragraph{{Starting Time} (ST)}
 \textit{ST} attribute represents the starting execution time of each operation by solving the problem using heuristics FIFO, MTWR and EST developed by \cite{tassel2021reinforcement,el2020job}.
 
-\subsection{\textit{Waiting Time} (WT)}
+\paragraph{{Waiting Time} (WT)}
 Since we used three different heuristics to mine information from the obtained solution, we will have three waiting times.
@@
 
-\section{Literature Review}
+\section{Related Work}
 \label{sec:literature}

### 2021-10-05 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
update in section 3 and 4

@@
 	\item All machines are available at all times.
-	\item Processing the the jobs starts at time 0.
+	\item Processing the jobs starts at time 0.
 \end{itemize}
@@
 \label{sec:features}
-In this section, we show which features have been extracted and how did we extract them.
-Based on the previous work, \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data}, the following features have been identified: priority, processing time, remaining processing time, machine load, and route position.
+Data mining merges with machine learning and statistical techniques to gain information and new knowledge from the databases\cite{ismail2012production,nasiri2019data}.
+Examples of data mining tasks are classification, clustering, and sequential pattern mining.
+In this work, we aim to use clustering to study the operations' pattern to find a good sequencing of the operations on each machine that leads us to obtain a near-optimal solution of the JSP.
+In order to use the clustering approach, we should have some features which describe the characteristics of the problem.
+These features would aid us in the decision-making process, hence, reaching a near-optimal solution.
+\\
+
+Some of the previous work \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} have extracted the following features: priority, processing time, remaining processing time, machine load, and route position.
 Most of the researchers converted the attributes into qualitative data.
@@
 We present the features that we considered:
-\paragraph{Process.}
-The two attributes \textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are dependent on the problem domain.
+%\paragraph{Process.}
+%The two attributes \textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are dependent on the problem domain.
 The \textit{PT} represents the execution time of a particular operation in a machine.
@@
 Example of values extracted for the sample instance in Table 1 would also improve the readability of the paper.}
-\todo[inline]{Why in all other paragraphs we are talking only about one feature at a time and here about two.
+%\todo[inline]{Why in all other paragraphs we are talking only about one feature at a time and here about two.
 Can we split them?
 Like \\paragraph\{Processing Time (PT)\} and \\paragraph\{Remain Processing time (RPT)\}}
+\paragraph{Processing Time (PT):}
+The \textit{PT} represents the execution time of a particular operation in a machine.
 
-\paragraph{Operation.}
+\paragraph{Remain Processing Time (RPT):}
+\textit{RPT} attribute is the aggregate processing time for all subsequent operations of that job.
+
+\paragraph{Operation:}
 The \textit{operation} attribute is an ordinal value to represent the rank of each operation in its job.
 
-\paragraph{Time Length of a Job (TLJ).}
+\paragraph{Time Length of a Job (TLJ):}
 \textit{TLJ} is an attribute that describes the total processing time needed to complete a particular job. It is the summation of processing time for all operations of that job.
 
-\paragraph{Earliest Starting Time (EST).}
+\paragraph{Earliest Starting Time (EST):}
 The \textit{EST} attribute represents the earliest possible time for a particular operation to be executed.
@@
 
-\paragraph{{Machine Load} (ML)}
+\paragraph{{Machine Load} (ML):}
 \textit{ML} is a property of a machine that shows how much time a particular machine has to execute all the operations assigned to it.
@@
 
-\paragraph{{Starting Time} (ST)}
+\paragraph{{Starting Time} (ST):}
 \textit{ST} attribute represents the starting execution time of each operation by solving the problem using heuristics FIFO, MTWR and EST developed by \cite{tassel2021reinforcement,el2020job}.
 
-\paragraph{{Waiting Time} (WT)}
+\paragraph{{Waiting Time} (WT):}
 Since we used three different heuristics to mine information from the obtained solution, we will have three waiting times.
@@
 
-Algorithm.~\ref{alg} shows the pseudocode of the proposed algorithm.
+Our main procedure starts with selecting a subset of the extracted features, which will be used for the clustering process to get the Time window assignment.
+As shown in algorithm~\ref{alg1}, the extracted features are given, and a set of combinations of the features are initialized.
+At each iteration, a subset of features is selected and run with the clustering process discussed later, then the Time Window assignment.
+This process is repeated until the termination condition is met.
+
+\begin{algorithm}
+\caption{Feature Selection Algorithm}\label{alg1}
+\begin{algorithmic}
+\Require $extracted\_features$
+\State $count\_feature \gets 0$
+\State $comb\_features$ \Comment{Define a set of combined features}
+\State $num\_subsets \gets \textbf{count}(comb\_features)$
+\While{$count\_feature < num\_subsets$}
+	\State $selected\_subset \gets comb\_features[count\_feature]$
+	\State Proposed Clustering Algorithm ($selected\_subset$) \Comment{\textbf{(Algorithm.~\ref{alg2})}} %\textbf{(Algorithm.~\ref{alg2})}
+	\State Store the Time Windows Assignment
+	\State $count\_feature \gets count\_feature + 1$
+\EndWhile
+
+\end{algorithmic}
+\end{algorithm}
+
+Algorithm.~\ref{alg2} shows the pseudocode of the proposed algorithm.
 As shown in the pseudocode, the number of clusters is given.
@@
 \begin{algorithm}
-\caption{Proposed Clustering Algorithm}\label{alg}
+\caption{Proposed Clustering Algorithm}\label{alg2}
 \begin{algorithmic}

### 2021-10-11 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
discussion

@@
 % \renewcommand\UrlFont{\color{blue}\rmfamily}
+
+\newcommand{\stest}{\textit{ST\_EST}}
 
@@
 %It can be assumed that knowledge management is a crucial issue in the industry.
-To extract implicit, unknown, and potentially useful information from data, we use data mining techniques that have been responsible for many of artificial intelligence’s recent successes.
+To extract implicit, unknown, and potentially useful information from data, we use data mining techniques that have been responsible for many of artificial intelligence's recent successes.
 %Clustering is one of these techniques, which is applied whenever the extracted data is not labeled, i.e., the semantics of this data is with respect to the application is unclear.
@@
 Like \\paragraph\{Processing Time (PT)\} and \\paragraph\{Remain Processing time (RPT)\}}
-\paragraph{Processing Time (PT):}
-The \textit{PT} represents the execution time of a particular operation in a machine.
+\begin{description}
+  \item[Processing Time (PT)] that represents the execution time of a particular operation in a machine.
 
-\paragraph{Remain Processing Time (RPT):}
-\textit{RPT} attribute is the aggregate processing time for all subsequent operations of that job.
+  \item[Remaining Processing Time (RPT)] attribute corresponds to the total processing time for all subsequent operations of a job.
 
-\paragraph{Operation:}
-The \textit{operation} attribute is an ordinal value to represent the rank of each operation in its job.
+  \item[Operation (OP)] feature is an ordinal value representing the rank of each operation in its job.
 
-\paragraph{Time Length of a Job (TLJ):}
-\textit{TLJ} is an attribute that describes the total processing time needed to complete a particular job. It is the summation of processing time for all operations of that job.
+  \item[Time Length of a Job (TLJ)] is an attribute describing the total processing time needed to execute all operations of a job. 
+  %It is the summation of processing time for all operations of that job.
 
-\paragraph{Earliest Starting Time (EST):}
-The \textit{EST} attribute represents the earliest possible time for a particular operation to be executed.
-It is calculated by accumulating the processing time of the predecessor(s) of that operation, and it is $Zero$, if the operation has no predecessor.
+  \item[Earliest Starting Time (EST)] attribute represents the earliest possible execution time for an operation.
+It is calculated by accumulating the processing time of all preceding operations.
+This feature has a specific value $\mathit{Zero}$ if the operation has no predecessors.
 
-\paragraph{{Machine Load} (ML):}
-\textit{ML} is a property of a machine that shows how much time a particular machine has to execute all the operations assigned to it.
-In the initial state, the \textit{ML} is the aggregate of processing time of all operations assigned to that machine.
+  \item[Machine Load (ML)] is a property of a machine that shows how much time this machine requires to execute all the operations assigned to it.
+In the initial state, \textit{ML} corresponds to the total processing time of all operations assigned to that machine.
 For each step, the load of a machine is decreased by the processing time of an operation assigned to the machine and has the smallest \textit{EST}.
 
-\paragraph{{Starting Time} (ST):}
-\textit{ST} attribute represents the starting execution time of each operation by solving the problem using heuristics FIFO, MTWR and EST developed by \cite{tassel2021reinforcement,el2020job}.
+  \item[Starting Time (ST)] is a set of features representing the starting execution time of each operation by solving the problem using a heuristic.
+In our work, we consider EST (\stest), First-In-First-Out (ST\_FIFO), and Most Total Work Remaining (ST\_MTWR), see \cite{overview} for an overview.
+  To compute the starting time, we use a simple greedy algorithm \cite{tassel2021reinforcement,el2020job}, that at each step selects a not scheduled operation according to the heuristics and schedules it to a machine.
+In the case of FIFO, the algorithm selects an operation that appeared first in the queue.
+Similarly, for MTWR ...
 
-\paragraph{{Waiting Time} (WT):}
-Since we used three different heuristics to mine information from the obtained solution, we will have three waiting times.
-\textit{WT} attribute is the difference between the starting time of a particular operation and the maximum value between the end time of its predecessor and when the machine became free.
-\\
+  \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used three different heuristics EST (WT\_EST), FIFO (WT\_FIFO), and MTWR (WT\_MTWR).
+Each feature in \textit{WT} set is the difference between the scheduled starting time of a particular operation and its earliest possible starting time.
+The latter is the maximum out of the end processing time of the operation's predecessor and the time at which the required machine became available.
+\end{description}
 
@@
 Each row in the table represents an operation, and it contains its value according to each feature.
-The second column shows the \textit{Remain Processing Time (RPT)} of a job.
+The second column shows the \textit{Remaining Processing Time (RPT)} of a job.
 For instance, the first job $J_1$ consists of $3$ operations with a total processing time of $17$.
-In the beginning, the \textit{Remain Processing Time} of $O_{1,1}$ is $17$ because no operations are executed.
+In the beginning, the \textit{RPT} of $O_{1,1}$ is $17$ because no operations are executed.
 After finishing the first operation of job $1$, the \textit{RPT} became $8$ because the processing time of the completed operation $O_{1,1}$ is excluded.

### 2021-10-11 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
update the results section

@@
 \newcommand{\stest}{\textit{ST\_EST}}
+\newcommand{\stmtwr}{\textit{ST\_MTWR}}
+\newcommand{\stfifo}{\textit{ST\_FIFO}}
+
+\newcommand{\wtest}{\textit{WT\_EST}}
+\newcommand{\wtmtwr}{\textit{WT\_MTWR}}
+\newcommand{\wtfifo}{\textit{WT\_FIFO}}
 
@@
   \item[Starting Time (ST)] is a set of features representing the starting execution time of each operation by solving the problem using a heuristic.
-In our work, we consider EST (\stest), First-In-First-Out (ST\_FIFO), and Most Total Work Remaining (ST\_MTWR), see \cite{overview} for an overview.
-  To compute the starting time, we use a simple greedy algorithm \cite{tassel2021reinforcement,el2020job}, that at each step selects a not scheduled operation according to the heuristics and schedules it to a machine.
+In our work, we consider EST (\stest), First-In-First-Out (\stfifo), and Most Total Work Remaining (\stmtwr), see \cite{jones1998survey} for an overview.
+To compute the starting time, we use a simple greedy algorithm \cite{el2020job}, that at each step selects a not scheduled operation according to the heuristics and schedules it to a machine.
 In the case of FIFO, the algorithm selects an operation that appeared first in the queue.
-Similarly, for MTWR ...
+Similarly, for MTWR, the algorithm selects the operation that belongs to the job with the longest processing time needed to be processed.
 
-  \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used three different heuristics EST (WT\_EST), FIFO (WT\_FIFO), and MTWR (WT\_MTWR).
+  \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used three different heuristics EST (\wtest), FIFO (\wtfifo), and MTWR (\wtmtwr).
 Each feature in \textit{WT} set is the difference between the scheduled starting time of a particular operation and its earliest possible starting time.
@@
 In order to know which features have a significant effect on the results, we did some combinations of the features to perform the decomposition process clustering-based and then solved the problem using the Answer Set Programming model developed in \cite{el2020job} with a timeout of 10 minutes.
+We started with a simple combination that consists of $3$ features; \stfifo, \stmtwr, and \stest.
+At each step, we added another feature to get a new combination.
 The combinations of the selected features are listed as follows:
@@
 \fontsize{9pt}{10pt}
-\item \textbf{F1} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST \}}
-\item \textbf{F2} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT \}} 
-\item \textbf{F3} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, EST \}} 
-\item \textbf{F4} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, ML \}} 
-\item \textbf{F5} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT, EST \}}
-\item \textbf{F6} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, RPT, ML \}} 
-\item \textbf{F7} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, EST, ML \}} 
-\item \textbf{F8} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST \}} 
-\item \textbf{F9} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT \}} 
-\item \textbf{F10} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, EST \}} 
-\item \textbf{F11} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, ML \}}
-\item \textbf{F12} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, EST \}}
-\item \textbf{F13} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, ML \}}
-\item \textbf{F14} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, EST, ML \}}
-\item \textbf{F15} \textrightarrow \textit{\{ ST\textunderscore FIFO, ST\textunderscore MTWR, ST\textunderscore EST, WT\textunderscore FIFO, WT\textunderscore MTWR, WT\textunderscore EST, RPT, EST, ML \}}
+\item \textbf{F1} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest \}}
+\item \textbf{F2} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, RPT \}}
+\item \textbf{F3} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, EST \}} 
+\item \textbf{F4} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, ML \}} 
+\item \textbf{F5} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, RPT, EST \}}
+\item \textbf{F6} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, RPT, ML \}} 
+\item \textbf{F7} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, EST, ML \}} 
+\item \textbf{F8} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest \}} 
+\item \textbf{F9} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT \}} 
+\item \textbf{F10} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST \}} 
+\item \textbf{F11} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, ML \}}
+\item \textbf{F12} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, EST \}}
+\item \textbf{F13} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, ML \}}
+\item \textbf{F14} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST, ML \}}
+\item \textbf{F15} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, EST, ML \}}
 \end{enumerate}
 We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments.
-The following Table~\ref{tab4} shows the selected features and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
+The following Table~\ref{tab4} shows the selected features, and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
 The following table shows the results of our experiments.
-The first column represents the instances we tested, and the next four columns show the makespan values obtained by FIFO, MTWR, and (EST with single and multi-shot) heuristics.
-From the sixth column, we show the makespan values obtained by combining the features mentioned above.
-Our experiments found that the first five combinations do not provide good solutions compared to the other heuristics.
-However, features (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}, and \textbf{F15}) have a good impact in the cost function in the most cases.
-Especially, combinations \textbf{F6} and \textbf{F15} provided good results.
+The first column shows the approaches used for solving the problem in the literature in addition to the combinations of features we presented previously.
+The next columns represent the instances of our benchmark set except for the last column, which shows the average of each approach through the whole benchmark set.
+The first $4$ rows in the table show the values of each makespan obtained by heuristics; FIFO, MTWR, and (EST with a single and multi-shot).
+We can say that FIFO and MTWR provided the best results compared to the other two heuristics.
+However for the features combinations, the combinations (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}and \textbf{F15}) provided good results according to the cost function in the most cases especially with combinations (\textbf{F6} and \textbf{F15}).
 It means that the most important features extracted from the problem itself and have a positive impact on the performance are \textit{Remaining Processing Time, Machine Load and Earliest Starting Time}.
@@
 More specifically, combining \textit{Remaining Processing Time} with \textit{Machine Load} provides a good decomposition with/without \textit{Waiting Time} attributes obtained by the other heuristics.
-In a few instances, our proposed method did not outperform other heuristics, and the average makespan of \textbf{F6} was better than the average makespan obtained by the other heuristics and close to the FIFO makespan.
+In a few instances, our proposed method did not outperform other heuristics, however, in the last column, the best average makespan of the other approaches was $3331$ by FIFO, and the average makespan of our feature selection algorithm was $3256$.
+We can conclude that the feature selection algorithm outperformed other heuristics in the literature.
 The following section illustrates the difference between our proposed algorithm and the methods implemented by other researchers.
@@
       \textbf{Approach}   & TA51                           & TA52                           & TA53                           & TA54 & TA55                           & TA56                           & TA57                           & TA58 & TA59                           & TA60 & AVG  \\ \hline
-      \textbf{FIFO}        & 3549                           & 3339                           & 3160                           & 3218 & 3291                           & 3325                           & 3654                           & 3299 & 3344                           & 3129 & 3331 \\
-      \textbf{MTWR}        & 3364                           & 3304                           & 3168                           & 3494 & 3237                           & 3287                           & 3633                           & 3591 & 3394                           & 3257 & 3373 \\
+      \textbf{FIFO}        & 3549                           & 3339                           & \textbf{3160}              & \textbf{3218} & 3291               & 3325                           & 3654                           & \textbf{3299} & \textbf{3344}                           & \textbf{3129} & \textbf{3331} \\
+      \textbf{MTWR}        & \textbf{3364}           & \textbf{3304}              & 3168                           & 3494 & \textbf{3237}                           & \textbf{3287}               & \textbf{3633}              & 3591 & 3394                           & 3257 & 3373 \\
       \textbf{Single-shot} & 3632                           & 3615                           & 3481                           & 3462 & 3552                           & 3610                           & 3778                           & 3718 & 3613                           & 3550 & 3601 \\
@@
       \textbf{F1}          & 3506                           & 3277                           & 3382                           & 3414 & 3308                           & 3353                           & 3605                           & 3352 & 3453                           & 3483 & 3413 \\
-      \textbf{F2}          & 3362                           & 3318                           & 3585                           & 3425 & 3441                           & 3380                           & 3573                           & 3412 & 3416                           & 3315 & 3423 \\
+      \textbf{F2}          & 3362                           & 3318                           & 3585                           & 3425 & 3441                           & 3380                           & 3573                           & 3412 & 3416                           & \textbf{3315} & 3423 \\
       \textbf{F3}          & 3324                           & 3330                           & 3347                           & 3425 & 3424                           & 3548                           & 3601                           & 3370 & 3301                           & 3617 & 3429 \\
@@
       \textbf{F5}          & 3346                           & 3243                           & 3517                           & 3397 & 3355                           & 3383                           & 3650                           & 3707 & 3508                           & 3591 & 3470 \\
-      \textbf{F6}          & \textbf{3294} & 3275                           & 3250                           & 3265 & 3386                           & 3366                           & \textbf{3480} & 3324 & 3430                           & 3327 & 3340 \\
+      \textbf{F6}          & \textbf{3294} & 3275                           & 3250                           & 3265 & 3386                           & 3366                           & \textbf{3480} & \textbf{3324} & 3430                           & 3327 & \textbf{3340} \\
       \textbf{F7}          & 3328                           & 3226                           & 3338                           & 3353 & 3301                           & 3564                           & 3592                           & 3496 & 3352                           & 3399 & 3395 \\
@@
       \textbf{F13}         & 3624                           & 3505                           & 3427                           & 3236 & 3370                           & 3338                           & 3749                           & 3437 & 3253                           & 3659 & 3460 \\
-      \textbf{F14}         & 3554                           & 3373                           & 3524                           & 3232 & 3385                           & 3483                           & 3573                           & 3365 & 3284                           & 3767 & 3454 \\
+      \textbf{F14}         & 3554                           & 3373                           & 3524                           & \textbf{3232} & 3385                           & 3483                           & 3573                           & 3365 & 3284                           & 3767 & 3454 \\
       \textbf{F15}         & 3509                           & \textbf{3184} & 3380                           & 3289 & \textbf{3218} & 3389                           & \textbf{3480} & 3481 & 3290                           & 3569 & 3379 \\ \hline
+      \textbf{Obtained} & 3294                           & 3184 & 3143                           & 3232 & 3218 & 3229                           & 3480 & 3324 & 3145                           & 3315 & \textbf{3256} \\
+\hline
       \end{tabular}

### 2021-10-12 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
authors update

@@
 Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958} \and
-Martin Gebser\inst{1,2}\orcidID{0000--0000-0000-0000}}
+Martin Gebser\inst{1,2}\orcidID{0000-0002-8010-4752}}
 %
@@
 %
-\institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
-\email{\{mohammed.el-kholany, konstantin.schekotihin and martin.gebser\}@aau.at}\\
-%\url{http://www.springer.com/gp/computer-science/lncs} \and
-Technische Universität Graz, Graz, Austria\\
-\email{mgebser@ist.tugraz.at}}
+\institute{Alpen-Adria-Universität Klagenfurt, % Klagenfurt, 
+Austria\\% 
+% \email{\{mohammed.el-kholany, konstantin.schekotihin and martin.gebser\}@aau.at}%\\
+%\url{http://www.springer.com/gp/computer-science/lncs}
+\and
+Technische Universität Graz, % Graz, 
+Austria\\%
+% \email{mgebser@ist.tugraz.at}
+\email{firstname.lastname@aau.at}}
 %

### 2021-10-12 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
abstract somewhat reformulated

@@
 \begin{abstract}
-Scheduling is a crucial problem appearing in various domains, such as manufacturing, transportation, or healthcare.
-In most problem definitions, the goal is to schedule a given set of operations on available resources to complete the operations as early as possible.
-Unfortunately, most scheduling problems cannot be solved efficiently.
-Therefore, the research of suitable approximation methods is of primary importance.
+Scheduling is a crucial problem appearing in various domains, such as manufacturing, transportation, or healthcare, where % . In most problem definitions, 
+the goal is to schedule % a given set of 
+given operations on available resources such that the operations are completed as soon as possible. Unfortunately, most scheduling problems cannot be solved efficiently, % Therefore, the 
+so that research on suitable approximation methods is of primary importance.
+%
+This work introduces a novel approximation approach based on problem decomposition with data mining methodologies.
+We propose a constrained clustering algorithm to group operations into clusters, corresponding to time windows in which the operations must be scheduled.
+The decomposition process consists of two main phases. 
+First, features are extracted, either from the problem itself or from solutions obtained by heuristic methods, to predict the execution sequence of operations on each resource.
+% The first phase is to extract features to predict the sequence of operations on each resource.
+These features are extracted either from the problem itself or from solutions obtained by other heuristics.
+The second phase % is to develop
+deploys our constrained clustering algorithm to assign each operation into a time window.
+We then schedule the operations by time windows using Answer Set Programming.
+Evaluation results show that our proposed approach outperforms other heuristic schedulers in most cases, where incorporating features like \textit{Remaining Processing Time}, \textit{Machine Load}, and \textit{Earliest Starting Time} significantly improves the solution quality.
 
-This work suggests a novel approximation approach based on problem decomposition with data mining methodologies.
-This study proposes a constrained clustering algorithm to group the operations into clusters corresponding to time windows in which these operations must be scheduled.
-The decomposition process depends on two main phases.
-The first phase is to extract features to predict the sequence of the operations on each resource.
-These features are extracted either from the problem itself or from solutions obtained by other heuristics.
-The second phase is to develop a constrained clustering algorithm to assign each operation into a time window.
-We solved the problem using the Answer Set Programming.
-Evaluation results show that our proposed outperformed other heuristic schedulers in most cases, where features, like \textit{Remaining Processing Time}, \textit{Machine Load}, and \textit{Earliest Starting Time}, contributed significantly to the solution quality.
-
-\keywords{Job-shop Scheduling  Problem \and Constrained Clustering \and Time Windows \and Answer Set Programming.}
+\keywords{Job-shop Scheduling  Problem \and Constrained Clustering \and Time Windows \and Answer Set Programming}
 \end{abstract}

### 2021-10-12 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
start of introduction checked

@@
  \cite{pezzella2008genetic,chaudhry2016research,nouiri2018effective,demirbilek2019dynamically,schoenfelder2020nurse,wang2019routing,janakbhai2021blockchain}. 
+%
+Such applications result in different scheduling problem definitions, with the Job-shop Scheduling Problem (JSP) \cite{taillard1993benchmarks} being one if not the most well-known variant.
+In JSP, operations of given jobs must be scheduled on available machines in an optimal way wrt.\ some pre-defined criteria.
+The latter include, for instance, minimization of makespan, i.e., the latest completion time of any job, or tardiness, i.e., the sum of delays over all jobs completed after their deadlines.
 
-These applications resulted in different definitions of scheduling problems, with Job-shop Scheduling Problem (JSP) being among the most popular ones.
-In JSP, operations of given jobs must be scheduled on available machines in an optimal way wrt.\ some pre-defined criteria.
-The latter include, for instance, minimization of makespan, i.e., the completion time of the last job, or tardiness, i.e., the sum of delays for all jobs completed after their deadlines.
-
-However, JSP is a NP-complete combinatorial optimization problem \cite{DBLP:journals/mor/GareyJS76,SOTSKOV1995237} for which no efficient algorithms are known.
-Therefore, searching for an optimal solution with modern approaches like Answer Set Programming (ASP), Mixed Integer Programming, or Constraint Programming \cite{meng2020mixed,el2020job,al2017job}
+However, JSP is an NP-complete combinatorial optimization problem \cite{DBLP:journals/mor/GareyJS76,SOTSKOV1995237} for which no efficient algorithms are known.
+Therefore, searching for an optimal solution with state-of-the-art solvers for Answer Set Programming (ASP), Mixed Integer Programming, or Constraint Programming \cite{meng2020mixed,coltep19a,el2020job,al2017job}
 %\todo{cite at least 1 paper solving JSP with these methods} 
-often takes too much computation time, even for small instances. Practical applications require solving JSP instances of vast scale with thousands of operations \cite{zhang2010hybrid}.
-As a result, many researchers focused on the development of efficient algorithms for finding high-quality approximations of the optimal schedule, including dispatching rules and other heuristic approaches~\cite{blackstone1982state}, or stochastic optimization techniques \cite{DBLP:journals/informs/VaessensAL96,DBLP:journals/jim/CalisB15}.
+often takes too much computation time, even for seemingly small instances.
+Practical applications instead necessitate solving JSP instances of large scale with thousands of operations~\cite{zhang2010hybrid}.
+As a result, much research work focuses on % the development of 
+efficient methods for finding high-quality approximations of optimal schedules, including dispatching rules and other heuristic approaches~\cite{blackstone1982state} as well as stochastic optimization techniques \cite{DBLP:journals/informs/VaessensAL96,DBLP:journals/jim/CalisB15}.
 The main issue of these approaches is that they require manual parametrization for a particular scheduling problem, which is tedious and error-prone.
-For instance, heuristic approaches might require the development of new heuristics or their combinations.
-Similarly, default parameters of stochastic methods might result in a finding of mediocre schedules.
-As a result, much focus of the recent research lies on the automation of the parametrization using machine learning techniques~\cite{bengio2020machine}.
+For instance, adapting heuristic methods might involve the development of new dispatching rules or specific combinations of existing ones.
+Similarly, merely utilizing default parameters for stochastic techniques might lead to mediocre schedules.
+In view of these challenges, recent research interest lies on the automated parametrization of approximation methods using machine learning methodologies~\cite{bengio2020machine}.
 
@@
 \section{Problem Definition}
-The Job-shop Scheduling Problem (JSP) is one of the well-known scheduling problems \cite{baker1974introduction,lenstra1979computational}.
+The Job-shop Scheduling Problem (JSP) is one of the well-known scheduling problems \cite{baker1974introduction,lenstra1979computational,taillard1993benchmarks}.
 An instance of a JSP comprises a set $J = \{ J_1, J_2, \dots ,J_n \}$  of jobs and a set $M = \{ M_1, M_2, \dots ,M_m \}$ of Machines.

### 2021-10-12 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
went over introduction

@@
 Similarly, merely utilizing default parameters for stochastic techniques might lead to mediocre schedules.
-In view of these challenges, recent research interest lies on the automated parametrization of approximation methods using machine learning methodologies~\cite{bengio2020machine}.
+In view of these challenges, recent research interest lies on the automatic parametrization of approximation methods using machine learning methodologies~\cite{bengio2020machine}.
 
@@
 
-In this work, we suggest a clustering-based method that automatically decomposes a JSP instance into sub-instances which can be tackled by 
-optimization algorithms of an ASP solver. 
-General clustering algorithms are unsupervised learning methods that partition a given set of objects into disjoint clusters, where each cluster comprises closest objects wrt.\ some distance function.
-In scheduling settings, however, clusters must satisfy additional constraints, which define precedence relations between operations of a job.
-Standard constrained clustering algorithms are not applicable in scheduling scenarios since they usually allow for the definition of disjointness constraints, which specify objects that cannot appear in one cluster \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}.
+In this work, we introduce a method based on clustering to automatically decomposes a JSP instance into several time windows that are small enough for optimization by an ASP solver.
+General clustering algorithms are unsupervised learning methods that partition a given set of objects into disjoint clusters, where each cluster comprises close objects wrt.\ some distance measure. 
+In scheduling settings, however, clusters must satisfy additional constraints implied by the precedence relation between operations of a job.
+Standard constrained clustering algorithms are not readily applicable to scheduling scenarios either, as they are limited to disjointness constraints specifying objects that must not appear together in a cluster \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}.
 Therefore, our clustering method implements a novel type of constraints that
 \begin{enumerate*}[label=\emph{(\roman*)}]
-  \item prevent the assignment of an operation to a cluster if its preceding operation is not yet assigned to the same or previous cluster, and
-  \item ensure balancing between the clusters according to the number of operations per cluster.
+  \item prevent the assignment of an operation to a cluster if its preceding operation is not yet assigned to the same or a previous cluster and
+  \item ensure balancing between clusters according to the target number of operations per cluster.
 \end{enumerate*}
-Further contributions of this paper can be summarized as follows:
+Further contributions of our work can be summarized as follows:
 \begin{itemize}
-  \item Since a typical JSP instance describing jobs, their operations, and available machines, does not provide enough information for a clustering algorithm to find a good decomposition, we suggest various heuristic feature extractors, such as First-In-First-Out or machine load.
-  \item We implement the suggested constrained clustering method and combine it with the forward selection of features.
-The latter is an automatic procedure aiming to find a subset of features allowing the clustering algorithm to compute a decomposition resulting in the best quality schedule.
-  \item Evaluation of our method on the Taillard benchmark instances \cite{taillard1993benchmarks} showed that it significantly outperforms baseline heuristic methods and pure ASP optimization algorithms wrt.\ the makespan optimization criteria within the timeout of 600 seconds.
+  \item Since a typical JSP instance describing jobs, their operations, and available machines does not provide sufficient information by itself for finding some promising decomposition by a clustering algorithm, we incorporate heuristic approaches like \textit{First-In-First-Out} and \textit{Machine Load} to extract features from their solutions.
+  \item We implement the proposed constrained clustering algorithm and combine it with the forward selection of features, which is an automatic method for identifying a subset of features allowing the clustering method to compute decompositions resulting in best-quality schedules.
+  \item The evaluation of our approach on Taillard's benchmark instances \cite{taillard1993benchmarks} shows that it significantly outperforms baseline heuristic methods and pure ASP optimization wrt.\ the makespan optimization criterion within a timeout limit of 600 seconds.
 \end{itemize}
@@
 
-The paper is organized as follows: the following section provides the definition of JSP and gives an illustrative example of input data.
-Section \ref{sec:features} discusses heuristic feature extractors, followed by Section \ref{sec:method}, which presents the proposed constrained clustering algorithm.
-Section \ref{sec:eval} presents an extensive computational study on benchmark instances, comparing our results with the baseline solvers.
-A literature review is given in Section \ref{sec:literature}, followed by the conclusions and future work.
+The rest of this paper is organized as follows.
+Section~\ref{sec:problem} introduces JSP along with a running example.
+In Section \ref{sec:features}, we describe the feature extraction, including heuristic methods to obtain corresponding reference solutions.
+Section~\ref{sec:method} presents our proposed constrained clustering algorithm.
+In Section~\ref{sec:eval}, we empirically evaluate our approach and compare it to baseline heuristic methods as well as pure ASP optimization.
+Section~\ref{sec:literature} then surveys related work, followed by conclusions and future work in Section~\ref{sec:conclusions}.
 
-\section{Problem Definition}
+\section{Problem Definition}\label{sec:problem}
 The Job-shop Scheduling Problem (JSP) is one of the well-known scheduling problems \cite{baker1974introduction,lenstra1979computational,taillard1993benchmarks}.

### 2021-10-12 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
Features

--- 2021-10-12 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958} \and
-Martin Gebser\inst{1,2}\orcidID{0000-0002-8010-4752}}
+Martin Gebser\inst{1,2}\orcidID{0000--0000-0000-0000}}
 %
@@
 %
-\institute{Alpen-Adria-Universität Klagenfurt, % Klagenfurt, 
-Austria\\% 
-% \email{\{mohammed.el-kholany, konstantin.schekotihin and martin.gebser\}@aau.at}%\\
-%\url{http://www.springer.com/gp/computer-science/lncs}
-\and
-Technische Universität Graz, % Graz, 
-Austria\\%
-% \email{mgebser@ist.tugraz.at}
-\email{firstname.lastname@aau.at}}
+\institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
+\email{\{mohammed.el-kholany, konstantin.schekotihin and martin.gebser\}@aau.at}\\
+%\url{http://www.springer.com/gp/computer-science/lncs} \and
+Technische Universität Graz, Graz, Austria\\
+\email{mgebser@ist.tugraz.at}}
 %
@@
 \begin{abstract}
-Scheduling is a crucial problem appearing in various domains, such as manufacturing, transportation, or healthcare, where % . In most problem definitions, 
-the goal is to schedule % a given set of 
-given operations on available resources such that the operations are completed as soon as possible. Unfortunately, most scheduling problems cannot be solved efficiently, % Therefore, the 
-so that research on suitable approximation methods is of primary importance.
-%
-This work introduces a novel approximation approach based on problem decomposition with data mining methodologies.
-We propose a constrained clustering algorithm to group operations into clusters, corresponding to time windows in which the operations must be scheduled.
-The decomposition process consists of two main phases. 
-First, features are extracted, either from the problem itself or from solutions obtained by heuristic methods, to predict the execution sequence of operations on each resource.
-% The first phase is to extract features to predict the sequence of operations on each resource.
+Scheduling is a crucial problem appearing in various domains, such as manufacturing, transportation, or healthcare.
+In most problem definitions, the goal is to schedule a given set of operations on available resources to complete the operations as early as possible.
+Unfortunately, most scheduling problems cannot be solved efficiently.
+Therefore, the research of suitable approximation methods is of primary importance.
+
+This work suggests a novel approximation approach based on problem decomposition with data mining methodologies.
+This study proposes a constrained clustering algorithm to group the operations into clusters corresponding to time windows in which these operations must be scheduled.
+The decomposition process depends on two main phases.
+The first phase is to extract features to predict the sequence of the operations on each resource.
 These features are extracted either from the problem itself or from solutions obtained by other heuristics.
-The second phase % is to develop
-deploys our constrained clustering algorithm to assign each operation into a time window.
-We then schedule the operations by time windows using Answer Set Programming.
-Evaluation results show that our proposed approach outperforms other heuristic schedulers in most cases, where incorporating features like \textit{Remaining Processing Time}, \textit{Machine Load}, and \textit{Earliest Starting Time} significantly improves the solution quality.
+The second phase is to develop a constrained clustering algorithm to assign each operation into a time window.
+We solved the problem using the Answer Set Programming.
+Evaluation results show that our proposed outperformed other heuristic schedulers in most cases, where features, like \textit{Remaining Processing Time}, \textit{Machine Load}, and \textit{Earliest Starting Time}, contributed significantly to the solution quality.
 
-\keywords{Job-shop Scheduling  Problem \and Constrained Clustering \and Time Windows \and Answer Set Programming}
+\keywords{Job-shop Scheduling  Problem \and Constrained Clustering \and Time Windows \and Answer Set Programming.}
 \end{abstract}
@@
 
-\section{Features extraction}
+\section{Feature extraction}
 \label{sec:features}
-Data mining merges with machine learning and statistical techniques to gain information and new knowledge from the databases\cite{ismail2012production,nasiri2019data}.
-Examples of data mining tasks are classification, clustering, and sequential pattern mining.
-In this work, we aim to use clustering to study the operations' pattern to find a good sequencing of the operations on each machine that leads us to obtain a near-optimal solution of the JSP.
-In order to use the clustering approach, we should have some features which describe the characteristics of the problem.
-These features would aid us in the decision-making process, hence, reaching a near-optimal solution.
-\\
+Application of machine learning methods to scheduling problems requires careful selection of data describing the hidden dependencies between operations of different jobs \cite{ismail2012production,nasiri2019data}.
+Clustering methods, which we aim to apply in our approach, are not an exception.
+Thus, a clustering method requires an informative set of features describing jobs, their operations, and machines of an instance to find patterns resulting in a beneficial decomposition of all operations in time windows.
 
-Some of the previous work \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} have extracted the following features: priority, processing time, remaining processing time, machine load, and route position.
-Most of the researchers converted the attributes into qualitative data.
-As mentioned in these studies, the main reason for this conversion is to make the decision rules more generic, which can be applied to several scheduling problems.
-However, in our case, we will not convert the data into classes because we will use these features to measure the similarities/dissimilarities between the operations in the clustering process, which requires having numerical values.
-We present the features that we considered:
-%\paragraph{Process.}
-%The two attributes \textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are dependent on the problem domain.
-The \textit{PT} represents the execution time of a particular operation in a machine.
-However, \textit{RPT} is the aggregate processing time for all subsequent operations of that job.
-\todo[inline]{we need to provide rationale behind the features, why do we think they are useful for the clusering algorithm.
-Example of values extracted for the sample instance in Table 1 would also improve the readability of the paper.}
-%\todo[inline]{Why in all other paragraphs we are talking only about one feature at a time and here about two.
-Can we split them?
-Like \\paragraph\{Processing Time (PT)\} and \\paragraph\{Remain Processing time (RPT)\}}
+Methods suggested in the previous work \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} used characterizations of scheduling instances based on the following features: priority, processing time, remaining processing time, machine load, and route position.
+Most of the approaches converted the quantitative values of the attributes into qualitative data.
+As mentioned in these studies, the main reason for this conversion was to make the heuristic dispatching rules more generic and allow their application to several scheduling problems.
+In this work, we suggest a method that can work directly on the sets of features and does not require any problem-specific transformations.
+However, our clustering method for JSP instance decomposition requires all features to have numerical values allowing distance measures to estimate similarities/dissimilarities between operations of jobs.
+%
+Therefore, we suggest the following features:
 \begin{description}
   \item[Processing Time (PT)] that represents the execution time of a particular operation in a machine.
-
-  \item[Remaining Processing Time (RPT)] attribute corresponds to the total processing time for all subsequent operations of a job.
+This feature is provided in every JSP instance, like in the sample instance presented in Table \ref{tab1}.
 
@@
 
+  \item[Remaining Processing Time (RPT)] attribute corresponds to the total processing time for all subsequent operations of a job.
+For example, Table~\ref{tab3} shows a set of \textit{RPT} values computed for the sample instance.
+Since the first job $J_1$ consists of $3$ operations with a total processing time of $17$, \textit{RPT} of $O_{1,1}$ is $17$ because it has no preceding operations.
+After finishing the first operation of $J_1$, the \textit{RPT} becomes $8$ because the processing time of the completed operation $O_{1,1}$ is subtracted.
+Similarly, when the second operation of $J_1$ is completed, the \textit{RPT} of the last operation is $5$.
+
   \item[Time Length of a Job (TLJ)] is an attribute describing the total processing time needed to execute all operations of a job.
+Since all operations of a job have the same feature value, this feature can only be used to differentiate between jobs.
   %It is the summation of processing time for all operations of that job.
@@
 It is calculated by accumulating the processing time of all preceding operations.
-This feature has a specific value $\mathit{Zero}$ if the operation has no predecessors.
+This feature has a specific value $\mathit{Zero}$ if the operation has no predecessors.\todo{This value does not appear in the table, why?}
 
@@
 In the initial state, \textit{ML} corresponds to the total processing time of all operations assigned to that machine.
-For each step, the load of a machine is decreased by the processing time of an operation assigned to the machine and has the smallest \textit{EST}.
+In each time step, the load of a machine is decreased by the processing time of an operation assigned to the machine and has the smallest \textit{EST}.
+For example, machine $3$ has to process operations $\{ O_{1,2}, O_{2,1}, O_{3,2} \}$.
+The load of machine $3$ at time $0$ is 10.
+We assumed that the machine would start processing the operation $O_{2,1}$ that has the smallest \textit{EST}.
+After completing the operation $O_{2,1}$, the machine will target the next operation, which has the smallest value of \textit{EST} and the load of the machine is decreased by the processing time of the completed operation $O_{2,1}$, and so on till completing all the operations assigned to a particular machine.
 
@@
 In our work, we consider EST (\stest), First-In-First-Out (\stfifo), and Most Total Work Remaining (\stmtwr), see \cite{jones1998survey} for an overview.
-To compute the starting time, we use a simple greedy algorithm \cite{el2020job}, that at each step selects a not scheduled operation according to the heuristics and schedules it to a machine.
+  To compute the starting time, we use a simple greedy algorithm \cite{el2020job} that at each step selects a not scheduled operation according to the heuristics and schedules it to a machine.
 In the case of FIFO, the algorithm selects an operation that appeared first in the queue.
-Similarly, for MTWR, the algorithm selects the operation that belongs to the job with the longest processing time needed to be processed.
+Similarly, for MTWR, the algorithm selects the operation that belongs to the job with the longest processing time from the queue.
 
-  \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used three different heuristics EST (\wtest), FIFO (\wtfifo), and MTWR (\wtmtwr).
+  \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used the same greedy algorithm as above with three different heuristics EST (\wtest), FIFO (\wtfifo), and MTWR (\wtmtwr).
 Each feature in \textit{WT} set is the difference between the scheduled starting time of a particular operation and its earliest possible starting time.
-The latter is the maximum out of the end processing time of the operation's predecessor and the time at which the required machine became available.
+The latter is the maximum out of the end processing time of the operation's predecessor and the earliest time at which the required machine became available.
+
+  \todo{It would be nice to add examples of other features to the table. If this is complicated, then only of \wtest and \stest.}
 \end{description}
 
-The following Table~\ref{tab3} shows a sample of the features that we extracted for the decomposition process.
-Each row in the table represents an operation, and it contains its value according to each feature.
-The second column shows the \textit{Remaining Processing Time (RPT)} of a job.
-For instance, the first job $J_1$ consists of $3$ operations with a total processing time of $17$.
-In the beginning, the \textit{RPT} of $O_{1,1}$ is $17$ because no operations are executed.
-After finishing the first operation of job $1$, the \textit{RPT} became $8$ because the processing time of the completed operation $O_{1,1}$ is excluded.
-When the second operation of job $1$ is completed, the \textit{RPT} of that job is $5$.
-\\
-
-The third column represents the \textit{Earliest Starting Time} which we described earlier.
-The last column is the \textit{Machine Load} which is the summation of the processing time for all operations assigned to a machine.
-For example, Machine $3$ has to process operations $\{ O_{1,2}, O_{2,1}, O_{3,2} \}$, the load of machine $3$ at time $0$ is 10.
-We assumed that the machine would start processing the operation $O_{2,1}$ that has the smallest \textit{EST}.
-After completing the operation $O_{2,1}$, the machine will target the next operation, which has the smallest value of \textit{EST} and the load of the machine is decreased by the processing time of the completed operation $O_{2,1}$, and so on till completing all the operations assigned to a particular machine.
-
 \begin{table}[h!]
-\fontsize{12}{9}
 \setlength{\tabcolsep}{10.0pt}
@@
   \begin{center}
-  \caption{Time Window Assignment.}
+  \caption{Features extracted for the sample problem instance.}
   \label{tab3}

### 2021-10-13 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
Updating features extraction section

@@
 It is calculated by accumulating the processing time of all preceding operations.
-This feature has a specific value $\mathit{Zero}$ if the operation has no predecessors.\todo{This value does not appear in the table, why?}
+This feature has a specific value $0$ if the operation has no predecessors.%\todo{This value does not appear in the table, why?}
 
@@
 In the case of FIFO, the algorithm selects an operation that appeared first in the queue.
+For example, $O_{2,2}$ and $O_{3,3}$ are in the queue at time $9$ on machine $2$.
+Since $O_{2,2}$ and $O_{3,3}$ were ready to be processed at time $4$ and $7$, respectively.
+The algorithm selects $O_{2,2}$ to process at time $9$ before $O_{3,3}$ which starts at time $15$.
 Similarly, for MTWR, the algorithm selects the operation that belongs to the job with the longest processing time from the queue.
@@
   \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used the same greedy algorithm as above with three different heuristics EST (\wtest), FIFO (\wtfifo), and MTWR (\wtmtwr).
-Each feature in \textit{WT} set is the difference between the scheduled starting time of a particular operation and its earliest possible starting time.
-The latter is the maximum out of the end processing time of the operation's predecessor and the earliest time at which the required machine became available.
+Each feature in \textit{WT} set is the difference between the scheduled starting time of a particular operation and ending time of its predecessor.
+%The latter is the maximum out of the end processing time of the operation's predecessor and the earliest time at which the required machine became available.
+  As shown in Table~\ref{tab3}, the predecessor of $O_{2,2}$ which is $O_{2,1}$ is finished at time $4$.
+However machine $2$ that executes $O_{2,2}$ was not free at that time, so $O_{2,2}$ had to wait $5$ time units to start the processing.
 
-  \todo{It would be nice to add examples of other features to the table. If this is complicated, then only of \wtest and \stest.}
+  %\todo{It would be nice to add examples of other features to the table. If this is complicated, then only of \wtest and \stest.}
 \end{description}
@@
  %\begin{adjustwidth}{0.0cm}{}
-    \begin{tabular}{c  c  c  c}
-      \textbf{Operation} & \textbf{RPT} & \textbf{EST} & \textbf{ML}\\
+    \begin{tabular}{c  c  c  c  c  c}
+      \textbf{Operation} & \textbf{RPT} & \textbf{EST} & \textbf{ML} & \textbf{\stfifo} & \textbf{\wtfifo}\\
       \hline
 										\\
-      $O_{1,1}$  & 17 & 0   & 20	\\
-      $O_{1,2}$  & 8  & 9   & 3		\\
-      $O_{1,3}$  & 5  & 12 & 5		\\
+      $O_{1,1}$  & 17 & 0   & 20	&  0  & 0\\
+      $O_{1,2}$  & 8  & 9   & 3		&  9  & 0\\
+      $O_{1,3}$  & 5  & 12  & 5	  &  12 & 0\\
 										\\
-      $O_{2,1}$  & 12 & 0   & 10	\\
-      $O_{2,2}$  & 8  & 4   & 11	\\
-      $O_{2,3}$  & 2  & 10 & 7		\\
+      $O_{2,1}$  & 12 & 0   & 10	&  0  & 0\\
+      $O_{2,2}$  & 8  & 4   & 11	&  9  & 5\\
+      $O_{2,3}$  & 2  & 10  & 7	  &  17 & 2\\
 					    					\\
-      $O_{3,1}$  & 12 & 0  & 11	\\
-      $O_{3,2}$  & 8  & 4  & 6		\\
-      $O_{3,3}$  & 5  & 7  & 5		\\
+      $O_{3,1}$  & 12 & 0  & 11	  &  0  & 0\\
+      $O_{3,2}$  & 8  & 4  & 6		&  4  & 0\\
+      $O_{3,3}$  & 5  & 7  & 5		&  15 & 8\\
     \end{tabular}

### 2021-10-13 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
started to look at Problem Definition

@@
 Such applications result in different scheduling problem definitions, with the Job-shop Scheduling Problem (JSP) \cite{taillard1993benchmarks} being one if not the most well-known variant.
-In JSP, operations of given jobs must be scheduled on available machines in an optimal way wrt.\ some pre-defined criteria.
+In JSP, operations of given jobs must be scheduled on available machines in an optimal way wrt.\ some predefined criteria.
 The latter include, for instance, minimization of makespan, i.e., the latest completion time of any job, or tardiness, i.e., the sum of delays over all jobs completed after their deadlines.
@@
 \section{Problem Definition}\label{sec:problem}
-The Job-shop Scheduling Problem (JSP) is one of the well-known scheduling problems \cite{baker1974introduction,lenstra1979computational,taillard1993benchmarks}.
-An instance of a JSP comprises a set $J = \{ J_1, J_2, \dots ,J_n \}$  of jobs and a set $M = \{ M_1, M_2, \dots ,M_m \}$ of Machines.
-Each job $J_i$ has a sequence $(O_{i,1}, O_{i,2}, \dots ,O_{i,m})$ of operations that must be processed in a predefined order.
-Each machine $M_i$ executes an operation of each job with a known and fixed processing time.
-The main objective of a JSP solver is to find a schedule that minimizes some criteria like makespan, i.e., the total completion time for all jobs.
-In addition, JSP makes the following assumptions:
-\begin{itemize}
-	\item Once an operation starts, it must be completed without interruptions.
-	\item Each job has no duplicated operations, i.e., visits a machine only once.
-	\item A machine cannot perform more than one operation at a time.
-	\item All machines are available at all times.
-	\item Processing the jobs starts at time 0.
-\end{itemize}
+The Job-shop Scheduling Problem (JSP) is one of the most well-known scheduling problems \cite{baker1974introduction,lenstra1979computational,taillard1993benchmarks}.
+A JSP instance comprises a set $J = \{ J_1, J_2, \dots ,J_n \}$  of jobs and a set $M = \{ M_1, M_2, \dots ,M_m \}$ of machines.
+Each job $J_i$ consists of a sequence $(O_{i,1},\linebreak[1] O_{i,2},\linebreak[1] \dots ,O_{i,m})$ of operations that must be processed in the given order.
+Each machine $M_i$ executes one operation per job with a predefined, fixed processing time.
+Once the execution of an operation is started, it cannot be interrupted,
+and each machine can perform at most one operation at a time.
+The main objective of JSP solving algorithms is to find a schedule that minimizes optimization criteria,
+where we focus on the makespan, i.e., the latest completion time of any job.
+% In addition, JSP solutions are subject to the following conditions:
+% \begin{itemize}
+% 	\item Once an operation starts, it must be completed without interruptions.
+%	\item Each job has no duplicated operations, i.e., visits a machine only once.
+% 	\item A machine cannot perform more than one operation at a time.
+% 	\item All machines are available at all times.
+% 	\item Processing the jobs starts at time 0.
+% \end{itemize}
 
-Let us illustrate the problem on a small example presented in Table~\ref{tab1}, which shows the information of a simple JSP comprising $3$ jobs and $3$ machines.
+Let us illustrate the problem on the small JSP instance specified in Table~\ref{tab1}, which provides parameters for $3$ jobs and $3$ machines.
 In this table, rows correspond to operations, whereas columns refer to the name of the operation, the machine assigned to an operation, and the processing time of a particular operation.
@@
 
-\begin{table}
-\begin{minipage}{.6\textwidth}
+\begin{table}[t]
+%  \centering
+\begin{minipage}{.56\textwidth}
 %\fontsize{12}{9}
@@
     \begin{tabular}{c c c}
-      \textbf{Operation} & \textbf{Machine} & \textbf{Procesing Time} \\
+      \textbf{Operation} & \textbf{Machine} & \textbf{Processing Time} \\
       \hline

### 2021-10-13 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
extended discussion of example

@@
 
-\section{Problem Definition}\label{sec:problem}
+\section{Job-shop Scheduling Problem}\label{sec:problem}
 The Job-shop Scheduling Problem (JSP) is one of the most well-known scheduling problems \cite{baker1974introduction,lenstra1979computational,taillard1993benchmarks}.
@@
 Let us illustrate the problem on the small JSP instance specified in Table~\ref{tab1}, which provides parameters for $3$ jobs and $3$ machines. 
-In this table, rows correspond to operations, whereas columns refer to the name of the operation, the machine assigned to an operation, and the processing time of a particular operation.
+The rows list operations along with their respective machines and processing times; e.g., the third operation of the first job, $O_{1,3}$, takes $5$ time units for execution by machine~$1$.
+The minimum makespan happens to be $20$, which matches the sum $9+6+5$ of processing times for the operations $O_{1,1}$, $O_{2,2}$, and $O_{3,3}$ executed by machine~$2$.
+While there are plenty, i.e., $234$, optimal schedules with makespan $20$,
+they agree on the execution orders $(O_{3,1}, O_{1,3}, O_{2,3})$ and $(O_{1,1}, O_{2,2}, O_{3,3})$
+of operations processed by machine~$1$ or~$2$, respectively.
+The two feasible execution orders for machine $3$ are
+$(O_{2,1}, O_{3,2}, O_{1,2})$ and
+$(O_{2,1}, O_{1,2}, O_{3,2})$.
+In both cases, the earliest eligible starting times for the operations
+$O_{1,1}$, $O_{1,2}$, and $O_{1,3}$ of job $J_1$ are
+$0$, $9$, and $12$,
+as well as
+$0$, $9$, and $17$ for $O_{2,1}$, $O_{2,2}$, and $O_{2,3}$
+belonging to the job~$J_2$.
+Regarding the operations of~$J_3$,
+the earliest starting time for $O_{3,1}$ is~$0$,
+$O_{3,3}$ can only be started at time~$15$ because its machine~$1$
+is occupied before,
+while either $4$ or $12$ is the earliest eligible starting time for $O_{3,2}$,
+depending on whether it is processed directly after completing $O_{3,1}$
+(and $O_{2,1}$ on its machine~$3$) or waits for the completion of $O_{1,2}$
+on machine~$3$.
+The latter option lets machine~$3$ idle unnecessarily 
+and may seem less attractive,
+yet it does not stretch the resulting makespan beyond time~$20$. 
+%
+% In this table, rows correspond to operations, whereas columns refer to the name of the operation, the machine assigned to an operation, and the processing time of a particular operation.
 For instance, in the third row of the table, $O_{1,3}$ is the third operation of the first job that will be executed by machine $1$ with processing time $5$.
-
+%
 \begin{table}[t]
@@
 								\\
-      $O_{1,1}$  & 2  &  9  	\\
-      $O_{1,2}$  & 3  &  3  	\\
-      $O_{1,3}$  & 1  &  5  	\\
+      $O_{1,1}$  & $2$  &  $9$  	\\
+      $O_{1,2}$  & $3$  &  $3$  	\\
+      $O_{1,3}$  & $1$  &  $5$ 	\\
 								\\
-      $O_{2,1}$  & 3  & 4  	\\
-      $O_{2,2}$  & 2  & 6  	\\
-      $O_{2,3}$  & 1  & 2  	\\
+      $O_{2,1}$  & $3$  & $4$  	\\
+      $O_{2,2}$  & $2$  & $6$  	\\
+      $O_{2,3}$  & $1$  & $2$  	\\
 								\\
-      $O_{3,1}$  & 1  & 4  	\\
-      $O_{3,2}$  & 3  & 3  	\\
-      $O_{3,3}$  & 2  & 5  	\\
+      $O_{3,1}$  & $1$  & $4$  	\\
+      $O_{3,2}$  & $3$  & $3$  	\\
+      $O_{3,3}$  & $2$  & $5$  	\\
     \end{tabular}
@@
 							\\
-      $O_{1,1}$  & 1    	\\
-      $O_{1,2}$  & 1    	\\
-      $O_{2,1}$  & 1    	\\
-      $O_{2,2}$  & 1   	\\
-      $O_{3,1}$  & 1		\\
+      $O_{1,1}$  & $1$    	\\
+      $O_{1,2}$  & $1$    	\\
+      $O_{2,1}$  & $1$    	\\
+      $O_{2,2}$  & $1$   	\\
+      $O_{3,1}$  & $1$		\\
 					    		\\
-      $O_{1,3}$  & 2  	\\
-      $O_{2,3}$  & 2		\\
-      $O_{3,2}$  & 2    	\\
-      $O_{3,3}$  & 2    	\\
+      $O_{1,3}$  & $2$  	\\
+      $O_{2,3}$  & $2$		\\
+      $O_{3,2}$  & $2$    	\\
+      $O_{3,3}$  & $2$    	\\
       \\

### 2021-10-13 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
looked through Section 2

@@
 
-However, JSP is an NP-complete combinatorial optimization problem \cite{DBLP:journals/mor/GareyJS76,SOTSKOV1995237} for which no efficient algorithms are known.
+However, JSP is an NP-hard combinatorial optimization problem \cite{garey1976complexity,SOTSKOV1995237} for which no efficient algorithms are known.
 Therefore, searching for an optimal solution with state-of-the-art solvers for Answer Set Programming (ASP), Mixed Integer Programming, or Constraint Programming \cite{meng2020mixed,coltep19a,el2020job,al2017job}
@@
 
-Since no efficient algorithm is known for JSP \cite{garey1976complexity}, 
-modern optimization algorithms often cannot reach a solution close to the optimal one in a reasonable time, even if the number of operations to be scheduled is low.
-In real-life applications, the number of operations in an instance might be up to tens of thousands \cite{zhang2010hybrid}.
-The JSP has attracted the interest of many researchers in the last decades to find effective methods to solve this problem.
-One of these methods is decomposition.
-The decomposition idea is to split the problem into sub-problems, solve them separately, and then integrate those solutions to solve the whole problem.
-Many decomposition strategies have been proposed for solving the JSP  \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.
-However, there is no guarantee of which method is the best because the performance varies according to the problem.
+Since JSP is NP-hard \cite{garey1976complexity,SOTSKOV1995237}
+and
+no efficient solving algorithms are known, 
+even state-of-the-art optimization methods can often not find (near-)optimal solutions in reasonable time, already for instances with a seemingly small number of operations.
+As the number of operations in real-life applications
+can easily reach tens of thousands~\cite{zhang2010hybrid},
+approximation methods have attracted particular research interest.
+One such approach is decomposition into easier to optimize parts,
+% . The decomposition idea is to split the problem into sub-problems,
+which can be solved separately and whose partial solutions are
+eventually combined into a joint schedule for the entire problem instance.
+While various decomposition strategies have been proposed in the literature \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance},
+none of them can provide solution quality guarantees
+or strictly dominates over heterogeneous JSP instances. 
+% there is no guarantee of which method is the best because the performance varies according to the problem.
 
-In our example in Table~\ref{tab1}, there are $9$ operations and $3$ machines.
-We aim to split the operations efficiently into two partitions and schedule them separately without violating the precedence constraint.
-More specifically, we should not assign an operation in the Time Window (TW) which precedes TW(s) of at least one of its predecessors.
-Table~\ref{tab2} shows one of the possible decompositions that we would like our method to return.
-Namely, operations $\{ O_{1,1}, O_{1,2}, O_{2,1}, O_{2,2}, O_{3,1} \}$ are assigned to the first TW and the rest to the second TW.
-Given this decomposition, a multi-shot optimization method, like ASP, will first find an optimal schedule of the operations in the first TW and provide it as a set of facts to the search of the optimal solution for the second TW.
-In this study, we investigate applying a constrained clustering algorithm to decompose the JSP, where we extract some features from the problem itself and the solutions of other heuristics.
-These features have been used for the splitting process, and we present them in detail in the following section.
+For our example in Table~\ref{tab1}, there are $9$ operations and $3$ machines.
+Assume that we aim to split the operations into two parts to be scheduled separately such that the precedence between operations belonging to the same job is preserved.
+That is, we should not assign a successor operation to an earlier Time Window (TW)
+than its predecessor. % operation.
+Table~\ref{tab2} shows one feasible decomposition that we may like to generate.
+The operations $\{ O_{1,1}, O_{1,2}, O_{2,1}, O_{2,2}, O_{3,1} \}$ are here assigned to TW~$1$, and the remaining operations to TW~$2$.
+Given this decomposition, a multi-shot optimization approach,
+as offered by ASP~\cite{gekakasc17a},
+can first optimize a schedule (wrt.\ to the makespan)
+for the operations in TW~$1$, and then extend the first part by
+additionally scheduling the operations in TW~$2$ in an optimal way.
+However, considering that the operation $O_{3,2}$ belongs to TW~$2$
+and should be executed later than $O_{1,2}$ of TW~$1$,
+the decomposition is incompatible with the execution order
+$(O_{2,1}, O_{3,2}, O_{1,2})$ for machine~$3$ and thus discards optimal schedules.
+% provide it as a set of facts to the search of the optimal solution for the second TW.
 
-\section{Feature extraction}
+In this work, we introduce and deploy a constrained clustering algorithm to decompose  JSP instances into time windows, where we extract some features from the problem itself and others from solutions obtained by heuristic methods.
+The features we consider for the decomposition process are explained in detail in the next section.
+
+\section{Feature Extraction}
 \label{sec:features}

### 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
section 4

--- 2021-10-13 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 \usepackage[inline]{enumitem}
+\usepackage{xspace}
 \captionsetup[table]{skip=10pt}
@@
 
-\newcommand{\stest}{\textit{ST\_EST}}
-\newcommand{\stmtwr}{\textit{ST\_MTWR}}
-\newcommand{\stfifo}{\textit{ST\_FIFO}}
+\newcommand{\stest}{\textit{ST\_EST}\xspace}
+\newcommand{\stmtwr}{\textit{ST\_MTWR}\xspace}
+\newcommand{\stfifo}{\textit{ST\_FIFO}\xspace}
 
-\newcommand{\wtest}{\textit{WT\_EST}}
-\newcommand{\wtmtwr}{\textit{WT\_MTWR}}
-\newcommand{\wtfifo}{\textit{WT\_FIFO}}
+\newcommand{\wtest}{\textit{WT\_EST}\xspace}
+\newcommand{\wtmtwr}{\textit{WT\_MTWR}\xspace}
+\newcommand{\wtfifo}{\textit{WT\_FIFO}\xspace}
 
@@
 However, our clustering method for JSP instance decomposition requires all features to have numerical values allowing distance measures to estimate similarities/dissimilarities between operations of jobs.
-%
-Therefore, we suggest the following features:
-\begin{description}
-  \item[Processing Time (PT)] that represents the execution time of a particular operation in a machine.
-This feature is provided in every JSP instance, like in the sample instance presented in Table \ref{tab1}.
 
-  \item[Operation (OP)] feature is an ordinal value representing the rank of each operation in its job.
-
-  \item[Remaining Processing Time (RPT)] attribute corresponds to the total processing time for all subsequent operations of a job.
-For example, Table~\ref{tab3} shows a set of \textit{RPT} values computed for the sample instance.
-Since the first job $J_1$ consists of $3$ operations with a total processing time of $17$, \textit{RPT} of $O_{1,1}$ is $17$ because it has no preceding operations.
-After finishing the first operation of $J_1$, the \textit{RPT} becomes $8$ because the processing time of the completed operation $O_{1,1}$ is subtracted.
-Similarly, when the second operation of $J_1$ is completed, the \textit{RPT} of the last operation is $5$.
-
-  \item[Time Length of a Job (TLJ)] is an attribute describing the total processing time needed to execute all operations of a job.
-Since all operations of a job have the same feature value, this feature can only be used to differentiate between jobs.
-  %It is the summation of processing time for all operations of that job.
-
-  \item[Earliest Starting Time (EST)] attribute represents the earliest possible execution time for an operation.
-It is calculated by accumulating the processing time of all preceding operations.
-This feature has a specific value $0$ if the operation has no predecessors.%\todo{This value does not appear in the table, why?}
-
-  \item[Machine Load (ML)] is a property of a machine that shows how much time this machine requires to execute all the operations assigned to it.
-In the initial state, \textit{ML} corresponds to the total processing time of all operations assigned to that machine.
-In each time step, the load of a machine is decreased by the processing time of an operation assigned to the machine and has the smallest \textit{EST}.
-For example, machine $3$ has to process operations $\{ O_{1,2}, O_{2,1}, O_{3,2} \}$.
-The load of machine $3$ at time $0$ is 10.
-We assumed that the machine would start processing the operation $O_{2,1}$ that has the smallest \textit{EST}.
-After completing the operation $O_{2,1}$, the machine will target the next operation, which has the smallest value of \textit{EST} and the load of the machine is decreased by the processing time of the completed operation $O_{2,1}$, and so on till completing all the operations assigned to a particular machine.
-
-  \item[Starting Time (ST)] is a set of features representing the starting execution time of each operation by solving the problem using a heuristic.
-In our work, we consider EST (\stest), First-In-First-Out (\stfifo), and Most Total Work Remaining (\stmtwr), see \cite{jones1998survey} for an overview.
-  To compute the starting time, we use a simple greedy algorithm \cite{el2020job} that at each step selects a not scheduled operation according to the heuristics and schedules it to a machine.
-In the case of FIFO, the algorithm selects an operation that appeared first in the queue.
-For example, $O_{2,2}$ and $O_{3,3}$ are in the queue at time $9$ on machine $2$.
-Since $O_{2,2}$ and $O_{3,3}$ were ready to be processed at time $4$ and $7$, respectively.
-The algorithm selects $O_{2,2}$ to process at time $9$ before $O_{3,3}$ which starts at time $15$.
-Similarly, for MTWR, the algorithm selects the operation that belongs to the job with the longest processing time from the queue.
-
-  \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used the same greedy algorithm as above with three different heuristics EST (\wtest), FIFO (\wtfifo), and MTWR (\wtmtwr).
-Each feature in \textit{WT} set is the difference between the scheduled starting time of a particular operation and ending time of its predecessor.
-%The latter is the maximum out of the end processing time of the operation's predecessor and the earliest time at which the required machine became available.
-  As shown in Table~\ref{tab3}, the predecessor of $O_{2,2}$ which is $O_{2,1}$ is finished at time $4$.
-However machine $2$ that executes $O_{2,2}$ was not free at that time, so $O_{2,2}$ had to wait $5$ time units to start the processing.
-
-  %\todo{It would be nice to add examples of other features to the table. If this is complicated, then only of \wtest and \stest.}
-\end{description}
-
-\begin{table}[h!]
+\begin{table}[tb]
   \setlength{\tabcolsep}{10.0pt}
@@
   %\end{adjustwidth}
-
     \end{center} 
@@
 
-\section{Proposed Algorithm}
+In our approach we evaluated the following features:
+\begin{description}
+  \item[Processing Time (PT)] that represents the execution time of a particular operation in a machine.
+This feature is provided in every JSP instance, like in the sample instance presented in Table \ref{tab1}.
+
+  \item[Operation (OP)] feature is an ordinal value representing the rank of each operation in its job.
+
+  \item[Remaining Processing Time (RPT)] attribute corresponds to the total processing time for all subsequent operations of a job.
+For example, Table~\ref{tab3} shows a set of \textit{RPT} values computed for the sample instance.
+Since the first job $J_1$ consists of $3$ operations with a total processing time of $17$, \textit{RPT} of $O_{1,1}$ is $17$ because it has no preceding operations.
+After finishing the first operation of $J_1$, the \textit{RPT} becomes $8$ because the processing time of the completed operation $O_{1,1}$ is subtracted.
+Similarly, when the second operation of $J_1$ is completed, the \textit{RPT} of the last operation is $5$.
+
+  \item[Time Length of a Job (TLJ)] is an attribute describing the total processing time needed to execute all operations of a job.
+Since all operations of a job have the same feature value, this feature can only be used to differentiate between jobs.
+  %It is the summation of processing time for all operations of that job.
+
+  \item[Earliest Starting Time (EST)] attribute represents the earliest possible execution time for an operation.
+It is calculated by accumulating the processing time of all preceding operations.
+This feature has zero value if the operation has no predecessors.
+
+  \item[Machine Load (ML)] is a property of a machine that shows how much time this machine requires to execute all the operations assigned to it.
+In the initial state, \textit{ML} corresponds to the total processing time of all operations assigned to that machine.
+In each time step, the load of a machine is decreased by the processing time of an operation with the smallest \textit{EST} assigned to the machine.
+For example, machine $3$ has to process operations $\{ O_{1,2}, O_{2,1}, O_{3,2} \}$.
+A load of machine $3$ at time $0$ is 10.
+The machine would start processing the operation $O_{2,1}$ that has the smallest \textit{EST}.
+After completing $O_{2,1}$, the load of the machine is decreased by the processing time of the completed operation $O_{2,1}$ and the machine will start the next of remaining operations, which has the smallest \textit{EST} value.
+The algorithm continues until a machine processes all operations assigned to it in the instance.
+
+  \item[Starting Time (ST)] is a set of features representing the starting execution time of each operation by solving the problem using a heuristic greedy search.
+In our work, we consider EST (\stest), First-In-First-Out (\stfifo), and Most Total Work Remaining (\stmtwr) heuristics, see \cite{jones1998survey} for an overview.
+  To compute the starting time, we use a simple greedy algorithm \cite{el2020job} that at each step selects a not scheduled operation according to the heuristic and schedules it to a machine.
+In the case of \stfifo, the algorithm selects an operation that appeared first in the queue for scheduling.
+For example, operations $O_{2,2}$ and $O_{3,3}$ appear in the processing queue at time points $6$ and $7$ respectively, see Table \ref{tab1}.
+Machine 2 required by both operations becomes available only at time point $9$ when the processing of the task $O_{1,1}$ is finished.
+According to FIFO heuristic, the algorithm selects $O_{2,2}$ to process at time $9$ in machine $2$ and schedules $O_{3,3}$ at time $15$.
+Similarly, using MTWR, the algorithm selects the operation that belongs to the job with the longest processing time from the queue, i.e., $O_{2,2}$ in our example.
+
+  \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used the same greedy algorithm as above with three different heuristics EST (\wtest), FIFO (\wtfifo), and MTWR (\wtmtwr).
+Each feature in \textit{WT} set is the difference between the scheduled starting time of a particular operation and the ending time of its predecessor.
+  As shown in Table~\ref{tab3}, the waiting time of $O_{2,2}$ is $5$ since the processing of its predecessor $O_{2,1}$ finishes at time $4$ and FIFO heuristic selects $O_{2,2}$ for scheduling to machine $2$ at time $9$, just as in the \stfifo case.
+\end{description}
+%
+All features extracted from a given JSP instance are provided to our decomposition method presented in the following section.
+
+\section{Method}
 \label{sec:method}
-In this section, we will show the implementation of our proposed model and how it works.
-The main idea of the proposed model is to split the problem into sub-problems, solve them sequentially and integrate the solutions to have the solution of the main problem.
-We aim to decompose the problem into TWs because this technique proved its effectiveness for solving the scheduling problems \cite{zhang2010hybrid,zhai2014decomposition}.
-Several ways have been presented during the last decades to decompose the problem \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.\\
+%In this section, we will show the implementation of our proposed model and how it works. 
+The proposed method comprises two main steps: 
+\begin{enumerate*}[label=\emph{(\roman*)}]
+  \item first, it decomposes a problem into a sequence of sub-problems, called \emph{time windows}; 
+  \item second, the sub-problems are solved one after another such that each subsequent sub-problem gets solutions of the previous sub-problems as input.
+\end{enumerate*}
+As a result, the solution of the last sub-problem provides a complete schedule for the given JSP instance.
+The computational efficiency of the approach and the quality of obtained schedules depend highly on the quality of decomposition obtained in the first step \cite{zhang2010hybrid,zhai2014decomposition}.
+%To address this problem, several decomposition approaches have been presented during the last decades \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.
+Our approach uses a novel constraint clustering method to find a favorable decomposition of JSP instances using features of operations provided in them.
+Each cluster represents a time window, i.e., a subset of all operations that a scheduling algorithm must consider for the assignment.
 
-The proposed model aims to apply clustering methods where each cluster will represent a TW, and each operation is a data object.
-The clustering algorithms are used to group a set of data objects into one cluster.
-It depends on measuring the similarity/dissimilarity between the data objects according to the features of these data objects.
-The similar objects will be in one cluster, and the dissimilar are in different clusters.
-The difference between the basic clustering and our proposed is the ordering of the clusters.
-In general, the cluster number does not have meaning; it just means that the data objects grouped in a particular cluster are similar.
-However, in our case, the cluster number is meaningful because the data objects grouped in a particular cluster $i$ should be scheduled before other data objects in the successor cluster(s).
-More specifically, it is not allowed to insert a data object (\textit{an operation}) in a particular cluster, and the predecessor of that operation is not assigned yet to the same cluster or the preceding cluster(s).
-Since we will solve each window separately, we should consider balancing the clusters concerning the number of operations assigned to each cluster.
-So that, we will have two types of constraints:
-\begin{enumerate}
-\item The predecessor(s) of an operation must be inserted into the same or previous cluster(s).
-\item The number of operations assigned to each cluster should be close to each other.
-\end{enumerate}
+The clustering algorithms are unsupervised learning methods which goal is to partition a set of objects into (disjoint) subsets, called clusters, such that each cluster contains the most similar objects.
+The similarity of objects is determined by some measure, like Euclidian distance, using features of each object, such as features of operations extracted in the previous section. 
+However, direct application of clustering algorithms, like K-Means \cite{Forgy1965ClusterAO}, in scheduling is limited since the partitioning cannot preserve the sequences of operations in a job.
+As a result, a clustering algorithm might place $O_{1,1}$ and $O_{1,3}$ in one cluster and $O_{1,2}$ in another.
+As a result, the sequence of sub-problems becomes inconsistent and no schedule can be found.
+
+Therefore, we propose a constrained clustering algorithm that preserves sequences of operations by considering the number of a cluster in which an operation is assigned.
+In particular, all predecessors of an operation placed in a cluster $i$ must be assigned to clusters $\{1, \dots, i\}$.
+%More specifically, it is not allowed to insert a data object (\textit{an operation}) in a particular cluster, and the predecessor of that operation is not assigned yet to the same cluster or the preceding cluster(s).
+Since our approach solves each sub-problem separately, the generation of large clusters might reduce its performance significantly.
+In the extreme case, finding a solution for a sub-problem might require the same time as the solution of the whole problem.
+Therefore, in addition to the similarity optimization criteria, our algorithm also aims to generate clusters with an equal number of operations assigned to each cluster.
 
 Our main procedure starts with selecting a subset of the extracted features, which will be used for the clustering process to get the Time window assignment.
-As shown in algorithm~\ref{alg1}, the extracted features are given, and a set of combinations of the features are initialized.
+As shown in Algorithm~\ref{alg1}, the extracted features are given, and a set of combinations of the features are initialized.
 At each iteration, a subset of features is selected and run with the clustering process discussed later, then the Time Window assignment.
@@
 \end{enumerate}
-We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments.
-The following Table~\ref{tab4} shows the selected features, and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
-The following table shows the results of our experiments.
-The first column shows the approaches used for solving the problem in the literature in addition to the combinations of features we presented previously.
-The next columns represent the instances of our benchmark set except for the last column, which shows the average of each approach through the whole benchmark set.
-The first $4$ rows in the table show the values of each makespan obtained by heuristics; FIFO, MTWR, and (EST with a single and multi-shot).
-We can say that FIFO and MTWR provided the best results compared to the other two heuristics.
-However for the features combinations, the combinations (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}and \textbf{F15}) provided good results according to the cost function in the most cases especially with combinations (\textbf{F6} and \textbf{F15}).
-It means that the most important features extracted from the problem itself and have a positive impact on the performance are \textit{Remaining Processing Time, Machine Load and Earliest Starting Time}.
-Also, The information obtained from the solutions of the other heuristics like (\textit{Starting Time} and \textit{Waiting Time}) led us to get better decomposition for the problem at hand.
-More specifically, combining \textit{Remaining Processing Time} with \textit{Machine Load} provides a good decomposition with/without \textit{Waiting Time} attributes obtained by the other heuristics.
-In a few instances, our proposed method did not outperform other heuristics, however, in the last column, the best average makespan of the other approaches was $3331$ by FIFO, and the average makespan of our feature selection algorithm was $3256$.
-We can conclude that the feature selection algorithm outperformed other heuristics in the literature.
-The following section illustrates the difference between our proposed algorithm and the methods implemented by other researchers.
 
-\begin{table}
+\begin{table}[bt]
   \begin{center}
@@
 
+We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments.
+The following Table~\ref{tab4} shows the selected features, and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
+The following table shows the results of our experiments.
+The first column shows the approaches used for solving the problem in the literature in addition to the combinations of features we presented previously.
+The next columns represent the instances of our benchmark set except for the last column, which shows the average of each approach through the whole benchmark set.
+The first $4$ rows in the table show the values of each makespan obtained by heuristics; FIFO, MTWR, and (EST with a single and multi-shot).
+We can say that FIFO and MTWR provided the best results compared to the other two heuristics.
+However for the features combinations, the combinations (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}and \textbf{F15}) provided good results according to the cost function in the most cases especially with combinations (\textbf{F6} and \textbf{F15}).
+It means that the most important features extracted from the problem itself and have a positive impact on the performance are \textit{Remaining Processing Time, Machine Load and Earliest Starting Time}.
+Also, The information obtained from the solutions of the other heuristics like (\textit{Starting Time} and \textit{Waiting Time}) led us to get better decomposition for the problem at hand.
+More specifically, combining \textit{Remaining Processing Time} with \textit{Machine Load} provides a good decomposition with/without \textit{Waiting Time} attributes obtained by the other heuristics.
+In a few instances, our proposed method did not outperform other heuristics, however, in the last column, the best average makespan of the other approaches was $3331$ by FIFO, and the average makespan of our feature selection algorithm was $3256$.
+We can conclude that the feature selection algorithm outperformed other heuristics in the literature.
+The following section illustrates the difference between our proposed algorithm and the methods implemented by other researchers.
+
 \section{Related Work}
@@
 The main idea of data mining is to explore the patterns from the obtained solutions by some heuristics and then develop some decision rules that approximate the order of the operations in each machine \cite{ismail2012production}.
-More precisely, it aims to get an order of the operations according to some characteristics.\\
+More precisely, it aims to get an order of the operations according to some characteristics.
 
@@
 The results showed that the generated rules were able to outperform the shortest processing time heuristic.
-One of the drawbacks is that the problem instances considered contain only 36 operations.\\
+One of the drawbacks is that the problem instances considered contain only 36 operations.
 
@@
 If two or more operations have the same processing time, they look at the time length of the job.
-If the job length is the same, sort them out based on the remaining processing time.\\
+If the job length is the same, sort them out based on the remaining processing time.
 
@@
 However, dynamic JSP became one of the most crucial problems in real-life applications.
-New dispatching rules for JSP are presented using a data mining approach to obtain an efficient schedule in a dynamic Job-shop Scheduling environment \cite{shahzad2010discovering}.
+New dispatching rules for JSP are presented using a data mining approach to obtain an efficient schedule in a dynamic job-shop scheduling environment \cite{shahzad2010discovering}.
 This study aimed to minimize makespan by creating new dispatching rules.
@@
 They considered some other features like (Earliest possible Starting Time EST, the difference in remaining processing time for a job, and the difference in the processing time of two operations) to create a rule-set for dispatching the operations.
-The experiments showed that the results of mined rule-set are superior or comparable to the heuristics in the literature.\\
+The experiments showed that the results of mined rule-set are superior or comparable to the heuristics in the literature.
 
@@
 The proposed method was compared to VNS and some common heuristics using a simulated job shop.
-the results indicated that the performance of the proposed model is better than the others, including some heuristics (FIFO, LIFO, and SPT).\\
+the results indicated that the performance of the proposed model is better than the others, including some heuristics (FIFO, LIFO, and SPT).
 
@@
 To evaluate the performance of the proposed method, they performed a series of computational experiments on several test instances.
-The results showed that the proposed method with data mining generated initial population outperforms the methods with randomly generated initial population.\\
+The results showed that the proposed method with data mining generated initial population outperforms the methods with randomly generated initial population.
 
-From the literature, we have found that most of the studies which used data mining techniques aimed to extract some features either from the problem instances or solutions obtained by some heuristics to build a decision-rules to order the operations assigned to a machine.\\
+From the literature, we have found that most of the studies which used data mining techniques aimed to extract some features either from the problem instances or solutions obtained by some heuristics to build a decision-rules to order the operations assigned to a machine.
 

### 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
went through start of Section 3

--- 2021-10-13 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 \label{sec:features}
-Application of machine learning methods to scheduling problems requires careful selection of data describing the hidden dependencies between operations of different jobs \cite{ismail2012production,nasiri2019data}.
-Clustering methods, which we aim to apply in our approach, are not an exception.
-Thus, a clustering method requires an informative set of features describing jobs, their operations, and machines of an instance to find patterns resulting in a beneficial decomposition of all operations in time windows.
+The application of machine learning methods to scheduling problems requires a careful selection of data describing the hidden dependencies between operations of different jobs \cite{ismail2012production,nasiri2019data}.
+Clustering methods, which we intend to apply in our approach, are not an exception to this.
+That is, a clustering method requires an informative set of features characterizing the jobs, their operations, and machines of a problem instance to identify patterns resulting in a beneficial decomposition of the operations into time windows.
 
-Methods suggested in the previous work \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} used characterizations of scheduling instances based on the following features: priority, processing time, remaining processing time, machine load, and route position.
-Most of the approaches converted the quantitative values of the attributes into qualitative data.
-As mentioned in these studies, the main reason for this conversion was to make the heuristic dispatching rules more generic and allow their application to several scheduling problems.
-In this work, we suggest a method that can work directly on the sets of features and does not require any problem-specific transformations.
-However, our clustering method for JSP instance decomposition requires all features to have numerical values allowing distance measures to estimate similarities/dissimilarities between operations of jobs.
-%
-Therefore, we suggest the following features:
+Methods suggested in the literature \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} characterize instances of scheduling problems based on the following features: priority, processing time, remaining processing time, machine load, and sequence position.
+Most of these approaches convert the quantitative feature values % of the attributes
+into qualitative attributes in order to obtain generic dispatching rules
+that remain applicable to instances of different size.
+% As mentioned in these studies, the main reason for this conversion was to make the heuristic dispatching rules more generic and allow their application to several scheduling problems. 
+In this work, we propose a method that can be applied to feature values directly and does not require any problem-specific transformations.
+However, our clustering method for JSP instance decomposition requires all features to have numerical values, which permit the calculation of distance measures for estimating (dis)similarities between operations.
+
+In detail, we consider the following features of jobs, operations, and machines:
 \begin{description}
@@
 
-\begin{table}[h!]
+\begin{table}[t]
 \setlength{\tabcolsep}{10.0pt}
@@
   \begin{center}
-  \caption{Features extracted for the sample problem instance.}
+  \caption{Features extracted for the sample problem instance from Table~\ref{tab1}.}
   \label{tab3}
@@
 
-\section{Proposed Algorithm}
+\section{Constrained Clustering Algorithm}
 \label{sec:method}

### 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex

--- 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 \label{sec:features}
-Application of machine learning methods to scheduling problems requires careful selection of data describing the hidden dependencies between operations of different jobs \cite{ismail2012production,nasiri2019data}.
-Clustering methods, which we aim to apply in our approach, are not an exception.
-Thus, a clustering method requires an informative set of features describing jobs, their operations, and machines of an instance to find patterns resulting in a beneficial decomposition of all operations in time windows.
+The application of machine learning methods to scheduling problems requires a careful selection of data describing the hidden dependencies between operations of different jobs \cite{ismail2012production,nasiri2019data}.
+Clustering methods, which we intend to apply in our approach, are not an exception to this.
+That is, a clustering method requires an informative set of features characterizing the jobs, their operations, and machines of a problem instance to identify patterns resulting in a beneficial decomposition of the operations into time windows.
 

### 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
Section 4

--- 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 
-Our main procedure starts with selecting a subset of the extracted features, which will be used for the clustering process to get the Time window assignment.
-As shown in Algorithm~\ref{alg1}, the extracted features are given, and a set of combinations of the features are initialized.
-At each iteration, a subset of features is selected and run with the clustering process discussed later, then the Time Window assignment.
-This process is repeated until the termination condition is met.
+Our main procedure starts with the selection of a subset of the extracted features, which allows the clustering algorithm to compute the best possible time window assignments. 
+The procedure follows the forward feature selection principle: start with a small set of features and in every iteration extends it with one feature.
+Run the clustering algorithm on the new set of features and solve the obtained sequence of time windows.
+The subset of features resulting in the best possible solution quality is preserved and provided for extension in the next loop of the feature selection algorithm.
+The latter continues until either
+\begin{enumerate*}[label=\emph{(\roman*)}]
+  \item all features are included in the target set, or 
+  \item any extension of the provided subset of features results in solutions of worse quality than the solution of the given set.
+\end{enumerate*}
+
+%\begin{algorithm}
+%\caption{Feature Selection Algorithm}\label{alg1}
+%\begin{algorithmic}
+%\Require $extracted\_features$
+%\State $count\_feature \gets 0$
+%\State $comb\_features \get \emptyset$ \Comment{Define a set of combined %features}
+%\State $num\_subsets \gets \textbf{count}(comb\_features)$
+%\While{$count\_feature < num\_subsets$}
+%	\State $selected\_subset \gets comb\_features[count\_feature]$
+%	\State Clustering Algorithm ($selected\_subset$) \Comment{\textbf{(Algorithm.%~\ref{alg2})}} %\textbf{(Algorithm.~\ref{alg2})}
+%	\State Store the Time Windows Assignment
+%	\State $count\_feature \gets count\_feature + 1$
+%\EndWhile
+%
+%\end{algorithmic}
+%\end{algorithm}
+
+Algorithm~\ref{alg2} shows the pseudocode of the proposed clustering algorithm.
+Similar to K-Means, we assume that the algorithm gets the number of clusters as input.
+We calculate the cluster capacity by dividing the total number of operations by the number of clusters.
+In the next step, the clustering process starts by generating the centroids, where the number of centroids corresponds to the number of clusters.
+Each centroid in our case corresponds to some randomly selected operation from a set of operations allowing for preserving of sequencing constraints.
+For example, let us assume that a job consists of 15 operations, and the number of clusters is $3$.
+The first centroid will be an operation of order $\{1, dots, 5\}$ in the sequence, the second centroid will be selected from operations with the order $\{6, dots, 10\}$, and so forth.
+After generating the centroids, we check for exceeding the number of clusters and the cluster capacity, respectively.
+%
+Next, the algorithm calculates the distance between each operation and the centroid of the current cluster using Euclidean distance and assigns the nearest operation to that cluster.
+In order to satisfy the precedence constraint, we check the assignments of predecessors.
+If the constraint is satisfied, the algorithm assigns the operation to the cluster, and the centroid is updated.
+Whenever the maximum number of operations is reached, the algorithm continues with the next cluster.
+These iterations continue until all operations are assigned to some cluster.
 
 \begin{algorithm}
-\caption{Feature Selection Algorithm}\label{alg1}
-\begin{algorithmic}
-\Require $extracted\_features$
-\State $count\_feature \gets 0$
-\State $comb\_features$ \Comment{Define a set of combined features}
-\State $num\_subsets \gets \textbf{count}(comb\_features)$
-\While{$count\_feature < num\_subsets$}
-	\State $selected\_subset \gets comb\_features[count\_feature]$
-	\State Proposed Clustering Algorithm ($selected\_subset$) \Comment{\textbf{(Algorithm.~\ref{alg2})}} %\textbf{(Algorithm.~\ref{alg2})}
-	\State Store the Time Windows Assignment
-	\State $count\_feature \gets count\_feature + 1$
-\EndWhile
-
-\end{algorithmic}
-\end{algorithm}
-
-Algorithm.~\ref{alg2} shows the pseudocode of the proposed algorithm.
-As shown in the pseudocode, the number of clusters is given.
-We calculate the cluster capacity by dividing the total number of operations by the number of clusters.
-In the next step, the clustering process starts by generating the centroids, where the number of centroids is the number of clusters.
-In general, the centroids are generated randomly as well as in our model.
-However, we selected some operations as centroids.
-Although the operations are selected randomly, we followed a rule to select the centroids.
-For example, let us assume that a job consists of 15 operations, and the number of clusters is $3$.
-The first centroid will be an operation of rank 1 - 5, the second centroid will be 6 - 10, and the third will be from the last five operations.
-We followed this strategy to make them far from each other because the operations grouped in the first clusters should be at the beginning of their jobs.
-After generating the centroids, we check for exceeding the number of clusters and the cluster capacity, respectively.
-After that, we calculate the distance between each operation and the centroid of the current cluster using Euclidean distance and assign the nearest operation to that cluster.
-In order to satisfy the precedence constraint, we check if the predecessor is not assigned to a cluster, it will be assigned to the current cluster, and the centroid is updated.
-If the maximum number of operations is reached, we will go to the next cluster.
-After assigning all operations into clusters, the algorithm terminates.
-
-\begin{algorithm}
-\caption{Proposed Clustering Algorithm}\label{alg2}
+\caption{Clustering Algorithm}\label{alg2}
 \begin{algorithmic}

### 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
put together modifications in Section 3

--- 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 
-Methods suggested in the previous work \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} used characterizations of scheduling instances based on the following features: priority, processing time, remaining processing time, machine load, and route position.
-Most of the approaches converted the quantitative values of the attributes into qualitative data.
-As mentioned in these studies, the main reason for this conversion was to make the heuristic dispatching rules more generic and allow their application to several scheduling problems.
-In this work, we suggest a method that can work directly on the sets of features and does not require any problem-specific transformations.
-However, our clustering method for JSP instance decomposition requires all features to have numerical values allowing distance measures to estimate similarities/dissimilarities between operations of jobs.
+Methods suggested in the literature \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} characterize instances of scheduling problems based on the following features: priority, processing time, remaining processing time, machine load, and sequence position.
+Most of these approaches convert the quantitative feature values % of the attributes
+into qualitative attributes in order to obtain generic dispatching rules
+that remain applicable to instances of different size.
+% As mentioned in these studies, the main reason for this conversion was to make the heuristic dispatching rules more generic and allow their application to several scheduling problems. 
+In this work, we propose a method that can be applied to feature values directly and does not require any problem-specific transformations.
+However, our clustering method for JSP instance decomposition requires all features to have numerical values, which permit the calculation of distance measures for estimating (dis)similarities between operations.
 
@@
     \begin{center}
-    \caption{Features extracted for the sample problem instance.}
+    \caption{Features extracted from the sample problem instance in Table~\ref{tab1}.}
     \label{tab3}
@@
   \end{table}
-
-In our approach we evaluated the following features:
+%
+% In our approach we evaluated the following features:
+In detail, we consider the following features of jobs, operations, and machines:
 \begin{description}
@@
 %
-All features extracted from a given JSP instance are provided to our decomposition method presented in the following section.
+All features extracted from a given JSP instance are provided to our decomposition method presented in the next section.
 
-\section{Method}
+\section{Constrained Clustering Algorithm}
 \label{sec:method}

### 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex

--- 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 \label{sec:features}
-Application of machine learning methods to scheduling problems requires careful selection of data describing the hidden dependencies between operations of different jobs \cite{ismail2012production,nasiri2019data}.
-Clustering methods, which we aim to apply in our approach, are not an exception.
-Thus, a clustering method requires an informative set of features describing jobs, their operations, and machines of an instance to find patterns resulting in a beneficial decomposition of all operations in time windows.
+The application of machine learning methods to scheduling problems requires a careful selection of data describing the hidden dependencies between operations of different jobs \cite{ismail2012production,nasiri2019data}.
+Clustering methods, which we intend to apply in our approach, are not an exception to this.
+That is, a clustering method requires an informative set of features characterizing the jobs, their operations, and machines of a problem instance to identify patterns resulting in a beneficial decomposition of the operations into time windows.
 
-Methods suggested in the previous work \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} used characterizations of scheduling instances based on the following features: priority, processing time, remaining processing time, machine load, and route position.
-Most of the approaches converted the quantitative values of the attributes into qualitative data.
-As mentioned in these studies, the main reason for this conversion was to make the heuristic dispatching rules more generic and allow their application to several scheduling problems.
-In this work, we suggest a method that can work directly on the sets of features and does not require any problem-specific transformations.
-However, our clustering method for JSP instance decomposition requires all features to have numerical values allowing distance measures to estimate similarities/dissimilarities between operations of jobs.
+Methods suggested in the literature \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} characterize instances of scheduling problems based on the following features: priority, processing time, remaining processing time, machine load, and sequence position.
+Most of these approaches convert the quantitative feature values % of the attributes
+into qualitative attributes in order to obtain generic dispatching rules
+that remain applicable to instances of different size.
+% As mentioned in these studies, the main reason for this conversion was to make the heuristic dispatching rules more generic and allow their application to several scheduling problems. 
+In this work, we propose a method that can be applied to feature values directly and does not require any problem-specific transformations.
+However, our clustering method for JSP instance decomposition requires all features to have numerical values, which permit the calculation of distance measures for estimating (dis)similarities between operations.
 

### 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
eval

--- 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 For example, let us assume that a job consists of 15 operations, and the number of clusters is $3$.
-The first centroid will be an operation of order $\{1, dots, 5\}$ in the sequence, the second centroid will be selected from operations with the order $\{6, dots, 10\}$, and so forth.
+The first centroid will be an operation of order $\{1, \dots, 5\}$ in the sequence, the second centroid will be selected from operations with the order $\{6, \dots, 10\}$, and so forth.
 After generating the centroids, we check for exceeding the number of clusters and the cluster capacity, respectively.
@@
 
-\section{Results}
+\section{Evaluation Results}
 \label{sec:eval}
-We use the benchmark instances presented in \cite{taillard1993benchmarks} as a dataset to work on.
-Many researchers have studied these instances and presented different approaches to solving them.
+We use a popular benchmark \cite{taillard1993benchmarks} comprising randomly generated JSP instances as an evaluation dataset for our approach.
+%Many researchers have studied these instances and presented different approaches to solving them.
 We also developed different decomposition strategies to solve those.
 Therefore, it will be possible to measure the success/failure of our proposed model by comparing the results of the strategies developed in the past with those obtained from the proposed algorithm.
-There are 12 features, as we described in the previous section.
-In order to know which features have a significant effect on the results, we did some combinations of the features to perform the decomposition process clustering-based and then solved the problem using the Answer Set Programming model developed in \cite{el2020job} with a timeout of 10 minutes.
-We started with a simple combination that consists of $3$ features; \stfifo, \stmtwr, and \stest.
-At each step, we added another feature to get a new combination.
-The combinations of the selected features are listed as follows:
-\begin{enumerate}
+For each instance, we extracted the 12 features described in Section \ref{sec:features}.
+In order to study the impact of the features on the clustering and, consequently, schedule quality, we store and present the output of the feature selection step and results of the decomposition-based solving using the Answer Set Programming model developed in \cite{el2020job}.
+All experiments were executed with a timeout of 10 minutes. 
+The feature selection algorithm was initialized with a simple combination that consists of $3$ features: \stfifo, \stmtwr, and \stest.
+The combinations of the selected features presented below.
+We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments.
+
+\begin{enumerate}[label=]
 \fontsize{9pt}{10pt}
@@
 
-We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments.
-The following Table~\ref{tab4} shows the selected features, and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
-The following table shows the results of our experiments.
-The first column shows the approaches used for solving the problem in the literature in addition to the combinations of features we presented previously.
+Table~\ref{tab4} shows the selected features and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
+The first column shows the approaches used for solving the problem in the literature in addition to the combinations of features considered by our feature selection algorithm.
 The next columns represent the instances of our benchmark set except for the last column, which shows the average of each approach through the whole benchmark set.
-The first $4$ rows in the table show the values of each makespan obtained by heuristics; FIFO, MTWR, and (EST with a single and multi-shot).
-We can say that FIFO and MTWR provided the best results compared to the other two heuristics.
-However for the features combinations, the combinations (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}and \textbf{F15}) provided good results according to the cost function in the most cases especially with combinations (\textbf{F6} and \textbf{F15}).
+The first $4$ rows in the table show the values of each makespan obtained by heuristics; FIFO, MTWR, and EST with a single and multi-shot ASP solving.
+We can conclude that FIFO and MTWR provided the best results compared to the other two heuristics.
+However for the features combinations, the sets (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12} and \textbf{F15}) provided good results according to the cost function in the most cases especially with combinations (\textbf{F6} and \textbf{F15}).
 It means that the most important features extracted from the problem itself and have a positive impact on the performance are \textit{Remaining Processing Time, Machine Load and Earliest Starting Time}.
-Also, The information obtained from the solutions of the other heuristics like (\textit{Starting Time} and \textit{Waiting Time}) led us to get better decomposition for the problem at hand.
+Also, the information obtained from the solutions of the other heuristics, like \textit{Starting Time} and \textit{Waiting Time}, led our algorithm to better decomposition for the problem at hand.
 More specifically, combining \textit{Remaining Processing Time} with \textit{Machine Load} provides a good decomposition with/without \textit{Waiting Time} attributes obtained by the other heuristics.
-In a few instances, our proposed method did not outperform other heuristics, however, in the last column, the best average makespan of the other approaches was $3331$ by FIFO, and the average makespan of our feature selection algorithm was $3256$.
-We can conclude that the feature selection algorithm outperformed other heuristics in the literature.
-The following section illustrates the difference between our proposed algorithm and the methods implemented by other researchers.
+In a few instances, our proposed method did not outperform other heuristics.
+However, in general, the best average makespan of the other approaches was $3331$ by FIFO, and the average makespan of our feature selection algorithm was $3256$.
+Therefore, the feature selection algorithm outperformed other heuristics in the literature by efficiently combining different heuristic information using clustering and feature selection.
 

### 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
rw and conclusions

@@
 \label{sec:literature}
-In this section, we will review some of the most related works about data mining approaches employed to solve scheduling problems.
-Over the last decades, solving the scheduling problem has attracted the interest of many researchers, and a wide range of approaches has been developed for solving the scheduling problem.
-One of the approaches that have been widely used to solve the problem is data mining.
-The main idea of data mining is to explore the patterns from the obtained solutions by some heuristics and then develop some decision rules that approximate the order of the operations in each machine \cite{ismail2012production}.
-More precisely, it aims to get an order of the operations according to some characteristics.
+In this section, we review some of the most related work using data mining approaches employed to solve scheduling problems.
+Over the last decades, solving the scheduling problem has attracted the interest of many researchers.The main idea of data mining in these applications is to explore the patterns from the obtained solutions by some heuristics and then develop decision rules that approximate the order of the operations in each machine \cite{ismail2012production}.
+More precisely, it aims to get an order of all operations in the given instance according to their characteristics.
 
-Data mining methodologies are applied to explore the patterns in data generated by Genetic Algorithm (GA) \cite{koonce2000using}.
-This work aimed to generate rules to determine the priority for each job with weight.
-They solved the scheduling problem using GA to create some rules.
-Some characteristics (rank of operations, processing time, machine load, and remaining processing time) have been considered.
-In order to evaluate their heuristic, they tested the generated rules on some other instances that were randomly generated.
-The results showed that the generated rules were able to outperform the shortest processing time heuristic.
-One of the drawbacks is that the problem instances considered contain only 36 operations.
+Data mining methodologies were applied to explore the patterns in data generated by genetic algorithms in \cite{koonce2000using}.
+This work aimed to generate rules determining the priority for each job.
+The authors solved the scheduling problem using a genetic algorithm and derived dispatching rules that considered characteristics like the rank of operations, processing time, machine load, and remaining processing time.
+The evaluation of the obtained rules was done on previously unused random instances.
+Obtained results show that the generated rules were able to outperform the shortest processing time heuristic.
+One of the drawbacks is that the problem instances considered contain only 36 operations, i.e., considered JSP instances were relatively small.
 
-Other researchers also implemented GA to get some information of the solution but considered other features to generate decision rules \cite{harrath2002genetic}.
-They considered processing time, job length, remaining processing time, the rank of the operation, and machine load.
-To build a generic model for scheduling problems, they discretized the data by converting the quantitative data into qualitative.
-They sorted the operations share the same machine according to processing time.
-If two or more operations have the same processing time, they look at the time length of the job.
-If the job length is the same, sort them out based on the remaining processing time.
+The authors in \cite{harrath2002genetic} also implemented a genetic algorithm to construct and analyze solutions of JSP instances for the feature extraction for the suggested dispatching rule.
+In particular, the approach uses processing time, job length, remaining processing time, the rank of the operation, and machine load.
+The authors preprocessed the data by converting it from quantitative into qualitative representation, which was then used to build a generic model for scheduling problems as follows: First, the authors sorted the operations share the same machine according to processing time.
+Then, if two or more operations have the same processing time, the heuristic preferred a job with the smallest length.
+Finally, if the jobs have the same length, the remaining processing time heuristic was used as a tie-breaker.
+In opposite, our approach does not depend on a set of features extracted from an instance.
+New features can easily be included in the set of observations and considered by the clustering method.
 
-The work mentioned above had considered static JSP.
-However, dynamic JSP became one of the most crucial problems in real-life applications.
-New dispatching rules for JSP are presented using a data mining approach to obtain an efficient schedule in a dynamic job-shop scheduling environment \cite{shahzad2010discovering}.
-This study aimed to minimize makespan by creating new dispatching rules.
-Firstly, they solved the scheduling problems using Tabu Search and used the obtained solution as a training set to get a rule-set capable of approximating efficient solutions.
-They considered some other features like (Earliest possible Starting Time EST, the difference in remaining processing time for a job, and the difference in the processing time of two operations) to create a rule-set for dispatching the operations.
+In addition to the work focusing on static JSP, there are approaches using data mining methods for dynamic JSP.
+For instance, \cite{shahzad2010discovering} presents new dispatching rules for JSP aiming to minimize makespan and which process a given JSP instance as follows:
+First, a scheduling instance is solved using Tabu Search and the obtained solutions are used as a training set to get a rule-set capable of approximating efficient solutions.
+This approach considers features like EST, the difference in remaining processing time for a job, and the difference in the processing time of two operations, to create a rule-set for dispatching the operations.
 The experiments showed that the results of mined rule-set are superior or comparable to the heuristics in the literature.
 
-The variable neighborhood search (VNS) is combined with the k-means algorithm to address Dynamic JSP \cite{adibi2014clustering}.
-The proposed algorithm depends on extracting information from the problem's input data when an event like (machine breakdown or adding new job(s)) occurred.
+The variable neighborhood search (VNS) is combined with the k-Means algorithm to address Dynamic JSP \cite{adibi2014clustering}.
+The proposed algorithm depends on extracting information from the problem's input data when an event, like a machine breakdown or an addition of a new job, occurs.
 At any rescheduling point, cluster analysis is performed to replace jobs according to their distances.
@@
 The proposed method was compared to VNS and some common heuristics using a simulated job shop.
-the results indicated that the performance of the proposed model is better than the others, including some heuristics (FIFO, LIFO, and SPT).
+the results indicated that the performance of the proposed model is better than the others, including heuristics like FIFO, LIFO, and SPT.
 
@@
 From the literature, we have found that most of the studies which used data mining techniques aimed to extract some features either from the problem instances or solutions obtained by some heuristics to build a decision-rules to order the operations assigned to a machine.
+In most cases, these approaches depend on the set of extracted features and cannot add new or remove the existing ones.
+Our approach is completely independent of the underlying set of features and, therefore, has broader and more flexible applications.
 

### 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
features section

--- 2021-10-14 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
                       \\
-        $O_{1,1}$  & 17 & 0   & 20	&  0  & 0\\
-        $O_{1,2}$  & 8  & 9   & 3		&  9  & 0\\
-        $O_{1,3}$  & 5  & 12  & 5	  &  12 & 0\\
+        $O_{1,1}$  & $17$ & $0$   & $20$	&  $0$  & $0$\\
+        $O_{1,2}$  & $8$  & $9$   & $3$		&  $9$  & $0$\\
+        $O_{1,3}$  & $5$  & $12$  & $5$	  &  $12$ & $0$\\
                       \\
-        $O_{2,1}$  & 12 & 0   & 10	&  0  & 0\\
-        $O_{2,2}$  & 8  & 4   & 11	&  9  & 5\\
-        $O_{2,3}$  & 2  & 10  & 7	  &  17 & 2\\
+        $O_{2,1}$  & $12$ & $0$   & $10$	&  $0$  & $0$\\
+        $O_{2,2}$  & $8$  & $4$   & $11$	&  $9$  & $5$\\
+        $O_{2,3}$  & $2$  & $10$  & $7$	  &  $17$ & $2$\\
                           \\
-        $O_{3,1}$  & 12 & 0  & 11	  &  0  & 0\\
-        $O_{3,2}$  & 8  & 4  & 6		&  4  & 0\\
-        $O_{3,3}$  & 5  & 7  & 5		&  15 & 8\\
+        $O_{3,1}$  & $12$ & $0$  & $11$	  &  $0$  & $0$\\
+        $O_{3,2}$  & $8$  & $4$  & $6$		&  $4$  & $0$\\
+        $O_{3,3}$  & $5$  & $7$  & $5$		&  $15$ & $8$\\
       \end{tabular}
@@
 \begin{description}
-  \item[Processing Time (PT)] that represents the execution time of a particular operation in a machine.
-This feature is provided in every JSP instance, like in the sample instance presented in Table \ref{tab1}.
+  \item[Operation (OP)] is the ordinal value for the position of an operation in its job.
 
-  \item[Operation (OP)] feature is an ordinal value representing the rank of each operation in its job.
+  \item[Processing Time (PT)] is the time for executing an operation on its machine. This feature is part of a JSP instance, such as the sample instance specified in Table~\ref{tab1}.
 
-  \item[Remaining Processing Time (RPT)] attribute corresponds to the total processing time for all subsequent operations of a job.
-For example, Table~\ref{tab3} shows a set of \textit{RPT} values computed for the sample instance.
-Since the first job $J_1$ consists of $3$ operations with a total processing time of $17$, \textit{RPT} of $O_{1,1}$ is $17$ because it has no preceding operations.
-After finishing the first operation of $J_1$, the \textit{RPT} becomes $8$ because the processing time of the completed operation $O_{1,1}$ is subtracted.
+  \item[Remaining Processing Time (RPT)] provides the total processing time for pending operations until the completion of a job.
+For example, Table~\ref{tab3} lists \textit{RPT} values for operations of the sample instance in Table~\ref{tab1}.
+The job $J_1$ consists of $3$ operations with a total processing time of $17$, which matches \textit{RPT} for the first operation $O_{1,1}$.
+% is $17$ because it has no preceding operations.
+  The \textit{RPT} for $O_{1,2}$ is obtained by subtracting the processing time of $O_{1,1}$, i.e., $17-9=8$, and it corresponds to the processing time $5$ for the last job $O_{1,3}$.
+  % After finishing the first operation of $J_1$, the \textit{RPT} becomes $8$ because the processing time of the completed operation $O_{1,1}$ is subtracted.
 Similarly, when the second operation of $J_1$ is completed, the \textit{RPT} of the last operation is $5$.
 
-  \item[Time Length of a Job (TLJ)] is an attribute describing the total processing time needed to execute all operations of a job.
+  \item[Time Length of a Job (TLJ)] is the total processing time for operations of a job, which coincides with the \textit{RPT} value of the job's first operation and is more coarse-grained than the operation-specific \textit{RPT} feature.
+  % is an attribute describing the total processing time needed to execute all operations of a job.
 Since all operations of a job have the same feature value, this feature can only be used to differentiate between jobs.
@@
 
-  \item[Earliest Starting Time (EST)] attribute represents the earliest possible execution time for an operation.
-It is calculated by accumulating the processing time of all preceding operations.
+  \item[Earliest Starting Time (EST)] represents the earliest possible time for executing an operation, given by the total processing time for the predecessor operations in its job.
+% It is calculated by accumulating the processing time of all preceding operations.
 This feature has zero value if the operation has no predecessors.
+  For the first operation of each job, the \textit{EST} value defaults to~$0$.
 
-  \item[Machine Load (ML)] is a property of a machine that shows how much time this machine requires to execute all the operations assigned to it.
-In the initial state, \textit{ML} corresponds to the total processing time of all operations assigned to that machine.
-In each time step, the load of a machine is decreased by the processing time of an operation with the smallest \textit{EST} assigned to the machine.
+  \item[Machine Load (ML)] is a property describing how much time it takes to execute the operations assigned to a machine.
+Initially, \textit{ML} corresponds to the total processing time for all operations to be executed by a machine.
+  Then the assumption is that the operations are processed in increasing order of their \textit{EST} values, and \textit{ML} is thus reduced by the processing times of preceeding operations.
+  For example, the \textit{EST} for the operations $O_{2,1}$, $O_{3,2}$, and $O_{1,2}$ assigned to machine~$3$ is $0$, $4$, or $9$, respectively.
+  Proceeding in this execution order, \textit{ML} is the total processing time $10$ for $O_{2,1}$, reduced by the processing time of $O_{2,1}$ to $10-4 = 6$ for $O_{3,2}$, and then we obtain the execution time $3$ for the last operation $O_{1,2}$.
+  % In each time step, the load of a machine is decreased by the processing time of an operation with the smallest \textit{EST} assigned to the machine.
 For example, machine $3$ has to process operations $\{ O_{1,2}, O_{2,1}, O_{3,2} \}$.
@@
 
-  \item[Starting Time (ST)] is a set of features representing the starting execution time of each operation by solving the problem using a heuristic greedy search.
-In our work, we consider EST (\stest), First-In-First-Out (\stfifo), and Most Total Work Remaining (\stmtwr) heuristics, see \cite{jones1998survey} for an overview.
-  To compute the starting time, we use a simple greedy algorithm \cite{el2020job} that at each step selects a not scheduled operation according to the heuristic and schedules it to a machine.
-In the case of \stfifo, the algorithm selects an operation that appeared first in the queue for scheduling.
-For example, operations $O_{2,2}$ and $O_{3,3}$ appear in the processing queue at time points $6$ and $7$ respectively, see Table \ref{tab1}.
+  \item[Starting Time (ST)] is a family of features providing the starting times of operations obtained by scheduling them with heuristic greedy search methods.
+In our work, we consider Earliest Starting Time (\stest), First-In-First-Out (\stfifo), and Most Total Work Remaining (\stmtwr) as heuristics for the greedy operation allocation; see \cite{jones1998survey} for an overview of such techniques.
+  At each step, a simple greedy algorithm~\cite{el2020job} selects some pending operation whose machine is available according to the heuristic and schedules it.
+In the case of \stfifo, the algorithm selects an operation waiting longest for its machine to become available.
+  For example, the first operations $O_{1,1}$, $O_{2,1}$, and $O_{3,1}$ are all scheduled at time~$0$ (on different machines),
+  then $O_{2,2}$ waits from time~$4$ for its machine~$2$ to get available,
+  while $O_{3,2}$ is started on machine~$3$ at time~$4$, so that
+  $O_{3,3}$ also waits for machine~$2$ from time~$7$.
+  When the machine~$2$ is at time~$9$ ready to start another operation,
+  the \stfifo heuristic thus selects the operation $O_{2,2}$, which waits
+  for longer, to be executed next.
+  With the \stest and \stmtwr heuristics,
+  the selection of the next operation is based on smaller \textit{EST} or
+  greater \textit{RPT} values, respectively, which in view of the attributes in Table~\ref{tab3} also leads to the result that $O_{2,2}$
+  is processed before $O_{3,3}$. % on machine~$2$.
+  % operations $O_{2,2}$ and $O_{3,3}$ appear in the processing queue at time points $6$ and $7$ respectively, see Table \ref{tab1}.
 Machine 2 required by both operations becomes available only at time point $9$ when the processing of the task $O_{1,1}$ is finished.
@@
 
-  \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used the same greedy algorithm as above with three different heuristics EST (\wtest), FIFO (\wtfifo), and MTWR (\wtmtwr).
+  \item[Waiting Time (WT)] is also a family of features, where variants denoted by \wtest, \wtfifo, and \wtmtwr rely on schedules obtained with the corresponding ST heuristic, i.e., \stest, \stfifo, or \stmtwr.
+  Given a schedule computed by the greedy algorithm,
+  the waiting time of an operation is determined by the difference
+  between its starting time and the time of completing the predecessor
+  operation, or simply the starting time for the first operation of each job.
+  For instance, the starting times with \stfifo listed in Table~\ref{tab3}
+  yield the waiting times given in the \wtfifo column.
+  In fact, $O_{2,2}$ waits for $5$ time units for its machine~$2$ to get available,
+  its processing time is included in the waiting time~$8$ of $O_{3,3}$, and
+  $O_{2,3}$ also needs to wait $2$ time units for the completion of $O_{1,3}$
+  before its execution by machine~$1$.
+  %
+  % for which computation we used the same greedy algorithm as above with three different heuristics EST (\wtest), FIFO (\wtfifo), and MTWR (\wtmtwr).
 Each feature in \textit{WT} set is the difference between the scheduled starting time of a particular operation and the ending time of its predecessor.
-  As shown in Table~\ref{tab3}, the waiting time of $O_{2,2}$ is $5$ since the processing of its predecessor $O_{2,1}$ finishes at time $4$ and FIFO heuristic selects $O_{2,2}$ for scheduling to machine $2$ at time $9$, just as in the \stfifo case.
+  % As shown in Table~\ref{tab3}, the waiting time of $O_{2,2}$ is $5$ since the processing of its predecessor $O_{2,1}$ finishes at time $4$ and FIFO heuristic selects $O_{2,2}$ for scheduling to machine $2$ at time $9$, just as in the \stfifo case.
 \end{description}
 %
-All features extracted from a given JSP instance are provided to our decomposition method presented in the next section.
+We extract all of the features described above from a given JSP instance
+and can thus use them as inputs to our decomposition method presented in the next section.
 

### 2021-10-22 Second paper/samplepaper.tex
adding a draft of a paper

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{helvet,times,courier}
\usepackage{listings}
\usepackage[inline]{enumitem}
\lstset{numberbychapter=false,numbers=left,numberblanklines=false,basicstyle=\footnotesize\ttfamily,xleftmargin=5ex}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-0002-1088-2081} \and
Martin Gebser\inst{1,2}\orcidID{0000-0002-8010-4752} \and
Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958}}
%
\authorrunning{M. El-Kholany et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
\email{\{mohammed.el-kholany, martin.gebser and konstantin.schekotihin\}@aau.at}\\
%\url{http://www.springer.com/gp/computer-science/lncs} \and
Technische Universität Graz, Graz, Austria\\
\email{mgebser@ist.tugraz.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{Job-shop Scheduling Problem  \and Answer Set Programming \and Decomposition.}
\end{abstract}
%
%
%
\section{Introduction}
In today's competitive markets, manufacturers have to respond quickly to orders and meet shipping dates committed to the customers. This requires the ability to schedule production activities to use the available scarce resources efficiently. Effective scheduling techniques are essential in complex manufacturing systems such as semiconductor manufacturing. In general, scheduling operations is one of the most critical issues in the planning and managing of the manufacturing processes\cite{uzsoy2000performance}.

One of the most challenging scheduling problems is the Jop-shop Scheduling Problem (JSP). A set of jobs needs to be processed on a set of machines while optimizing a performance indicator, minimizing makespan, the time needed to complete all the jobs, or tardiness, i.e., the summation of the delays in executing all jobs according to their deadlines. Each job has a set of consecutive operations; each operation requires exactly one machine; machines are continuously available and can process only one operation at a time. The main purpose is to determine the sequence of the operations on each machine to optimize a performance indicator. The JSP is an NP-complete combinatorial optimization problem where obtaining the optimal solution is challenging regardless of the problem scale\cite{garey1976complexity,lenstra1977complexity,liu2008prediction}. For instance, benchmark (FT10) consists of $10$ jobs and $10$ machines took researchers roughly $20$ years to find the optimal solution\cite{adams1988shifting,zhang2010hybrid}.

In reality, the number of operations is reaching thousands of operations. Therefore, the exact methods Answer Set Programming (ASP), Constraint Programming and Branch and Bound cannot find the optimal solutions in a reasonable time\cite{daneshamooz2021mathematical,shi2021solving,francescutto2021solving}. Since the performance of the exact methods will be hardly satisfactory, several studies have presented one of the most effective methods for solving the JSP, which is decomposition \cite{zhang2010hybrid}. The decomposition aims to split the problem into a series of subproblems based on a particular decomposition policy and then solve each part separately and obtain the final solution by integrating all of these solutions. Several strategies have been proposed to decompose the JSP, and each varies according to the problem features. More specifically, there is no evidence that one of the introduced strategies is the best for solving all JSP(s) \cite{ovacik2012decomposition}.

In this study, we aim to introduce a new decomposition strategy for solving the JSP. Our study proposes a new policy to efficiently divide the problem into subproblems, aiming to reach a near-optimal solution. Then each part is solved using an ASP scheduler-based, and the solutions are integrated to obtain the solution of the whole problem. We develop and implement our model using ASP, which is considered as one of the most popular paradigms for knowledge representation and reasoning, especially in combinatorial optimization problems \cite{abseher2016shift}.

The paper is organized as follows. The following section shows the most related work and which decomposition techniques have been introduced in the past, followed by the problem formulation section illustrates the problem in detail. Section $4$ presents our proposed model, which shows the decomposition techniques we used and describes one in more detail. In addition, we present the schedule model using Answer Set Programming. Our model is tested by performing extensive experiments on a set of benchmark instances in Section $5$. Section $6$ summarizes this work and provides some ideas to extend the current work.

\section{Literature Review}
This section will review some of the work that has applied the decomposition approach for solving scheduling problems. As mentioned in the previous section, no particular decomposition method has proved its efficiency in solving all scheduling problems. However, many articles have introduced efficient decomposition methods for solving particular scheduling problems with specific features. The decomposition idea for solving JSP has been initially suggested by introducing a \textit{Shifting bottleneck} (SB) procedure which decomposes the problem into parts; each part contains only one machine \cite{adams1988shifting}. At each iteration, the bottleneck machine is determined from a set of unsequenced machines and then scheduled. Afterward, all the previously established sequences are locally reoptimized and iterate until all the machines have been sequenced. The idea of SB had been later improved by integrating an optimization algorithm for solving the one-machine problem with the delayed precedence constraints procedure \cite{balas1995one}.

Another study investigated the performance of a new decomposition procedure based on SB in the Job-shop and flow shop with different levels of bottleneck machines. The computational experiments showed that the proposed method obtained better solutions in a shorter time than for the problems in which the machine's workload is identical. Other different methods have been developed to decompose the JSP. For instance, a rolling horizon heuristic has been presented \cite{singer2001decomposition} to solve a large-scale job shop. The authors have decomposed the problem and solve each independently while minimizing the total weighted tardiness. They tested their model on a set of benchmark instances, and the results showed that the proposed model is superior to large instances.

The rolling horizon procedure has been extended by constructing a prediction model to obtain the scheduling characteristics values, including the information of the bottleneck jobs. The obtained information aided in decomposing the problem efficiently. In addition, they have proposed a genetic algorithm to solve each sub-problem. In order to evaluate the performance of the proposed model, they performed numerical computational experiments that showed the effectiveness of the model \cite{liu2008prediction}. A decomposition-based hybrid optimization algorithm has been introduced for the JSP, where the total weighted tardiness is minimized. A Simulated Annealing is used to define subproblems iteratively and then is solved by a Genetic Algorithm. They have developed a fuzzy system that provides information about bottleneck jobs. This information is used to guide the process of subproblem-solving to promote the optimization efficiency \cite{zhang2010hybrid}. The numerical computational results showed that the proposed algorithm is effective for large-scale scheduling problems.

A decomposition method based on bottleneck machines has been presented for solving the JSP. The proposed model decomposes the problems into subproblems and then detects the multi-bottleneck machines using a critical path method. The information of the bottleneck machines has been employed to improve the solution quality by splitting the operations into bottleneck operations and non-bottleneck operations, where the bottleneck operations are scheduled by the Genetic Algorithm and the non-bottleneck operations are scheduled by dispatching rules \cite{zhai2014decomposition}.

From the literature, we can find that most of the related work focused on applying the decomposition approach while minimizing the total tardiness, and the number of work tackled large-scale instances to minimize the makespan is relatively low. This paper proposes different decomposition strategies for solving the JSP with larger instances using the ASP model while minimizing the makespan.

\section{Preliminaries}
\paragraph{Answer Set Programming (ASP)} is a declarative programming language for solving hard combinatorial optimization problems. ASP has become an established paradigm for knowledge representation and reasoning. ASP has proved its efficiency for solving the combinatorial optimization problem in different applications such as bioinformatics\cite{erdem2015generating,koponen2015optimizing}, databses\cite{caniupan2010consistency} and scheduling and industrial applications~\cite{dodaro2015allotment,dodaro2016combining,fabricius2020towards,dodaro2021operating}.

A logic program is a finite set of rules of the form 
\begin{equation}
	a_0 \gets a_1, \ldots, a_m, \sim a_{m+1}, \dots, \sim a_n
	\label{rule1}
\end{equation}
where $a_i$ is an atom for $0 \leq i \leq n$ and ``$\sim$''  is a default negation. If $n$ quals to $0$, then a rule \eqref{rule1} is a fact. If $a_0$ is omitted, the rule is an integrity constraint. An \emph{atom} is an expression of the form $p(t_1,\ldots,t_l)$, where $p$ is a predicate and $t_1,\ldots,t_l$ are \emph{terms}. Each term can be a variable or a constant. A \emph{literal} is either an atom or its negation. 
Given a rule $r$ of the form \eqref{rule1}, the set $H(r)=a_0$ denotes the \emph{head} atom and the set $B(r) = B^+(r) \cup B^-(r) = \{a_1,\dots,a_m\} \cup \{\sim a_{m+1}, \dots \sim a_n\}$ is the \emph{body} of $r$ where $B^+(r)$ and $B^+(r)$ represent the positive and negative body literals, respectively.
\paragraph{Multi-shot Solving:}
Multi-shot solving ASP allows solving logic programs dynamically in sequential order. This can be manipulated using APIs implementation via an imperative programming language. Such programming language is used to control the grounding and solving processes and allows the usage of external variables that could be set to $True$ or $False$ to control some logic rules. The main advantage of \emph{multi-shot} solving is to avoid the re-grounding and exploit conflicts learned over time.
%ASP is a paradigm that deals with continuously changing logic programs. In a single-shot approach, an ASP system takes a logic program, computes answer sets and exits. However, the idea of Multi-shot is to consider evolving grounding and solving processes. \emph{clingo} systems enhances the ASP declaritive language with control capacities. 
\paragraph{ASP Difference Logic:}
\emph{clingo}[DL] is an extension of the input language \emph{clingo} by theory atoms representing difference constraints \cite{gebser2016theory,janhunen2017clingo}. Difference constraints are defined by specific constraint atoms of the form $\text{\lstinline{&diff}}\{u-v\} \leq k$ where $u$ and $v$ are terms which are interpreted as integer variables and $k$ is an integer constant. \emph{clingo}[DL] provides the following extension of the normal form \eqref{rule1}: 
\begin{align*}
	\text{\lstinline{&diff}}\{u-v\} \leq k \gets a_1, \ldots, a_m, \sim a_{m+1}, \ldots, \sim a_n.
\end{align*}
 This rule shows that whenever the body is $True$ the inequality constraints must be satisfied.

\section{Problem Formulation}
The Job-shop Scheduling Problem (JSP) has been a complex and combinatorial optimization problem since the 1950s, and it was shown to be NP-hard. In the JSP, there are a set of jobs to be processed on a set of machines. The number of jobs is $n$, and the number of machines is $m$. Each job $J_i$ contains a chain $ O_{i,1}, O_{i,2},...,O_{i,m} $ of operations must be executed on an known order. The routings of the operations are deterministic and known a priori, as are the processing times of each operation on each machine. We consider here the problem of minimizing the total completion time (makespan). The task of the scheduling is to determine the starting time for each operation while optimizing the makespan. The basic assumptions are as follows:

\begin{enumerate}
	\item The processing time of each operation is fixed.
	\item Machine breakdown preemption of the operations are not allowed.
	\item Each machine can process only one operation at a time.
	\item Each operation can be executed by only one machine.
	\item A machine cannot process more than one operation at a time.
	\item The jobs and the machines are available at time $0$.
\end{enumerate}

\subsection{Modelling JSP with Hybrid ASP}
ASP has been applied to solve different scheduling problems. For example, ASP is used to solve scheduling problems in the healthcare systems \cite{dodaro2017nurse,dodaro2019asp}. One of these studies aimed to assign nurses to shifts according to some constraints\cite{dodaro2017nurse}. In addition, ASP has been proposed to solve a train scheduling problem while minimizing the train delay\cite{abels2019train}. One of the limitations of ASP mentioned in these studies is the grounding issues while solving a large number of instances. One of the recent works has presented a multi-shot solving for scheduling problem \cite{francescutto2021solving}. This paper aims to use different decomposition methods to solve the JSP using ASP with Difference Logic.

\paragraph{Problem Instances.}

In order to encode the problem using ASP, we should define a set of predicates that represents the problem instances. For instance, the operations are encoded with the predicate \lstinline{operation/2} where the first variable denotes the job number and the second is the operation number, see lines \ref{prg:facts:ops:begin}-\ref{prg:facts:ops:end} in Listing \ref{prg:facts}. The time needed to finish an operation is represented by the predicate \lstinline{pro/3} in which the operation is determined by the first two terms and the third represents the processing time in lines \ref{prg:facts:pro:begin}-\ref{prg:facts:pro:end}. For the machine assignment, the predicate \lstinline{assign/3} shows that a particular operation is executed by a machine in lines \ref{prg:facts:mach:begin}-\ref{prg:facts:mach:end} For instance, \emph{operation(1,1)} is processed by machine 2. 
\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:facts,caption={Problem instance}, linerange={1-11} ]{listing/facts.lp}

\subsection{Decomposition strategies}
As we mentioned in the previous sections, it is hard to solve the JSP and reach a near-optimal solution in a reasonable time, especially when the number of operations to be processed is high. The decomposition approach has been widely used to solve the JSP by many researchers and proved its efficiency.

This section will mention $4$ different decomposition strategies that we developed to split the problem into parts, describe one of them in detail, and show the encoding. The decomposition strategies can be classified into two categories; time-based and machine-based. The time-based decomposition aims to prioritize the operations based on the processing time. However, machine-based decomposition methods consider the processing time of the operations and the machine's workload. The main idea behind the decomposition procedure is to find a criterion that ranks the operations to assign in the proper time window to obtain higher-quality schedules without violating the precedence constraints. Earliest Starting Time (EST) and Most Total Work Remaining (MTWR) procedures have been applied to rank the order the operations. The decomposition procedures will be shown as follows:

\paragraph{\textbf{EST Time-based}:} it ranks the operations based on calculating the earliest possible starting time for each operation. It is calculated by aggregating its predecessor(s) processing time. The operation with smaller EST will be assigned to a time window before the others with greater EST.

\paragraph{\textbf{MTWR Time-based}:} the operation rank is determined based on the work remaining of a job to be completed. The operation belongs to a job with a high remaining processing time, will be assigned earlier to a time window.

\paragraph{\textbf{EST Machine-based}:} it applies the same idea as EST Time-based. However, the bottleneck machine is taken into account throughout the decomposition process. More specifically, the operation with smaller EST and executed by a bottleneck machine will be assigned earlier to a time window.

\paragraph{\textbf{MTWR Machine-based}:} we calculate the MTWR of each operation, and the operation with the Largest MTWR and processed by a bottleneck machine will be assigned earlier to a time window.

In this study, we will describe in detail the decomposition of the operations based on the \textbf{EST Time-based}. We have split the decomposition encoding into two parts; the first part is for the pre-decomposition phase, which is in Listing \ref{prg:predeco}. The first two lines are to determine the number of time windows, which is two in this example. The lines \ref{prg:predeco:jmo:begin}-\ref{prg:predeco:jmo:end} are to calculate the total number of jobs and machines of a particular instance, respectively. Therefore, the total number of operations to be scheduled is computed by the rule in lines \ref{prg:predeco:o:begin}-\ref{prg:predeco:o:end}. In line \ref{prg:predeco:optw:begin}, the number of operations assigned to a time window is determined.

\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:predeco,caption={Pre-decomposition}, linerange={1-15} ]{listing/pre_deco.lp}
Listing \ref{prg:deco} shows the decomposition using \emph{EST Time-based} strategy. The first two rules in the lines \ref{prg:assignment_process:est:begin}-\ref{prg:assignment_process:est:end} calculate the earliest possible start time of each operation, where the EST for the first operation of all jobs are $0$ in line \ref{prg:assignment_process:est:begin}. The second rule calculates the other operations by aggregating the processing time of their predecessors.

We prioritize the operations based on the estimated starting time, calculating an index for each operation. The operation with a shorter estimated start time will get a smaller index than the others with a higher estimated starting time. If the estimated starting time between two operations or more is similar, we check the processing time; the higher the processing time, the smaller index, and, therefore, the higher priority to be assigned earlier. If the processing time is equal, we look at the operation number and then the job number. This strategy is encoded in lines \ref{prg:assignment_process:rank:begin}-\ref{prg:assignment_process:rank:end}. The fourth rule in lines \ref{prg:assignment_process:assign:begin}-\ref{prg:assignment_process:assign:end} is to assign each operation to a time winodw.

\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:deco,caption={EST-decomposition}, linerange={1-20} ]{listing/assignment_process.lp}

After running the decomposition encoding, we described above; we get an assignment of each operation to a TW. This assignment is represented by a set of atoms shown in the Listing \ref{prg:tw}. We can see that the operations operations $\{ O_{1,1}, O_{2,1}, O_{2,2}, O_{3,1}, O_{3,2} \}$ are assigned to the first TW and the rest to the second TW.

\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:tw,caption={Time Window Assignment}, linerange={1-7} ]{listing/TW.lp}

\subsection{Multi-shot with JSP}
This section will show how to solve the problem dynamically, where the output of each iteration is an input to the next and merging them to obtain the solution of the whole problem. In the first step, we consider the output of the decomposition phase, which is the assignment of the operations to TW as facts. For the multi-shot solving, the encoding is split into two parts; the base, which is run and grounded one time at the beginning of the optimization process presented in Listing \ref{prg:base}. The first rule is to determine the logical sequence between operations of same job in lines \ref{prg:base:same_job:begin}-\ref{prg:base:same_job:end} which is represented in predicate \lstinline{seqL/3}. On the other hand, Lines \ref{prg:base:sharing_machine:begin}-\ref{prg:base:sharing_machine:end} identify the sequence of the operations assigned to the same machine with predicate \lstinline{sameMach/4}.

\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:base,caption={base-prog}, linerange={1-12} ]{listing/base.lp}

The next iteration is to optimize each TW by solving the subproblem(t) where $t$ refers to the current TW we aim to optimize. The model starts to schedule and optimize the first TW and after a fixed amount of time, the solver stops and obtain the makespan of the current TW and the execution time of each operation represented by an atom \lstinline{startTime((Job, Step), ST, t)} where the first term is an operation, the second is the starting time and the last is the current time window. The solver moves to the next TW, where the starting time of the scheduled operations obtained from the first TW is sent entirely as an input to the next TW. Listing \ref{prg:sub} handles the sequence between the operations in different cases. The first rule in lines \ref{prg:sub_pro:logic_seq:begin}-\ref{prg:sub_pro:logic_seq:end} derives a new atom with a predicate \lstinline{seq/4} represents the logical sequence between two operations of a same job. In this rule, the first term of the head is the predecessor of a particular operation followed by the successor, the third term defines the minimum waiting time between the starting of the both, and the last term defines the TW of the successor. The second and third rules handle the case of two operations assigned to the same machine and in the same TW. On the other hand, if two operations are assigned to the same machine and different TW, the fourth and fifth rules ensure that the operation assigned to the earlier TW will be executed before the other is assigned to the subsequent TW. For instance, the fourth rule assumes that the \lstinline{operation (Job1, Step1)} is assigned to an earlier TW than \lstinline{operation (Job2, Step2)}.

\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:sub,caption={sub-prog}, linerange={1-40} ]{listing/sub_pro.lp}

The third part of the scheduler is shown in Listing \ref{prg:diff} that handles the difference constraints between the operations. The first rule in lines \ref{prg:diff_log:non_nega:begin}-\ref{prg:diff_log:non_nega:end} ensures that all the operations will be processed starting from time $0$. The second and the third rule fix the execution time of the operations scheduled in the previous TW. The fourth rule aimed to avoid overlapping the operations that either belongs to the same job or are assigned to the same machine. The fifth rule ensures that the ending time of all operations will not exceed the bound variable, which is the makespan in our case. 
\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:diff,caption={diff-log}, linerange={1-20} ]{listing/diff_log.lp}

\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation, etc., do not need, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\bibliographystyle{splncs04}
\bibliography{Ref}

\end{document}

### 2021-11-21 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
minor text modifications suggested by Reviewer 2 applied

@@
 Let us illustrate the problem on the small JSP instance specified in Table~\ref{tab1}, which provides parameters for $3$ jobs and $3$ machines. 
-The rows list operations along with their respective machines and processing times; e.g., the third operation of the first job, $O_{1,3}$, takes $5$ time units for execution by machine~$1$.
+The rows list operations along with their respective machines and processing times, e.g., the third operation of the first job, $O_{1,3}$, takes $5$ time units for execution by machine~$1$.
 The minimum makespan happens to be $20$, which matches the sum $9+6+5$ of processing times for the operations $O_{1,1}$, $O_{2,2}$, and $O_{3,3}$ executed by machine~$2$.
@@
 
-The clustering algorithms are unsupervised learning methods which goal is to partition a set of objects into (disjoint) subsets, called clusters, such that each cluster contains the most similar objects.
+Clustering algorithms are unsupervised learning methods whose goal is to partition a set of objects into (disjoint) subsets, called clusters, such that each cluster contains the most similar objects. 
 The similarity of objects is determined by some measure, like Euclidian distance, using features of each object, such as features of operations extracted in the previous section. 
@@
 In this section, we review some of the most related work using data mining approaches employed to solve scheduling problems.
-Over the last decades, solving the scheduling problem has attracted the interest of many researchers.The main idea of data mining in these applications is to explore the patterns from the obtained solutions by some heuristics and then develop decision rules that approximate the order of the operations in each machine \cite{ismail2012production}.
+Over the last decades, solving the scheduling problem has attracted the interest of many researchers.
+The main idea of data mining in these applications is to explore the patterns from the obtained solutions by some heuristics and then develop decision rules that approximate the order of the operations in each machine \cite{ismail2012production}.
 More precisely, it aims to get an order of all operations in the given instance according to their characteristics.

### 2021-11-21 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
working on Section 4

@@
 %In this section, we will show the implementation of our proposed model and how it works. 
-The proposed method comprises two main steps: 
+Our approximation approach comprises two main steps: 
 \begin{enumerate*}[label=\emph{(\roman*)}]
-  \item first, it decomposes a problem into a sequence of sub-problems, called \emph{time windows}; 
-  \item second, the sub-problems are solved one after another such that each subsequent sub-problem gets solutions of the previous sub-problems as input.
+  \item first, we decompose a problem instance into a sequence of sub-problems, called \emph{time windows}, and
+  \item second, the time windows are solved one after another,
+        where optimized solutions for preceding time windows are taken as input for solving the next.
 \end{enumerate*}
-As a result, the solution of the last sub-problem provides a complete schedule for the given JSP instance.
-The computational efficiency of the approach and the quality of obtained schedules depend highly on the quality of decomposition obtained in the first step \cite{zhang2010hybrid,zhai2014decomposition}.
+As a result, the solution obtained for the last time window provides a complete schedule for the given JSP instance.
+The computational efficiency of this approach and the quality of obtained schedules rely on the decomposition performed in the first step \cite{zhang2010hybrid,zhai2014decomposition}. 
 %To address this problem, several decomposition approaches have been presented during the last decades \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.
-Our approach uses a novel constraint clustering method to find a favorable decomposition of JSP instances using features of operations provided in them.
-Each cluster represents a time window, i.e., a subset of all operations that a scheduling algorithm must consider for the assignment.
+Our approach incorporates a novel constrained clustering method for the favorable decomposition of JSP instances based on features of their comprised operations.
+Each cluster represents a time window, i.e., a subset of operations to be scheduled together by an iterative solving algorithm.
 

### 2021-11-21 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
some more progress in Section 4

@@
 
-Clustering algorithms are unsupervised learning methods whose goal is to partition a set of objects into (disjoint) subsets, called clusters, such that each cluster contains the most similar objects. 
-The similarity of objects is determined by some measure, like Euclidian distance, using features of each object, such as features of operations extracted in the previous section. 
-However, direct application of clustering algorithms, like K-Means \cite{Forgy1965ClusterAO}, in scheduling is limited since the partitioning cannot preserve the sequences of operations in a job.
-As a result, a clustering algorithm might place $O_{1,1}$ and $O_{1,3}$ in one cluster and $O_{1,2}$ in another.
-As a result, the sequence of sub-problems becomes inconsistent and no schedule can be found.
+Clustering algorithms are unsupervised learning methods whose goal is to partition a set of objects into (disjoint) clusters, such that each cluster gathers objects of high similarity. 
+Such similarity is determined by some measure, e.g., Euclidean distance, based on features of each object, like the features of operations described in the previous section. 
+However, the direct application of common clustering algorithms, such as K-Means~\cite{Forgy1965ClusterAO}, to scheduling problems is impractical since the partitioning does not take the sequence of operations in a job into account.
+For instance, a clustering algorithm may put $O_{1,1}$ and $O_{1,3}$ into the same and $O_{1,2}$ into another cluster.
+As a result, the sequence of time windows becomes inconsistent, and no compatible schedule exists.
 

### 2021-11-22 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
Contiuing on Section 4

@@
 
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\algnewcommand{\To}{{\normalfont\bfseries to }}
+
 \newcommand{\stest}{\textit{ST\_EST}\xspace}
@@
 %
-\titlerunning{Decomposition-based JSP with Constrained Clustering}
+% \titlerunning{Decomposition-based JSP with Constrained Clustering}
 % If the paper title is too long for the running head, you can set
@@
 Our approach incorporates a novel constrained clustering method for the favorable decomposition of JSP instances based on features of their comprised operations.
-Each cluster represents a time window, i.e., a subset of operations to be scheduled together by an iterative solving algorithm.
+Each cluster represents a time window, i.e., a subset of operations to be scheduled in one go of an iterative solving algorithm.
 
-Clustering algorithms are unsupervised learning methods whose goal is to partition a set of objects into (disjoint) clusters, such that each cluster gathers objects of high similarity. 
+Clustering algorithms are unsupervised learning methods whose goal is to partition a set of data objects into (disjoint) clusters such that each cluster gathers objects of high similarity. 
 Such similarity is determined by some measure, e.g., Euclidean distance, based on features of each object, like the features of operations described in the previous section. 
@@
 
-Therefore, we propose a constrained clustering algorithm that preserves sequences of operations by considering the number of a cluster in which an operation is assigned.
-In particular, all predecessors of an operation placed in a cluster $i$ must be assigned to clusters $\{1, \dots, i\}$.
+We thus propose a constrained clustering algorithm that preserves sequences of operations by considering their order in the assignment to clusters.
+That is, the predecessors of an operation to be put into the $n$th cluster must be assigned to clusters $1,\dots,n$.
 %More specifically, it is not allowed to insert a data object (\textit{an operation}) in a particular cluster, and the predecessor of that operation is not assigned yet to the same cluster or the preceding cluster(s).
-Since our approach solves each sub-problem separately, the generation of large clusters might reduce its performance significantly.
-In the extreme case, finding a solution for a sub-problem might require the same time as the solution of the whole problem.
-Therefore, in addition to the similarity optimization criteria, our algorithm also aims to generate clusters with an equal number of operations assigned to each cluster.
-
-Our main procedure starts with the selection of a subset of the extracted features, which allows the clustering algorithm to compute the best possible time window assignments. 
-The procedure follows the forward feature selection principle: start with a small set of features and in every iteration extends it with one feature.
-Run the clustering algorithm on the new set of features and solve the obtained sequence of time windows.
-The subset of features resulting in the best possible solution quality is preserved and provided for extension in the next loop of the feature selection algorithm.
-The latter continues until either
-\begin{enumerate*}[label=\emph{(\roman*)}]
-  \item all features are included in the target set, or 
-  \item any extension of the provided subset of features results in solutions of worse quality than the solution of the given set.
-\end{enumerate*}
+Considering that our approach involves cluster-wise combinatorial optimization, the generation of large clusters might deteriorate the solving performance significantly.
+In the extreme case, all operations could be put into a single cluster representing the entire problem instance.
+% finding a solution for a sub-problem might require the same time as the solution of the whole problem.
+Hence, in addition to the similarity of operations, our decomposition method also aims at balancing the number of operations per cluster.
 
@@
 %\end{algorithm}
+%
+\begin{algorithm}[t]
+  \caption{Constrained Clustering Algorithm}\label{alg2}
+  \begin{algorithmic}
+  \Require $\mathit{operations}$, $\mathit{num\_clusters}$
+  %\Require $n \geq 0$
+  %\Ensure $y = x^n$
+  %\State $count\_clus \gets 0$  
+  \State $\mathit{cluster\_capacity} \gets \ceil*{\frac{|\mathit{operations}|}{\mathit{num\_clusters}}}$
+  \State Generate $\mathit{num\_clusters}$ many centroids
+%  \While{$count\_clus < num\_clus$}
+  \For{$n=1$ \To $\mathit{num\_clusters}$}
+      \State $\mathit{clusters}[n] \gets \emptyset$
+%      \State $count\_cap \gets 0$
+      \State $\mathit{current\_capacity} \gets \mathit{cluster\_capacity}$
+%      \State $curr\_clus \gets count\_clus + 1$
+%    \While{$count\_cap < clus\_capacity$}
+    \While{$0 < \mathit{current\_capacity}$}
+            \State Calculate distance between data objects and $n$th centroid   \Comment{Using Euclidean distance}
+            \State $O_{i,j} \gets \text{Nearest data object from }\mathit{operations}$ % $selected\_data\_obj \gets nearest\_data\_obj$
+            \Repeat
+            \State $\mathit{current\_capacity} \gets \mathit{current\_capacity} - 1$
+            \State $\mathit{operations} \gets \mathit{operations} \setminus \{O_{i,j}\}$
+            \State $\mathit{clusters}[n] \gets \mathit{clusters}[n] \cup \{O_{i,j}\}$ \Comment{Assigning operation $O_{i,j}$ to $n$th TW}
+            \State $j \gets j-1$
+            \Until{$O_{i,j}\notin \mathit{operations}$} \Comment{Satisfying the precedence constraint}
+            %\State $selected\_data\_obj \in curr\_clus$
+%            \State $curr\_clus \gets selected\_data\_obj$    \Comment{Assigning an operation to a TW}
 
-Algorithm~\ref{alg2} shows the pseudocode of the proposed clustering algorithm.
-Similar to K-Means, we assume that the algorithm gets the number of clusters as input.
-We calculate the cluster capacity by dividing the total number of operations by the number of clusters.
-In the next step, the clustering process starts by generating the centroids, where the number of centroids corresponds to the number of clusters.
-Each centroid in our case corresponds to some randomly selected operation from a set of operations allowing for preserving of sequencing constraints.
-For example, let us assume that a job consists of 15 operations, and the number of clusters is $3$.
-The first centroid will be an operation of order $\{1, \dots, 5\}$ in the sequence, the second centroid will be selected from operations with the order $\{6, \dots, 10\}$, and so forth.
-After generating the centroids, we check for exceeding the number of clusters and the cluster capacity, respectively.
+%            \If{$pred\_selected\_data\_obj \notin clus$}     \Comment{Satisfying the precedence constraint}
+%                \State $curr\_clus \gets pred\_selected\_data\_obj$
+%            \EndIf
+            \State Update the $n$th centroid % $update\_centroid$
+%            \State $count\_cap \gets count\_cap + 1 $
+    \EndWhile
+%      \State $count\_clus \gets count\_clus + 1 $
+%  \EndWhile
+  \EndFor
+%  \State \Return $\mathit{clusters}$
+  \end{algorithmic}
+  \end{algorithm}
+%
+Algorithm~\ref{alg2} provides a pseudocode description of our constrained clustering algorithm.
+Similar to K-Means, we assume that the algorithm gets the target number of clusters 
+into which the operations shall be partitioned as input.
+The cluster capacity, used for balancing the operations per cluster, is then obtained by dividing the total number of operations by the number of clusters.
+Moreover, the clustering process takes care of generating one initial centroid per cluster, given by randomly selected operations that are compatible with the precedence relation.
+For example, when each job consists of $15$ operations and the target number of clusters is $3$,
+the first centroid will be an operation at the first to fifth place of its job,
+the second an operation from place six to ten, and the third an operation at the eleventh
+or later place.
+% After generating the centroids, we check for exceeding the number of clusters and the cluster capacity, respectively. 
 %
@@
 These iterations continue until all operations are assigned to some cluster.
-
-\begin{algorithm}
-\caption{Clustering Algorithm}\label{alg2}
-\begin{algorithmic}
-\Require $num\_clus$
-%\Require $n \geq 0$
-%\Ensure $y = x^n$
-\State $count\_clus \gets 0$
-
-\State $clus\_capacity \gets \ceil*{\frac{num\_oper}{num\_clus}}$
-\State Generate centroids
-\While{$count\_clus < num\_clus$}
-    \State $count\_cap \gets 0$
-    \State $curr\_clus \gets count\_clus + 1$
-	\While{$count\_cap < clus\_capacity$}
-          \State Calculate distance between data objects and a centroid   \Comment{Using Euclidean distance}
-          \State $selected\_data\_obj \gets nearest\_data\_obj$
-          %\State $selected\_data\_obj \in curr\_clus$
-          \State $curr\_clus \gets selected\_data\_obj$    \Comment{Assigning an operation to a TW}
-
-          \If{$pred\_selected\_data\_obj \notin clus$}     \Comment{Satisfying the precedence constraint}
-              \State $curr\_clus \gets pred\_selected\_data\_obj$
-          \EndIf
-          \State $update\_centroid$
-          \State $count\_cap \gets count\_cap + 1 $
-	\EndWhile
-    \State $count\_clus \gets count\_clus + 1 $
-\EndWhile
-\end{algorithmic}
-\end{algorithm}
-
+%
 \iffalse
@@
 \fi
+
+Our main procedure starts with the selection of a subset of the extracted features, which allows the clustering algorithm to compute the best possible time window assignments. 
+The procedure follows the forward feature selection principle: start with a small set of features and in every iteration extends it with one feature.
+Run the clustering algorithm on the new set of features and solve the obtained sequence of time windows.
+The subset of features resulting in the best possible solution quality is preserved and provided for extension in the next loop of the feature selection algorithm.
+The latter continues until either
+\begin{enumerate*}[label=\emph{(\roman*)}]
+  \item all features are included in the target set, or 
+  \item any extension of the provided subset of features results in solutions of worse quality than the solution of the given set.
+\end{enumerate*}
 

### 2021-11-22 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
went through Section 4

@@
 We thus propose a constrained clustering algorithm that preserves sequences of operations by considering their order in the assignment to clusters.
-That is, the predecessors of an operation to be put into the $n$th cluster must be assigned to clusters $1,\dots,n$.
+That is, the predecessors of an operation to be put into the $n$th cluster must be assigned within the clusters $1,\dots,n$.
 %More specifically, it is not allowed to insert a data object (\textit{an operation}) in a particular cluster, and the predecessor of that operation is not assigned yet to the same cluster or the preceding cluster(s).
-Considering that our approach involves cluster-wise combinatorial optimization, the generation of large clusters might deteriorate the solving performance significantly.
+Also considering that our approach involves cluster-wise combinatorial optimization, the generation of large clusters risks to deteriorate the solving performance significantly.
 In the extreme case, all operations could be put into a single cluster representing the entire problem instance.
@@
 The cluster capacity, used for balancing the operations per cluster, is then obtained by dividing the total number of operations by the number of clusters.
-Moreover, the clustering process takes care of generating one initial centroid per cluster, given by randomly selected operations that are compatible with the precedence relation.
+Moreover, the clustering algorithm takes care of generating one initial centroid per cluster, given by randomly selected operations that are compatible with the precedence relation.
 For example, when each job consists of $15$ operations and the target number of clusters is $3$,
@@
 %
-Next, the algorithm calculates the distance between each operation and the centroid of the current cluster using Euclidean distance and assigns the nearest operation to that cluster.
-In order to satisfy the precedence constraint, we check the assignments of predecessors.
-If the constraint is satisfied, the algorithm assigns the operation to the cluster, and the centroid is updated.
-Whenever the maximum number of operations is reached, the algorithm continues with the next cluster.
-These iterations continue until all operations are assigned to some cluster.
+In order to populate each cluster, the algorithm inspects features to determine the Euclidean distance of each yet unassigned operation to the centroid of the current cluster and assigns the nearest operation to the cluster.
+To also preserve the precedence between operations, we additionally include any yet unassigned predecessor operations in the current cluster, and then update its centroid with the features of newly assigned operations.
+Whenever the cluster capacity is reached, the algorithm proceeds to the next cluster, and this decomposition process continues until all operations are assigned to clusters.
 %
@@
 
-Our main procedure starts with the selection of a subset of the extracted features, which allows the clustering algorithm to compute the best possible time window assignments. 
-The procedure follows the forward feature selection principle: start with a small set of features and in every iteration extends it with one feature.
-Run the clustering algorithm on the new set of features and solve the obtained sequence of time windows.
+In order to identify the most promising features for distance calculation among those introduced in the previous section,
+we applied the following forward selection principle:
+% Our main procedure starts with the selection of a subset of the extracted features, which allows the clustering algorithm to compute the best possible time window assignments. 
+% The procedure follows the forward feature selection principle:
+start with a small set of features, perform decomposition by Algorithm~\ref{alg2},
+and iteratively solve JSP instances.
+Then, we evaluate the possible extensions by one more feature,
+compare the quality of resulting schedules, and pick the best set of features. 
+% and in every iteration extends it with one feature.
+% Run the clustering algorithm on the new set of features and solve the obtained sequence of time windows.
 The subset of features resulting in the best possible solution quality is preserved and provided for extension in the next loop of the feature selection algorithm.
-The latter continues until either
+This process continues until either 
 \begin{enumerate*}[label=\emph{(\roman*)}]
-  \item all features are included in the target set, or 
-  \item any extension of the provided subset of features results in solutions of worse quality than the solution of the given set.
+  \item all features are selected, or 
+  \item any extension by another feature leads to solutions of lower quality.
 \end{enumerate*}

### 2021-11-23 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
middle of experiments section

@@
 \label{sec:eval}
-We use a popular benchmark \cite{taillard1993benchmarks} comprising randomly generated JSP instances as an evaluation dataset for our approach.
+We use a subset of Taillard's benchmark suite \cite{taillard1993benchmarks}, comprising $10$ randomly generated JSP instances known to be particularly hard for greedy and stochastic optimization techniques, for the empirical evaluation of our approach.
 %Many researchers have studied these instances and presented different approaches to solving them.
@@
 Therefore, it will be possible to measure the success/failure of our proposed model by comparing the results of the strategies developed in the past with those obtained from the proposed algorithm.
-For each instance, we extracted the 12 features described in Section \ref{sec:features}.
-In order to study the impact of the features on the clustering and, consequently, schedule quality, we store and present the output of the feature selection step and results of the decomposition-based solving using the Answer Set Programming model developed in \cite{el2020job}.
+For each instance, we extracted the $12$ features described in Section~\ref{sec:features}. In order to assess their impact % of the features
+on the clustering and, consequently, the quality of schedules, we make fix use of the features \stfifo, \stmtwr, and \stest, which provide the starting times obtained by
+heuristic greedy search methods, as these three features promised to be informative
+in preliminary experiments.
+In contrast to that, the position of operations in jobs (\textit{OP}),
+processing times (\textit{PT}), and jobs' time length (\textit{TLJ})
+CONTINUE HERE
+% we store and present the output of the feature selection step and results of the decomposition-based solving using the Answer Set Programming model developed in \cite{el2020job}.
 All experiments were executed with a timeout of 10 minutes. 
-The feature selection algorithm was initialized with a simple combination that consists of $3$ features: \stfifo, \stmtwr, and \stest.
+% The feature selection algorithm was initialized with a simple combination that consists of $3$ features: \stfifo, \stmtwr, and \stest.
 The combinations of the selected features presented below.
@@
 \fontsize{9pt}{10pt}
-\item \textbf{F1} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest \}}
-\item \textbf{F2} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, RPT \}}
-\item \textbf{F3} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, EST \}} 
-\item \textbf{F4} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, ML \}} 
-\item \textbf{F5} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, RPT, EST \}}
-\item \textbf{F6} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, RPT, ML \}} 
-\item \textbf{F7} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, EST, ML \}} 
-\item \textbf{F8} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest \}} 
-\item \textbf{F9} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT \}} 
-\item \textbf{F10} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST \}} 
-\item \textbf{F11} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, ML \}}
-\item \textbf{F12} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, EST \}}
-\item \textbf{F13} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, ML \}}
-\item \textbf{F14} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST, ML \}}
-\item \textbf{F15} \textrightarrow \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, EST, ML \}}
+\item \textbf{F1}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest \}}
+\item \textbf{F2}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, RPT \}}
+\item \textbf{F3}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, EST \}} 
+\item \textbf{F4}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, ML \}} 
+\item \textbf{F5}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, RPT, EST \}}
+\item \textbf{F6}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, RPT, ML \}} 
+\item \textbf{F7}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, EST, ML \}} 
+\item \textbf{F8}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest \}} 
+\item \textbf{F9}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT \}} 
+\item \textbf{F10} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST \}} 
+\item \textbf{F11} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, ML \}}
+\item \textbf{F12} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, EST \}}
+\item \textbf{F13} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, ML \}}
+\item \textbf{F14} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST, ML \}}
+\item \textbf{F15} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, EST, ML \}}
 \end{enumerate}
@@
   \begin{center}
-    \caption{The performance of selected features.}
+    \caption{Solution quality comparison for different sets of features.}
     \label{tab4}
-
-    \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|} \hline
-      & \multicolumn{11}{|c|}{\textbf{Instances}} \\ \cline{2-12}
-      \textbf{Approach}   & TA51                           & TA52                           & TA53                           & TA54 & TA55                           & TA56                           & TA57                           & TA58 & TA59                           & TA60 & AVG  \\ \hline
+    \addtolength{\tabcolsep}{1pt}
+    \begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|} \hline
+%      & \multicolumn{11}{|c|}{\textbf{Instances}} \\ \cline{2-12}
+      \textbf{Approach}   & TA51                           & TA52                           & TA53                           & TA54 & TA55                           & TA56                           & TA57                           & TA58 & TA59                           & TA60 & AVG  \\ \hline\hline
       \textbf{FIFO}        & 3549                           & 3339                           & \textbf{3160}              & \textbf{3218} & 3291               & 3325                           & 3654                           & \textbf{3299} & \textbf{3344}                           & \textbf{3129} & \textbf{3331} \\
       \textbf{MTWR}        & \textbf{3364}           & \textbf{3304}              & 3168                           & 3494 & \textbf{3237}                           & \textbf{3287}               & \textbf{3633}              & 3591 & 3394                           & 3257 & 3373 \\
-      \textbf{Single-shot} & 3632                           & 3615                           & 3481                           & 3462 & 3552                           & 3610                           & 3778                           & 3718 & 3613                           & 3550 & 3601 \\
-      \textbf{EST}         & 3506                           & 3773                           & 3478                           & 3497 & 3482                           & 3605                           & 3753                           & 3731 & 3398                           & 3247 & 3547 \\ \hline
+      \textbf{Single-shot} & 3632                           & 3615                           & 3481                           & 3462 & 3552                           & 3610                           & 3778                           & 3718 & 3613                           & 3550 & 3601 \\ % \hline
+%      \textbf{EST}         
+      \textbf{Multi-shot}         & 3506                           & 3773                           & 3478                           & 3497 & 3482                           & 3605                           & 3753                           & 3731 & 3398                           & 3247 & 3547 \\ \hline
       \textbf{F1}          & 3506                           & 3277                           & 3382                           & 3414 & 3308                           & 3353                           & 3605                           & 3352 & 3453                           & 3483 & 3413 \\
@@
       \textbf{F14}         & 3554                           & 3373                           & 3524                           & \textbf{3232} & 3385                           & 3483                           & 3573                           & 3365 & 3284                           & 3767 & 3454 \\
-      \textbf{F15}         & 3509                           & \textbf{3184} & 3380                           & 3289 & \textbf{3218} & 3389                           & \textbf{3480} & 3481 & 3290                           & 3569 & 3379 \\ \hline
-      \textbf{Obtained} & 3294                           & 3184 & 3143                           & 3232 & 3218 & 3229                           & 3480 & 3324 & 3145                           & 3315 & \textbf{3256} \\
+      \textbf{F15}         & 3509                           & \textbf{3184} & 3380                           & 3289 & \textbf{3218} & 3389                           & \textbf{3480} & 3481 & 3290                           & 3569 & 3379 \\ % \hline
+      \textbf{Virtual Best} & \textbf{3294}                           & \textbf{3184} & \textbf{3143}                           & \textbf{3232} & \textbf{3218} & \textbf{3229}                           & \textbf{3480} & \textbf{3324} & \textbf{3145}                           & \textbf{3315} & \textbf{3256} \\
 \hline

### 2021-11-23 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
pass through experiments sections

@@
 \usepackage{caption}
+\usepackage{url}
 \usepackage{todonotes}
@@
 In order to identify the most promising features for distance calculation among those introduced in the previous section,
-we applied the following forward selection principle:
+we suggest the following forward selection principle:
 % Our main procedure starts with the selection of a subset of the extracted features, which allows the clustering algorithm to compute the best possible time window assignments. 
@@
 \label{sec:eval}
-We use a subset of Taillard's benchmark suite \cite{taillard1993benchmarks}, comprising $10$ randomly generated JSP instances known to be particularly hard for greedy and stochastic optimization techniques, for the empirical evaluation of our approach.
+We use a subset of Taillard's benchmark suite \cite{taillard1993benchmarks}, comprising $10$ randomly generated JSP instances 
+with $50$ jobs, $15$ machines, and $50\times 15=750$ operations each that are
+known to be particularly hard for greedy and stochastic optimization techniques, for the empirical evaluation of our approach.
 %Many researchers have studied these instances and presented different approaches to solving them.
@@
 For each instance, we extracted the $12$ features described in Section~\ref{sec:features}. In order to assess their impact % of the features
-on the clustering and, consequently, the quality of schedules, we make fix use of the features \stfifo, \stmtwr, and \stest, which provide the starting times obtained by
+on the clustering and, consequently, the quality of schedules, we make fix use of the features \stfifo, \stmtwr, and \stest, which provide the starting times of operations obtained by
 heuristic greedy search methods, as these three features promised to be informative
 in preliminary experiments.
-In contrast to that, the position of operations in jobs (\textit{OP}),
+In contrast to that, considering the position of operations in jobs (\textit{OP}),
 processing times (\textit{PT}), and jobs' time length (\textit{TLJ})
-CONTINUE HERE
+turned out to be counterproductive in our preliminary investigation,
+so that we disregard such features in the following.
+The six leftover features, i.e., remaining processing time (\textit{RPT}),
+earliest starting time (\textit{EST}), machine load (\textit{ML})
+as well as the waiting times \wtfifo, \wtmtwr, and \wtest
+based on schedules computed by corresponding greedy algorithms,
+are added and combined to feature sets as follows:
+\todo{The feature set
+\textit{\{ \stfifo, \stmtwr, \stest, RPT, EST, ML \}} would be logical as \textbf{F8},
+but it's missing.}
 % we store and present the output of the feature selection step and results of the decomposition-based solving using the Answer Set Programming model developed in \cite{el2020job}.
-All experiments were executed with a timeout of 10 minutes. 
 % The feature selection algorithm was initialized with a simple combination that consists of $3$ features: \stfifo, \stmtwr, and \stest.
@@
 
-Table~\ref{tab4} shows the selected features and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
+We utilized the above feature sets for partitioning the operations of JSP instances into
+\todo{Please insert number of time windows.}$X$ clusters or time windows, respectively,
+and then ran the multi-shot optimization approach introduced in~\cite{el2020job} for
+$10$ minutes in total, i.e., \todo{Please insert minutes per time windows.}$X$ minutes per time window,
+on an
+\todo{Please check and update the machine and operating system setup.}
+Intel\textsuperscript{\textregistered} Xeon\textsuperscript{\textregistered} E5520 machine under Linux (Ubuntu 18.04.3).
+The resulting quality of schedules in terms of the achieved makespan for each instance is reported in Table~\ref{tab4},
+where the first four rows include the greedy FIFO and MTWR search methods supplied by the environment in~\cite{tagesc21a}
+as well as single-shot and multi-shot pure ASP optimization, the latter taking earliest starting times instead of 
+clustering for the decomposition into \todo{Please insert number of time windows.}$X$ time windows,
+for reference.
+Among these reference methods, FIFO and MTWR are on par, with a slight advantage of FIFO indicated
+by the average given in the last column, while combinatorial optimization by neither single- nor multi-shot ASP 
+solving is able to keep step and leads to greater makespans within the time limit for all instances.
+
+Among our feature sets for clustering, in preparation of multi-shot optimization by ASP,
+we do not identify any clear winner dominating over all instances.
+However, the feature sets \textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}, and \textbf{F15}
+yield significant makespan improvements relative to the other alternatives on particular instances.
+This means that considering the features \textit{RPT}, \textit{EST}, \textit{ML}, or
+\wtfifo, \wtmtwr, and \wtest in addition to starting times \stfifo, \stmtwr, and \stest by
+greedy algorithms is beneficial, while the particular combination of features achieving best schedules varies.
+Moreover, our clustering method for JSP instance decomposition consistently outperforms pure ASP optimization,
+and specific feature sets are ahead of the greedy FIFO and MTWR algorithms on seven out of ten instances,
+as indicated by the virtual best clustering results in the last row of Table~\ref{tab4}.
+We thus conclude that the instance decomposition by clustering can enable iterative constraint optimization
+to find better schedules than greedy algorithms in a short amount of time, where alternative feature sets deserve
+attention and trying several of them in parallel seems advisable.
+
+% Table~\ref{tab4} shows the selected features and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}.
 The first column shows the approaches used for solving the problem in the literature in addition to the combinations of features considered by our feature selection algorithm.

### 2021-11-26 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
adding anoher test set

@@
 \label{sec:eval}
-We use a subset of Taillard's benchmark suite \cite{taillard1993benchmarks}, comprising $10$ randomly generated JSP instances 
+We use a subset of Taillard's and Demirkol's benchmark suite \cite{taillard1993benchmarks,demirkol1998benchmarks}, comprising $10$ randomly generated JSP instances 
 with $50$ jobs, $15$ machines, and $50\times 15=750$ operations each that are
@@
 are added and combined to feature sets as follows:
-\todo{The feature set
-\textit{\{ \stfifo, \stmtwr, \stest, RPT, EST, ML \}} would be logical as \textbf{F8},
-but it's missing.}
+%\todo{The feature set
+%\textit{\{ \stfifo, \stmtwr, \stest, RPT, EST, ML \}} would be logical as \textbf{F8},
+%but it's missing.}
 % we store and present the output of the feature selection step and results of the decomposition-based solving using the Answer Set Programming model developed in \cite{el2020job}.
@@
 \item \textbf{F7}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, EST, ML \}}
-\item \textbf{F8}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest \}} 
-\item \textbf{F9}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT \}} 
-\item \textbf{F10} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST \}} 
-\item \textbf{F11} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, ML \}}
-\item \textbf{F12} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, EST \}}
-\item \textbf{F13} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, ML \}}
-\item \textbf{F14} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST, ML \}}
-\item \textbf{F15} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, EST, ML \}}
+\item  \textbf{F8}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, RPT, EST, ML \}}
+\item \textbf{F9}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest \}} 
+\item \textbf{F10}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT \}} 
+\item \textbf{F11} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST \}} 
+\item \textbf{F12} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, ML \}}
+\item \textbf{F13} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, EST \}}
+\item \textbf{F14} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, ML \}}
+\item \textbf{F15} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST, ML \}}
+\item \textbf{F16} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT, EST, ML \}}
 \end{enumerate}
@@
   \begin{center}
-    \caption{Solution quality comparison for different sets of features.}
+    \caption{Solution quality comparison for different sets of features (Taillard's instances).}
     \label{tab4}
@@
       \textbf{F7}          & 3328                           & 3226                           & 3338                           & 3353 & 3301                           & 3564                           & 3592                           & 3496 & 3352                           & 3399 & 3395 \\
-      \textbf{F8}          & 3588                           & 3397                           & 3251                           & 3563 & 3498                           & \textbf{3229} & 3621                           & 3517 & 3258                           & 3374 & 3430 \\
-      \textbf{F9}          & 3588                           & 3373                           & 3522                           & 3443 & 3385                           & 3306                           & 3543                           & 3716 & 3381                           & 3526 & 3475 \\
-      \textbf{F10}         & 3568                           & 3533                           & 3373                           & 3386 & 3530                           & 3362                           & 3608                           & 3644 & 3402                           & 3637 & 3504 \\
-      \textbf{F11}         & 3514                           & 3256                           & 3585                           & 3352 & 3365                           & 3344                           & 3670                           & 3611 & \textbf{3145} & 3425 & 3427 \\
-      \textbf{F12}         & 3682                           & 3494                           & \textbf{3143} & 3270 & 3309                           & 3332                           & 3632                           & 3471 & 3250                           & 3676 & 3426 \\
-      \textbf{F13}         & 3624                           & 3505                           & 3427                           & 3236 & 3370                           & 3338                           & 3749                           & 3437 & 3253                           & 3659 & 3460 \\
-      \textbf{F14}         & 3554                           & 3373                           & 3524                           & \textbf{3232} & 3385                           & 3483                           & 3573                           & 3365 & 3284                           & 3767 & 3454 \\
-      \textbf{F15}         & 3509                           & \textbf{3184} & 3380                           & 3289 & \textbf{3218} & 3389                           & \textbf{3480} & 3481 & 3290                           & 3569 & 3379 \\ % \hline
+
+      \textbf{F8}          & 3330                           & 3273                           & 3277                           & 3492 & 3375                           & 3469                           & 3760                           & 3410 & 3485                           & 3573 & 3444 \\
+
+      \textbf{F9}          & 3588                           & 3397                           & 3251                           & 3563 & 3498                           & \textbf{3229} & 3621                           & 3517 & 3258                           & 3374 & 3430 \\
+      \textbf{F10}          & 3588                           & 3373                           & 3522                           & 3443 & 3385                           & 3306                           & 3543                           & 3716 & 3381                           & 3526 & 3475 \\
+      \textbf{F11}         & 3568                           & 3533                           & 3373                           & 3386 & 3530                           & 3362                           & 3608                           & 3644 & 3402                           & 3637 & 3504 \\
+      \textbf{F12}         & 3514                           & 3256                           & 3585                           & 3352 & 3365                           & 3344                           & 3670                           & 3611 & \textbf{3145} & 3425 & 3427 \\
+      \textbf{F13}         & 3682                           & 3494                           & \textbf{3143} & 3270 & 3309                           & 3332                           & 3632                           & 3471 & 3250                           & 3676 & 3426 \\
+      \textbf{F14}         & 3624                           & 3505                           & 3427                           & 3236 & 3370                           & 3338                           & 3749                           & 3437 & 3253                           & 3659 & 3460 \\
+      \textbf{F15}         & 3554                           & 3373                           & 3524                           & \textbf{3232} & 3385                           & 3483                           & 3573                           & 3365 & 3284                           & 3767 & 3454 \\
+      \textbf{F16}         & 3509                           & \textbf{3184} & 3380                           & 3289 & \textbf{3218} & 3389                           & \textbf{3480} & 3481 & 3290                           & 3569 & 3379 \\ % \hline
       \textbf{Virtual Best} & \textbf{3294}                           & \textbf{3184} & \textbf{3143}                           & \textbf{3232} & \textbf{3218} & \textbf{3229}                           & \textbf{3480} & \textbf{3324} & \textbf{3145}                           & \textbf{3315} & \textbf{3256} \\
@@
 
-We utilized the above feature sets for partitioning the operations of JSP instances into
-\todo{Please insert number of time windows.}$X$ clusters or time windows, respectively,
+We utilized the above feature sets for partitioning the operations of JSP instances into $3$ clusters or time windows, respectively,
 and then ran the multi-shot optimization approach introduced in~\cite{el2020job} for
-$10$ minutes in total, i.e., \todo{Please insert minutes per time windows.}$X$ minutes per time window,
+$10$ minutes in total, i.e., $3.3$ minutes per time window,
 on an
-\todo{Please check and update the machine and operating system setup.}
-Intel\textsuperscript{\textregistered} Xeon\textsuperscript{\textregistered} E5520 machine under Linux (Ubuntu 18.04.3).
-The resulting quality of schedules in terms of the achieved makespan for each instance is reported in Table~\ref{tab4},
+%\todo{Please check and update the machine and operating system setup.}
+Intel\textsuperscript{\textregistered} Core(TM) i7-8650U CPU\textsuperscript{\textregistered} Dell Latitude 5590 under Windows 10.
+
+The resulting quality of schedules in terms of the achieved makespan for each instance is reported in Table~\ref{tab4} and Table~\ref{tab5},
 where the first four rows include the greedy FIFO and MTWR search methods supplied by the environment in~\cite{tagesc21a}
 as well as single-shot and multi-shot pure ASP optimization, the latter taking earliest starting times instead of 
-clustering for the decomposition into \todo{Please insert number of time windows.}$X$ time windows,
+clustering for the decomposition into $3$ time windows,
 for reference.
+\begin{table}[bt]
+  \begin{center}
+    \caption{Solution quality comparison for different sets of features (Demirkol's instances).}
+    \label{tab5}
+
+    \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|} \hline
+      & \multicolumn{11}{|c|}{\textbf{Instances}} \\ \cline{2-12}
+      \textbf{Approach}  & Dmu31  & Dmu32 & Dmu33 & Dmu34 & Dmu35 & Dmu71 & Dmu72 & Dmu73 & Dmu74 & Dmu75 & AVG  \\ \hline
+
+      \textbf{FIFO}  & \textbf{6817} & \textbf{6318} & \textbf{6029} & \textbf{6395} & \textbf{6409} & 9678 & 10349 & 9617 & 9847 & 9479 & 8094 \\
+      \textbf{MTWR}  & 7155 & 6042 & 6819 & 6570 & 6881 & 7764 & 8407 & 8411 & 8321 & 7893 & \textbf{7426} \\
+      \textbf{Single-shot} & 10370 & 10214 & 10064 & 10682 & 10203 & 6848 & 28369 & 32954 & 35325 & 6992 & 16202 \\
+      \textbf{EST} & 10318  & 9522  & 9785 & 10775 & 10398 & \textbf{6492}  & \textbf{6935} & \textbf{7202} & \textbf{7128}  & \textbf{6243} & 8480 \\ \hline
+      \textbf{F1}  & 8488  & 8325  & 7831 & 8356 & 8088 & 8593  & 7188 & 7826 & 6611 & 7662 & 7897 \\
+      \textbf{F2}  & 8440  & 8298  & \textbf{7663} & 8479 & 7986 & 6994  & 7606 & 7209 & 6627 & 7069 & 7637 \\
+      \textbf{F3}  & 8234  & 8401  & 8037 & 8444 & 7780 & \textbf{6590}  & 7785 & 7302 & 6690 & 7051 & 7631 \\
+      \textbf{F4}  & 7991  & 8589  & 7996 & 8471 & 8207 & 8091  & 7834 & 6972 & 6708 & 6859 & 7771 \\
+      \textbf{F5}  & 8164  & 8103  & 7835 & 8482 & 7802 & 6807  & 8357 & 7328 & 6774 & 7192 & 7684\\
+      \textbf{F6}  & \textbf{7709} & 8265 & 7795 & 8475 & \textbf{7678}  & 6611 & 6708 & 7030 &6569 & 6867 & \textbf{7371} \\
+      \textbf{F7}  & 8268  & 8553  & 7892 & 8359 & 8032 & 7603  & 6855 & 7223 & 6609 & 7200 & 7659 \\
+      \textbf{F8}  & 8873 & 8879 & 8201 & 9126 & 7982 & 6843 & 7971 & 7106 & 7545 & 6970 & 7950 \\
+      \textbf{F9}  & 8500  & 8185  & 8533 & 8772 & 8535 & 6726 & 6518& 7130 & 6592 & 7262 & 7675 \\
+      \textbf{F10}  & 8322  & 8119  & 8433 & 8910 & 8233 & 6868 & 6764 & 7183 & 6267 & 7459 & 7656 \\
+      \textbf{F11} & 8617  & 8383  & 8614 & 8755 & 7969 & 7587 & \textbf{6512} & \textbf{6762}& 6505 & 7376 & 7708 \\
+      \textbf{F12} & 8400  & 8124  & 8043 & \textbf{8185} & 8799 & 7849 & 6618 & 6876 & 6397 & 7746 & 7704 \\
+      \textbf{F13} & 8916 & 8343  & 8441 & 8331 & 8153 & 7464 & 6757 & 6994 & 6434 & 6961 & 7679 \\
+      \textbf{F14} & 8392  & \textbf{7937} & 7943 & 8375 & 8049 & 7310 & 7703 & 6827 & \textbf{6126} & 6893 & 7555 \\
+      \textbf{F15} & 8412  & 8210  & 8117 & 8563 & 9068 & 7853 & 6968 & 6909 & 6350  & 7011 & 7746 \\
+      \textbf{F16} & 8184 & 8150 & 8295 & 8721 & 8003 & 6997 & 7156 & 6991 & 6355 & \textbf{6584} & 7544 \\ 
+      \textbf{Virtual Best} & \textbf{6817} & \textbf{6318} & \textbf{6029} & \textbf{6395} & \textbf{6409} & \textbf{6492} & \textbf{6512} & \textbf{6762} & \textbf{6126} & \textbf{6243} & \textbf{6410} \\
+\hline
+      \end{tabular}
+  \end{center}
+\end{table}
+
 Among these reference methods, FIFO and MTWR are on par, with a slight advantage of FIFO indicated
@@
 we do not identify any clear winner dominating over all instances.
-However, the feature sets \textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}, and \textbf{F15}
+However, the feature sets \textbf{F6}, \textbf{F11}, \textbf{F14}, and \textbf{F16}
 yield significant makespan improvements relative to the other alternatives on particular instances.

### 2021-11-27 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
reformulations on first part of related work section

@@
 \label{sec:literature}
-In this section, we review some of the most related work using data mining approaches employed to solve scheduling problems.
+
+Data mining approaches~\cite{ismail2012production} have been extensively proposed and studied to generate % so-called
+dispatching rules~\cite{blackstone1982state} for scheduling problems.
+Unlike combinatorial and stochastic optimization methods, dispatching rules do not take
+complete schedules into account, but address the prioritization of operations for making
+local allocation decisions by greedy scheduling algorithms.
+In addition to Earliest Starting Time (EST), First-In-First-Out (FIFO), and
+Most Total Work Remaining (MTWR), which we also use for obtaining starting times as features,
+common dispatching rules include Last-In-First-Out (LIFO) and policies incorporating
+Job Length (JL) in terms of the number of operations as well as further features
+like those described in Section~\ref{sec:features}:
+Operation Position (OP), Processing Time (PT), Remaining Processing Time (RPT), 
+Time Length of a Job (TLJ), and Machine Load (ML).
+While dispatching rules enable an efficient operation allocation,
+even under real-time conditions, the quality of their decisions heavily depends on the
+problem instances under consideration, where data mining techniques come in to
+improve over tedious and ad hoc manual tuning.
+
+The approach of~\cite{koonce2000using} incorporates data mining methods,
+using OP, PT, RPT, and ML as features, for generating dispatching rules that approximate
+operation priorities found in schedules optimized by means of a genetic algorithm.
+On randomly generated test instances of comparably small size, i.e., $36$ operations
+per instance, the generated dispatching rules were shown to outperform greedy search
+with shortest PT as a rigid criterion for making allocation decisions.
+Genetic algorithms are also utilized in~\cite{harrath2002genetic} to obtain
+promising solutions for JSP instance and extract features to consider for the
+construction of dispatching rules.
+The specifically generated three-tiered policy comprises shortest PT, smallest JL, and longest RPT as
+dispatching criteria in the order of significance.
+Note that such a composite dispatching rule needs to reconstructed from scratch when
+new features or problem instances are analyzed, while our constrained clustering method
+can readily be applied to different instance or feature sets, respectively.
+
+Beyond static scheduling, dynamic JSP~\cite{french82a} deals with events, such as 
+the release of new jobs or sudden machine breakdowns, along with rescheduling or
+updating dispatching rules, respectively.
+In analogy to static approaches,
+the processing method of~\cite{shahzad2010discovering} applies tabu search to
+training instances and generates allocation policies approximating the optimized
+complete schedules based on features like EST, PT, and RPT of operations.
+Applied in the dynamic JSP setting, the dispatching rules obtained by means of data mining
+were shown to yield better or comparable performance to rigid allocation strategies from the
+literature.
+A rescheduling approach combining Variable Neighborhood Search (VNS)
+with K-Means clustering is presented in~\cite{adibi2014clustering},
+where clusters are updated to reconfigure the VNS on each event.
+The performance of this combined rescheduling strategy in the dynamic JSP setting turned out
+to be better than pure VNS and greedy algorithms with the FIFO, LIFO, or
+shortest PT allocation policy.
+
+% In this section, we review some of the most related work using data mining approaches employed to solve scheduling problems.
 Over the last decades, solving the scheduling problem has attracted the interest of many researchers.
@@
 
-Data mining methodologies were applied to explore the patterns in data generated by genetic algorithms in \cite{koonce2000using}.
+% Data mining methodologies were applied to explore the patterns in data generated by genetic algorithms in \cite{koonce2000using}.
 This work aimed to generate rules determining the priority for each job.
@@
 
-The authors in \cite{harrath2002genetic} also implemented a genetic algorithm to construct and analyze solutions of JSP instances for the feature extraction for the suggested dispatching rule.
+% The authors in \cite{harrath2002genetic} also implemented a genetic algorithm to construct and analyze solutions of JSP instances for the feature extraction for the suggested dispatching rule.
 In particular, the approach uses processing time, job length, remaining processing time, the rank of the operation, and machine load.
@@
 
-In addition to the work focusing on static JSP, there are approaches using data mining methods for dynamic JSP.
+% In addition to the work focusing on static JSP, there are approaches using data mining methods for dynamic JSP.
 For instance, \cite{shahzad2010discovering} presents new dispatching rules for JSP aiming to minimize makespan and which process a given JSP instance as follows:
-First, a scheduling instance is solved using Tabu Search and the obtained solutions are used as a training set to get a rule-set capable of approximating efficient solutions.
+% First, a scheduling instance is solved using Tabu Search and the obtained solutions are used as a training set to get a rule-set capable of approximating efficient solutions.
 This approach considers features like EST, the difference in remaining processing time for a job, and the difference in the processing time of two operations, to create a rule-set for dispatching the operations.
@@
 
-The variable neighborhood search (VNS) is combined with the k-Means algorithm to address Dynamic JSP \cite{adibi2014clustering}.
+% The variable neighborhood search (VNS) is combined with the k-Means algorithm to address Dynamic JSP \cite{adibi2014clustering}.
 The proposed algorithm depends on extracting information from the problem's input data when an event, like a machine breakdown or an addition of a new job, occurs.

### 2021-11-28 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
summary of metaheuristic approach in related work reformulated

@@
 
+The integration of dispatching rules generated with data mining techniques
+into stochastic optimization in~\cite{nasiri2019data} works by
+running a greedy algorithm called ``assignment procedure'' to obtain promising
+JSP solutions and mine dispatching rules approximating them.
+But instead of relying on the dispatching rules alone for computing optimized
+schedules or performing efficient operation allocation on new instances,
+the high-quality draft solutions from greedy search with the dispatching rules
+are taken as inputs to population-based stochastic optimization.
+This metaheuristic optimization was shown to yield improved outcomes when launched
+with schedules generated by means of suitable dispatching rules rather than with
+a randomly generated initial population.
+
 % In this section, we review some of the most related work using data mining approaches employed to solve scheduling problems.
@@
 
-Another study has presented a data mining approach to generate an initial population for population-based heuristics/metaheuristics to solve JSP \cite{nasiri2019data}.
+% Another study has presented a data mining approach to generate an initial population for population-based heuristics/metaheuristics to solve JSP \cite{nasiri2019data}.
 They extracted information from optimal or near-optimal solutions obtained by a heuristic procedure called "Assignment Procedure" to build decision rules to determine the priority of the operations.

### 2021-11-28 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
related work section extended

@@
 %
-\title{Decomposition-based Job-shop Scheduling with Constrained Clustering}%\thanks{Supported by organization x.}}
+\title{Decomposition-based Job-shop Scheduling with Constrained Clustering
+\thanks{This work was partially funded by
+KWF project 28472,
+cms electronics GmbH,
+FunderMax GmbH,
+Hirsch Armbänder GmbH,
+incubed IT GmbH,
+Infineon Technologies Austria AG,
+Isovolta AG,
+Kostwein Holding GmbH, and
+Privatstiftung Kärntner Sparkasse.}}%\thanks{Supported by organization x.}}
 %
@@
 as indicated by the virtual best clustering results in the last row of Table~\ref{tab4}.
-We thus conclude that the instance decomposition by clustering can enable iterative constraint optimization
+We thus conclude that the instance decomposition by clustering can enable iterative combinatorial optimization
 to find better schedules than greedy algorithms in a short amount of time, where alternative feature sets deserve
@@
 
+Similarities of our cluster-wise JSP solving approach to the discussed methods include
+that we apply (fixed) dispatching rules to extract the starting times of operations
+in corresponding schedules as instance features.
+Notably, the constrained clustering algorithm in Section~\ref{sec:method} can readily
+be applied to arbitrary feature sets for decomposing the operations of JSP instances
+into balanced time windows, i.e., our clustering method does not presuppose tuning to
+specific instances or features under consideration.
+The iterative combinatorial optimization by time windows can be viewed as a trade-off
+between efficient dispatching rules, aiming to optimize local allocation decisions by
+prioritizing operations, and the stochastic or combinatorial optimization of
+complete schedules for problem instances.
+While instance size and complexity can make the search for optimal solutions
+with state-of-the-art solvers for ASP, Mixed Integer Programming, or Constraint Programming
+prohibitive~\cite{zhang2010hybrid},
+their real-world application areas in scheduling include, e.g.,
+industrial printing~\cite{balduccini11a},
+team-building~\cite{rigralmaliiile12a},
+shift design~\cite{abgemuscwo15c},
+course timetabling~\cite{bainkaokscsotawa18a},
+lab resource allocation~\cite{frscel21a},
+and
+medical treatment planning~\cite{dogagrmamopo21a}.
+
 % In this section, we review some of the most related work using data mining approaches employed to solve scheduling problems.
@@
 
-From the literature, we have found that most of the studies which used data mining techniques aimed to extract some features either from the problem instances or solutions obtained by some heuristics to build a decision-rules to order the operations assigned to a machine.
+% From the literature, we have found that most of the studies which used data mining techniques aimed to extract some features either from the problem instances or solutions obtained by some heuristics to build a decision-rules to order the operations assigned to a machine.
 In most cases, these approaches depend on the set of extracted features and cannot add new or remove the existing ones.

### 2021-11-29 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
Adding my Second affiliation

@@
 %
-\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-0002-1088-2081} \and
+\author{Mohammed M. S. El-Kholany\inst{1,3}\orcidID{0000-0002-1088-2081} \and
 Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958} \and
@@
 Austria\\%
+\and
+Faculty of Computers and Artificial Intelligence, Cairo University, 
+Egypt\\%
 % \email{mgebser@ist.tugraz.at}

### 2021-11-29 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
unifying tables for experiments

--- 2021-11-28 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 \item \textbf{F9}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest \}} 
-\item \textbf{F10}\phantom{\textbf{0}} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT \}} 
+\item \textbf{F10} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, RPT \}} 
 \item \textbf{F11} \textrightarrow{} \textit{\{ \stfifo, \stmtwr, \stest, \wtfifo, \wtmtwr, \wtest, EST \}} 
@@
   \begin{center}
-    \caption{Solution quality comparison for different sets of features (Taillard's instances).}
+    \caption{Solution quality comparison for different feature sets (Taillard's instances).}
     \label{tab4}
-    \addtolength{\tabcolsep}{1pt}
+    \addtolength{\tabcolsep}{0.9pt}
     \begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|} \hline
@@
   \begin{center}
-    \caption{Solution quality comparison for different sets of features (Demirkol's instances).}
+    \caption{Solution quality comparison for different feature sets (Demirkol's instances).}
     \label{tab5}
+    \addtolength{\tabcolsep}{0.9pt}
+    \begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|} \hline
+%      & \multicolumn{11}{|c|}{\textbf{Instances}} \\ \cline{2-12}
+      \textbf{Approach}  & DE31  & DE32 & DE33 & DE34 & DE35 & DE71 & DE72 & DE73 & DE74 & DE75 & AVG  \\ \hline
 
-    \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|} \hline
-      & \multicolumn{11}{|c|}{\textbf{Instances}} \\ \cline{2-12}
-      \textbf{Approach}  & Dmu31  & Dmu32 & Dmu33 & Dmu34 & Dmu35 & Dmu71 & Dmu72 & Dmu73 & Dmu74 & Dmu75 & AVG  \\ \hline
-
-      \textbf{FIFO}  & \textbf{6817} & \textbf{6318} & \textbf{6029} & \textbf{6395} & \textbf{6409} & 9678 & 10349 & 9617 & 9847 & 9479 & 8094 \\
-      \textbf{MTWR}  & 7155 & 6042 & 6819 & 6570 & 6881 & 7764 & 8407 & 8411 & 8321 & 7893 & \textbf{7426} \\
+      \textbf{FIFO}  & \textbf{6817} & 6318 & \textbf{6029} & \textbf{6395} & \textbf{6409} & 9678 & 10349 & 9617 & 9847 & 9479 & 8094 \\
+      \textbf{MTWR}  & 7155 & \textbf{6042} & 6819 & 6570 & 6881 & 7764 & 8407 & 8411 & 8321 & 7893 & \textbf{7426} \\
       \textbf{Single-shot} & 10370 & 10214 & 10064 & 10682 & 10203 & 6848 & 28369 & 32954 & 35325 & 6992 & 16202 \\
-      \textbf{EST} & 10318  & 9522  & 9785 & 10775 & 10398 & \textbf{6492}  & \textbf{6935} & \textbf{7202} & \textbf{7128}  & \textbf{6243} & 8480 \\ \hline
+      \textbf{Multi-shot} & 10318  & 9522  & 9785 & 10775 & 10398 & \textbf{6492}  & \textbf{6935} & \textbf{7202} & \textbf{7128}  & \textbf{6243} & 8480 \\ \hline
       \textbf{F1}  & 8488  & 8325  & 7831 & 8356 & 8088 & 8593  & 7188 & 7826 & 6611 & 7662 & 7897 \\
@@
       \textbf{F16} & 8184 & 8150 & 8295 & 8721 & 8003 & 6997 & 7156 & 6991 & 6355 & \textbf{6584} & 7544 \\ 
-      \textbf{Virtual Best} & \textbf{6817} & \textbf{6318} & \textbf{6029} & \textbf{6395} & \textbf{6409} & \textbf{6492} & \textbf{6512} & \textbf{6762} & \textbf{6126} & \textbf{6243} & \textbf{6410} \\
+      \textbf{Virtual Best} & \textbf{7709} & \textbf{7937} & \textbf{7663} & \textbf{8185} & \textbf{7678} & \textbf{6590} & \textbf{6512} & \textbf{6762} & \textbf{6126} & \textbf{6584} & \textbf{7175} \\
 \hline

### 2021-11-29 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
re-reading the whole paper

--- 2021-11-29 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 
-In this work, we introduce a method based on clustering to automatically decomposes a JSP instance into several time windows that are small enough for optimization by an ASP solver.
+In this work, we introduce a method based on clustering to automatically decompose a JSP instance into several time windows that are small enough for optimization by an ASP solver.
 General clustering algorithms are unsupervised learning methods that partition a given set of objects into disjoint clusters, where each cluster comprises close objects wrt.\ some distance measure. 
@@
   \item We implement the proposed constrained clustering algorithm and combine it with the forward selection of features, which is an automatic method for identifying a subset of features allowing the clustering method to compute decompositions resulting in best-quality schedules.
-  \item The evaluation of our approach on Taillard's benchmark instances \cite{taillard1993benchmarks} shows that it significantly outperforms baseline heuristic methods and pure ASP optimization wrt.\ the makespan optimization criterion within a timeout limit of 600 seconds.
+  \item The evaluation of our approach on Taillard's and Demirkol’s benchmark instances \cite{taillard1993benchmarks, demirkol1998benchmarks} shows that it significantly outperforms baseline heuristic methods and pure ASP optimization wrt.\ the makespan optimization criterion within a timeout limit of 600 seconds.
 \end{itemize}
@@
 as indicated by the virtual best clustering results in the last row of Table~\ref{tab4}.
+In Table~\ref{tab5}, the feature sets outperformed the greedy on only three out of ten instances.
+However, feature set \textbf{F6} provided the best obtained average.
 We thus conclude that the instance decomposition by clustering can enable iterative combinatorial optimization

### 2021-11-29 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
checking experiments

--- 2021-11-29 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
@@
 \and
-Faculty of Computers and Artificial Intelligence, Cairo University, 
+% Faculty of Computers and Artificial Intelligence,
+Cairo University, 
 Egypt\\%
@@
 \label{sec:eval}
-We use a subset of Taillard's and Demirkol's benchmark suite \cite{taillard1993benchmarks,demirkol1998benchmarks}, comprising $10$ randomly generated JSP instances 
-with $50$ jobs, $15$ machines, and $50\times 15=750$ operations each that are
-known to be particularly hard for greedy and stochastic optimization techniques, for the empirical evaluation of our approach.
+We use randomly generated JSP instances
+with $50$ jobs, $15$ machines, and $50\times 15=750$ operations,
+part of which are known to be challenging for greedy and stochastic optimization techniques,
+from Taillard's and Demirkol's benchmark suites \cite{taillard1993benchmarks,demirkol1998benchmarks} for the empirical evaluation of our approach.
 %Many researchers have studied these instances and presented different approaches to solving them.

### 2021-11-29 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
rechecked experiments

@@
   \item We implement the proposed constrained clustering algorithm and combine it with the forward selection of features, which is an automatic method for identifying a subset of features allowing the clustering method to compute decompositions resulting in best-quality schedules.
-  \item The evaluation of our approach on Taillard's and Demirkol’s benchmark instances \cite{taillard1993benchmarks, demirkol1998benchmarks} shows that it significantly outperforms baseline heuristic methods and pure ASP optimization wrt.\ the makespan optimization criterion within a timeout limit of 600 seconds.
+  \item The evaluation of our approach on Taillard's and Demirkol's benchmark instances \cite{taillard1993benchmarks, demirkol1998benchmarks} shows that it significantly outperforms baseline heuristic methods and pure ASP optimization wrt.\ the makespan optimization criterion within a timeout limit of 600 seconds.
 \end{itemize}
@@
 and then ran the multi-shot optimization approach introduced in~\cite{el2020job} for
-$10$ minutes in total, i.e., $3.3$ minutes per time window,
+$10$ minutes in total, i.e., $200$ seconds per time window,
 on an
 %\todo{Please check and update the machine and operating system setup.}
-Intel\textsuperscript{\textregistered} Core(TM) i7-8650U CPU\textsuperscript{\textregistered} Dell Latitude 5590 under Windows 10.
-
+Intel\textsuperscript{\textregistered} Core\texttrademark{} i7-8650U CPU % \textsuperscript{\textregistered}
+Dell Latitude 5590 machine under Windows 10.
+%
 The resulting quality of schedules in terms of the achieved makespan for each instance is reported in Table~\ref{tab4} and Table~\ref{tab5},
@@
 for reference.
+%
+Regarding these reference methods, FIFO and MTWR are more or less on par, with a slight advantage of FIFO on Taillard's instances and an edge for MTWR on Demirkol's instances, 
+indicated
+by the average makespans given in the last column of each table.
+While combinatorial optimization by neither single- nor multi-shot ASP 
+solving is able to keep step within the time limit and leads to greater averages over all instances, the multi-shot pure ASP optimization approach turns out to be comparably robust
+on half of the instances from Demirkol's benchmark suite,
+which points out the strong impact of instance patterns.
+%
 \begin{table}[bt]
@@
 
-Among these reference methods, FIFO and MTWR are on par, with a slight advantage of FIFO indicated
-by the average given in the last column, while combinatorial optimization by neither single- nor multi-shot ASP 
-solving is able to keep step and leads to greater makespans within the time limit for all instances.
-
 Among our feature sets for clustering, in preparation of multi-shot optimization by ASP,
 we do not identify any clear winner dominating over all instances.
-However, the feature sets \textbf{F6}, \textbf{F11}, \textbf{F14}, and \textbf{F16}
+However, the feature sets \textbf{F2}, \textbf{F6}, \textbf{F11}, \textbf{F12}, \textbf{F14}, and \textbf{F16}
 yield significant makespan improvements relative to the other alternatives on particular instances.
@@
 greedy algorithms is beneficial, while the particular combination of features achieving best schedules varies.
-Moreover, our clustering method for JSP instance decomposition consistently outperforms pure ASP optimization,
-and specific feature sets are ahead of the greedy FIFO and MTWR algorithms on seven out of ten instances,
-as indicated by the virtual best clustering results in the last row of Table~\ref{tab4}.
-In Table~\ref{tab5}, the feature sets outperformed the greedy on only three out of ten instances.
-However, feature set \textbf{F6} provided the best obtained average.
+Moreover, our clustering method for JSP instance decomposition outperforms (multi-shot) pure ASP optimization,
+and specific feature sets are ahead of the greedy FIFO and MTWR algorithms on seven out of ten Taillard's instances as well as half of Demirkol's instances,
+as indicated by the virtual best clustering results in the last row of Table~\ref{tab4}
+and Table~\ref{tab5}.
+Notably, the feature set \textbf{F6} leads to the shortest average makespans for
+our cluster-wise JSP solving approach on both kinds of instances, comes close to
+FIFO on Taillard's and is most robust on Demirkol's instances, but the
+virtual best clustering results still improve significantly.
+% In Table~\ref{tab5}, the feature sets outperformed the greedy on only three out of ten instances. However, feature set \textbf{F6} provided the best obtained average. 
 We thus conclude that the instance decomposition by clustering can enable iterative combinatorial optimization

### 2021-11-29 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
reread Sections 5 and 6

@@
   \item We implement the proposed constrained clustering algorithm and combine it with the forward selection of features, which is an automatic method for identifying a subset of features allowing the clustering method to compute decompositions resulting in best-quality schedules.
-  \item The evaluation of our approach on Taillard's and Demirkol's benchmark instances \cite{taillard1993benchmarks, demirkol1998benchmarks} shows that it significantly outperforms baseline heuristic methods and pure ASP optimization wrt.\ the makespan optimization criterion within a timeout limit of 600 seconds.
+  \item The evaluation of our approach on Taillard's and Demirkol's benchmark instances \cite{taillard1993benchmarks, demirkol1998benchmarks} shows that it significantly outperforms baseline heuristic methods and pure ASP optimization wrt.\ the makespan optimization criterion within a short solving time limit of $10$ minutes.
 \end{itemize}
@@
 on half of the instances from Demirkol's benchmark suite,
-which points out the strong impact of instance patterns.
+which points out the critical impact of instance patterns.
 %
@@
 we do not identify any clear winner dominating over all instances.
-However, the feature sets \textbf{F2}, \textbf{F6}, \textbf{F11}, \textbf{F12}, \textbf{F14}, and \textbf{F16}
+However, the feature sets \textbf{F2}, \textbf{F6}, \textbf{F11}, \textbf{F12}, \textbf{F13}, \textbf{F14}, and \textbf{F16}
 yield significant makespan improvements relative to the other alternatives on particular instances.
@@
 % In Table~\ref{tab5}, the feature sets outperformed the greedy on only three out of ten instances. However, feature set \textbf{F6} provided the best obtained average. 
-We thus conclude that the instance decomposition by clustering can enable iterative combinatorial optimization
+We thus conclude that the instance decomposition by clustering can empower iterative combinatorial optimization
 to find better schedules than greedy algorithms in a short amount of time, where alternative feature sets deserve
@@
 
-Data mining approaches~\cite{ismail2012production} have been extensively proposed and studied to generate % so-called
+Data mining approaches~\cite{ismail2012production} have been extensively proposed and applied to generate % so-called
 dispatching rules~\cite{blackstone1982state} for scheduling problems.
@@
 Genetic algorithms are also utilized in~\cite{harrath2002genetic} to obtain
-promising solutions for JSP instance and extract features to consider for the
+promising solutions for JSP instances and extract features to consider for the
 construction of dispatching rules.
@@
 updating dispatching rules, respectively.
-In analogy to static approaches,
-the processing method of~\cite{shahzad2010discovering} applies tabu search to
+In analogy to the static setting,
+the processing approach of~\cite{shahzad2010discovering} applies tabu search to
 training instances and generates allocation policies approximating the optimized
@@
 into balanced time windows, i.e., our clustering method does not presuppose tuning to
-specific instances or features under consideration.
+specific instance or feature sets under consideration.
 The iterative combinatorial optimization by time windows can be viewed as a trade-off
@@
 with state-of-the-art solvers for ASP, Mixed Integer Programming, or Constraint Programming
-prohibitive~\cite{zhang2010hybrid},
-their real-world application areas in scheduling include, e.g.,
+prohibitive~\cite{coltep19a,zhang2010hybrid},
+successful % their real-world 
+application areas in scheduling include, e.g.,
 industrial printing~\cite{balduccini11a},

### 2021-11-29 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
revised conclusions

@@
 as offered by ASP~\cite{gekakasc17a},
-can first optimize a schedule (wrt.\ to the makespan)
+can first optimize a schedule (wrt.\ the makespan)
 for the operations in TW~$1$, and then extend the first part by
@@
 %In this section, we will show the implementation of our proposed model and how it works. 
-Our approximation approach comprises two main steps: 
+Our approximation approach comprises two main phases: 
 \begin{enumerate*}[label=\emph{(\roman*)}]
@@
 As a result, the solution obtained for the last time window provides a complete schedule for the given JSP instance.
-The computational efficiency of this approach and the quality of obtained schedules rely on the decomposition performed in the first step \cite{zhang2010hybrid,zhai2014decomposition}. 
+The computational efficiency of this approach and the quality of obtained schedules rely on the decomposition performed in the first phase \cite{zhang2010hybrid,zhai2014decomposition}. 
 %To address this problem, several decomposition approaches have been presented during the last decades \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.
@@
 \label{sec:conclusions}
-Data mining and machine learning is an area of interest in many manufacturing companies and other fields.
-In order to increase productivity and decrease the operation cost, there should be high-quality schedules to be used.
-This study focused on data mining and clustering approaches for solving the Job-shop Scheduling Problem.
-The first issue in this work is to explore the knowledge in good schedules and try to use this knowledge to get more efficient schedules.
+Data mining and machine learning are of wide interest in scientific, business, and further fields.
+% In order to increase productivity and decrease the operation cost, there should be high-quality schedules to be used.
+Our work investigates data mining and clustering methods to optimize solutions for the Job-shop Scheduling Problem within short solving times.
+The idea is to associate clusters of operations with time windows subject to iterative combinatorial optimization, eventually leading to a complete schedule.
+In contrast to usual clustering tasks, however,
+the time windows need to take the sequence of operations in a job into account,
+and the number of operations per cluster should also be balanced for partitioning
+a given problem instance into time windows of roughly similar solving complexity.
+We have thus devised a novel constrained clustering method that takes both concerns
+into account for gathering operations with similar feature values in a consistent and
+balanced way.
+Our experiments show that the decomposition into time windows obtained by clustering
+usually leads to significantly improved outcomes of multi-shot ASP solving in comparison
+to taking earliest starting times as simple (and consistent) partitioning criterion.
+While no single feature set for clustering turns out to be strictly more successful than greedy search
+with FIFO or MTWR allocation policy, schedules of comparable or better quality can be found
+with some feature set in the given solving time for most instances.
+Moreover, the iterative combinatorial optimization by time windows helps to increase
+the robustness of schedules obtained for different kinds of instances.
+
+While search for optimal solutions is beyond reach due to the size and complexity of
+practical scheduling applications,
+reliably computing high-quality approximations is an important subject of future work.
+This includes automatic methods for deciding about the features to consider for
+partitioning operations into clusters, i.e., determining whether an eligible feature is
+informative or noisy, respectively.
+As we can hardly expect that even moderately sized time windows can be solved to
+optimality in a short amount of time or eventual optimization criteria values be
+accurately predicted from solutions for sub-problems,
+it can still be worthwhile to incorporate measures allowing for corrections of 
+partitioning decisions, which can consist of overlapping time windows to enable
+revisions or dynamically relaunching the clustering algorithm wrt.\ computed
+partial schedules.
+Instead of obtaining features by greedy search methods only, we may also take
+the outcomes of one or several multi-shot optimization runs into account as
+features for another clustering and solving round until a fixpoint without
+further improvement is reached.
+That is, there are various opportunities to extend our approach in the future,
+where going beyond the classical Job-shop Scheduling Problem, e.g., by
+considering flexible machine allocations or partially ordered operations,
+can also be relevant for practical application scenarios.
+
+% The first issue in this work is to explore the knowledge in good schedules and try to use this knowledge to get more efficient schedules.
 We extracted some features from the problem at hand and other features from schedules obtained by heuristics.

### 2021-11-30 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
small rewordings in conclusions

@@
 it can still be worthwhile to incorporate measures allowing for corrections of 
-partitioning decisions, which can consist of overlapping time windows to enable
+partitioning decisions, which could consist of overlapping time windows to enable
 revisions or dynamically relaunching the clustering algorithm wrt.\ computed
@@
 where going beyond the classical Job-shop Scheduling Problem, e.g., by
-considering flexible machine allocations or partially ordered operations,
-can also be relevant for practical application scenarios.
+considering flexible resource allocations or partially ordered operations,
+can also be of interest for practical application scenarios.
 

### 2021-11-30 Artificial Intelligence and Machine Learning/PADL_conference/bbl.tex
pulled references inline

\begin{thebibliography}{10}
\providecommand{\url}[1]{\texttt{#1}}
\providecommand{\urlprefix}{URL }
\providecommand{\doi}[1]{https://doi.org/#1}

\bibitem{abgemuscwo15c}
Abseher, M., Gebser, M., Musliu, N., Schaub, T., Woltran, S.: Shift design with
  answer set programming. Fundamenta Informaticae  \textbf{147}(1),  1--25
  (2016)

\bibitem{adibi2014clustering}
Adibi, M., Shahrabi, J.: A clustering-based modified variable neighborhood
  search algorithm for a dynamic job shop scheduling problem. International
  Journal of Advanced Manufacturing Technology  \textbf{70}(9-12),  1955--1961
  (2014)

\bibitem{al2017job}
Al-Ashhab, M., Munshi, S., Oreijah, M., Ghulman, H.: Job shop scheduling using
  mixed integer programming. International Journal Of Modern Engineering
  Research  \textbf{7}(3),  2:23--2:29 (2017)

\bibitem{baker1974introduction}
Baker, K.: Introduction to Sequencing and Scheduling. John Wiley \& Sons (1974)

\bibitem{balduccini11a}
Balduccini, M.: Industrial-size scheduling with {ASP}+{CP}. In: Proceedings of
  the Eleventh International Conference on Logic Programming and Nonmonotonic
  Reasoning (LPNMR'11), pp. 284--296. Springer-Verlag (2011)

\bibitem{bainkaokscsotawa18a}
Banbara, M., Inoue, K., Kaufmann, B., Okimoto, T., Schaub, T., Soh, T., Tamura,
  N., Wanko, P.: teaspoon: Solving the curriculum-based course timetabling
  problems with answer set programming. Annals of Operations Research
  \textbf{275}(1),  3--37 (2019)

\bibitem{bengio2020machine}
Bengio, Y., Lodi, A., Prouvost, A.: Machine learning for combinatorial
  optimization: A methodological tour d'horizon. European Journal of
  Operational Research  \textbf{290}(2),  405--421 (2021)

\bibitem{blackstone1982state}
Blackstone, J., Phillips, D., Hogg, G.: A state-of-the-art survey of
  dispatching rules for manufacturing job shop operations. International
  Journal of Production Research  \textbf{20}(1),  27--45 (1982)

\bibitem{DBLP:journals/jim/CalisB15}
{\c{C}}alis, B., Bulkan, S.: A research survey: Review of {AI} solution
  strategies of job shop scheduling problem. Journal of Intelligent
  Manufacturing  \textbf{26}(5),  961--973 (2015)

\bibitem{chaudhry2016research}
Chaudhry, I., Khan, A.: A research survey: Review of flexible job shop
  scheduling techniques. International Transactions in Operational Research
  \textbf{23}(3),  551--591 (2016)

\bibitem{coltep19a}
{Da Col}, G., Teppan, E.: Industrial size job shop scheduling tackled by
  present day {CP} solvers. In: Proceedings of the Twenty-fifth International
  Conference on Principles and Practice of Constraint Programming (CP'19), pp.
  144--160. Springer-Verlag (2019)

\bibitem{demirbilek2019dynamically}
Demirbilek, M., Branke, J., Strauss, A.: Dynamically accepting and scheduling
  patients for home healthcare. Health Care Management Science  \textbf{22}(1),
   140--155 (2019)

\bibitem{demirkol1998benchmarks}
Demirkol, E., Mehta, S., Uzsoy, R.: Benchmarks for shop scheduling problems.
  European Journal of Operational Research  \textbf{109}(1),  137--141 (1998)

\bibitem{ding2020unified}
Ding, H., Xu, J.: A unified framework for clustering constrained data without
  locality property. Algorithmica  \textbf{82}(4),  808--852 (2020)

\bibitem{dogagrmamopo21a}
Dodaro, C., Galat{\`{a}}, G., Grioni, A., Maratea, M., Mochi, M., Porro, I.: An
  {ASP}-based solution to the chemotherapy treatment scheduling problem. Theory
  and Practice of Logic Programming  \textbf{First
  View}(\url{https://doi.org/10.1017/S1471068421000363}),  1--17 (2021)

\bibitem{el2020job}
El-Kholany, M., Gebser, M.: Job shop scheduling with multi-shot {ASP}. In:
  Proceedings of the Fourth Workshop on Trends and Applications of Answer Set
  Programming (TAASP'20).
  \url{http://www.kr.tuwien.ac.at/events/taasp20/papers/TAASP_2020_paper_4.pdf}
  (2020)

\bibitem{Forgy1965ClusterAO}
Forgy, E.: Cluster analysis of multivariate data: Efficiency versus
  interpretability of classifications. Biometrics  \textbf{21},  768--769
  (1965)

\bibitem{frscel21a}
Francescutto, G., Schekotihin, K., El{-}Kholany, M.: Solving a multi-resource
  partial-ordering flexible variant of the job-shop scheduling problem with
  hybrid {ASP}. In: Proceedings of the Seventeenth European Conference on
  Logics in Artificial Intelligence (JELIA'21), pp. 313--328. Springer-Verlag
  (2021)

\bibitem{french82a}
French, S.: Sequencing and Scheduling: An Introduction to the Mathematics of
  the Job-shop. John Wiley \& Sons (1982)

\bibitem{garey1976complexity}
Garey, M., Johnson, D., Sethi, R.: The complexity of flowshop and jobshop
  scheduling. Mathematics of Operations Research  \textbf{1}(2),  117--129
  (1976)

\bibitem{gekakasc17a}
Gebser, M., Kaminski, R., Kaufmann, B., Schaub, T.: Multi-shot {ASP} solving
  with clingo. Theory and Practice of Logic Programming  \textbf{19}(1),
  27--82 (2019)

\bibitem{harrath2002genetic}
Harrath, Y., Chebel-Morello, B., Zerhouni, N.: A genetic algorithm and data
  mining based meta-heuristic for job shop scheduling problem. In: Proceedings
  of the IEEE International Conference on Systems, Man and Cybernetics
  (SMC'02). IEEE (2002)

\bibitem{ismail2012production}
Ismail, R., Othman, Z., Bakar, A.: A production schedule generator framework
  for pattern sequential mining. In: Proceedings of the Seventh International
  Conference on Computing and Convergence Technology (ICCCT'12), pp. 784--788.
  IEEE (2012)

\bibitem{janakbhai2021blockchain}
Janakbhai, N., Saurin, M., Patel, M.: Blockchain-based intelligent
  transportation system with priority scheduling. In: Data Science and
  Intelligent Applications, pp. 311--317. Springer-Verlag (2021)

\bibitem{jones1998survey}
Jones, A., Rabelo, L., Sharawi, A.: Survey of job shop scheduling techniques.
  National Institute of Standards and Technology  \textbf{Encyclopedia of
  Electrical and Electronics
  Engineering}(\url{https://tsapps.nist.gov/publication/get_pdf.cfm?pub_id=821200})
  (1998)

\bibitem{koonce2000using}
Koonce, D., Tsai, S.: Using data mining to find patterns in genetic algorithm
  solutions to a job shop schedule. Computers \& Industrial Engineering
  \textbf{38}(3),  361--374 (2000)

\bibitem{lenstra1979computational}
Lenstra, J., Rinnooy~Kan, A.: Computational complexity of discrete optimization
  problems. Annals of Discrete Mathematics  \textbf{4},  121--140 (1979)

\bibitem{meng2020mixed}
Meng, L., Zhang, C., Ren, Y., Zhang, B., Lv, C.: Mixed-integer linear
  programming and constraint programming formulations for solving distributed
  flexible job shop scheduling problem. Computers \& Industrial Engineering
  \textbf{142},  Article 106347 (2020)

\bibitem{nasiri2019data}
Nasiri, M., Salesi, S., Rahbari, A., Meydani, N., Abdollai, M.: A data mining
  approach for population-based methods to solve the {JSSP}. Soft Computing
  \textbf{23}(21),  11107--11122 (2019)

\bibitem{nouiri2018effective}
Nouiri, M., Bekrar, A., Jemai, A., Niar, S., Ammari, A.: An effective and
  distributed particle swarm optimization algorithm for flexible job-shop
  scheduling problem. Journal of Intelligent Manufacturing  \textbf{29}(3),
  603--615 (2018)

\bibitem{ovacik2012decomposition}
Ovacik, I., Uzsoy, R.: Decomposition Methods for Complex Factory Scheduling
  Problems. Kluwer Academic Publishers (1997)

\bibitem{pezzella2008genetic}
Pezzella, F., Morganti, G., Ciaschetti, G.: A genetic algorithm for the
  flexible job-shop scheduling problem. Computers \& Operations Research
  \textbf{35}(10),  3202--3212 (2008)

\bibitem{rigralmaliiile12a}
Ricca, F., Grasso, G., Alviano, M., Manna, M., Lio, V., Iiritano, S., Leone,
  N.: Team-building with answer set programming in the {G}ioia-{T}auro seaport.
  Theory and Practice of Logic Programming  \textbf{12}(3),  361--381 (2012)

\bibitem{schoenfelder2020nurse}
Schoenfelder, J., Bretthauer, K., Wright, D., Coe, E.: Nurse scheduling with
  quick-response methods: Improving hospital performance, nurse workload, and
  patient experience. European Journal of Operational Research
  \textbf{283}(1),  390--403 (2020)

\bibitem{shahzad2010discovering}
Shahzad, A., Mebarki, N.: Discovering dispatching rules for job shop scheduling
  problem through data mining. In: Proceedings of the Eighth International
  Conference of Modeling and Simulation (MOSIM'10).
  \url{https://www.academia.edu/3068769/DISCOVERING_DISPATCHING_RULES_FOR_JOB_SHOP_SCHEDULING_PROBLEM_THROUGH_DATA_MINING}
  (2010)

\bibitem{singer2001decomposition}
Singer, M.: Decomposition methods for large job shops. Computers \& Operations
  Research  \textbf{28}(3),  193--207 (2001)

\bibitem{SOTSKOV1995237}
Sotskov, Y., Shakhlevich, N.: {NP}-hardness of shop-scheduling problems with
  three jobs. Discrete Applied Mathematics  \textbf{59}(3),  237--266 (1995)

\bibitem{taillard1993benchmarks}
Taillard, E.: Benchmarks for basic scheduling problems. European Journal of
  Operational Research  \textbf{64}(2),  278--285 (1993)

\bibitem{tagesc21a}
Tassel, P., Gebser, M., Schekotihin, K.: A reinforcement learning environment
  for job-shop scheduling. In: Proceedings of the ICAPS 2021 Workshop on
  Planning and Reinforcement Learning (PRL'21).
  \url{https://prl-theworkshop.github.io/prl2021/papers/PRL2021_paper_9.pdf}
  (2021)

\bibitem{uzsoy2000performance}
Uzsoy, R., Wang, C.: Performance of decomposition procedures for job shop
  scheduling problems with bottleneck machines. International Journal of
  Production Research  \textbf{38}(6),  1271--1286 (2000)

\bibitem{DBLP:journals/informs/VaessensAL96}
Vaessens, R., Aarts, E., Lenstra, J.: Job shop scheduling by local search.
  INFORMS Journal on Computing  \textbf{8}(3),  302--317 (1996)

\bibitem{wagstaff2001constrained}
Wagstaff, K., Cardie, C., Rogers, S., Schr{\"{o}}dl, S.: Constrained k-means
  clustering with background knowledge. In: Proceedings of the Eighteenth
  International Conference on Machine Learning (ICML'01), pp. 577--584. Morgan
  Kaufmann (2001)

\bibitem{wang2019routing}
Wang, H.: Routing and scheduling for a last-mile transportation system.
  Transportation Science  \textbf{53}(1),  131--147 (2019)

\bibitem{zhai2014decomposition}
Zhai, Y., Liu, C., Chu, W., Guo, R., Liu, C.: A decomposition heuristics based
  on multi-bottleneck machines for large-scale job shop scheduling problems.
  Journal of Industrial Engineering and Management  \textbf{7}(5),  1397--1414
  (2014)

\bibitem{zhang2019framework}
Zhang, H., Basu, S., Davidson, I.: A framework for deep constrained clustering:
  Algorithms and advances. In: Proceedings of the Joint European Conference on
  Machine Learning and Knowledge Discovery in Databases (ECML/PKDD'19), pp.
  57--72. Springer-Verlag (2019)

\bibitem{zhang2010hybrid}
Zhang, R., Wu, C.: A hybrid approach to large-scale job shop scheduling.
  Applied Intelligence  \textbf{32}(1),  47--59 (2010)

\end{thebibliography}

### 2021-11-30 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
pulled references inline
link to repo

@@
 \maketitle              % typeset the header of the contribution
+\setcounter{footnote}{0}
 %
@@
 part of which are known to be challenging for greedy and stochastic optimization techniques,
-from Taillard's and Demirkol's benchmark suites \cite{taillard1993benchmarks,demirkol1998benchmarks} for the empirical evaluation of our approach.
+from Taillard's and Demirkol's benchmark suites \cite{taillard1993benchmarks,demirkol1998benchmarks} for the empirical evaluation of our approach.%
+\footnote{The benchmarks and our implementation are available at: \url{https://github.com/Sa3doun13/PADL-2022}}
 %Many researchers have studied these instances and presented different approaches to solving them.
@@
 
-\bibliographystyle{splncs04}
-\bibliography{references}
+% \bibliographystyle{splncs04}
+% \bibliography{references}
+\input{bbl}
 

### 2022-01-11 ICLP Paper/samplepaper3.tex
First draft of ICLP Paper

--- 2021-10-22 Second paper/samplepaper.tex
@@
 \usepackage[inline]{enumitem}
-\lstset{numberbychapter=false,numbers=left,numberblanklines=false,basicstyle=\footnotesize\ttfamily,xleftmargin=5ex}
+\usepackage{adjustbox}
+\usepackage{array}
+\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
+\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
+\lstset{numberbychapter=false,numbers=left,numberblanklines=false,basicstyle=\footnotesize\ttfamily,xleftmargin=5ex,breaklines=false}
+\usepackage{multirow}
 % Used for displaying a sample figure. If possible, figure files should
@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
-
+    Scheduling plays a significant role in the success of any manufacturing system where the tasks are assigned to a limited number of resources to be scheduled.
+Job-shop Scheduling Problem (JSP) is one of the most challenging problems in production systems.
+JSP deals with assigning the tasks to a set of machines and determining the sequence of the tasks on each machine while completing them as early as possible.
+This paper applied a decomposition approach that splits the problem into a pre-defined set of small problems called Time Windows (TWs) and solves each separately in sequential order.
+Our proposed model is split into two main phases for solving the (JSP).
+The decomposition phase assigns each operation to a particular TW while satisfying the precedence constraint.
+The second phase is the scheduling phase, in which the model starts to optimize the first TW and continues until finishing the last TW.
+We built our model using Answer Set Programming (ASP).
+A Multi-shot solving paradigm is used to deal with the decomposition approach and differenc logic to handle the constraints.
+To evaluate the proposed work, we tested our model on a set of benchmark instances with different sizes and the results showed that the multi-shot solving is better than the single-shot when the number of operations increases.
 \keywords{Job-shop Scheduling Problem  \and Answer Set Programming \and Decomposition.}
@@
 The time needed to finish an operation is represented by the predicate \lstinline{pro/3} in which the operation is determined by the first two terms and the third represents the processing time in lines \ref{prg:facts:pro:begin}-\ref{prg:facts:pro:end}.
-For the machine assignment, the predicate \lstinline{assign/3} shows that a particular operation is executed by a machine in lines \ref{prg:facts:mach:begin}-\ref{prg:facts:mach:end} For instance, \emph{operation(1,1)} is processed by machine 2.
+For the machine assignment, the predicate \lstinline{assign/3} shows that a particular operation is executed by a machine in lines \ref{prg:facts:mach:begin}-\ref{prg:facts:mach:end}.
+For instance, \emph{operation(1,1)} is processed by machine 2.
 \lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:facts,caption={Problem instance}, linerange={1-11} ]{listing/facts.lp}
@@
 This strategy is encoded in lines \ref{prg:assignment_process:rank:begin}-\ref{prg:assignment_process:rank:end}.
-The fourth rule in lines \ref{prg:assignment_process:assign:begin}-\ref{prg:assignment_process:assign:end} is to assign each operation to a time winodw.
+The fourth rule in lines \ref{prg:assignment_process:assign:begin}-\ref{prg:assignment_process:assign:end} is to assign each operation to a time window.
 
@@
 
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:base,caption={base-prog}, linerange={1-12} ]{listing/base.lp}
+\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:base,caption={base-prog} ]{listing/base.lp}
 
@@
 
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:sub,caption={sub-prog}, linerange={1-40} ]{listing/sub_pro.lp}
+% \lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:sub,caption={sub-prog}, linerange={1-40} ]{listing/sub_pro.lp}
 
@@
 The fifth rule ensures that the ending time of all operations will not exceed the bound variable, which is the makespan in our case.
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:diff,caption={diff-log}, linerange={1-20} ]{listing/diff_log.lp}
+\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:diff,caption={diff-log}]{listing/diff_log.lp}
 
-\subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is
-not indented. The first paragraph that follows a table, figure,
-equation, etc., do not need, either.
+\section{Results}
+This section will show the computational experiments that we conducted to test the performance of our proposed model.
+In order to evaluate our model, we performed several experiments on a set of benchmark instances with a different number of jobs and machines.
+The benchmark instances we considered are generated in 1993 by E.
+Taillard \cite{taillard1993benchmarks}, and the size of the problems variate starting from $6$ jobs and $6$ machines till $100$ jobs and $20$ machines.
+We have decided to focus on $3$ different sizes of the problem, which are $50$ jobs, $15$ machines, $50$ jobs and $20$ machines and $100$ jobs and $20$ machines.
+Our experiments had been conducted on a laptop with Intel(R) Core(TM) i7-8650U CPU @ 1.90GHz 2.11 GHz, 8th generation, and $16$ GB.
+We split the problem into a different number of TWs starting from $2$ - $10$ throughout each size of the instances we focused on to investigate the most appropriate number of TWs.
+We run the model with a fixed amount of time $1000$ seconds.
+We set this amount because we observed that increasing the time more did not provide better results, and when we reduced it, the results became worse.
+The next table shows the obtained results with solving $50 \times 15$.
 
-Subsequent paragraphs, however, are indented.
+\begin{table}[ht]
+	\fontsize{19pt}{30pt}
+	\caption{Results for JSP $50 \times 15$ instances with different TWs} \label{tab: Table01}
+	\centering
+	%\resizebox{\columnwidth}{!}{%
+	\rotatebox{90}{
+	\begin{adjustbox}{width=1.7\textwidth}
+	\begin{tabular}{|M{2.0cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|}
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
+	%\begin{tabular}{l c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c}
+		\hline
+		\multirow{3}*{\textbf{{\large Instances}}}  &  \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 01}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 02}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 03}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 04}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 05}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 10}}}} \\[1cm]
 
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
+		\hline
+		& {\large Makespan} & {\Large Timeout}  & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} &  {\large Makespan} & {\Large Timeout} & {\large Interr-calls}\\[1cm]
 
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
+		{\LARGE TA51} & $3594.0$ & $1000.0$ & $1.0$ & $\mathbf{3442.0}$ & $1000.0$ & $2.0$ & $3843.0$ & $1000.0$ & $3.0$ & $4049.0$ & $1000.0$ & $4.0$ & $3965.0$ & $994.5$ & $4.0$ & $4269.0$ & $287.7$ & $2.0$\\[1cm]
+
+		{\LARGE TA52} & $3638.0$ & $1000.0$ & $1.0$ & $3750.0$ & $1000.0$ & $2.0$ & $\mathbf{3613.0}$ & $1000.0$ & $3.0$ & $3850.0$ & $1000.0$ & $4.0$ & $4022.0$ & $1000.0$ & $5.0$ & $4256.0$ & $241.1$ & $1.0$\\[1cm]
+
+		{\LARGE TA53} & $\mathbf{3388.0}$ & $1000.0$ & $1.0$ & $3497.0$ & $1000.0$ & $2.0$ & $3566.0$ & $1000.0$ & $3.0$ & $3618.0$ & $1000.0$ & $4.0$ & $3803.0$ & $1000.0$ & $5.0$ & $4065.0$ & $206.6$ & $2.0$\\[1cm]
+
+		{\LARGE TA54} & $3434.0$ & $1000.0$ & $1.0$ & $3489.0$ & $1000.0$ & $2.0$ & $\mathbf{3433.0}$ & $1000.0$ & $3.0$ & $3766.0$ & $1000.0$ & $4.0$ & $3732.0$ & $1000.0$ & $5.0$ & $4079.0$ & $241.5$ & $2.0$\\[1cm]
+
+		{\LARGE TA55} & $3599.0$ & $1000.0$ & $1.0$ & $\mathbf{3439.0}$  & $1000.0$ & $2.0$ & $3745.0$ & $1000.0$ & $3.0$ & $3878.0$ & $1000.0$ & $4.0$ & $4046.0$ & $1000.0$ & $5.0$ & $4336.0$ & $424.2$ & $4.0$\\	[1cm]
+
+		{\LARGE TA56} & $\mathbf{3465.0}$ & $1000.0$ & $1.0$ & $3471.0$ & $1000.0$ & $2.0$ & $3583.0$ & $1000.0$ & $3.0$ & $4117.0$ & $1000.0$ & $4.0$ & $4069.0$ & $1000.0$ & $5.0$ & $4147.0$ & $212.0$ & $2.0$\\[1cm]
+
+		{\LARGE TA57} & $3695.0$ & $1000.0$ & $1.0$ & $3779.0$ & $1000.0$ & $2.0$ & $\mathbf{3665.0}$ & $1000.0$ & $3.0$ & $4199.0$ & $1000.0$ & $4.0$ & $4096.0$ & $1000.0$ & $5.0$ & $3960.0$ & $208.8$ & $2.0$\\[1cm]
+
+		{\LARGE TA58} & $\mathbf{3642.0}$ & $1000.0$ & $1.0$ & $3777.0$ & $1000.0$ & $2.0$ & $3787.0$ & $1000.0$ & $3.0$ & $3765.0$ & $1000.0$ & $4.0$ & $3826.0$ & $810.3$ & $4.0$ & $4109.0$ & $137.9$ & $1.0$\\[1cm]
+
+		{\LARGE TA59} & $3497.0$ & $1000.0$ & $1.0$ & $\mathbf{3385.0}$ & $1000.0$ & $2.0$ & $3642.0$ & $1000.0$ & $3.0$ & $4032.0$ & $1000.0$ & $4.0$ & $3909.0$ & $1000.0$ & $5.0$ & $4227.0$ & $404.1$ & $3.0$\\[1cm]
+
+		{\LARGE TA60} & $3469.0$ & $1000.0$ & $1.0$ & $\mathbf{3219.0}$ & $1000.0$ & $2.0$ & $3338.0$ & $1000.0$ & $3.0$ & $3426.0$ & $1000.0$ & $4.0$ & $3622.0$ & $1000.0$ & $5.0$ & $3825.0$ & $32.7$ & $0.0$\\ [1cm]
         \hline
-Heading level &  Example & Font size and style\\
+
+		\multirow{3}{*}{{\large AVG}} & \multirow{3}{*}{$3542.1$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$1.0$} & \multirow{3}{*}{$\mathbf{3524.8}$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$2.0$} & \multirow{3}{*}{$3621.5$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$3.0$} & \multirow{3}{*}{$3870.0$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$4.0$} & \multirow{3}{*}{$3908.1$} & \multirow{3}{*}{$978.7$} & \multirow{3}{*}{$4.8$} &  \multirow{3}{*}{$4127.3$} & \multirow{3}{*}{$239.7$} & \multirow{3}{*}{$1.9$}\\ [1cm]
+
 		\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
+
 	\end{tabular}
+    \end{adjustbox}
+	}
+
 \end{table}
 
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
+\begin{table}[ht]
+	\fontsize{19pt}{30pt}
+	\caption{Results for JSP $50 \times 20$ instances with different TWs} \label{tab: Table02}
+	\centering
+	%\resizebox{\columnwidth}{!}{%
+	\rotatebox{90}{
+	\begin{adjustbox}{width=1.7\textwidth}
+    \begin{tabular}{|M{2.0cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|}
 
-\begin{figure}
-\includegraphics[width=\textwidth]{fig1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig1}
-\end{figure}
+	%\begin{tabular}{l c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c}
+		\hline
+		\multirow{3}*{\textbf{{\large Instances}}}  &  \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 01}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 02}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 03}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 04}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 05}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 10}}}} \\[1cm]
 
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
+		\hline
+		& {\large Makespan} & {\Large Timeout}  & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} &  {\large Makespan} & {\Large Timeout} & {\large Interr-calls}\\[1cm]
+
+		{\LARGE TA61} & $3919.0$ & $1000.0$ & $1.0$ & $\mathbf{3537.0}$ & $1000.0$ & $2.0$ & $3604.0$ & $1000.0$ & $3.0$ & $4231.0$ & $1000.0$ & $4.0$ & $4051.0$ & $1000.0$ & $5.0$ & $4463.0$ & $253.0$ & $2.0$\\[1cm]
+
+		{\LARGE TA62} & $3982.0$ & $1000.0$ & $1.0$ & $3610.0$ & $1000.0$ & $2.0$ & $\mathbf{3602.0}$ & $1000.0$ & $3.0$ & $3848.0$ & $1000.0$ & $4.0$ & $3955.0$ & $925.4$ & $4.0$  & $4445.0$ & $326.9$ & $3.0$\\[1cm]
+
+		{\LARGE TA63} & $3746.0$ & $1000.0$ & $1.0$ & $\mathbf{3285.0}$ & $1000.0$ & $2.0$ & $3410.0$ & $1000.0$ & $3.0$ & $3654.0$ & $1000.0$ & $4.0$ & $4295.0$ & $1000.0$ & $5.0$ & $4717.0$ & $311.3$ & $3.0$\\[1cm]
+
+		{\LARGE TA64} & $3610.0$ & $1000.0$ & $1.0$ & $\mathbf{3362.0}$ & $1000.0$ & $2.0$ & $3463.0$ & $1000.0$ & $3.0$ & $3670.0$ & $1000.0$ & $4.0$ & $3929.0$ & $821.3$ & $4.0$ & $4111.0$ & $357.4$ & $2.0$\\[1cm]
+
+		{\LARGE TA65} & $3781.0$ & $1000.0$ & $1.0$ & $\mathbf{3434.0}$ & $1000.0$ & $2.0$ & $3550.0$ & $1000.0$ & $3.0$ & $3639.0$ & $1000.0$ & $4.0$ & $3805.0$ & $1000.0$ & $5.0$ & $4406.0$ & $81.6$ & $0.0$\\	[1cm]
+
+		{\LARGE TA66} & $3895.0$ & $1000.0$ & $1.0$ & $3582.0$ & $1000.0$ & $2.0$ & $\mathbf{3504.0}$ & $1000.0$ & $3.0$ & $4082.0$ & $1000.0$ & $4.0$ & $4077.0$ & $1000.0$ & $5.0$ & $4210.0$ & $249.2$ & $2.0$\\[1cm]
+
+		{\LARGE TA67} & $3866.0$ & $1000.0$ & $1.0$ & $\mathbf{3547.0}$ & $1000.0$ & $2.0$ & $3571.0$ & $1000.0$ & $3.0$ & $4041.0$ & $1000.0$ & $4.0$ & $4115.0$ & $1000.0$ & $5.0$ & $4387.0$ & $285.5$ & $2.0$\\[1cm]
+
+		{\LARGE TA68} & $3763.0$ & $1000.0$ & $1.0$ & $\mathbf{3476.0}$ & $1000.0$ & $2.0$ & $3759.0$ & $1000.0$ & $3.0$ & $3897.0$ & $1000.0$ & $4.0$ & $3855.0$ & $1000.3$ & $5.0$ & $4063.0$ & $210.6$ & $2.0$\\[1cm]
+
+		{\LARGE TA69} & $4121.0$ & $1000.0$ & $1.0$ & $\mathbf{3748.0}$ & $1000.0$ & $2.0$ & $3803.0$ & $1000.0$ & $3.0$ & $4116.0$ & $1000.0$ & $4.0$ & $3937.0$ & $1000.0$ & $5.0$ & $4518.0$ & $111.0$ & $1.0$\\[1cm]
+
+		{\LARGE TA70} & $4111.0$ & $1000.0$ & $1.0$ & $3827.0$ & $1000.0$ & $2.0$ & $\mathbf{3791.0}$ & $1000.0$ & $3.0$ & $4008.0$ & $1000.0$ & $4.0$ & $4044.0$ & $1000.0$ & $5.0$ & $4651.0$ & $365.2$ & $3.0$\\ [1cm]
+        \hline
+
+		\multirow{3}{*}{{\LARGE AVG}} & \multirow{3}{*}{$3879.4$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$1.0$} & \multirow{3}{*}{$\mathbf{3540.8}$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$2.0$} & \multirow{3}{*}{$3605.7$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$3.0$} & \multirow{3}{*}{$3918.6$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$4.0$} & \multirow{3}{*}{$4006.3$} & \multirow{3}{*}{$974.3$} & \multirow{3}{*}{$4.8$} & \multirow{3}{*}{$4397.1$} & \multirow{3}{*}{$245.2$} & \multirow{3}{*}{$2.0$}\\ [1cm]
+
+		\hline
+
+	\end{tabular}
+    \end{adjustbox}
+	}
+
+\end{table}
+
+\begin{table}[ht]
+	\fontsize{19pt}{30pt}
+	\caption{Results for JSP $100 \times 20$ instances with different TWs} \label{tab: Table03}
+	\centering
+	%\resizebox{\columnwidth}{!}{%
+	\rotatebox{90}{
+	\begin{adjustbox}{width=1.7\textwidth}
+		%\begin{tabular}[m]{|m{2cm}|*{17}{c}|c|}
+    \begin{tabular}{|M{2.0cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|}
+
+        %\begin{tabular}{l c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c}
+        \hline
+        \multirow{3}*{\textbf{{\large Instances}}}  &  \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 01}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 02}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 03}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 04}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 05}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 10}}}} \\[1cm]
+
+        \hline
+        & {\large Makespan} & {\Large Timeout}  & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} &  {\large Makespan} & {\Large Timeout} & {\large Interr-calls}\\[1cm]
+
+		{\LARGE TA71} & $52147.0$ & $1000.0$ & $1.0$ & $21345.0$ & $1000.0$ & $2.0$ & $7641.0$ & $1000.0$ & $3.0$ & $7471.0$ & $1000.0$ & $4.0$ & $\mathbf{7226.0}$ & $1000.0$ & $5.0$ & $8478.0$ & $1000.0$ & $10.0$\\[1cm]
+
+		{\LARGE TA72} & $45196.0$ & $1000.0$ & $1.0$ & $23505.0$ & $1000.0$ & $2.0$ & $\mathbf{6750.0}$ & $1000.0$ & $3.0$ & $7134.0$ & $1000.0$ & $4.0$ & $7300.0$ & $1000.0$ & $5.0$ & $8164.0$ & $1000.0$ & $10.0$\\[1cm]
+
+		{\LARGE TA73} & $36548.0$ & $1000.0$ & $1.0$ & $24995.0$ & $1000.0$ & $2.0$ & $7121.0$ & $1000.0$ & $3.0$ & $\mathbf{7027.0}$ & $1000.0$ & $4.0$ & $7416.0$ & $1000.0$ & $5.0$ & $7975.0$ & $1000.0$ & $10.0$\\[1cm]
+
+		{\LARGE TA74} & $52973.0$ & $1000.0$ & $1.0$ & $25230.0$ & $1000.0$ & $2.0$ & $6703.0$ & $1000.0$ & $3.0$ & $\mathbf{6634.0}$ & $1000.0$ & $4.0$ & $7062.0$ & $1000.0$ & $4.0$ & $7994.0$ & $1000.0$ & $10.0$\\[1cm]
+
+		{\LARGE TA75} & $45316.0$ & $1000.0$ & $1.0$ & $23651.0$  & $1000.0$ & $2.0$ & $\mathbf{7042.0}$ & $1000.0$ & $3.0$ & $7511.0$ & $1000.0$ & $4.0$ & $7070.0$ & $1000.0$ & $5.0$ & $7934.0$ & $1000.0$ & $10.0$\\[1cm]
+
+		{\LARGE TA76} & $53532.0$ & $1000.0$ & $1.0$ & $40944.0$ & $1000.0$ & $2.0$ & $7067.0$ & $1000.0$ & $3.0$ & $\mathbf{6879.0}$ & $1000.0$ & $4.0$ & $7061.0$ & $1000.0$ & $5.0$ & $7938.0$ & $1000.0$ & $10.0$\\[1cm]
+
+		{\LARGE TA77} & $45003.0$ & $1000.0$ & $1.0$ & $24517.0$ & $1000.0$ & $2.0$ & $6947.0$ & $1000.0$ & $3.0$ & $\mathbf{6929.0}$ & $1000.0$ & $4.0$ & $7307.0$ & $1000.0$ & $5.0$ & $7773.0$ & $1000.0$ & $10.0$\\[1cm]
+
+		{\LARGE TA78} & $41380.0$ & $1000.0$ & $1.0$ & $24686.0$ & $1000.0$ & $2.0$ & $6917.0$ & $1000.0$ & $3.0$ & $\mathbf{6793.0}$ & $1000.0$ & $4.0$ & $7215.0$ & $1000.0$ & $5.0$ & $7986.0$ & $1000.0$ & $10.0$\\[1cm]
+
+		{\LARGE TA79} & $48827.0$ & $1000.0$ & $1.0$ & $24401.0$ & $1000.0$ & $2.0$ & $6913.0$ & $1000.0$ & $3.0$ & $\mathbf{6621.0}$ & $1000.0$ & $4.0$ & $6645.0$ & $1000.0$ & $5.0$ & $7405.0$ & $1000.0$ & $10.0$\\[1cm]
+
+		{\LARGE TA80} & $46939.0$ & $1000.0$ & $1.0$ & $22833.0$ & $1000.0$ & $2.0$ & $6925.0$ & $1000.0$ & $3.0$ & $\mathbf{6645.0}$ & $1000.0$ & $4.0$ & $6647.0$ & $1000.0$ & $5.0$ & $7609.0$ & $1000.0$ & $10.0$\\ [1cm]
+        \hline
+
+		\multirow{3}{*}{{\LARGE AVG}} & \multirow{3}{*}{$46786.1$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$1.0$} & \multirow{3}{*}{$25610.7$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$2.0$} & \multirow{3}{*}{$7002.6$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$3.0$} & \multirow{3}{*}{$\mathbf{6964.4}$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$4.0$} & \multirow{3}{*}{$7094.9$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$5.0$} & \multirow{3}{*}{$7925.6$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$10.0$}\\ [1cm]
+
+		\hline
+
+	\end{tabular}
+    \end{adjustbox}
+	}
+
+\end{table}
+
 %

### 2022-01-12 Second paper/samplepaper.tex
changed the style

@@
 %
-\documentclass[runningheads]{llncs}
+\documentclass{tlp} % \documentclass[runningheads]{llncs}
 %
@@
 
+\newtheorem{theorem}{Theorem}
+
 \begin{document}
+
+\lefttitle{Mohammed M.\ S.\ El-Kholany, Martin Gebser, Konstantin Schekotihin}
+
+\jnlPage{1}{8}
+\jnlDoiYr{2021}
+\doival{10.1017/xxxxx}
+
 %
-\title{A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}}
+\title{A decomposition Approach for solving the JSP using multi-shot ASP}
+% \thanks{This work was partially funded by
+% KWF project 28472,
+% cms electronics GmbH,
+% FunderMax GmbH,
+% Hirsch Armbänder GmbH,
+% incubed IT GmbH,
+% Infineon Technologies Austria AG,
+% Isovolta AG,
+% Kostwein Holding GmbH, and
+% Privatstiftung Kärntner Sparkasse.}}
 %
@@
 %
-\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-0002-1088-2081} \and
-Martin Gebser\inst{1,2}\orcidID{0000-0002-8010-4752} \and
-Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958}}
+\begin{authgrp}
+\author{\sn{Mohammed M.\ S.} \gn{El-Kholany}} % \inst{1}\orcidID{0000-0002-1088-2081} \and
+\affiliation{University of Klagenfurt, Austria \and Cairo University, Egypt}
+\author{\sn{Martin} \gn{Gebser}} % \inst{1,2}\orcidID{0000-0002-8010-4752} \and
+\affiliation{University of Klagenfurt, Austria \and Graz University of Technology, Austria}
+\author{\sn{Konstantin} \gn{Schekotihin}} % \inst{1}\orcidID{0000-0002-0286-0958}
+\affiliation{University of Klagenfurt, Austria}
+\end{authgrp}
 %
-\authorrunning{M. El-Kholany et al.}
+%%% \authorrunning{M. El-Kholany et al.}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
-\email{\{mohammed.el-kholany, martin.gebser and konstantin.schekotihin\}@aau.at}\\
+%%% \institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
+%%% \email{\{mohammed.el-kholany, martin.gebser and konstantin.schekotihin\}@aau.at}\\
 %\url{http://www.springer.com/gp/computer-science/lncs} \and
-Technische Universität Graz, Graz, Austria\\
-\email{mgebser@ist.tugraz.at}}
-%
+%%% Technische Universität Graz, Graz, Austria\\
+%%% \email{mgebser@ist.tugraz.at}}
+
+\history{\sub{xx xx xxxx;} \rev{xx xx xxxx;} \acc{xx xx xxxx}}
+
 \maketitle              % typeset the header of the contribution
-%
+
 \begin{abstract}
@@
 
-\keywords{Job-shop Scheduling Problem  \and Answer Set Programming \and Decomposition.}
+% \keywords{Job-shop Scheduling Problem  \and Answer Set Programming \and Decomposition.}
 \end{abstract}
+%
+\begin{keywords}
+Job-shop Scheduling Problem, Answer Set Programming, Decomposition
+\end{keywords}
 %
@@
 %
-\bibliographystyle{splncs04}
+\bibliographystyle{tlplike} % \bibliographystyle{splncs04}
 \bibliography{Ref}

### 2022-01-12 ICLP Paper/samplepaper.tex
adjusted paper style

--- 2022-01-11 ICLP Paper/samplepaper3.tex
@@
 \begin{document}
+
+\lefttitle{Mohammed M.\ S.\ El-Kholany, Martin Gebser, Konstantin Schekotihin}
+
+\jnlPage{1}{8}
+\jnlDoiYr{2021}
+\doival{10.1017/xxxxx}
 %
-\title{A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}}
+\title{A decomposition Approach for solving the JSP using multi-shot ASP%
+% \thanks{This work was partially funded by
+% KWF project 28472,
+% cms electronics GmbH,
+% FunderMax GmbH,
+% Hirsch Armbänder GmbH,
+% incubed IT GmbH,
+% Infineon Technologies Austria AG,
+% Isovolta AG,
+% Kostwein Holding GmbH, and
+% Privatstiftung Kärntner Sparkasse.}%
+}
 %
@@
 %
-\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-0002-1088-2081} \and
-Martin Gebser\inst{1,2}\orcidID{0000-0002-8010-4752} \and
-Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958}}
-%
-\authorrunning{M. El-Kholany et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
-\email{\{mohammed.el-kholany, martin.gebser and konstantin.schekotihin\}@aau.at}\\
-%\url{http://www.springer.com/gp/computer-science/lncs} \and
-Technische Universität Graz, Graz, Austria\\
-\email{mgebser@ist.tugraz.at}}
-%
+% \author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-0002-1088-2081} \and
+% Martin Gebser\inst{1,2}\orcidID{0000-0002-8010-4752} \and
+% Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958}}
+% %
+% \authorrunning{M. El-Kholany et al.}
+% % First names are abbreviated in the running head.
+% % If there are more than two authors, 'et al.' is used.
+% %
+% \institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
+% \email{\{mohammed.el-kholany, martin.gebser and konstantin.schekotihin\}@aau.at}\\
+% %\url{http://www.springer.com/gp/computer-science/lncs} \and
+% Technische Universität Graz, Graz, Austria\\
+% \email{mgebser@ist.tugraz.at}}
+\begin{authgrp}
+\author{\sn{Mohammed M.\ S.} \gn{El-Kholany}} % \inst{1}\orcidID{0000-0002-1088-2081} \and
+\affiliation{University of Klagenfurt, Austria \and Cairo University, Egypt}
+\author{\sn{Martin} \gn{Gebser}} % \inst{1,2}\orcidID{0000-0002-8010-4752} \and
+\affiliation{University of Klagenfurt, Austria \and Graz University of Technology, Austria}
+\author{\sn{Konstantin} \gn{Schekotihin}} % \inst{1}\orcidID{0000-0002-0286-0958}
+\affiliation{University of Klagenfurt, Austria}
+\end{authgrp}
+
+\history{\sub{xx xx xxxx;} \rev{xx xx xxxx;} \acc{xx xx xxxx}}
+
 \maketitle              % typeset the header of the contribution
@@
 To evaluate the proposed work, we tested our model on a set of benchmark instances with different sizes and the results showed that the multi-shot solving is better than the single-shot when the number of operations increases.
-\keywords{Job-shop Scheduling Problem  \and Answer Set Programming \and Decomposition.}
+% \keywords{Job-shop Scheduling Problem  \and Answer Set Programming \and Decomposition.}
 \end{abstract}
 %
+\begin{keywords}
+Job-shop Scheduling Problem, Answer Set Programming, Decomposition
+\end{keywords}
 %

### 2022-01-12 ICLP Paper/samplepaper.tex
still working on abstract

@@
 %
-\title{A decomposition Approach for solving the JSP using multi-shot ASP%
+\title{Problem Decomposition and Multi-shot ASP Solving for Job-shop Scheduling%
 % \thanks{This work was partially funded by
@@
 \begin{abstract}
-    Scheduling plays a significant role in the success of any manufacturing system where the tasks are assigned to a limited number of resources to be scheduled.
-Job-shop Scheduling Problem (JSP) is one of the most challenging problems in production systems.
-JSP deals with assigning the tasks to a set of machines and determining the sequence of the tasks on each machine while completing them as early as possible.
+Scheduling methods are important for effective production and logistics management, where tasks need to be allocated to and performed by limited resources.
+In particular, the Job-shop Scheduling Problem (JSP) is a well-known and challenging
+combinatorial optimization problem in which tasks sharing a machine are to be arranged
+in sequence such that jobs encompassing them can be completed as early as possible.
+Given that already moderately sized JSP instances can turn out as highly combinatorial,
+so that neither optimal schedules nor the runtime to termination of complete optimization methods
+is known, efficient approaches to approximate good-quality schedules are of interest.
+
 This paper applied a decomposition approach that splits the problem into a pre-defined set of small problems called Time Windows (TWs) and solves each separately in sequential order.

### 2022-01-13 ICLP Paper/samplepaper.tex
went through abstract

@@
 \begin{abstract}
-Scheduling methods are important for effective production and logistics management, where tasks need to be allocated to and performed by limited resources.
+Scheduling methods are important for effective production and logistics management, where tasks need to be allocated and performed by limited resources.
 In particular, the Job-shop Scheduling Problem (JSP) is a well-known and challenging
 combinatorial optimization problem in which tasks sharing a machine are to be arranged
-in sequence such that jobs encompassing them can be completed as early as possible.
+in sequence such that encompassing jobs can be completed as early as possible.
 Given that already moderately sized JSP instances can turn out as highly combinatorial,
@@
 is known, efficient approaches to approximate good-quality schedules are of interest.
-
-This paper applied a decomposition approach that splits the problem into a pre-defined set of small problems called Time Windows (TWs) and solves each separately in sequential order.
+In this paper, we propose problem decomposition into time windows whose operations
+can be successively scheduled and optimized by means of multi-shot ASP solving.
+From a computational perspective, decomposition aims to split highly complex scheduling tasks into better manageable subproblems with a balanced number of 
+operations, so that good-quality or even optimal partial solutions can be
+reliably found in a small fraction of runtime.
+Regarding the feasibility and quality of solutions, problem decomposition must
+respect the precedence of operations within their jobs, and partial schedules
+optimized by time windows should yield better global solutions than obtainable
+in similar runtime on the full problem.
+We devise and investigate a variety of decomposition strategies in terms of the number and size of time windows as well as heuristics for choosing their operations.
+Moreover, we incorporate time window overlapping and compression techniques
+into the iterative scheduling process in order to counteract limitations of window-wise optimization restricted to partial schedules.
+Our experiments on JSP benchmark sets of several sizes show that
+successive optimization by multi-shot ASP solving leads to substantially better schedules within the runtime limit than global optimization on the full problem,
+where the gap increases with the number of operations to schedule.
+%
+% This paper applied a decomposition approach that splits the problem into a pre-defined set of small problems called Time Windows (TWs) and solves each separately in sequential order.
 Our proposed model is split into two main phases for solving the (JSP).

### 2022-01-14 ICLP Paper/samplepaper.tex
looking at intro

@@
 \section{Introduction}
-In today's competitive markets, manufacturers have to respond quickly to orders and meet shipping dates committed to the customers.
+% In today's competitive markets, manufacturers have to respond quickly to orders and meet shipping dates committed to the customers.
 This requires the ability to schedule production activities to use the available scarce resources efficiently.
@@
 In general, scheduling operations is one of the most critical issues in the planning and managing of the manufacturing processes\cite{uzsoy2000performance}.
+Effective scheduling methods are essential for complex manufacturing and logistics systems,
+where allocating and performing diverse tasks within resource capacity limits 
+is one of the most critical challenges for production management \citep{uzsoy2000performance}.
+The Jop-shop Scheduling Problem (JSP) \citep{baker1974introduction,taillard1993benchmarks}
+constitutes a well-known mathematical abstraction of industrial production scheduling in
+which sequences of operations need to be processed by machines such that a given objective
+like the makespan for completing all jobs or their tardiness w.r.t.\ deadlines is minimized.
+Finding optimal JSP solutions, determined by a sequence of operations for each machine,
+is an NP-hard combinatorial problem
+\citep{garey1976complexity,lenstra1977complexity,liu2008prediction},
+and optimal schedules as well as termination guarantees can be extremely challenging
+or beyond reach 
+for complete optimization methods already for moderately sized instances.
+For example, it took about $20$ years to develop a search procedure able to find a
+(provably) optimal solution for an instance called FT10 with $10$ jobs
+\citep{adams1988shifting,zhang2010hybrid},
+each consisting of a sequence of $10$ operations to be processed by $10$ machines.
 
-One of the most challenging scheduling problems is the Jop-shop Scheduling Problem (JSP).
+% One of the most challenging scheduling problems is the Jop-shop Scheduling Problem (JSP).
 A set of jobs needs to be processed on a set of machines while optimizing a performance indicator, minimizing makespan, the time needed to complete all the jobs, or tardiness, i.e., the summation of the delays in executing all jobs according to their deadlines.

### 2022-01-16 ICLP Paper/samplepaper.tex
update tables

--- 2022-01-13 ICLP Paper/samplepaper.tex
@@
 
-\begin{table}[ht]
-	\fontsize{19pt}{30pt}
-	\caption{Results for JSP $50 \times 15$ instances with different TWs} \label{tab: Table01}
+\begin{table}[t]
+    \caption{Results for three sizes of job shop scheduling instances with different numbers of time windows using static machine-based EST decomposition \label{tab:Table01}}%
+%    \mbox{~}\hfill
+%    \resizebox{0.8\textwidth}{!}{%
+%    \begin{minipage}{\textwidth}
     \centering
-	%\resizebox{\columnwidth}{!}{%
-	\rotatebox{90}{
-	\begin{adjustbox}{width=1.7\textwidth}
-	\begin{tabular}{|M{2.0cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|}
+    \begin{tabular}{l l r r r r r}
+    \hline
+%       & {}        & \multicolumn{6}{c}{Group of Instances}  \\
+    Window(s) & Comparison & 50$\times$15 & 50$\times$20 & 100$\times$20 \\
 
-	%\begin{tabular}{l c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c}
+    \hline\\[-2.75mm]
+    1                & Makespan             & 3542.1   & 3879.4	     &  46786.1 \\
+                     & CPU(Sec)             & 1000.0	& 1000.0   & 1000.0	     	    \\
+                     & Interrupted Calls    & 1.0	    & 1.0      & 1.0	       \\[1.5mm]
+
+    2                & Makespan             & 3149.7   & 3479.4	      & 23977.5 \\
+                     & CPU(Sec)             & 1000.0	    & 1000.0	   & 1000.0	      \\
+                     & Interrupted Calls    & 2.0	    & 2.0      & 2.0	      \\[1.5mm]
+
+    3                & Makespan             & \textbf{3083.8}   & 3290.5	    & 7201.5 \\
+                     & CPU(Sec)             & 1000.0	    & 1000.0	   & 1000.0	      \\
+                     & Interrupted Calls    & 3.0	    & 3.0      & 3.0	      \\[1.5mm]
+
+    4                & Makespan             & 3110.7   & \textbf{3260.3}	   & 6288.7 \\
+                     & CPU(Sec)             & 934.4	    & 944.9	   & 1000.0	      \\
+                     & Interrupted Calls    & 3.6	    & 3.5      & 4.0	      \\[1.5mm]
+
+    5                & Makespan             & 3213.6   & 3355.5    & 6106.1 \\
+                     & CPU(Sec)             & 615.5	    & 583.4	   & 1000.0	      \\
+                     & Interrupted Calls    & 2.7	    & 2.5      & 5.0	      \\[1.5mm]
+
+    6                & Makespan             & 3225.3   & 3414.8	      & \textbf{6031.9} \\
+                     & CPU(Sec)             & 243.8	    & 321.5	   & 1000.0	      \\
+                     & Interrupted Calls    & 1.0	    & 1.6      & 6.0	      \\[1.5mm]
+
+    7                & Makespan             & 3350.5   & 3540.3	      & 6094.1 \\
+                     & CPU(Sec)             & 146.7	    & 161.4	   & 1000.0	      \\
+                     & Interrupted Calls    & 0.6	    & 0.9      & 7.0	      \\[1.5mm]
+
+    8                & Makespan             & 3378.5   & 3561.5	      & 6053.1 \\
+                     & CPU(Sec)             & 81.0	    & 54.9	   & 1000.0	      \\
+                     & Interrupted Calls    & 0.3	    & 0.2      & 8.0	      \\[1.5mm]
+
+    9                & Makespan             & 3495.0   & 3720.7	      & 6095.9 \\
+                     & CPU(Sec)             & 45.6	    & 38.0	   & 977.0	      \\
+                     & Interrupted Calls    & 0.1	    & 0.2      & 8.2	      \\[1.5mm]
+
+    10               & Makespan             & 3524.9   & 3793.1	      & 6234.0 \\
+                     & CPU(Sec)             & 18.5	    & 33.3	   & 861.8	      \\
+                     & Interrupted Calls    & 0.1	    & 0.1      & 7.9	      \\[1.5mm]
+
+    \\[-2.75mm]
     \hline
-		\multirow{3}*{\textbf{{\large Instances}}}  &  \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 01}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 02}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 03}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 04}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 05}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 10}}}} \\[1cm]
+    \end{tabular}
+   % }
+%    \end{minipage}
+% }
+%    \hfill\hfill\mbox{~}
+\end{table}
 
-		\hline
-		& {\large Makespan} & {\Large Timeout}  & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} &  {\large Makespan} & {\Large Timeout} & {\large Interr-calls}\\[1cm]
-
-		{\LARGE TA51} & $3594.0$ & $1000.0$ & $1.0$ & $\mathbf{3442.0}$ & $1000.0$ & $2.0$ & $3843.0$ & $1000.0$ & $3.0$ & $4049.0$ & $1000.0$ & $4.0$ & $3965.0$ & $994.5$ & $4.0$ & $4269.0$ & $287.7$ & $2.0$\\[1cm]
-
-		{\LARGE TA52} & $3638.0$ & $1000.0$ & $1.0$ & $3750.0$ & $1000.0$ & $2.0$ & $\mathbf{3613.0}$ & $1000.0$ & $3.0$ & $3850.0$ & $1000.0$ & $4.0$ & $4022.0$ & $1000.0$ & $5.0$ & $4256.0$ & $241.1$ & $1.0$\\[1cm]
-
-		{\LARGE TA53} & $\mathbf{3388.0}$ & $1000.0$ & $1.0$ & $3497.0$ & $1000.0$ & $2.0$ & $3566.0$ & $1000.0$ & $3.0$ & $3618.0$ & $1000.0$ & $4.0$ & $3803.0$ & $1000.0$ & $5.0$ & $4065.0$ & $206.6$ & $2.0$\\[1cm]
-
-		{\LARGE TA54} & $3434.0$ & $1000.0$ & $1.0$ & $3489.0$ & $1000.0$ & $2.0$ & $\mathbf{3433.0}$ & $1000.0$ & $3.0$ & $3766.0$ & $1000.0$ & $4.0$ & $3732.0$ & $1000.0$ & $5.0$ & $4079.0$ & $241.5$ & $2.0$\\[1cm]
-
-		{\LARGE TA55} & $3599.0$ & $1000.0$ & $1.0$ & $\mathbf{3439.0}$  & $1000.0$ & $2.0$ & $3745.0$ & $1000.0$ & $3.0$ & $3878.0$ & $1000.0$ & $4.0$ & $4046.0$ & $1000.0$ & $5.0$ & $4336.0$ & $424.2$ & $4.0$\\	[1cm]
-
-		{\LARGE TA56} & $\mathbf{3465.0}$ & $1000.0$ & $1.0$ & $3471.0$ & $1000.0$ & $2.0$ & $3583.0$ & $1000.0$ & $3.0$ & $4117.0$ & $1000.0$ & $4.0$ & $4069.0$ & $1000.0$ & $5.0$ & $4147.0$ & $212.0$ & $2.0$\\[1cm]
-
-		{\LARGE TA57} & $3695.0$ & $1000.0$ & $1.0$ & $3779.0$ & $1000.0$ & $2.0$ & $\mathbf{3665.0}$ & $1000.0$ & $3.0$ & $4199.0$ & $1000.0$ & $4.0$ & $4096.0$ & $1000.0$ & $5.0$ & $3960.0$ & $208.8$ & $2.0$\\[1cm]
-
-		{\LARGE TA58} & $\mathbf{3642.0}$ & $1000.0$ & $1.0$ & $3777.0$ & $1000.0$ & $2.0$ & $3787.0$ & $1000.0$ & $3.0$ & $3765.0$ & $1000.0$ & $4.0$ & $3826.0$ & $810.3$ & $4.0$ & $4109.0$ & $137.9$ & $1.0$\\[1cm]
-
-		{\LARGE TA59} & $3497.0$ & $1000.0$ & $1.0$ & $\mathbf{3385.0}$ & $1000.0$ & $2.0$ & $3642.0$ & $1000.0$ & $3.0$ & $4032.0$ & $1000.0$ & $4.0$ & $3909.0$ & $1000.0$ & $5.0$ & $4227.0$ & $404.1$ & $3.0$\\[1cm]
-
-		{\LARGE TA60} & $3469.0$ & $1000.0$ & $1.0$ & $\mathbf{3219.0}$ & $1000.0$ & $2.0$ & $3338.0$ & $1000.0$ & $3.0$ & $3426.0$ & $1000.0$ & $4.0$ & $3622.0$ & $1000.0$ & $5.0$ & $3825.0$ & $32.7$ & $0.0$\\ [1cm]
+\begin{table}[t]
+    \caption{Comparing different static/dynamic decomposition strategies \label{tab:Table02}}%
+%    \mbox{~}\hfill
+%    \resizebox{0.8\textwidth}{!}{%
+%    \begin{minipage}{\textwidth}
+    \centering
+    \begin{tabular}{c c c c c c c c c c}
     \hline
 
-		\multirow{3}{*}{{\large AVG}} & \multirow{3}{*}{$3542.1$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$1.0$} & \multirow{3}{*}{$\mathbf{3524.8}$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$2.0$} & \multirow{3}{*}{$3621.5$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$3.0$} & \multirow{3}{*}{$3870.0$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$4.0$} & \multirow{3}{*}{$3908.1$} & \multirow{3}{*}{$978.7$} & \multirow{3}{*}{$4.8$} &  \multirow{3}{*}{$4127.3$} & \multirow{3}{*}{$239.7$} & \multirow{3}{*}{$1.9$}\\ [1cm]
+\multirow{2}*{Strategy} & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
+    & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC \\
+    \hline\\[-2.75mm]
+    Static machine-based EST              & \boldmath{$3083.8$}  & $1000.0$   & $3.0$	& \boldmath{$3260.3$}      & $944.9$ & $3.5$ & \boldmath{$6031.9$} & $1000.0$ & $6.0$\\
+    [1.5mm]
 
+    Static time-based EST             & $3524.8$ & $1000.0$   & $2.0$	   & $3540.8$      & $1000.0$ & $2.0$ & $6964.4$ & $1000.0$ & $4.0$\\ 
+    [1.5mm]
+
+    Static time-based MTWR             & $3456.2$  & $1000.0$   & $2.0$	& $3561.0$      & $1000.0$ & $2.0$ & $7149.6$ & $1000.0$ & $4.0$\\ 
+    [1.5mm]
+
+    Dynamic machine-based EST             & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $3.3$ & $6115.1$ & $1000.0$ & $6.9$\\ 
+    [1.5mm]
+
+    \\[-2.75mm]
     \hline
@@
     \end{tabular}
-    \end{adjustbox}
-	}
-
+   % }
+%    \end{minipage}
+% }
+%    \hfill\hfill\mbox{~}
 \end{table}
 
-\begin{table}[ht]
-	\fontsize{19pt}{30pt}
-	\caption{Results for JSP $50 \times 20$ instances with different TWs} \label{tab: Table02}
+\begin{table}[t]
+    \caption{Comparing results of best decomposition strategy with/without dompression and overlapping \label{tab:Table03}}%
+%    \mbox{~}\hfill
+%    \resizebox{0.8\textwidth}{!}{%
+%    \begin{minipage}{\textwidth}
     \centering
-	%\resizebox{\columnwidth}{!}{%
-	\rotatebox{90}{
-	\begin{adjustbox}{width=1.7\textwidth}
-    \begin{tabular}{|M{2.0cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|}
-
-	%\begin{tabular}{l c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c}
-		\hline
-		\multirow{3}*{\textbf{{\large Instances}}}  &  \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 01}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 02}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 03}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 04}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 05}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 10}}}} \\[1cm]
+    \begin{tabular}{c c c c c c c c c c}
 
     \hline
-		& {\large Makespan} & {\Large Timeout}  & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} &  {\large Makespan} & {\Large Timeout} & {\large Interr-calls}\\[1cm]
+    \multicolumn{10}{c}{\textbf{Without Compressing Time Windows}}\\        
+\multirow{2}*{Overlap} & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
+    & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC \\
+    \hline\\[-2.75mm]
+    0\%              & $3083.8$  & $1000.0$   & $3.0$	& $3260.3$      & $944.9$ & $3.5$ & $6031.9$ & $1000.0$ & $6.0$\\
+    [1.5mm]
 
-		{\LARGE TA61} & $3919.0$ & $1000.0$ & $1.0$ & $\mathbf{3537.0}$ & $1000.0$ & $2.0$ & $3604.0$ & $1000.0$ & $3.0$ & $4231.0$ & $1000.0$ & $4.0$ & $4051.0$ & $1000.0$ & $5.0$ & $4463.0$ & $253.0$ & $2.0$\\[1cm]
+    10\%             & $3050.0$  & $1000.0$   & $3.0$	& \boldmath{$3194.9$}     & $1000.0$ & $4.0$ & $5979.9$ & $1000.0$ & $6.0$\\
 
-		{\LARGE TA62} & $3982.0$ & $1000.0$ & $1.0$ & $3610.0$ & $1000.0$ & $2.0$ & $\mathbf{3602.0}$ & $1000.0$ & $3.0$ & $3848.0$ & $1000.0$ & $4.0$ & $3955.0$ & $925.4$ & $4.0$  & $4445.0$ & $326.9$ & $3.0$\\[1cm]
+    20\%             & \boldmath{$3046.0$} & $1000.0$   & $3.0$	   & $3200.6$      & $1000.0$ & $4.0$ & \boldmath{$5954.3$} & $1000.0$ & $6.0$\\
 
-		{\LARGE TA63} & $3746.0$ & $1000.0$ & $1.0$ & $\mathbf{3285.0}$ & $1000.0$ & $2.0$ & $3410.0$ & $1000.0$ & $3.0$ & $3654.0$ & $1000.0$ & $4.0$ & $4295.0$ & $1000.0$ & $5.0$ & $4717.0$ & $311.3$ & $3.0$\\[1cm]
+    30\%             & $3051.2$  & $1000.0$   & $3.0$	& $3451.8$      & $1000.0$ & $4.0$ & $6933.3$ & $1000.0$ & $6.0$\\ 
+                     [1.5mm]
 
-		{\LARGE TA64} & $3610.0$ & $1000.0$ & $1.0$ & $\mathbf{3362.0}$ & $1000.0$ & $2.0$ & $3463.0$ & $1000.0$ & $3.0$ & $3670.0$ & $1000.0$ & $4.0$ & $3929.0$ & $821.3$ & $4.0$ & $4111.0$ & $357.4$ & $2.0$\\[1cm]
+    \\[-2.75mm]
 
-		{\LARGE TA65} & $3781.0$ & $1000.0$ & $1.0$ & $\mathbf{3434.0}$ & $1000.0$ & $2.0$ & $3550.0$ & $1000.0$ & $3.0$ & $3639.0$ & $1000.0$ & $4.0$ & $3805.0$ & $1000.0$ & $5.0$ & $4406.0$ & $81.6$ & $0.0$\\	[1cm]
+    \multicolumn{10}{c}{\textbf{Compressing Time Windows}}\\
+    \hline
+    \multirow{2}*{Overlap} & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
+    & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC \\
+    \hline\\[-2.75mm]
+    0\%              & $3013.3$  & $572.3$   & $3.7$	   & $3186.8$      & $540.8$ & $3.3$ & $5713.0$ & $589.2$ & $8.6$\\
+    [1.5mm]
 
-		{\LARGE TA66} & $3895.0$ & $1000.0$ & $1.0$ & $3582.0$ & $1000.0$ & $2.0$ & $\mathbf{3504.0}$ & $1000.0$ & $3.0$ & $4082.0$ & $1000.0$ & $4.0$ & $4077.0$ & $1000.0$ & $5.0$ & $4210.0$ & $249.2$ & $2.0$\\[1cm]
+    10\%             & $2970.2$  & $600.0$   & $4.0$	   & $3149.8$      & $600.0$ & $4.0$ & $5656.0$ & $600.0$ & $9.0$\\
 
-		{\LARGE TA67} & $3866.0$ & $1000.0$ & $1.0$ & $\mathbf{3547.0}$ & $1000.0$ & $2.0$ & $3571.0$ & $1000.0$ & $3.0$ & $4041.0$ & $1000.0$ & $4.0$ & $4115.0$ & $1000.0$ & $5.0$ & $4387.0$ & $285.5$ & $2.0$\\[1cm]
+    20\%             & $2952.2$  & $600.0$   & $4.0$	   & \boldmath{$3147.7$}      & $600.0$ & $4.0$ & \boldmath{$5619.2$} & $600.0$ & $9.0$\\
 
-		{\LARGE TA68} & $3763.0$ & $1000.0$ & $1.0$ & $\mathbf{3476.0}$ & $1000.0$ & $2.0$ & $3759.0$ & $1000.0$ & $3.0$ & $3897.0$ & $1000.0$ & $4.0$ & $3855.0$ & $1000.3$ & $5.0$ & $4063.0$ & $210.6$ & $2.0$\\[1cm]
-
-		{\LARGE TA69} & $4121.0$ & $1000.0$ & $1.0$ & $\mathbf{3748.0}$ & $1000.0$ & $2.0$ & $3803.0$ & $1000.0$ & $3.0$ & $4116.0$ & $1000.0$ & $4.0$ & $3937.0$ & $1000.0$ & $5.0$ & $4518.0$ & $111.0$ & $1.0$\\[1cm]
-
-		{\LARGE TA70} & $4111.0$ & $1000.0$ & $1.0$ & $3827.0$ & $1000.0$ & $2.0$ & $\mathbf{3791.0}$ & $1000.0$ & $3.0$ & $4008.0$ & $1000.0$ & $4.0$ & $4044.0$ & $1000.0$ & $5.0$ & $4651.0$ & $365.2$ & $3.0$\\ [1cm]
+    30\%             & \boldmath{$2943.1$}  & $600.0$   & $4.0$	   & $3162.2$      & $600.0$ & $4.0$ & $5857.0$ & $600.0$ & $9.0$\\ 
+                     [1.5mm]
     \hline
-
-		\multirow{3}{*}{{\LARGE AVG}} & \multirow{3}{*}{$3879.4$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$1.0$} & \multirow{3}{*}{$\mathbf{3540.8}$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$2.0$} & \multirow{3}{*}{$3605.7$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$3.0$} & \multirow{3}{*}{$3918.6$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$4.0$} & \multirow{3}{*}{$4006.3$} & \multirow{3}{*}{$974.3$} & \multirow{3}{*}{$4.8$} & \multirow{3}{*}{$4397.1$} & \multirow{3}{*}{$245.2$} & \multirow{3}{*}{$2.0$}\\ [1cm]
-
-		\hline
-
     \end{tabular}
-    \end{adjustbox}
-	}
-
-\end{table}
-
-\begin{table}[ht]
-	\fontsize{19pt}{30pt}
-	\caption{Results for JSP $100 \times 20$ instances with different TWs} \label{tab: Table03}
-	\centering
-	%\resizebox{\columnwidth}{!}{%
-	\rotatebox{90}{
-	\begin{adjustbox}{width=1.7\textwidth}
-		%\begin{tabular}[m]{|m{2cm}|*{17}{c}|c|}
-    \begin{tabular}{|M{2.0cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|M{1.9cm}|M{1.9cm}|M{1.2cm}|}
-
-        %\begin{tabular}{l c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c | c c c}
-        \hline
-        \multirow{3}*{\textbf{{\large Instances}}}  &  \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 01}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 02}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 03}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 04}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 05}}}} & \multicolumn{3}{|c|}{\multirow{3}{*}{\textbf{{\LARGE TW 10}}}} \\[1cm]
-
-        \hline
-        & {\large Makespan} & {\Large Timeout}  & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} & {\large Makespan} & {\Large Timeout} & {\large Interr-calls} &  {\large Makespan} & {\Large Timeout} & {\large Interr-calls}\\[1cm]
-
-		{\LARGE TA71} & $52147.0$ & $1000.0$ & $1.0$ & $21345.0$ & $1000.0$ & $2.0$ & $7641.0$ & $1000.0$ & $3.0$ & $7471.0$ & $1000.0$ & $4.0$ & $\mathbf{7226.0}$ & $1000.0$ & $5.0$ & $8478.0$ & $1000.0$ & $10.0$\\[1cm]
-
-		{\LARGE TA72} & $45196.0$ & $1000.0$ & $1.0$ & $23505.0$ & $1000.0$ & $2.0$ & $\mathbf{6750.0}$ & $1000.0$ & $3.0$ & $7134.0$ & $1000.0$ & $4.0$ & $7300.0$ & $1000.0$ & $5.0$ & $8164.0$ & $1000.0$ & $10.0$\\[1cm]
-
-		{\LARGE TA73} & $36548.0$ & $1000.0$ & $1.0$ & $24995.0$ & $1000.0$ & $2.0$ & $7121.0$ & $1000.0$ & $3.0$ & $\mathbf{7027.0}$ & $1000.0$ & $4.0$ & $7416.0$ & $1000.0$ & $5.0$ & $7975.0$ & $1000.0$ & $10.0$\\[1cm]
-
-		{\LARGE TA74} & $52973.0$ & $1000.0$ & $1.0$ & $25230.0$ & $1000.0$ & $2.0$ & $6703.0$ & $1000.0$ & $3.0$ & $\mathbf{6634.0}$ & $1000.0$ & $4.0$ & $7062.0$ & $1000.0$ & $4.0$ & $7994.0$ & $1000.0$ & $10.0$\\[1cm]
-
-		{\LARGE TA75} & $45316.0$ & $1000.0$ & $1.0$ & $23651.0$  & $1000.0$ & $2.0$ & $\mathbf{7042.0}$ & $1000.0$ & $3.0$ & $7511.0$ & $1000.0$ & $4.0$ & $7070.0$ & $1000.0$ & $5.0$ & $7934.0$ & $1000.0$ & $10.0$\\[1cm]
-
-		{\LARGE TA76} & $53532.0$ & $1000.0$ & $1.0$ & $40944.0$ & $1000.0$ & $2.0$ & $7067.0$ & $1000.0$ & $3.0$ & $\mathbf{6879.0}$ & $1000.0$ & $4.0$ & $7061.0$ & $1000.0$ & $5.0$ & $7938.0$ & $1000.0$ & $10.0$\\[1cm]
-
-		{\LARGE TA77} & $45003.0$ & $1000.0$ & $1.0$ & $24517.0$ & $1000.0$ & $2.0$ & $6947.0$ & $1000.0$ & $3.0$ & $\mathbf{6929.0}$ & $1000.0$ & $4.0$ & $7307.0$ & $1000.0$ & $5.0$ & $7773.0$ & $1000.0$ & $10.0$\\[1cm]
-
-		{\LARGE TA78} & $41380.0$ & $1000.0$ & $1.0$ & $24686.0$ & $1000.0$ & $2.0$ & $6917.0$ & $1000.0$ & $3.0$ & $\mathbf{6793.0}$ & $1000.0$ & $4.0$ & $7215.0$ & $1000.0$ & $5.0$ & $7986.0$ & $1000.0$ & $10.0$\\[1cm]
-
-		{\LARGE TA79} & $48827.0$ & $1000.0$ & $1.0$ & $24401.0$ & $1000.0$ & $2.0$ & $6913.0$ & $1000.0$ & $3.0$ & $\mathbf{6621.0}$ & $1000.0$ & $4.0$ & $6645.0$ & $1000.0$ & $5.0$ & $7405.0$ & $1000.0$ & $10.0$\\[1cm]
-
-		{\LARGE TA80} & $46939.0$ & $1000.0$ & $1.0$ & $22833.0$ & $1000.0$ & $2.0$ & $6925.0$ & $1000.0$ & $3.0$ & $\mathbf{6645.0}$ & $1000.0$ & $4.0$ & $6647.0$ & $1000.0$ & $5.0$ & $7609.0$ & $1000.0$ & $10.0$\\ [1cm]
-        \hline
-
-		\multirow{3}{*}{{\LARGE AVG}} & \multirow{3}{*}{$46786.1$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$1.0$} & \multirow{3}{*}{$25610.7$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$2.0$} & \multirow{3}{*}{$7002.6$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$3.0$} & \multirow{3}{*}{$\mathbf{6964.4}$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$4.0$} & \multirow{3}{*}{$7094.9$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$5.0$} & \multirow{3}{*}{$7925.6$} & \multirow{3}{*}{$1000.0$} & \multirow{3}{*}{$10.0$}\\ [1cm]
-
-		\hline
-
-	\end{tabular}
-    \end{adjustbox}
-	}
-
+   % }
+%    \end{minipage}
+% }
+%    \hfill\hfill\mbox{~}
 \end{table}

### 2022-01-16 ICLP Paper/samplepaper.tex
added some literature

--- 2022-01-14 ICLP Paper/samplepaper.tex
@@
 For instance, benchmark (FT10) consists of $10$ jobs and $10$ machines took researchers roughly $20$ years to find the optimal solution\cite{adams1988shifting,zhang2010hybrid}.
+
+In real-world production scheduling, the number of operations to process can easily go
+into tens of thousands \citep{coltep19a,kohakamo20a,kotakoscge21a}, which exceeds exact
+optimization capacities even of state-of-the-art solvers for Answer Set Programming (ASP),
+Mixed Integer Programming (MIP), or Constraint Programming (CP)
+\citep{daneshamooz2021mathematical,francescutto2021solving,shi2021solving}.
 

### 2022-01-17 ICLP Paper/samplepaper.tex
update tables_2

--- 2022-01-16 ICLP Paper/samplepaper.tex
@@
 \usepackage{multirow}
+\usepackage{comment}
 % Used for displaying a sample figure. If possible, figure files should
@@
 
+\begin{comment}
 \begin{table}[t]
@@
     \centering
-    \begin{tabular}{l l r r r r r}
+    \begin{tabular}{l l r r r r r r r r r r}
     \hline
 %       & {}        & \multicolumn{6}{c}{Group of Instances}  \\
-    Window(s) & Comparison & 50$\times$15 & 50$\times$20 & 100$\times$20 \\
+    &  &  \multicolumn{10}{c}{Number of Time Windows} \\
+    Instances & Comparison & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
 
     \hline\\[-2.75mm]
-    1                & Makespan             & 3542.1   & 3879.4	     &  46786.1 \\
-                     & CPU(Sec)             & 1000.0	& 1000.0   & 1000.0	     	    \\
-                     & Interrupted Calls    & 1.0	    & 1.0      & 1.0	       \\[1.5mm]
+    {50$\times$15}       & Makespan             & $3542.1$   & $3149.7$	     &  \boldmath{$3083.8$} & $3110.7$ & $3213.6$ & $3225.3$ & $3350.5$ & $3378.5$ & $3495.0$ & $3524.9$ \\
+    & CPU(Sec)             & $1000.0$	& $1000.0$   & $1000.0$	 & $934.4$ & $615.5$  & $243.8$  & $146.7$ 	& $81.0$ & $45.6$ & $18.5$  \\
+    & Interrupted Calls    & $1.0$	    & $2.0$      & $3.0$	& $3.6$  & $2.7$  & $1.0$  &  $0.6$  & $0.3$ & $0.1$ & $0.1$       \\[1.5mm]
 
-    2                & Makespan             & 3149.7   & 3479.4	      & 23977.5 \\
-                     & CPU(Sec)             & 1000.0	    & 1000.0	   & 1000.0	      \\
-                     & Interrupted Calls    & 2.0	    & 2.0      & 2.0	      \\[1.5mm]
+    {50$\times$20}       & Makespan             & $3879.4$   & $3479.4$	      & $3290.5$  &  \boldmath{$3260.3$} &  $3355.5$ & $3414.8$ & $3540.3$  &  $3561.5$  & $3720.7$  & $3793.1$ \\
+    & CPU(Sec)             & $1000.0$	    & $1000.0$	   & $1000.0$	  & $944.9$  & $583.4$  & $321.5$  &  $161.4$  & $54.9$  & $38.0$  & $33.3$     \\
+    & Interrupted Calls    & $1.0$	    & $2.0$      & $3.0$	& $3.5$ & $2.5$  & $1.6$  & $0.9$  & $0.2$ & $0.2$  & $0.1$      \\[1.5mm]
 
-    3                & Makespan             & \textbf{3083.8}   & 3290.5	    & 7201.5 \\
-                     & CPU(Sec)             & 1000.0	    & 1000.0	   & 1000.0	      \\
-                     & Interrupted Calls    & 3.0	    & 3.0      & 3.0	      \\[1.5mm]
+    {100$\times$20}      & Makespan             & $46786.1$  &  $23977.5$  & $7201.5$ &  $6288.7$                & $6106.1$  &  \boldmath{$6031.9$} & $6094.1$ & $6053.1$ & $6095.9$  & $6234.0$  \\
+    & CPU(Sec) & $1000.0$	    & $1000.0$	   & $1000.0$	   & $1000.0$	    & $1000.0$	& $1000.0$	& $1000.0$ & $1000.0$	    & $1000.0$	& $1000.0$            \\
+    & Interrupted Calls    & $1.0$	    & $2.0$      & $3.0$	& $4.0$ & $5.0$  & $6.0$ & $7.0$ & $8.0$ & $8.2$ & $7.9$      \\[1.5mm]
 
-    4                & Makespan             & 3110.7   & \textbf{3260.3}	   & 6288.7 \\
-                     & CPU(Sec)             & 934.4	    & 944.9	   & 1000.0	      \\
-                     & Interrupted Calls    & 3.6	    & 3.5      & 4.0	      \\[1.5mm]
+    \\[-2.75mm]
+    \hline
+    \end{tabular}
+   % }
+%    \end{minipage}
+% }
+%    \hfill\hfill\mbox{~}
+\end{table}}
 
-    5                & Makespan             & 3213.6   & 3355.5    & 6106.1 \\
-                     & CPU(Sec)             & 615.5	    & 583.4	   & 1000.0	      \\
-                     & Interrupted Calls    & 2.7	    & 2.5      & 5.0	      \\[1.5mm]
+\end{comment}
 
-    6                & Makespan             & 3225.3   & 3414.8	      & \textbf{6031.9} \\
-                     & CPU(Sec)             & 243.8	    & 321.5	   & 1000.0	      \\
-                     & Interrupted Calls    & 1.0	    & 1.6      & 6.0	      \\[1.5mm]
+\begin{table}[t]
+    \caption{Results for three sizes of job shop scheduling instances with different numbers of time windows using static machine-based EST decomposition \label{tab:Table01}}%
+%    \mbox{~}\hfill
+%    \resizebox{0.8\textwidth}{!}{%
+%    \begin{minipage}{\textwidth}
+    \centering
+    \begin{tabular}{l l r r r r r r r}
+    \hline
+%       & {}        & \multicolumn{6}{c}{Group of Instances}  \\
+    &  &  \multicolumn{7}{c}{Number of Time Windows} \\
+    Instances & Comparison & 1 & 2 & 3 & 4 & 5 & 6 & 10 \\
 
-    7                & Makespan             & 3350.5   & 3540.3	      & 6094.1 \\
-                     & CPU(Sec)             & 146.7	    & 161.4	   & 1000.0	      \\
-                     & Interrupted Calls    & 0.6	    & 0.9      & 7.0	      \\[1.5mm]
+    \hline\\[-2.75mm]
 
-    8                & Makespan             & 3378.5   & 3561.5	      & 6053.1 \\
-                     & CPU(Sec)             & 81.0	    & 54.9	   & 1000.0	      \\
-                     & Interrupted Calls    & 0.3	    & 0.2      & 8.0	      \\[1.5mm]
+    {50$\times$15}       & Makespan             & $3542.1$   & $3149.7$	     &  \boldmath{$3083.8$} & $3110.7$ & $3213.6$ & $3225.3$ & $3524.9$ \\
+    & CPU(Sec)             & $1000.0$	& $1000.0$   & $1000.0$	 & $934.4$ & $615.5$  & $243.8$  & $18.5$  \\
+    & Interrupted Calls    & $1.0$	    & $2.0$      & $3.0$	& $3.6$  & $2.7$  & $1.0$  & $0.1$       \\[1.5mm]
 
-    9                & Makespan             & 3495.0   & 3720.7	      & 6095.9 \\
-                     & CPU(Sec)             & 45.6	    & 38.0	   & 977.0	      \\
-                     & Interrupted Calls    & 0.1	    & 0.2      & 8.2	      \\[1.5mm]
+    {50$\times$20}       & Makespan             & $3879.4$   & $3479.4$	      & $3290.5$  &  \boldmath{$3260.3$} &  $3355.5$ & $3414.8$ & $3793.1$ \\
+    & CPU(Sec)             & $1000.0$	    & $1000.0$	   & $1000.0$	  & $944.9$  & $583.4$  & $321.5$   & $33.3$     \\
+    & Interrupted Calls    & $1.0$	    & $2.0$      & $3.0$	& $3.5$ & $2.5$  & $1.6$  & $0.1$      \\[1.5mm]
 
-    10               & Makespan             & 3524.9   & 3793.1	      & 6234.0 \\
-                     & CPU(Sec)             & 18.5	    & 33.3	   & 861.8	      \\
-                     & Interrupted Calls    & 0.1	    & 0.1      & 7.9	      \\[1.5mm]
+    {100$\times$20}      & Makespan             & $46786.1$  &  $23977.5$  & $7201.5$ &  $6288.7$                & $6106.1$  &  \boldmath{$6031.9$}   & $6234.0$  \\
+    & CPU(Sec) & $1000.0$	    & $1000.0$	   & $1000.0$	   & $1000.0$	    & $1000.0$	& $1000.0$	& $861.8$            \\
+    & Interrupted Calls    & $1.0$	    & $2.0$      & $3.0$	& $4.0$ & $5.0$  & $6.0$ & $7.9$      \\[1.5mm]
 
@@
 \multirow{2}*{Strategy} & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
-    & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC \\
+    & Makespan & Time & IC & Makespan & Time & IC & Makespan & Time & IC \\
     \hline\\[-2.75mm]
-    Static machine-based EST              & \boldmath{$3083.8$}  & $1000.0$   & $3.0$	& \boldmath{$3260.3$}      & $944.9$ & $3.5$ & \boldmath{$6031.9$} & $1000.0$ & $6.0$\\
+    M-Based EST (S)             & $3083.8$  & $1000.0$   & $3.0$	& \boldmath{$3260.3$}      & $944.9$ & $3.5$ & $6031.9$ & $1000.0$ & $6.0$\\
     [1.5mm]
 
-    Static time-based EST             & $3524.8$ & $1000.0$   & $2.0$	   & $3540.8$      & $1000.0$ & $2.0$ & $6964.4$ & $1000.0$ & $4.0$\\ 
+    T-based EST (S)            & $3524.8$ & $1000.0$   & $2.0$	   & $3540.8$      & $1000.0$ & $2.0$ & $6964.4$ & $1000.0$ & $4.0$\\ 
     [1.5mm]
 
-    Static time-based MTWR             & $3456.2$  & $1000.0$   & $2.0$	& $3561.0$      & $1000.0$ & $2.0$ & $7149.6$ & $1000.0$ & $4.0$\\ 
+    M-based MTWR (S)        & \boldmath{$3060.3$}  & $1000.0$   & $3.0$	& $3287.9$      & $986.1$ & $3.9$ & \boldmath{$5948.6$} & $1000.0$ & $7.0$\\ 
     [1.5mm]
 
-    Dynamic machine-based EST             & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $3.3$ & $6115.1$ & $1000.0$ & $6.9$\\ 
+    T-based MTWR (S)            & $3456.2$  & $1000.0$   & $2.0$	& $3561.0$      & $1000.0$ & $2.0$ & $7149.6$ & $1000.0$ & $4.0$\\ 
+    [1.5mm]
+
+    M-based EST (D)            & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $3.3$ & $6115.1$ & $1000.0$ & $6.9$\\ 
     [1.5mm]

### 2022-01-18 ICLP Paper/samplepaper.tex
about integrating related work into introduction

@@
 \begin{keywords}
-Job-shop Scheduling Problem, Answer Set Programming, Decomposition
+Job-shop Scheduling Problem, Answer Set Programming, Problem Decomposition
 \end{keywords}
@@
 In general, scheduling operations is one of the most critical issues in the planning and managing of the manufacturing processes\cite{uzsoy2000performance}.
+Effective scheduling methods are essential for complex manufacturing and logistics systems,
+where allocating and performing diverse tasks within resource capacity limits 
+is one of the most critical challenges for production management \citep{uzsoy2000performance}.
+The Job-shop Scheduling Problem (JSP) \citep{baker1974introduction,taillard1993benchmarks}
+constitutes a well-known mathematical abstraction of industrial production scheduling in
+which sequences of operations need to be processed by machines such that a given objective
+like the makespan for completing all jobs or their tardiness w.r.t.\ deadlines is minimized.
+Finding optimal JSP solutions, determined by a sequence of operations for each machine,
+is an NP-hard combinatorial problem
+\citep{garey1976complexity,lenstra1977complexity,liu2008prediction},
+and optimal schedules as well as termination guarantees can be extremely challenging
+or beyond reach 
+of complete optimization methods already for moderately sized instances.
+For example, it took about $20$ years to develop a search procedure able to find a
+(provably) optimal solution for an instance called FT10 with $10$ jobs
+\citep{adams1988shifting,zhang2010hybrid},
+each consisting of a sequence of $10$ operations to be processed by $10$ machines.
 
-One of the most challenging scheduling problems is the Jop-shop Scheduling Problem (JSP).
+% One of the most challenging scheduling problems is the Jop-shop Scheduling Problem (JSP).
 A set of jobs needs to be processed on a set of machines while optimizing a performance indicator, minimizing makespan, the time needed to complete all the jobs, or tardiness, i.e., the summation of the delays in executing all jobs according to their deadlines.
@@
 \citep{daneshamooz2021mathematical,francescutto2021solving,shi2021solving}.
+Hence, more efficient approaches to approximate good-quality schedules 
+instead of striving for optimal solutions have attracted wide research interest.
+On the one hand, respective methods include greedy and local search techniques such as
+dispatching rules \citep{blackstone1982state}, shifting bottleneck \citep{adams1988shifting} and
+genetic algorithms \citep{pezzella2008genetic}.
+On the other hand, problem decomposition strategies based on a
+rolling horizon \citep{singer2001decomposition,liu2008prediction} or
+bottleneck operations \citep{zhang2010hybrid,zhai2014decomposition} 
+have been proposed to partition large-scale instances into better manageable subproblems,
+where no single strategy strictly dominates \citep{ovacik2012decomposition}.
 
-In reality, the number of operations is reaching thousands of operations.
+% In reality, the number of operations is reaching thousands of operations.
 Therefore, the exact methods Answer Set Programming (ASP), Constraint Programming and Branch and Bound cannot find the optimal solutions in a reasonable time\cite{daneshamooz2021mathematical,shi2021solving,francescutto2021solving}.

### 2022-01-18 ICLP Paper/samplepaper.tex
work on intro

@@
 \usepackage{multirow}
+\usepackage{xspace}
 \usepackage{comment}
@@
 % \renewcommand\UrlFont{\color{blue}\rmfamily}
+
+\newcommand{\clingo}{\emph{clingo}\xspace}
+\newcommand{\clingodl}{\emph{clingo}[DL]\xspace}
 
@@
 More specifically, there is no evidence that one of the introduced strategies is the best for solving all JSP(s) \cite{ovacik2012decomposition}.
+
+While decision versions of scheduling problems can be successfully modeled and
+solved by an extension of ASP with Difference Logic (DL) constraints
+\citep{gebser2016theory}, implemented by \clingodl on top of the
+(multi-shot) ASP system \clingo \citep{gekakasc17a},
+the optimization capacities of \clingodl come to their limits on moderately sized
+yet highly combinatorial JSP instances \citep{elkgeb20a}.
 

### 2022-01-19 ICLP Paper/samplepaper.tex
progress on intro

--- 2022-01-18 ICLP Paper/samplepaper.tex
@@
 More specifically, there is no evidence that one of the introduced strategies is the best for solving all JSP(s) \cite{ovacik2012decomposition}.
+
+While decision versions of scheduling problems can be successfully modeled and
+solved by an extension of ASP with Difference Logic (DL) constraints
+\citep{gebser2016theory}, implemented by \clingodl on top of the
+(multi-shot) ASP system \clingo \citep{gekakasc17a},
+the optimization capacities of \clingodl come to their limits on moderately sized
+yet highly combinatorial JSP instances \citep{elkgeb20a}, for some of which optimal solutions are so far unknown \citep{shysha18a}.
+In this paper, we greatly extend our preliminary study \citep{elkgeb20a} 
+on problem decomposition into time windows and successive schedule optimization
+by means of multi-shot ASP modulo DL solving with \clingodl.
+The goal of the decomposition is to split highly complex scheduling tasks into
+balanced portions for which partial schedules of good quality can be reliably
+found within tight runtime limits, and then be merged into a global solution of
+significantly better quality than obtainable in similar runtime 
+with single-shot optimization on the full problem.
+In this process, problem decomposition must respect the precedence of operations
+within their encompassing jobs to guarantee the feasibility of schedules, and as
+a secondary objective, the order by time windows of operations sharing a machine should come close to the sequences of operations in (unknown) optimal schedules.
+We address computational efficiency as well as solution quality by devising and investigating decomposition strategies regarding the size of
+time windows and heuristics to choose their operations.
+
+The contributions of our work going beyond the study in \citep{elkgeb20a} are:
+\begin{enumerate}
+\item In addition to problem decomposition based on the earliest starting times
+      of operations, we consider the most total work remaining criterion and
+      refinements of both strategies by bottleneck machines.
+      We encode static as well as dynamic decomposition variants, the latter taking
+      partial schedules for operations of previous time windows into account, by stratified
+      ASP programs (without DL constraints).
+\item Considering that a decomposition into time windows may be incompatible with
+      the optimal sequences of operations sharing a machine, we incorporate
+      time window overlapping into the iterative scheduling process to
+      offer a chance for revising "decomposition mistakes".
+      Moreover, as the makespan objective we apply for the optimization of
+      (partial) schedules tolerates unnecessary idle times of machines as long as
+      they do not stretch the scheduling horizon,
+      we devise a stratified ASP encoding to postprocess and compress partial
+      schedules by reassigning operations to earlier idle times of their machines.
+\end{enumerate}
+
+% From a computational perspective, decomposition aims to split highly complex scheduling tasks into better manageable subproblems with a balanced number of 
+% operations, so that good-quality or even optimal partial solutions can be
+% reliably found in a small fraction of runtime.
+% Regarding the feasibility and quality of solutions, problem decomposition must
+% respect the precedence of operations within their jobs, and partial schedules
+% optimized by time windows should yield better global solutions than obtainable
+% in similar runtime on the full problem.
+% We devise and investigate a variety of decomposition strategies in terms of the number and size of time windows as well as heuristics for choosing their operations.
+% Moreover, we incorporate time window overlapping and compression techniques
+% into the iterative scheduling process in order to counteract limitations of window-wise optimization restricted to partial schedules.
+% Our experiments on JSP benchmark sets of several sizes show that
+% successive optimization by multi-shot ASP solving leads to substantially better schedules within the runtime limit than global optimization on the full problem,
+% where the gap increases with the number of operations to schedule.
 

### 2022-01-19 ICLP Paper/samplepaper.tex
completed pass through introduction

@@
 \usepackage{xspace}
+\usepackage{url}
 \usepackage{comment}
@@
 yet highly combinatorial JSP instances \citep{elkgeb20a}, for some of which optimal solutions are so far unknown \citep{shysha18a}.
+Successful application areas beyond JSP include 
+industrial printing \citep{balduccini11a},
+team-building \citep{rigralmaliiile12a},
+shift design \citep{abseher2016shift},
+course timetabling \citep{bainkaokscsotawa18a},
+lab resource allocation \citep{francescutto2021solving},
+and
+medical treatment planning \citep{dogagrmamopo21a},
+pointing out the general attractiveness of ASP for modeling and solving scheduling problems.
+
 In this paper, we greatly extend our preliminary study \citep{elkgeb20a} 
@@
 The contributions of our work going beyond the study in \citep{elkgeb20a} are:
-\begin{enumerate}
+\begin{itemize}
 \item In addition to problem decomposition based on the earliest starting times
@@
       (partial) schedules tolerates unnecessary idle times of machines as long as
-      they do not stretch the scheduling horizon,
+      they do not yield a greater scheduling horizon,
       we devise a stratified ASP encoding to postprocess and compress partial
-      schedules by reassigning operations to earlier idle times of their machines.
-\end{enumerate}
+      schedules by reassigning operations to earlier idle slots available on their machines.
+\item We experimentally evaluate decomposition strategies varying the number of
+      operations per time window as well as multi-shot ASP solving
+      augmented with overlapping and compression techniques
+      on JSP benchmark sets of several sizes.
+      In particular, our experiments demonstrate that successive optimization
+      by multi-shot ASP solving leads to substantially better schedules within
+      tight runtime limits than global optimization on the full problem,
+      where the gap increases with the number of operations to schedule.
+\end{itemize}
 
-% From a computational perspective, decomposition aims to split highly complex scheduling tasks into better manageable subproblems with a balanced number of 
-% operations, so that good-quality or even optimal partial solutions can be
-% reliably found in a small fraction of runtime.
-% Regarding the feasibility and quality of solutions, problem decomposition must
-% respect the precedence of operations within their jobs, and partial schedules
-% optimized by time windows should yield better global solutions than obtainable
-% in similar runtime on the full problem.
-% We devise and investigate a variety of decomposition strategies in terms of the number and size of time windows as well as heuristics for choosing their operations.
-% Moreover, we incorporate time window overlapping and compression techniques
-% into the iterative scheduling process in order to counteract limitations of window-wise optimization restricted to partial schedules.
-% Our experiments on JSP benchmark sets of several sizes show that
-% successive optimization by multi-shot ASP solving leads to substantially better schedules within the runtime limit than global optimization on the full problem,
-% where the gap increases with the number of operations to schedule.
-
-In this study, we aim to introduce a new decomposition strategy for solving the JSP.
+% In this study, we aim to introduce a new decomposition strategy for solving the JSP.
 Our study proposes a new policy to efficiently divide the problem into subproblems, aiming to reach a near-optimal solution.
@@
 
-The paper is organized as follows.
-The following section shows the most related work and which decomposition techniques have been introduced in the past, followed by the problem formulation section illustrates the problem in detail.
+The rest of this 
+paper is organized as follows.
+Section~\ref{sec:preliminaries} briefly introduces ASP along with the relevant
+extensions of multi-shot solving and DL constraints.
+In Section~\ref{sec:problem}, we present our successive optimization approach
+and detail the ASP programs encoding problem decomposition or iterative scheduling by
+time windows, respectively.
+Section~\ref{sec:experiments} provides experimental results on JSP benchmark sets, % of several sizes,
+assessing different decomposition strategies as well as the impact of overlapping and
+compression techniques.
+Conclusions and future work are discussed in Section~\ref{sec:conclusions}.  
+% The following section shows the most related work and which decomposition techniques have been introduced in the past, followed by the problem formulation section illustrates the problem in detail.
 Section $4$ presents our proposed model, which shows the decomposition techniques we used and describes one in more detail.
@@
 
-\section{Literature Review}
-This section will review some of the work that has applied the decomposition approach for solving scheduling problems.
+% \section{Literature Review}
+% This section will review some of the work that has applied the decomposition approach for solving scheduling problems.
 As mentioned in the previous section, no particular decomposition method has proved its efficiency in solving all scheduling problems.
@@
 
-Another study investigated the performance of a new decomposition procedure based on SB in the Job-shop and flow shop with different levels of bottleneck machines.
+% Another study investigated the performance of a new decomposition procedure based on SB in the Job-shop and flow shop with different levels of bottleneck machines.
 The computational experiments showed that the proposed method obtained better solutions in a shorter time than for the problems in which the machine's workload is identical.
@@
 
-The rolling horizon procedure has been extended by constructing a prediction model to obtain the scheduling characteristics values, including the information of the bottleneck jobs.
+% The rolling horizon procedure has been extended by constructing a prediction model to obtain the scheduling characteristics values, including the information of the bottleneck jobs.
 The obtained information aided in decomposing the problem efficiently.
@@
 
-A decomposition method based on bottleneck machines has been presented for solving the JSP.
+% A decomposition method based on bottleneck machines has been presented for solving the JSP.
 The proposed model decomposes the problems into subproblems and then detects the multi-bottleneck machines using a critical path method.
@@
 
-From the literature, we can find that most of the related work focused on applying the decomposition approach while minimizing the total tardiness, and the number of work tackled large-scale instances to minimize the makespan is relatively low.
+% From the literature, we can find that most of the related work focused on applying the decomposition approach while minimizing the total tardiness, and the number of work tackled large-scale instances to minimize the makespan is relatively low.
 This paper proposes different decomposition strategies for solving the JSP with larger instances using the ASP model while minimizing the makespan.
 
-\section{Preliminaries}
+\section{Preliminaries}\label{sec:preliminaries}
 \paragraph{Answer Set Programming (ASP)} is a declarative programming language for solving hard combinatorial optimization problems.
@@
 
-\section{Problem Formulation}
+\section{Multi-shot JSP Solving}\label{sec:problem}
 The Job-shop Scheduling Problem (JSP) has been a complex and combinatorial optimization problem since the 1950s, and it was shown to be NP-hard.
@@
 
-\subsection{Modelling JSP with Hybrid ASP}
+\subsection{Modeling JSP with Hybrid ASP}
 ASP has been applied to solve different scheduling problems.
@@
 
-\section{Results}
+\section{Experiments}\label{sec:experiments}
 This section will show the computational experiments that we conducted to test the performance of our proposed model.
@@
 
+\section{Conclusions}\label{sec:conclusions}
+
 %

### 2022-01-22 ICLP Paper/samplepaper.tex
working on instance

@@
 \documentclass{tlp} % \documentclass[runningheads]{llncs}
-%
+
 \usepackage{graphicx}
@@
 \usepackage{helvet,times,courier}
-\usepackage{listings}
 \usepackage[inline]{enumitem}
@@
 \newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
-\lstset{numberbychapter=false,numbers=left,numberblanklines=false,basicstyle=\footnotesize\ttfamily,xleftmargin=5ex,breaklines=false}
+
+\usepackage{calc}
+\newlength\listingnumberwidth
+\setlength\listingnumberwidth{\widthof{\scriptsize 00} + 10pt}
+% \setlength{\mathindent}{\listingnumberwidth}
+
+\usepackage{listings}
+\lstset{numberbychapter=false,numbers=left,numberblanklines=false,basicstyle=\ttfamily\small,%xleftmargin=5ex,
+xleftmargin=\listingnumberwidth,captionpos=b,breaklines=false,frame=single}
+
 \usepackage{multirow}
@@
 \usepackage{url}
+
 \usepackage{comment}
@@
 \section{Multi-shot JSP Solving}\label{sec:problem}
-The Job-shop Scheduling Problem (JSP) has been a complex and combinatorial optimization problem since the 1950s, and it was shown to be NP-hard.
+This section describes our successive optimization approach to
+JSP solving by means of multi-shot ASP with \clingodl.
+We start with specifying the fact format for JSP instances,
+then detail problem decomposition based on earliest starting times of operations,
+present our ASP encoding with DL constraints for optimizing the makespan of partial schedules, and
+finally outline the incorporation of time window overlapping and compression
+into the iterative scheduling process.   
+% The Job-shop Scheduling Problem (JSP) has been a complex and combinatorial optimization problem since the 1950s, and it was shown to be NP-hard.
 In the JSP, there are a set of jobs to be processed on a set of machines.
@@
 The basic assumptions are as follows:
+%
+% \begin{enumerate}
+% 	\item The processing time of each operation is fixed.
+% 	\item Machine breakdown preemption of the operations are not allowed.
+% 	\item Each machine can process only one operation at a time.
+% 	\item Each operation can be executed by only one machine.
+% 	\item A machine cannot process more than one operation at a time.
+% 	\item The jobs and the machines are available at time $0$.
+% \end{enumerate}
 
-\begin{enumerate}
-	\item The processing time of each operation is fixed.
-	\item Machine breakdown preemption of the operations are not allowed.
-	\item Each machine can process only one operation at a time.
-	\item Each operation can be executed by only one machine.
-	\item A machine cannot process more than one operation at a time.
-	\item The jobs and the machines are available at time $0$.
-\end{enumerate}
-
-\subsection{Modeling JSP with Hybrid ASP}
-ASP has been applied to solve different scheduling problems.
+\subsection{Instance Format}\label{subsec:instance}
+% ASP has been applied to solve different scheduling problems.
 For example, ASP is used to solve scheduling problems in the healthcare systems \cite{dodaro2017nurse,dodaro2019asp}.
@@
 This paper aims to use different decomposition methods to solve the JSP using ASP with Difference Logic.
-
-\paragraph{Problem Instances.}
+% 
+% \paragraph{Problem Instances.} 
+%
+Each job in a JSP instance is a sequence of operations with associated
+machines and processing times.
+Corresponding facts for an example instance with three jobs and three
+machines are displayed in Listing~\ref{prg:facts}.
 
@@
 For instance, \emph{operation(1,1)} is processed by machine 2.
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:facts,caption={Problem instance}, linerange={1-11} ]{listing/facts.lp}
+\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:facts,caption={Example JSP instance},linerange={1-3}]{listing/facts.lp}
 

### 2022-01-22 ICLP Paper/samplepaper.tex
working on instance format

@@
 
-\subsection{Instance Format}\label{subsec:instance}
+\subsection{Problem Instances}\label{subsec:instance}
 % ASP has been applied to solve different scheduling problems.
@@
 machines are displayed in Listing~\ref{prg:facts}.
-
-In order to encode the problem using ASP, we should define a set of predicates that represents the problem instances.
+An atom of the form
+\lstinline{operation(}$j$\lstinline{,}$s$\lstinline{,}$m$\lstinline{,}$p$\lstinline{)}
+denotes that the step~$s$ of job~$j$ needs to be processed by machine~$m$ for $p$ time units.
+For example, the second operation of job~\lstinline{3} has a processing time of
+\lstinline{3} time units on machine~\lstinline{1},
+as specified by the fact
+\lstinline{operation(3,}\linebreak[1]\lstinline{2,}\linebreak[1]\lstinline{1,3).}
+The operation cannot be performed before the first operation of job~\lstinline{3}
+is completed,
+and its execution must not intersect with the first operation of job~\lstinline{1}
+or the second operation of job~\lstinline{2},
+which need to be processed by machine~\lstinline{1} as well.
+That is, a schedule for the example instance must determine a sequence in which
+to process the three mentioned operations on machine~\lstinline{1},
+and likewise for operations sharing machine~\lstinline{2} or~\lstinline{3}, respectively.
+%
+% In order to encode the problem using ASP, we should define a set of predicates that represents the problem instances.
 For instance, the operations are encoded with the predicate \lstinline{operation/2} where the first variable denotes the job number and the second is the operation number, see lines \ref{prg:facts:ops:begin}-\ref{prg:facts:ops:end} in Listing \ref{prg:facts}.
@@
 
-\subsection{Decomposition strategies}
+\subsection{Problem Decomposition}\label{subsec:decomposition}
 As we mentioned in the previous sections, it is hard to solve the JSP and reach a near-optimal solution in a reasonable time, especially when the number of operations to be processed is high.

### 2022-01-22 ICLP Paper/samplepaper.tex
added example schedule for the instance

@@
 
-\usepackage{graphicx}
+% \usepackage{graphicx}
 \usepackage{amsmath}
-\usepackage{amssymb}
+% \usepackage{amssymb}
 \usepackage{helvet,times,courier}
-\usepackage[inline]{enumitem}
-\usepackage{adjustbox}
-\usepackage{array}
-\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
-\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
+% \usepackage[inline]{enumitem}
+% \usepackage{adjustbox}
+% \usepackage{array}
+% \newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
+% \newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
+
+\usepackage{tikz}
+\usetikzlibrary{positioning,patterns}
+\newlength{\scaledx}
+\newlength{\scaledy}
+\newcommand\SetScales{%
+  \pgfextractx{\scaledx}{\pgfpointxy{1}{0}}%
+  \pgfextracty{\scaledy}{\pgfpointxy{0}{1}}%
+}
 
@@
 \newcommand{\clingodl}{\emph{clingo}[DL]\xspace}
+\newcommand{\figspace}{\vspace{-1ex}}
+
+\makeatletter
+\def\fcapsize@figure{\normalfont\normalsize\rmfamily}
+\def\fcapstyle@figure{\normalfont\normalsize\itshape}
+\makeatother
 
@@
 
-\subsection{Problem Instances}\label{subsec:instance}
+\subsection{Problem Instance}\label{subsec:instance}
 % ASP has been applied to solve different scheduling problems.
@@
 \lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:facts,caption={Example JSP instance},linerange={1-3}]{listing/facts.lp}
+
+Figure~\ref{fig:schedule} depicts an optimal schedule in terms of makespan,
+i.e., the latest completion time of any job/operation,
+for the JSP instance from Listing~\ref{prg:facts}.
+The \textbf{J}-\textbf{S} pairs in horizontal bars indicated for the
+machines~\lstinline{1}, \lstinline{2}, and~\lstinline{3} identify operations
+by their job~\textbf{J} and step number~\textbf{S}.
+For each machine, observe that the bars for operations it processes do not
+intersect, so that the operations are performed in sequential order.
+Moreover, operations belonging to the same job are scheduled one after the other.
+For example,
+the second operation of job~\lstinline{3} is only started after the
+completion of the predecessor operation at time~$9$,
+regardless of the availability of its machine~\lstinline{1}
+from time~$3$ on.
+As the precedence of operations within their jobs must be respected
+and the sum of processing times for operations of job~\lstinline{3}
+matches the makespan~$20$, it is impossible to reduce the scheduling
+horizon any further,
+which in turn means that the schedule shown in Figure~\ref{fig:schedule} is optimal.
+%
+\begin{figure}[t]
+\begin{tikzpicture}[x=3.5ex,y=3.5ex,label distance=-1ex,thick]
+\SetScales
+\foreach \i in {1,...,3} {
+  \node[label=left:Machine \i] at (0,3.5-\i) {};
+}
+\foreach \i in {0,...,20} {
+  \draw[dotted] (\i,0) -- (\i,3) node [below] at (\i,0) {$\i$};
+}
+\foreach \i in {0,...,3} {
+  \draw[dotted] (0,\i) -- (20,\i);
+}
+\node[minimum width=3\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=north west lines] at (0,2) {\textbf{1}-\textbf{1}};
+\node[minimum width=3\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=crosshatch] at (9,2) {\textbf{3}-\textbf{2}};
+\node[minimum width=6\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=grid] at (12,2) {\textbf{2}-\textbf{2}};
+\node[minimum width=4\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=grid] at (0,1) {\textbf{2}-\textbf{1}};
+\node[minimum width=3\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=north west lines] at (4,1) {\textbf{1}-\textbf{2}};
+\node[minimum width=8\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=crosshatch] at (12,1) {\textbf{3}-\textbf{3}};
+\node[minimum width=9\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=crosshatch] at (0,0) {\textbf{3}-\textbf{1}};
+\node[minimum width=1\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=north west lines] at (9,0) {\textbf{1}-\textbf{3}};
+\node[minimum width=2\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=grid] at (18,0) {\textbf{2}-\textbf{3}};
+\end{tikzpicture}
+\figspace
+\caption{Optimal schedule for example JSP instance\label{fig:schedule}}% in Listing~\ref{prg:facts}\label{fig:schedule}}
+\end{figure}
 

### 2022-01-25 ICLP Paper/samplepaper.tex
simplifying decomposition encoding
extending the results and writing the conclusions

@@
 The decomposition approach has been widely used to solve the JSP by many researchers and proved its efficiency.
+%
+\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:decomposition,caption={Time-based EST decomposition}]{listing/decomposition.lp}
 
@@
 
+Table \ref{tab:Table01} shows the results of solving the mentioned different sizes before with a machine-based EST decomposition method.
+The columns represent how many windows split the original problem where TW $1$ refers that the problem is not decomposed.
+The results of $7, 8$ and $9$ TWs have not been presented because they have not changed the tendency of the results.
+Every three rows show the average makespan, running time and the number of interrupted calls of a particular size of the instances.
+We run the model with a fixed amount of time $1000$ seconds divided by TWs for each window.
+It is observed that increasing the timeout does not provide better results, and when we decrease it, the results become worse.
+The best-obtained average makespan for the $50\times15$ instances is $3083.8$ with $3$ TWs where the number of operations assigned to each window is $250$.
+In addition, increasing the TWs supported the optimizer to reach an optimal of a sub-problem, as the number of operations assigned to a TW decreased.
+Therefore, The average running time for 10 TWs is $18.5$ seconds.
+However, the average makespan is getting worse.\\
+
+The number of windows with the best obtained average makespan $3260.3$ for scheduling $50\times20$ is $4$ with an average timeout of $944.9$.
+We observe that assigning $250$ operations to each window leads to getting the best average makespan for the first two sets.
+For the $100\times20$ set, we got an average makespan of $6031.9$ with $6$ TWs and $1000$ seconds timeout, where the optimizer could not reach a sub-optimal for any TW.
+
 \begin{table}[t]
@@
 
+Table \ref{tab:Table02} demonstrates the results of the four static decomposition strategies and the machine-based MTWR with the dynamic assignment.
+Every three sub-columns have illustrated the results of a set of the benchmark instances and each row refers to a decomposition strategy with static (S) or dynamic (D) assignment.
+The results of this table conclude that the most effective decomposition strategies are machine-based regardless of whether the operations are ordered based on EST or MTWR; however, the dynamic assignment has no positive effect than the static.
+%The best average makespan of $50\times15$ and $100\times20$ are  $3060.3$, and $5948.6$.
+We applied the overlap between the windows to repair the mistakes generated by the decomposition; in addition, we compressed each optimized window to ensure that all the operations had been started as early as possible.
+Table \ref{tab:Table03} summarizes the results with/without the compression and the overlap.
+The first column refers to the overlap percentage where $0\%$ means no overlap between TWs.
+Every three columns have presented the results of a particular set of instances.
+We applied the overlap and compression with the solutions of the number of windows, which provided a better average makespan using a machine-based EST strategy.
+In the non-compression part, the overlap has improved the results slightly through the different sizes where $10\%$ and $20\%$ of overlap provided the best results.
+However, the compression has shown a significant improvement wrt the running time and the average makespan, especially for the $100\times20$ set.
+The compression demonstrated effectiveness as it minimized the average makespan and decreased the running time compared to the model without compression.
+Therefore, the overlap and compression have improved the makespan results by $4.5\%, 3.4\%$ and $6.8\%$ and the running time by $40\%$.
+
 \section{Conclusions}\label{sec:conclusions}
+In this work, different decomposition strategies have been proposed for solving the Job-shop Scheduling Problem (JSP) using a multi-shot Answer Set Programming with Difference Logic constraints to handle the decomposition approach and the problem constraints.
+The proposed model split the problem into sub-problems using different decomposition strategies and then optimized each separately while minimizing the Makespan.
+The proposed model has been developed to improve the performance by applying the overlap between TWs to fix the decomposition mistakes.
+In addition, the compression of each window is to fulfill the gaps that occurred during the optimization process.
+We tested our model on a set of benchmark instances and the experiments have shown that i) considering the bottleneck machines during the decomposition process is effective, ii) The overlap between TWs improves the solution quality, and iii) the compression of TWs is super practical, as it aided the proposed model to obtain better solutions in less computational time.
+Our work can be extended by linking the optimization model with a simulator that considers more realistic factors such as machine breakdowns, new job arrivals and order cancellations.
 

### 2022-01-25 ICLP Paper/samplepaper.tex
Update table01

@@
 
-    {100$\times$20}      & Makespan             & $46786.1$  &  $23977.5$  & $7201.5$ &  $6288.7$                & $6106.1$  &  \boldmath{$6031.9$}   & $6234.0$  \\
+    {100$\times$20}      & Makespan             & $46786.1$  &  $26210.4$  & $7201.5$ &  $6288.7$                & $6106.1$  &  \boldmath{$6031.9$}   & $6234.0$  \\
     & CPU(Sec) & $1000.0$	    & $1000.0$	   & $1000.0$	   & $1000.0$	    & $1000.0$	& $1000.0$	& $861.8$            \\

### 2022-01-25 ICLP Paper/samplepaper.tex
encodings update

@@
 \newlength\listingnumberwidth
-\setlength\listingnumberwidth{\widthof{\scriptsize 00} + 10pt}
+\setlength\listingnumberwidth{\widthof{\scriptsize 00} + 7pt}
 % \setlength{\mathindent}{\listingnumberwidth}
@@
 \usepackage{listings}
-\lstset{numberbychapter=false,numbers=left,numberblanklines=false,basicstyle=\ttfamily\small,%xleftmargin=5ex,
+\lstset{numberbychapter=false,numbers=left,numberblanklines=false,basicstyle=\ttfamily\small,numbersep=7pt,%xleftmargin=5ex,
 xleftmargin=\listingnumberwidth,captionpos=b,breaklines=false,frame=single}
@@
 \subsection{Problem Decomposition}\label{subsec:decomposition}
-As we mentioned in the previous sections, it is hard to solve the JSP and reach a near-optimal solution in a reasonable time, especially when the number of operations to be processed is high.
-The decomposition approach has been widely used to solve the JSP by many researchers and proved its efficiency.
 %
@@
 \subsection{Multi-shot with JSP}
+%
+\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:encoding,caption={Multi-shot ASP modulo DL encoding}]{listing/encoding.lp}
+
 This section will show how to solve the problem dynamically, where the output of each iteration is an input to the next and merging them to obtain the solution of the whole problem.

### 2022-01-25 ICLP Paper/samplepaper.tex
Section 3.2 revised

@@
 \newlength\listingnumberwidth
-\setlength\listingnumberwidth{\widthof{\scriptsize 00} + 7pt}
+\setlength\listingnumberwidth{\widthof{\scriptsize 00} + 5pt}
 % \setlength{\mathindent}{\listingnumberwidth}
@@
 \usepackage{listings}
-\lstset{numberbychapter=false,numbers=left,numberblanklines=false,basicstyle=\ttfamily\small,numbersep=7pt,%xleftmargin=5ex,
-xleftmargin=\listingnumberwidth,captionpos=b,breaklines=false,frame=single}
+\lstset{numberbychapter=false,numbers=left,numberblanklines=false,basicstyle=\ttfamily\small,numbersep=5pt,mathescape=true,escapeinside={\#(}{\#)},%xleftmargin=5ex,
+xleftmargin=\listingnumberwidth,xrightmargin=3pt,captionpos=b,breaklines=false,frame=single}
 
@@
 For instance, \emph{operation(1,1)} is processed by machine 2.
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:facts,caption={Example JSP instance},linerange={1-3}]{listing/facts.lp}
+\lstinputlisting[float=t,label=prg:facts,caption={Example JSP instance},linerange={1-3},numbers=none,xleftmargin=3pt]{listing/facts.lp}
 
@@
 %
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:decomposition,caption={Time-based EST decomposition}]{listing/decomposition.lp}
+Since JSP instances are highly combinatorial and the ground representation size
+can also become problematic for large real-world scheduling problems, achieving
+scalability of complete optimization methods necessitates problem decomposition.
+To this end, we consider strategies for partitioning the operations of JSP instances
+into balanced time windows, which respect the operation precedence within jobs and
+enable a successive extension of good-quality partial schedules to a global solution.
+In the following, we first detail problem decomposition based on earliest starting
+times of operations, and then outline further strategies that can be encoded
+by stratified ASP programs as well.
 
-This section will mention $4$ different decomposition strategies that we developed to split the problem into parts, describe one of them in detail, and show the encoding.
+\lstinputlisting[float=t,label=prg:decomposition,caption={Job-based EST decomposition encoding}]{listing/decomposition.lp}
+%
+Our encoding for \emph{Job-based Earliest Starting Time (EST)} decomposition
+in Listing~\ref{prg:decomposition} takes a JSP instance specified by facts over
+\lstinline{operation/4} as input.
+In addition, a constant~\lstinline{n},
+set to the default value~\lstinline{2} in line~\ref{prg:predeco:tw:begin},
+determines the number of time windows into which the given operations
+shall be split.
+As we aim at time windows of (roughly) similar size,
+the target number of operations per time window is in line~\ref{prg:predeco:optw:begin}
+calculated by $\lceil \text{\lstinline{N}} / \text{\lstinline{n}}\rceil$,
+where \lstinline{N} is the total number of operations.
+For example, we obtain \lstinline{width(5)} for partitioning the nine operations
+of the JSP instance in Listing~\ref{prg:facts} into two time windows.
+
+The rules in lines~\ref{prg:assignment_process:est:begin} and~\ref{prg:assignment_process:est:end}
+encode the EST calculation per operation of a job,
+given by the sum of processing times for predecessor operations belonging to the same job.
+This yields, e.g., 
+\lstinline{est(3,1,9,0)},
+\lstinline{est(3,2,3,9)}, and
+\lstinline{est(3,3,8,12)}
+for the three operations of job~\lstinline{3} in our example instance,
+where the third argument of an atom over \lstinline{est/4} provides
+the processing time and the fourth the earliest starting time of an operation.
+Note that the obtained earliest starting times match the first feasible time
+points for scheduling operations and do thus constitute an optimistic
+estimation of when to process the operations.
+
+With the earliest starting times of operations at hand,
+the rule in lines \ref{prg:assignment_process:rank:begin}-\ref{prg:assignment_process:rank:end}
+determines a \emph{total order} of operations in terms of consecutive indexes ranging
+from~\lstinline{0}.
+That is, each operation is mapped to the number of operations with
+(i) a smaller earliest starting time,
+(ii) the same earliest starting time and shorter processing time, or
+(iii) a smaller job identifier as tie breaker in case of similar
+earliest starting and processing times.
+For the example JSP instance in Listing~\ref{prg:facts},
+we obtain the indexes~\lstinline{0} to~\lstinline{2} for the first operations of the
+three jobs,
+the indexes~\lstinline{3} and~\lstinline{4} for the second operation of
+job~\lstinline{1} or~\lstinline{2}, respectively,
+in view of their earliest starting times~\lstinline{3} and~\lstinline{4},
+and indexes from~\lstinline{5} to~\lstinline{8} for the remaining operations.
+A relevant condition that is guaranteed by such a total order is that
+indexes increase according to the precedence of operations within their jobs,
+given that earliest starting times grow along the sequence of operations in a job.
+
+The last rule in line~\ref{prg:assignment_process:assign:begin} inspects the total
+order of operations to partition them into time windows of the size~\lstinline{W}
+determined by \lstinline{width(W)},
+where only the last time window may possibly include fewer operations in case the
+split is uneven.
+As the ASP program encoding problem decomposition is stratified, its ground
+instantiation can be simplified to (derived) facts, as shown in Listing~\ref{prg:tw}
+for our example instance.
+Time window numbers from~\lstinline{1} to~\lstinline{n = 2} are given by
+the third argument of atoms over \lstinline{window/3}, so that the second operation
+of job~\lstinline{3} and the third operation of each job form the time window~\lstinline{2},
+while time window~\lstinline{1} consists of the five other operations.
+%
+\lstinputlisting[float=t,label=prg:tw,caption={Example time windows},numbers=none,xleftmargin=3pt]{listing/TW.lp}
+
+In addition to Job-based EST decomposition,
+we have devised a similar ASP program for
+\emph{Job-based Most Total Work Remaining (MTWR)} decomposition,
+where the total order of operations is decreasing by the sum of processing
+times for an operation and its successors in a job.
+For example, we obtain the MTWR values~\lstinline{7}, \lstinline{12}, and~\lstinline{20}
+for the first operation of job~\lstinline{1}, \lstinline{2}, or~\lstinline{3},
+respectively, for the JSP instance in Listing~\ref{prg:facts},
+matching the time for executing all three operations of each job.
+Hence, the first operation of job~\lstinline{3} is considered most
+important and associated with the index~\lstinline{0}, and the other
+operations follow in the total order taken for partitioning into time windows.
+As MTWR values are decreasing along the sequence of operations in a job, the obtained time windows
+also respect the precedence of operations.
+
+Instead of assessing EST and MTWR values in a purely job-based fashion,
+we have encoded \emph{Machine-based} decomposition in which an operation
+from a bottleneck machine with greatest sum of processing times for yet
+unordered operations is considered next.
+For our example instance, the sum of processing times for operations 
+is~\lstinline{12}, \lstinline{15}, or again \lstinline{12}, respectively,
+for the machines~\lstinline{1}, \lstinline{2}, and~\lstinline{3}.
+Hence, the operation to be inserted into the total order first is picked from
+machine~\lstinline{2}, where the smallest EST or greatest MTWR value is
+used for choosing a particular job/operation.
+Unlike Job-based decomposition, this may lead to the choice of an operation
+such that predecessor operations processed by other machines are yet unordered.
+In this case, they are inserted into the total order directly before the
+operation from a bottleneck machine,
+and all machine loads are reduced by processing times
+before determining the next machine to consider.
+
+Moreover, we have devised dynamic versions of the Job- and Machine-based 
+decomposition strategies in which partial schedules are taken into account,
+and already scheduled operations from previous time windows may increase the
+EST and MTWR values used to order the yet unscheduled operations.
+However, the principle of applying a stratified ASP program to determine
+operations for the time window to schedule next is similar for all decomposition
+strategies, and we experimentally evaluate them in Section~\ref{sec:experiments}.
+
+% This section will mention $4$ different decomposition strategies that we developed to split the problem into parts, describe one of them in detail, and show the encoding.
 The decomposition strategies can be classified into two categories; time-based and machine-based.
@@
 
-\paragraph{\textbf{EST Time-based}:} it ranks the operations based on calculating the earliest possible starting time for each operation.
+% \paragraph{\textbf{EST Time-based}:} it ranks the operations based on calculating the earliest possible starting time for each operation.
 It is calculated by aggregating its predecessor(s) processing time.
@@
 
-\paragraph{\textbf{MTWR Time-based}:} the operation rank is determined based on the work remaining of a job to be completed.
+% \paragraph{\textbf{MTWR Time-based}:} the operation rank is determined based on the work remaining of a job to be completed.
 The operation belongs to a job with a high remaining processing time, will be assigned earlier to a time window.
 
-\paragraph{\textbf{EST Machine-based}:} it applies the same idea as EST Time-based.
+% \paragraph{\textbf{EST Machine-based}:} it applies the same idea as EST Time-based.
 However, the bottleneck machine is taken into account throughout the decomposition process.
@@
 
-\paragraph{\textbf{MTWR Machine-based}:} we calculate the MTWR of each operation, and the operation with the Largest MTWR and processed by a bottleneck machine will be assigned earlier to a time window.
+% \paragraph{\textbf{MTWR Machine-based}:} we calculate the MTWR of each operation, and the operation with the Largest MTWR and processed by a bottleneck machine will be assigned earlier to a time window.
 
-In this study, we will describe in detail the decomposition of the operations based on the \textbf{EST Time-based}.
+% In this study, we will describe in detail the decomposition of the operations based on the \textbf{EST Time-based}.
 We have split the decomposition encoding into two parts; the first part is for the pre-decomposition phase, which is in Listing \ref{prg:predeco}.
@@
 
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:predeco,caption={Pre-decomposition}, linerange={1-15} ]{listing/pre_deco.lp}
-Listing \ref{prg:deco} shows the decomposition using \emph{EST Time-based} strategy.
+% % \lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:predeco,caption={Pre-decomposition}, linerange={1-15} ]{listing/pre_deco.lp}
+% Listing \ref{prg:deco} shows the decomposition using \emph{EST Time-based} strategy.
 The first two rules in the lines \ref{prg:assignment_process:est:begin}-\ref{prg:assignment_process:est:end} calculate the earliest possible start time of each operation, where the EST for the first operation of all jobs are $0$ in line \ref{prg:assignment_process:est:begin}.
@@
 
-We prioritize the operations based on the estimated starting time, calculating an index for each operation.
+% We prioritize the operations based on the estimated starting time, calculating an index for each operation.
 The operation with a shorter estimated start time will get a smaller index than the others with a higher estimated starting time.
@@
 
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:deco,caption={EST-decomposition}, linerange={1-20} ]{listing/assignment_process.lp}
+% \lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:deco,caption={EST-decomposition}, linerange={1-20} ]{listing/assignment_process.lp}
 
-After running the decomposition encoding, we described above; we get an assignment of each operation to a TW.
+% After running the decomposition encoding, we described above; we get an assignment of each operation to a TW.
 This assignment is represented by a set of atoms shown in the Listing \ref{prg:tw}.
@@
 
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:tw,caption={Time Window Assignment}, linerange={1-7} ]{listing/TW.lp}
-
-\subsection{Multi-shot with JSP}
+\subsection{Problem Encoding}\label{subsec:encoding}
 %
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:encoding,caption={Multi-shot ASP modulo DL encoding}]{listing/encoding.lp}
+\lstinputlisting[float=t,label=prg:encoding,caption={Multi-shot ASP modulo DL encoding}]{listing/encoding.lp}
 
@@
 
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:base,caption={base-prog} ]{listing/base.lp}
+% \lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:base,caption={base-prog} ]{listing/base.lp}
 
@@
 The fifth rule ensures that the ending time of all operations will not exceed the bound variable, which is the makespan in our case.
-\lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:diff,caption={diff-log}]{listing/diff_log.lp}
+% \lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:diff,caption={diff-log}]{listing/diff_log.lp}
+%
+\begin{figure}[t]
+\begin{tikzpicture}[x=3.5ex,y=3.5ex,label distance=-1ex,thick]
+\SetScales
+\foreach \i in {1,...,3} {
+  \node[label=left:Machine \i] at (0,3.5-\i) {};
+}
+\foreach \i in {0,...,20} {
+  \draw[dotted] (\i,0) -- (\i,3) node [below] at (\i,0) {$\i$};
+}
+\foreach \i in {0,...,3} {
+  \draw[dotted] (0,\i) -- (20,\i);
+}
+\node[minimum width=3\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=north west lines] at (0,2) {\textbf{1}-\textbf{1}};
+\node[minimum width=3\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=crosshatch] at (10,2) {\textbf{3}-\textbf{2}};
+\node[minimum width=6\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=grid] at (4,2) {\textbf{2}-\textbf{2}};
+\node[minimum width=4\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=grid] at (0,1) {\textbf{2}-\textbf{1}};
+\node[minimum width=3\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=north west lines] at (4,1) {\textbf{1}-\textbf{2}};
+\node[minimum width=8\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=crosshatch] at (13,1) {\textbf{3}-\textbf{3}};
+\node[minimum width=9\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=crosshatch] at (0,0) {\textbf{3}-\textbf{1}};
+\node[minimum width=1\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=north west lines] at (12,0) {\textbf{1}-\textbf{3}};
+\node[minimum width=2\scaledx-0.05\scaledx,minimum height=1\scaledy-0.05\scaledy,inner sep=0,rectangle,draw,anchor=south west,pattern color=lightgray,pattern=grid] at (10,0) {\textbf{2}-\textbf{3}};
+\draw[ultra thick,double] (9,0) -- (9,1);
+% \draw[ultra thick] (7,1) -- (9,1);
+\draw[ultra thick,double] (7,1) -- (7,2);
+% \draw[ultra thick] (7,2) -- (10,2);
+\draw[ultra thick,double] (10,2) -- (10,3);
+\end{tikzpicture}
+\figspace
+\caption{Decomposed schedule for example JSP instance\label{fig:decomposed}}% in Listing~\ref{prg:facts}\label{fig:schedule}}
+\end{figure}
 

### 2022-01-26 ICLP Paper/samplepaper.tex
progress on multi-shot encoding

@@
 \subsection{Problem Encoding}\label{subsec:encoding}
+
+Given a JSP instance as in Listing~\ref{prg:facts} along with facts like those
+in Listing~\ref{prg:tw} providing a decomposition into time windows,
+the idea of successive schedule optimization is to consider time windows
+one after the other and gradually extend a partial schedule that fixes the
+operations of previous time windows.
+In this process, we adopt the makespan as optimization objective for scheduling
+the operations of each time window,
+thus applying the rule of thumb that small scheduling horizons for
+partial schedules are likely to lead towards a global solution with short makespan.
+While we use DL variables to compactly represent the starting times of operations to schedule,
+we assume that a partial schedule for the operations of previous time windows is reified
+in terms of additional input facts of the form 
+\lstinline{start((}$j$\lstinline{,}$s$\lstinline{),}$t$\lstinline{,}$w$\lstinline{).},
+where $t$ is the starting time scheduled for the step~$s$ of job~$j$ at the previous
+time window indicated by~$w$.
+
+The \lstinline{step(w)} subprogram until line~\ref{prg:diff_log:oper_limit:end} 
+constitutes the central part of our multi-shot ASP modulo DL encoding in Listing~\ref{prg:encoding}, whose parameter~\lstinline{w} stands for consecutive integers from~\lstinline{1}
+identifying time windows to schedule.
+Auxiliary atoms of the form \lstinline{use(}$m$\lstinline{,}$w'$\lstinline{,}$w$\lstinline{)},
+supplied by the rules in lines~\ref{prg:encoding:use1} and~\ref{prg:encoding:use2},
+indicate the latest time window $1\leq w'\leq w$ including some operation that needs
+to be processed by machine~$m$.
+The next rule in lines \ref{prg:base:sharing_machine:begin}-\ref{prg:base:sharing_machine:end}
+identifies pairs \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} and
+\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)} of operations sharing
+the same machine~$m$,
+where
+\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)} belongs to the time window~$w$ and
+\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} is either 
+(i)
+contained in the latest time window $1\leq w'< w$ indicated by
+\lstinline{use(}$m$\lstinline{,}$w'$\lstinline{,}$w-1$\lstinline{)} or
+(ii)
+also part of the time window~$w$, in which case $j_1 < j_2$ establishes
+an asymmetric representation for the pair of operations in derived atoms
+\lstinline{share((}$j_1$\lstinline{,}$s_1$\lstinline{),(}$j_2$\lstinline{,}$s_2$\lstinline{),}$p_1$\lstinline{,}$p_2$\lstinline{,}$x$\lstinline{,}$w$\lstinline{)}.
+If the flag $x=\text{\lstinline{1}}$ signals that 
+\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)}
+belongs to a previous time window~$w'$,
+the rule in line~\ref{prg:base:ordered} % applies and 
+derives the atom
+\lstinline{order((}$j_1$\lstinline{,}$s_1$\lstinline{),(}$j_2$\lstinline{,}$s_2$\lstinline{),}$p_1$\lstinline{,}$w$\lstinline{)}
+to signal that \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} needs to be completed
+before performing \lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}, i.e.,
+the execution order must comply with the decomposition into time windows.
+The rule in lines \ref{prg:base:same_job:begin}-\ref{prg:base:same_job:end} yields
+a similar atom when \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)}
+is the predecessor operation $s_1 = s_2-1$ of \lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}
+in the same job $j_1=j_2$.
+In contrast to the cases in which 
+\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} must be processed before
+\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)},
+the choice rule in line~\ref{prg:base:order}
+allows for performing two operations sharing a machine in the lexicographic order
+of their jobs if the operations belong to the same time window.
+In case the atom representing execution in lexicographic order is not chosen,
+the rule in lines \ref{prg:base:order:begin}-\ref{prg:base:order:end}
+derives an atom signaling the inverse, given that the operations must not intersect and some sequence has to be determined.%
 %

### 2022-01-26 ICLP Paper/samplepaper.tex
encoding description extended

@@
 
-This section will show how to solve the problem dynamically, where the output of each iteration is an input to the next and merging them to obtain the solution of the whole problem.
+While the rules up to line~\ref{prg:base:order:end} yield atoms of the form
+\lstinline{order(}$o_1$\lstinline{,}$o_2$\lstinline{,}$p_1$\lstinline{,}$w$\lstinline{)},
+expressing the hard requirement or choice to perform an operation~$o_1$
+with processing time~$p_1$ before the operation~$o_2$ belonging to time window~$w$,
+the remaining rules assert corresponding DL constraints. % on the starting time of~$o_2$.
+To begin with,
+the starting times of operations~$o$ from the previous time window $w-1$ (if any)
+are in the lines~\ref{prg:diff_log:freez1:begin} and~\ref{prg:diff_log:freez2:begin}
+fixed to the value~$t$ in an (optimized) partial schedule for time window $w-1$, 
+as supplied by reified facts \lstinline{start(}$o${,}$t$\lstinline{,}$w-1$\lstinline{).}
+In line~\ref{prg:diff_log:non_nega:begin},
+the lower bound~$0$ is asserted for the starting time of the first operation of some job,
+included in the time window~$w$ for which a partial schedule is to be determined next.
+In addition, constraints reflecting the order of performing operations are imposed 
+in line~\ref{prg:diff_log:oper_const:begin},
+which concerns operations sharing a machine as well as successor operations within jobs.
+Since such constraints trace the sequence of operations in a job, they establish the
+earliest starting time, considered for problem decomposition in Section~\ref{subsec:decomposition},
+as lower bound for scheduling an operation, and the execution order on the machine
+processing the operation can increase its starting time further.
+The last rule of the \lstinline{step(w)} subprogram in lines
+\ref{prg:diff_log:oper_limit:begin}-\ref{prg:diff_log:oper_limit:end} asserts
+that the value for the DL variable \lstinline{makespan} cannot be less than the
+completion time of any operation of the time window~$w$.
+That is, the least feasible \lstinline{makespan} value provides the
+scheduling horizon of a partial schedule for operations of time windows up to~$w$.
+
+The task of optimizing the horizon of a (partial) schedule means choosing
+the execution order of operations of the latest time window sharing a machine
+such that the value for the \lstinline{makespan} variable is minimized.
+In single-shot ASP modulo DL solving with \clingodl, this can be accomplished
+via the command-line option \lstinline{--minimize-variable=makespan}.
+For the successive optimization by time windows, where the scheduling
+horizon gradually increases, we require a dedicated treatment of DL constraints
+limiting the value for \lstinline{makespan}.
+To this end, the \lstinline{optimize(m)} subprogram below line~\ref{prg:diff_log:oper_min}
+declares an external atom \lstinline{horizon(m)} for controlling whether a DL
+constraint asserted in line~\ref{prg:diff_log:oper_min:end} is active and limits
+the \lstinline{makespan} value to an integer supplied for the parameter~\lstinline{m}.
+When launching the optimization process for a new time window, any instance of the
+\lstinline{horizon(m)} atom introduced before is set to false for making sure that
+some (partial) schedule is feasible.
+Once some schedule with a horizon $h+1$ is found,
+the subprogram \lstinline{optimize(}$h$\lstinline{)} is instantiated on demand, i.e.,
+in case $h$ has not been passed as a value for~\lstinline{m} before, and
+the corresponding \lstinline{horizon(}$h$\lstinline{)} atom is set to true for activating the DL
+constraint reducing the admitted scheduling horizon to~$h$.
+The process of successively reducing the horizon~$h$ in order to find better partial schedules
+is repeated until the imposed \lstinline{makespan} value turns out to be infeasible,
+meaning that an optimal partial schedule has been found.
+As already mentioned above, the introduced instances of the external \lstinline{horizon(m)} atom
+are then set to false and the successive optimization proceeds by instantiating the
+\lstinline{step(}$w+1$\lstinline{)} subprogram for the next time window $w+1$ (if any), and
+also supplying the determined starting times of operations from time window~$w$
+by reified facts.
+The described control loop for successively extending good-quality partial schedules to
+a global solution is implemented by means of the Python interface of \clingodl,
+where a time limit can be configured to restrict the optimization efforts per time window
+and thus make sure that the iterative scheduling progresses.
+
+% This section will show how to solve the problem dynamically, where the output of each iteration is an input to the next and merging them to obtain the solution of the whole problem.
 In the first step, we consider the output of the decomposition phase, which is the assignment of the operations to TW as facts.
@@
 
-The next iteration is to optimize each TW by solving the subproblem(t) where $t$ refers to the current TW we aim to optimize.
+% The next iteration is to optimize each TW by solving the subproblem(t) where $t$ refers to the current TW we aim to optimize.
 The model starts to schedule and optimize the first TW and after a fixed amount of time, the solver stops and obtain the makespan of the current TW and the execution time of each operation represented by an atom \lstinline{startTime((Job, Step), ST, t)} where the first term is an operation, the second is the starting time and the last is the current time window.
@@
 
-The third part of the scheduler is shown in Listing \ref{prg:diff} that handles the difference constraints between the operations.
+% The third part of the scheduler is shown in Listing \ref{prg:diff} that handles the difference constraints between the operations.
 The first rule in lines \ref{prg:diff_log:non_nega:begin}-\ref{prg:diff_log:non_nega:end} ensures that all the operations will be processed starting from time $0$.

### 2022-01-26 ICLP Paper/samplepaper.tex
paragraph on schedule example

@@
 % \lstinputlisting[float=bt,mathescape=true,escapeinside={\#(}{\#)},basicstyle={\ttfamily\small},label=prg:diff,caption={diff-log}]{listing/diff_log.lp}
+
+For illustrating some phenomena going along with problem decomposition
+and iterative scheduling, let us inspect the schedule in Figure~\ref{fig:decomposed}
+that can be obtained with the decomposition into windows given in Listing~\ref{prg:tw}.
+The separation between the two time windows is indicated by bold double lines marking
+the completion of the latest operation of the first time window on each of the three machines.
+Notably, the partial schedule for the first time window as well as its extension to the
+second time window are optimal in terms of makespan.
+However, the obtained global solution has a makespan of~$21$ rather than just~$20$
+as for the optimal schedule in Figure~\ref{fig:schedule}.
+The reason is that the second operation of job~\lstinline{3} would need to be
+scheduled before the second operation of job~\lstinline{2} on machine~\lstinline{1},
+while the decomposition into time windows dictates the inverse order and necessitates
+the later completion of job~\lstinline{3}.
+Given the resulting buffer for scheduling operations with comparably short processing
+times on machine~\lstinline{3},
+the third operation of job~\lstinline{1} can be performed after the third operation
+of job~\lstinline{2} without deteriorating the makespan,
+yet introducing an unnecessary idle time of machine~\lstinline{3} from~$9$ to~$10$
+that could be avoided by choosing the inverse execution order.
+Even though this may seem negligible for the example instance at hand,
+idle slots can potentially propagate when a partial schedule gets extended to later time windows.
 %

### 2022-01-26 ICLP Paper/samplepaper.tex
implementation section completed

@@
 Even though this may seem negligible for the example instance at hand,
-idle slots can potentially propagate when a partial schedule gets extended to later time windows.
+idle slots can potentially propagate when a partial schedule gets extended to later time windows.%
 %
@@
 
+In order to counteract limitations of window-wise successive optimization
+due to "decomposition mistakes" and unnecessary idle times that do not directly
+affect the makespan to some extent,
+we have devised two additional techniques that can be incorporated into the
+iterative scheduling process.
+The first extension is time window \emph{overlapping}, where a configurable percentage
+of the operations per time window can still be rescheduled when proceeding
+to the next time window.
+To this end, an (optimized) partial schedule is postprocessed and the configured
+number of operations to overlap are picked in decreasing order of starting times.
+For example, if one operation from the first time window is supposed overlap for the
+decomposed schedule in Figure~\ref{fig:decomposed}, the second operation of job~\lstinline{2}
+whose starting time~\lstinline{4} is latest may be chosen,
+which in turn enables its processing after the second operation of job~\lstinline{3}
+by rescheduling together with operations of the second time window.
+The addition of overlapping operations to the encoding in Listing~\ref{prg:encoding}
+requires handling them similar to operations of the time window to schedule,
+i.e., enabling the choice of an execution order by the rules in lines
+\ref{prg:base:order}-\ref{prg:base:order:end} rather than fixing the order
+by the rule in line~\ref{prg:base:ordered}.
+As the second extension, we can make use of a stratified ASP program to
+postprocess an (optimized) partial schedule by inspecting operations of the
+latest time window in the order of starting times whether idle slots
+on their machines allow for an earlier execution.
+The starting times obtained by the \emph{compression} are then taken instead
+of those calculated in the actual schedule optimization, and, e.g.,
+the starting time~\lstinline{12} of the third operation of job~\lstinline{1}
+would be turned into~\lstinline{9} to fill the idle slot available on
+machine~\lstinline{3} when compressing the decomposed schedule displayed in Figure~\ref{fig:decomposed}.
+
 \section{Experiments}\label{sec:experiments}

### 2022-01-27 ICLP Paper/samplepaper.tex
table format adjusted (still partially)

@@
     \centering
-    \begin{tabular}{l l r r r r r r r r r r}
+    \begin{tabular}{r l r r r r r r r r r r}
     \hline
@@
     &  &  \multicolumn{10}{c}{Number of Time Windows} \\
-    Instances & Comparison & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
+    Size & Comparison & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
 
@@
 \begin{table}[t]
-    \caption{Results for three sizes of job shop scheduling instances with different numbers of time windows using static machine-based EST decomposition \label{tab:Table01}}%
+    \caption{%
+      Experiments varying the number of time windows on JSP benchmark sets of three sizes\label{tab:Table01}}%
+      % Results for three sizes of job shop scheduling instances with different numbers of time windows using static machine-based EST decomposition \label{tab:Table01}}%
 %    \mbox{~}\hfill
@@
     \centering
-    \begin{tabular}{l l r r r r r r r}
-    \hline
+    \begin{tabular}{r l r r r r r r r}
+    \topline%\hline
 %       & {}        & \multicolumn{6}{c}{Group of Instances}  \\
-    &  &  \multicolumn{7}{c}{Number of Time Windows} \\
-    Instances & Comparison & 1 & 2 & 3 & 4 & 5 & 6 & 10 \\
-
-    \hline\\[-2.75mm]
+%    &  &  \multicolumn{7}{c}{Number of Time Windows} \\
+    Size & Result & 1 & 2 & 3 & 4 & 5 & 6 & 10 % \\
+
+    \midline%\hline% \\[-2.75mm]
 
     {50$\times$15}       & Makespan             & $3542.1$   & $3149.7$	     &  \boldmath{$3083.8$} & $3110.7$ & $3213.6$ & $3225.3$ & $3524.9$ \\
-    & CPU(Sec)             & $1000.0$	& $1000.0$   & $1000.0$	 & $934.4$ & $615.5$  & $243.8$  & $18.5$  \\
-    & Interrupted Calls    & $1.0$	    & $2.0$      & $3.0$	& $3.6$  & $2.7$  & $1.0$  & $0.1$       \\[1.5mm]
+    & Time             & $1000.0$	& $1000.0$   & $1000.0$	 & $934.4$ & $615.5$  & $243.8$  & $18.5$  \\
+    & Interrupts    & $1.0$	    & $2.0$      & $3.0$	& $3.6$  & $2.7$  & $1.0$  & $0.1$       \\[1.5mm]
 
     {50$\times$20}       & Makespan             & $3879.4$   & $3479.4$	      & $3290.5$  &  \boldmath{$3260.3$} &  $3355.5$ & $3414.8$ & $3793.1$ \\
-    & CPU(Sec)             & $1000.0$	    & $1000.0$	   & $1000.0$	  & $944.9$  & $583.4$  & $321.5$   & $33.3$     \\
-    & Interrupted Calls    & $1.0$	    & $2.0$      & $3.0$	& $3.5$ & $2.5$  & $1.6$  & $0.1$      \\[1.5mm]
+    & Time             & $1000.0$	    & $1000.0$	   & $1000.0$	  & $944.9$  & $583.4$  & $321.5$   & $33.3$     \\
+    & Interrupts    & $1.0$	    & $2.0$      & $3.0$	& $3.5$ & $2.5$  & $1.6$  & $0.1$      \\[1.5mm]
 
     {100$\times$20}      & Makespan             & $46786.1$  &  $26210.4$  & $7201.5$ &  $6288.7$                & $6106.1$  &  \boldmath{$6031.9$}   & $6234.0$  \\
-    & CPU(Sec) & $1000.0$	    & $1000.0$	   & $1000.0$	   & $1000.0$	    & $1000.0$	& $1000.0$	& $861.8$            \\
-    & Interrupted Calls    & $1.0$	    & $2.0$      & $3.0$	& $4.0$ & $5.0$  & $6.0$ & $7.9$      \\[1.5mm]
-
-    \\[-2.75mm]
-    \hline
+    & Time & $1000.0$	    & $1000.0$	   & $1000.0$	   & $1000.0$	    & $1000.0$	& $1000.0$	& $861.8$            \\
+    & Interrupts    & $1.0$	    & $2.0$      & $3.0$	& $4.0$ & $5.0$  & $6.0$ & $7.9$      % \\%[1.5mm]
+
+    % \\[-2.75mm]
+    \botline%\hline
     \end{tabular}
@@
     \centering
-    \begin{tabular}{c c c c c c c c c c}
-    \hline
-
-\multirow{2}*{Strategy} & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
-    & Makespan & Time & IC & Makespan & Time & IC & Makespan & Time & IC \\
-    \hline\\[-2.75mm]
-    M-Based EST (S)             & $3083.8$  & $1000.0$   & $3.0$	& \boldmath{$3260.3$}      & $944.9$ & $3.5$ & $6031.9$ & $1000.0$ & $6.0$\\
+    \begin{tabular}{r r r r r r r r r r}
+    \topline% \hline
+
+Size & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
+Strategy & Makespan & Time & IC & Makespan & Time & IC & Makespan & Time & IC % \\
+    \midline% \hline\\[-2.75mm]
+    J-EST            & $3524.8$ & $1000.0$   & $2.0$	   & $3540.8$      & $1000.0$ & $2.0$ & $6964.4$ & $1000.0$ & $4.0$\\ 
+    J-MTWR            & $3456.2$  & $1000.0$   & $2.0$	& $3561.0$      & $1000.0$ & $2.0$ & $7149.6$ & $1000.0$ & $4.0$\\ 
     [1.5mm]
-
-    T-based EST (S)            & $3524.8$ & $1000.0$   & $2.0$	   & $3540.8$      & $1000.0$ & $2.0$ & $6964.4$ & $1000.0$ & $4.0$\\ 
+    M-EST             & $3083.8$  & $1000.0$   & $3.0$	& \boldmath{$3260.3$}      & $944.9$ & $3.5$ & $6031.9$ & $1000.0$ & $6.0$\\
+    M-MTWR        & \boldmath{$3060.3$}  & $1000.0$   & $3.0$	& $3287.9$      & $986.1$ & $3.9$ & \boldmath{$5948.6$} & $1000.0$ & $7.0$\\ 
     [1.5mm]
-
-    M-based MTWR (S)        & \boldmath{$3060.3$}  & $1000.0$   & $3.0$	& $3287.9$      & $986.1$ & $3.9$ & \boldmath{$5948.6$} & $1000.0$ & $7.0$\\ 
-    [1.5mm]
-
-    T-based MTWR (S)            & $3456.2$  & $1000.0$   & $2.0$	& $3561.0$      & $1000.0$ & $2.0$ & $7149.6$ & $1000.0$ & $4.0$\\ 
-    [1.5mm]
-
-    M-based EST (D)            & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $3.3$ & $6115.1$ & $1000.0$ & $6.9$\\ 
-    [1.5mm]
-
-    \\[-2.75mm]
-    \hline
-
+    M-EST$^\text{\,dyn}$            & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $3.3$ & $6115.1$ & $1000.0$ & $6.9$% \\[1.5mm]
+
+    % \\[-2.75mm]
+    \botline% \hline
     \end{tabular}

### 2022-01-27 ICLP Paper/samplepaper.tex
table formatting

@@
 %    &  &  \multicolumn{7}{c}{Number of Time Windows} \\
-    Size & Result & 1 & 2 & 3 & 4 & 5 & 6 & 10 % \\
+    Size & Average & 1 & 2 & 3 & 4 & 5 & 6 & 10 % \\
 
@@
 
-    {50$\times$15}       & Makespan             & $3542.1$   & $3149.7$	     &  \boldmath{$3083.8$} & $3110.7$ & $3213.6$ & $3225.3$ & $3524.9$ \\
+    {$50\times15$}       & Makespan             & $3542.1$   & $3149.7$	     &  \boldmath{$3083.8$} & $3110.7$ & $3213.6$ & $3225.3$ & $3524.9$ \\
     & Time             & $1000.0$	& $1000.0$   & $1000.0$	 & $934.4$ & $615.5$  & $243.8$  & $18.5$  \\
@@
 
-    {50$\times$20}       & Makespan             & $3879.4$   & $3479.4$	      & $3290.5$  &  \boldmath{$3260.3$} &  $3355.5$ & $3414.8$ & $3793.1$ \\
+    {$50\times20$}       & Makespan             & $3879.4$   & $3479.4$	      & $3290.5$  &  \boldmath{$3260.3$} &  $3355.5$ & $3414.8$ & $3793.1$ \\
     & Time             & $1000.0$	    & $1000.0$	   & $1000.0$	  & $944.9$  & $583.4$  & $321.5$   & $33.3$     \\
@@
 
-    {100$\times$20}      & Makespan             & $46786.1$  &  $26210.4$  & $7201.5$ &  $6288.7$                & $6106.1$  &  \boldmath{$6031.9$}   & $6234.0$  \\
+    {$100\times20$}      & Makespan             & $46786.1$  &  $26210.4$  & $7201.5$ &  $6288.7$                & $6106.1$  &  \boldmath{$6031.9$}   & $6234.0$  \\
     & Time & $1000.0$	    & $1000.0$	   & $1000.0$	   & $1000.0$	    & $1000.0$	& $1000.0$	& $861.8$            \\
@@
 \begin{table}[t]
-    \caption{Comparing different static/dynamic decomposition strategies \label{tab:Table02}}%
+    \caption{Experiments comparing Job- and Machine-based problem decomposition strategies\label{tab:Table02}}%
 %    \mbox{~}\hfill
@@
 
-Size & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
-Strategy & Makespan & Time & IC & Makespan & Time & IC & Makespan & Time & IC % \\
+Size & \multicolumn{3}{c}{$50\times15$} & \multicolumn{3}{c}{$50\times20$} & \multicolumn{3}{c}{$100\times20$}\\
+Strategy & Makespan & Time & Ints & Makespan & Time & Ints & Makespan & Time & Ints % \\
     \midline% \hline\\[-2.75mm]
@@
     [1.5mm]
-    M-EST$^\text{\,dyn}$            & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $3.3$ & $6115.1$ & $1000.0$ & $6.9$% \\[1.5mm]
+    M-EST$^\text{\,dyn}$            & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $4.0$ & $6115.1$ & $1000.0$ & $7.0$% \\[1.5mm]
 
@@
 \begin{table}[t]
-    \caption{Comparing results of best decomposition strategy with/without dompression and overlapping \label{tab:Table03}}%
+    \caption{Experiments comparing time window overlapping and compression techniques\label{tab:Table03}}%
 %    \mbox{~}\hfill
@@
     \centering
-    \begin{tabular}{c c c c c c c c c c}
-
-    \hline
-    \multicolumn{10}{c}{\textbf{Without Compressing Time Windows}}\\        
-\multirow{2}*{Overlap} & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
-    & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC \\
-    \hline\\[-2.75mm]
+    \begin{tabular}{r r r r r r r r r r}
+
+    \topline%\hline
+%     \multicolumn{10}{c}{\textbf{Without Compressing Time Windows}}\\        
+% \multirow{2}*{Overlap} & 
+    Size & \multicolumn{3}{c}{$50\times15$} & \multicolumn{3}{c}{$50\times20$} & \multicolumn{3}{c}{$100\times20$} \\
+    Overlap & Makespan & Time & Ints & Makespan & Time & Ints & Makespan & Time & Ints % \\
+    \midline % \hline\\[-2.75mm]
     0\%              & $3083.8$  & $1000.0$   & $3.0$	& $3260.3$      & $944.9$ & $3.5$ & $6031.9$ & $1000.0$ & $6.0$\\
-    [1.5mm]
+     %[1.5mm]
 
@@
 
-    30\%             & $3051.2$  & $1000.0$   & $3.0$	& $3451.8$      & $1000.0$ & $4.0$ & $6933.3$ & $1000.0$ & $6.0$\\ 
-                     [1.5mm]
-
-    \\[-2.75mm]
-
-    \multicolumn{10}{c}{\textbf{Compressing Time Windows}}\\
-    \hline
-    \multirow{2}*{Overlap} & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
-    & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC \\
-    \hline\\[-2.75mm]
+    30\%             & $3051.2$  & $1000.0$   & $3.0$	& $3451.8$      & $1000.0$ & $4.0$ & $6933.3$ & $1000.0$ & $6.0$ % \\ 
+%                     [1.5mm]
+%                     
+%    \\[-2.75mm]
+%    
+%    \multicolumn{10}{c}{\textbf{Compressing Time Windows}}\\
+%    \hline
+%    \multirow{2}*{Overlap} & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
+%     & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC \\
+%     \hline\\[-2.75mm]
+    \midline
     0\%              & $3013.3$  & $572.3$   & $3.7$	   & $3186.8$      & $540.8$ & $3.3$ & $5713.0$ & $589.2$ & $8.6$\\
-    [1.5mm]
+    % [1.5mm]
 
@@
 
-    30\%             & \boldmath{$2943.1$}  & $600.0$   & $4.0$	   & $3162.2$      & $600.0$ & $4.0$ & $5857.0$ & $600.0$ & $9.0$\\ 
-                     [1.5mm]
-    \hline
+    30\%             & \boldmath{$2943.1$}  & $600.0$   & $4.0$	   & $3162.2$      & $600.0$ & $4.0$ & $5857.0$ & $600.0$ & $9.0$%\\ 
+                     % [1.5mm]
+    \botline %\hline
     \end{tabular}

### 2022-01-27 ICLP Paper/samplepaper.tex
update 2nd table

@@
     M-EST             & $3083.8$  & $1000.0$   & $3.0$	& \boldmath{$3260.3$}      & $944.9$ & $3.5$ & $6031.9$ & $1000.0$ & $6.0$\\
-    M-MTWR        & \boldmath{$3060.3$}  & $1000.0$   & $3.0$	& $3287.9$      & $986.1$ & $3.9$ & \boldmath{$5948.6$} & $1000.0$ & $7.0$\\ 
+    M-MTWR        & $3060.3$  & $1000.0$   & $3.0$	& $3287.9$      & $986.1$ & $3.9$ & \boldmath{$5948.6$} & $1000.0$ & $7.0$\\ 
     [1.5mm]
-    M-EST$^\text{\,dyn}$            & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $4.0$ & $6115.1$ & $1000.0$ & $7.0$% \\[1.5mm]
+    M-EST$^\text{\,dyn}$            & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $4.0$ & $6115.1$ & $1000.0$ & $7.0$ \\
+    M-MTWR$^\text{\,dyn}$            & \boldmath{$3041.6$}  & $1000.0$   & $3.0$	& $3287.0$     & $980.4$ & $3.9$ & $6167.3$ & $1000.0$ & $5.0$
 

### 2022-01-27 ICLP Paper/Implementation/dlO.py

@@ -1,0 +1,129 @@
+#!/usr/bin/python
+import sys
+import clingo
+import theory
+import time
+from clingo import Function
+
+NUM_OF_TIME_WINDOWS = 2
+MAX_TIMEOUT = 1000
+
+class Application:
+    def __init__(self, name):
+        self.program_name = name
+        self.version = "1.0"
+        self.__theory = theory.Theory("clingodl", "clingo-dl")
+
+    def __on_model(self, model):
+        self.__theory.on_model(model)
+
+    def register_options(self, options):
+        self.__theory.register_options(options)
+
+    def validate_options(self):
+        self.__theory.validate_options()
+        return True
+
+    def __on_statistics(self, step, accu):
+        self.__theory.on_statistics(step, accu)
+        pass
+    
+    # get the Time out per Time Window
+    def get_timeout(self):
+        return MAX_TIMEOUT/NUM_OF_TIME_WINDOWS
+    # ************************************************
+
+    # get the assignment of the operations in a string format to be sent as facts for the next Time Window
+    def get_total_facts(self, assignment, i):
+        total_facts = ''
+        to_join = []
+        for name, value in assignment:
+            if str(name) != "makespan":
+                facts_format = "startTime({}, {}, {}). ".format(name, value, i)
+                to_join.append(facts_format)
+            else:
+                makespan = int(value)
+        total_facts = ''.join(to_join)
+        return total_facts, makespan
+    # ****************************************************************************************************
+
+    # get the part that should be grounded and solved
+    def step_to_ground(self, prg, step, total_facts):
+        parts = []
+        if step > 0:
+            parts.append(("subproblem", [step]))
+            if step > 1:
+                parts.append(("solutionTimeWindow", []))
+                prg.add("solutionTimeWindow", [], total_facts)
+        else:
+            parts.append(("base", []))
+        return parts
+    # ***********************************************
+
+    # add a new constraint to get lower value of makespan (Optimization Part)
+    def add_new_constraint(self, prg, makespan):
+        prg.cleanup()
+        prg.ground([("opt", [makespan-1])])
+        prg.assign_external(Function("makespan", [makespan-1]), True)
+    # ********************************************************************
+
+    def main(self, prg, files):
+        self.__theory.configure("propagate", "full,1")
+        self.__theory.register(prg)
+        if not files:
+            files.append("-")
+        for f in files:
+            prg.load(f)
+        timeout_for_window = self.get_timeout()
+        i, ret = 0, None
[…]

### 2022-01-27 ICLP Paper/samplepaper.tex
ASP paragraph in background

--- 2022-01-27 ICLP Paper/samplepaper.tex
@@
 \section{Preliminaries}\label{sec:preliminaries}
-\paragraph{Answer Set Programming (ASP)} is a declarative programming language for solving hard combinatorial optimization problems.
-ASP has become an established paradigm for knowledge representation and reasoning.
-ASP has proved its efficiency for solving the combinatorial optimization problem in different applications such as bioinformatics\cite{erdem2015generating,koponen2015optimizing}, databses\cite{caniupan2010consistency} and scheduling and industrial applications~\cite{dodaro2015allotment,dodaro2016combining,fabricius2020towards,dodaro2021operating}.
-
-A logic program is a finite set of rules of the form 
-\begin{equation}
-	a_0 \gets a_1, \ldots, a_m, \sim a_{m+1}, \dots, \sim a_n
-	\label{rule1}
-\end{equation}
-where $a_i$ is an atom for $0 \leq i \leq n$ and ``$\sim$''  is a default negation.
+
+\paragraph{Answer Set Programming (ASP)} \citep{lifschitz19a}
+is a knowledge representation and reasoning paradigm geared for the effective 
+modeling and solving of combinatorial (optimization) problems. % ASP has become an established paradigm for knowledge representation and reasoning. 
+% ASP has proved its efficiency for solving the combinatorial optimization problem in different applications such as bioinformatics\cite{erdem2015generating,koponen2015optimizing}, databses\cite{caniupan2010consistency} and scheduling and industrial applications~\cite{dodaro2015allotment,dodaro2016combining,fabricius2020towards,dodaro2021operating}.
+A (first-order) ASP program consists of \emph{rules} of the form
+% \begin{lstlisting}[numbers=none,frame=none]
+% h :- b#($_1$#), #(\dots#), b#($_n$#).
+% \end{lstlisting}
+\lstinline{h :- b}$_1$\lstinline{, }$\dots$\lstinline{, b}$_n$\lstinline{.},
+in which the head \lstinline{h} is an atom
+\lstinline{p(t}$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$\lstinline{)} or a
+choice
+\lstinline{{p(t}$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$\lstinline|)}|
+and each body literal \lstinline{b}$_i$ is an atom
+\lstinline{p(t}$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$\lstinline{)},
+possibly preceded by the default negation connective \lstinline{not} and/or
+followed by a condition
+\lstinline|: c|$_1$\lstinline{, }$\dots$\lstinline{, c}$_l$, 
+a built-in comparison \lstinline{t}$_1\circ{}$\lstinline{t}$_2$
+with $\circ\in\{\text{\lstinline{<}},\text{\lstinline{=}}\}$,
+or an aggregate
+\lstinline|t|$_0$\lstinline| = #count{t|$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$%
+\lstinline| : c|$_1$\lstinline{, }$\dots$\lstinline{, c}$_l$\lstinline|}|.
+Each \lstinline|t|$_j$ denotes a term, i.e., a constant, variable, tuple,
+or arithmetic expression, and each element \lstinline|c|$_k$ of a condition is
+an atom that may be preceded by \lstinline{not} or a built-in comparison.
+Roughly speaking, an ASP program is a shorthand for its ground instantiation,
+obtainable by substituting variables with all of the available constants
+and evaluating arithmetic expressions,
+and the semantics is given by \emph{answer sets}, i.e.,
+sets of (true) ground atoms such that all rules of the ground instantiation
+are satisfied and allow for deriving each of the ground atoms in the head of
+some rule whose body is satisfied.
+The syntax of the considered ASP programs is a fragment of the modeling languages
+described in \citep{cafageiakakrlemarisc20a,gehakalisc15a},
+the ground instantiation process is detailed in \citep{kamsch21a},
+and the answer set semantics is further elaborated in \citep{gehakalisc15a,lifschitz19a}.
+%
+% A logic program is a finite set of rules of the form 
+% \begin{equation}
+% 	a_0 \gets a_1, \ldots, a_m, \sim a_{m+1}, \dots, \sim a_n
+% 	\label{rule1}
+% \end{equation}
+% where $a_i$ is an atom for $0 \leq i \leq n$ and ``$\sim$''  is a default negation.
 If $n$ quals to $0$, then a rule \eqref{rule1} is a fact.
@@
 A \emph{literal} is either an atom or its negation.
-Given a rule $r$ of the form \eqref{rule1}, the set $H(r)=a_0$ denotes the \emph{head} atom and the set $B(r) = B^+(r) \cup B^-(r) = \{a_1,\dots,a_m\} \cup \{\sim a_{m+1}, \dots \sim a_n\}$ is the \emph{body} of $r$ where $B^+(r)$ and $B^+(r)$ represent the positive and negative body literals, respectively.
+% Given a rule $r$ of the form \eqref{rule1}, the set $H(r)=a_0$ denotes the \emph{head} atom and the set $B(r) = B^+(r) \cup B^-(r) = \{a_1,\dots,a_m\} \cup \{\sim a_{m+1}, \dots \sim a_n\}$ is the \emph{body} of $r$ where $B^+(r)$ and $B^+(r)$ represent the positive and negative body literals, respectively.
+
 \paragraph{Multi-shot Solving:}
@@
 \emph{clingo} systems enhances the ASP declaritive language with control capacities.
+
 \paragraph{ASP Difference Logic:}

### 2022-01-28 ICLP Paper/samplepaper.tex
working state to continue from

--- 2022-01-27 ICLP Paper/samplepaper.tex
@@
 
-\paragraph{Multi-shot Solving:}
+\paragraph{Multi-shot ASP solving} \citep{gekakasc17a}
+allows for iterative reasoning processes by controlling and interleaving
+the grounding and search phases of a stateful ASP system.
+For identifying a collection of rules to instantiate,
+the input language of \clingo supports \lstinline{#program name(c).}
+directives, where the constant~\lstinline{c} is a placeholder for some value, e.g.,
+the current time step in case of a planning problem,
+supplied upon instantiating the rules below such a directive.
+
 Multi-shot solving ASP allows solving logic programs dynamically in sequential order.
@@
     M-EST             & $3083.8$  & $1000.0$   & $3.0$	& \boldmath{$3260.3$}      & $944.9$ & $3.5$ & $6031.9$ & $1000.0$ & $6.0$\\
-    M-MTWR        & \boldmath{$3060.3$}  & $1000.0$   & $3.0$	& $3287.9$      & $986.1$ & $3.9$ & \boldmath{$5948.6$} & $1000.0$ & $7.0$\\ 
+    M-MTWR        & $3060.3$  & $1000.0$   & $3.0$	& $3287.9$      & $986.1$ & $3.9$ & \boldmath{$5948.6$} & $1000.0$ & $7.0$\\ 
     [1.5mm]
-    M-EST$^\text{\,dyn}$            & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $4.0$ & $6115.1$ & $1000.0$ & $7.0$% \\[1.5mm]
+    M-EST$^\text{\,d}$            & $3098.4$  & $1000.0$   & $3.0$	& $3309.5$     & $1000.0$ & $4.0$ & $6115.1$ & $1000.0$ & $7.0$ \\
+    M-MTWR$^\text{\,d}$            & \boldmath{$3041.6$}  & $1000.0$   & $3.0$	& $3287.0$     & $980.4$ & $3.9$ & $6167.3$ & $1000.0$ & $5.0$
 

### 2022-01-28 ICLP Paper/samplepaper.tex
background tuned

@@
 the grounding and search phases of a stateful ASP system.
-For identifying a collection of rules to instantiate,
+For referring to a collection of rules to instantiate,
 the input language of \clingo supports \lstinline{#program name(c).}
-directives, where the constant~\lstinline{c} is a placeholder for some value, e.g.,
-the current time step in case of a planning problem,
-supplied upon instantiating the rules below such a directive.
-
-Multi-shot solving ASP allows solving logic programs dynamically in sequential order.
+directives, where \lstinline{name} denotes a \emph{subprogram} comprising
+the rules below such a directive and the parameter~\lstinline{c} is a
+placeholder for some value, e.g., the current time step in case of a
+planning problem, supplied upon instantiating the subprogram.
+Moreover, \lstinline{#external h : b}$_1$\lstinline{, }$\dots$\lstinline{, b}$_n$\lstinline{.}
+statements are formed similar to rules yet declare an atom~\lstinline{h} as
+\emph{external} when the body is satisfied:
+such an external atom can be freely set to true or false by means of the
+Python interface of \clingo,
+so that rules including it in the body can be selectively (de)activated
+to direct the search.
+%
+% Multi-shot solving ASP allows solving logic programs dynamically in sequential order.
 This can be manipulated using APIs implementation via an imperative programming language.
@@
 
-\paragraph{ASP Difference Logic:}
-\emph{clingo}[DL] is an extension of the input language \emph{clingo} by theory atoms representing difference constraints \cite{gebser2016theory,janhunen2017clingo}.
+\paragraph{ASP modulo DL}
+integrates DL constraints \citep{cotmal06a}, i.e., expressions written as
+\lstinline|&diff{t|$_1$\lstinline| - t|$_2$\lstinline|} <= t|$_3$, in the head of rules.
+With the exception of the constant~\lstinline{0}, which denotes the number zero,
+the terms \lstinline|t|$_1$ and \lstinline|t|$_2$ represent DL variables that can be
+assigned any integer value.
+However, the difference \lstinline|t|$_1$\lstinline| - t|$_2$ must not exceed the
+integer constant~\lstinline|t|$_3$ if the body of a rule with the DL constraint in the
+head is satisfied.
+That is, the DL constraints asserted by rules whose body is satisfied restrict the
+feasible values for DL variables,
+and the \clingodl system extends \clingo by checking the consistency of DL constraints
+imposed by an answer set.
+If these DL constraints are satisfiable, a canonical assignment of smallest feasible
+integer values to DL variables can be determined in polynomial time and is output
+together with the answer set.
+%
+% \emph{clingo}[DL] is an extension of the input language \emph{clingo} by theory atoms representing difference constraints \cite{gebser2016theory,janhunen2017clingo}.
 Difference constraints are defined by specific constraint atoms of the form $\text{\lstinline{&diff}}\{u-v\} \leq k$ where $u$ and $v$ are terms which are interpreted as integer variables and $k$ is an integer constant.
 \emph{clingo}[DL] provides the following extension of the normal form \eqref{rule1}:
-\begin{align*}
-	\text{\lstinline{&diff}}\{u-v\} \leq k \gets a_1, \ldots, a_m, \sim a_{m+1}, \ldots, \sim a_n.
-\end{align*}
- This rule shows that whenever the body is $True$ the inequality constraints must be satisfied.
+% \begin{align*}
+% 	\text{\lstinline{&diff}}\{u-v\} \leq k \gets a_1, \ldots, a_m, \sim a_{m+1}, \ldots, \sim a_n.
+% \end{align*}
+%  This rule shows that whenever the body is $True$ the inequality constraints must be satisfied.
 

### 2022-01-28 ICLP Paper/samplepaper.tex
started with experiments section

@@
 \usepackage{url}
+\urldef\myurl\url{https://github.com/prosysscience/Job-Shop-Scheduling/tree/master/ICLP%20Paper/Implementation}
 
@@
 \section{Experiments}\label{sec:experiments}
-This section will show the computational experiments that we conducted to test the performance of our proposed model.
-In order to evaluate our model, we performed several experiments on a set of benchmark instances with a different number of jobs and machines.
+% This section will show the computational experiments that we conducted to test the performance of our proposed model.
+In order to evaluate our multi-shot ASP modulo DL approach to JSP solving,
+we ran experiments on JSP benchmark sets by \cite{taillard1993benchmarks}
+of several sizes,
+considering ten instances each with
+$50\times15$, $50\times20$, or $100\times20$ jobs and machines.
+The instances are generated such that each job includes one operation per machine, e.g.,
+$15$ operations in case of the $50\times15$ instances,
+where the sequence in which the operations of a job are allocated to machines varies.
+In our experiments, we assess decomposition strategies presented in Section~\ref{subsec:decomposition}
+with different number and size of time windows as well as the impact of
+time window overlapping and compression techniques described in Section~\ref{subsec:encoding}.
+For the comparability of results between runs with a different number of time windows
+and respective optimization subproblems addressed by multi-shot solving,
+we divide the total runtime limit of $1000$ seconds for \clingodl (version 1.3.0) by the number
+of time windows to evenly spend optimization efforts on subproblems.
+Our experiments have been run on an 
+Intel\textsuperscript{\textregistered} Core\texttrademark{} i7-8650U CPU % \textsuperscript{\textregistered}
+Dell Latitude 5590 machine under Windows 10.%
+\footnote{The benchmarks and our implementation are available at: \myurl}
+
+% a set of benchmark instances with a different number of jobs and machines.
 The benchmark instances we considered are generated in 1993 by E.

### 2022-01-29 ICLP Paper/samplepaper.tex
extended text on results in Table 1

@@
 \end{table}
-
-Table \ref{tab:Table01} shows the results of solving the mentioned different sizes before with a machine-based EST decomposition method.
-The columns represent how many windows split the original problem where TW $1$ refers that the problem is not decomposed.
+%
+Our first series of experiments shown in Table~\ref{tab:Table01} concerns the number
+of time windows into which the $750$, $1000$, or $2000$ operations, respectively,
+of the instances with $50\times15$, $50\times20$, and $100\times20$ jobs and machines
+are split.
+We use the static Machine-based EST decomposition strategy, which augments the encoding
+given in Listing~\ref{prg:decomposition} with bottleneck machines for inserting 
+their operations into time windows, as baseline because it leads to more robust
+solution quality than Job-based problem decomposition.
+For each of the three problem sizes,
+we report the average makespan over ten instances,
+where smaller values indicate better schedules,
+the average runtime of \clingodl, and the
+average number of interrupted optimization runs on subproblems.
+We gradually increase the number of time windows from~$1$ to~$6$ and additionally
+include results for~$10$ time windows to outline the trend of degrading
+solution quality when the partition into time windows becomes too fragmented
+for merging (optimized) partial schedules into proficient global solutions.
+In fact, the shortest makespans, highlighted in boldface,
+are obtained with problem decomposition into
+$3$, $4$, or $6$ time windows, respectively, for the instances of size 
+$50\times15$, $50\times20$, and $100\times20$.
+Compared to the results for $1$ time window,
+which represent global optimization on the full problem up to the
+runtime limit of $1000$ seconds,
+we observe substantial makespan improvements, amounting to almost one order
+of magnitude for the largest instances with $100\times20$ jobs and machines.
+These advantages do not come as a surprise considering that the JSP instances
+are highly combinatorial \citep{shysha18a} and each global optimization run
+times out with a more or less optimized solution, 
+where a good-quality schedule is the more difficult to find the greater the
+number of operations gets.
+Except for about 12\% completed optimization runs in multi-shot solving with
+$4$ time windows on the $50\times20$ instances,
+the best schedules are obtained with optimization runs timing out on subproblems,
+yet with good-quality partial schedules that can be merged into better global
+solutions than achievable with more time windows, which
+become smaller and easier to solve but also cut off good-quality
+schedules.
+
+% Table \ref{tab:Table01} shows the results of solving the mentioned different sizes before with a machine-based EST decomposition method.
+% The columns represent how many windows split the original problem where TW $1$ refers that the problem is not decomposed.
 The results of $7, 8$ and $9$ TWs have not been presented because they have not changed the tendency of the results.
@@
 Therefore, The average running time for 10 TWs is $18.5$ seconds.
-However, the average makespan is getting worse.\\
-
-The number of windows with the best obtained average makespan $3260.3$ for scheduling $50\times20$ is $4$ with an average timeout of $944.9$.
+However, the average makespan is getting worse.
+% 
+% The number of windows with the best obtained average makespan $3260.3$ for scheduling $50\times20$ is $4$ with an average timeout of $944.9$.
 We observe that assigning $250$ operations to each window leads to getting the best average makespan for the first two sets.

### 2022-01-29 ICLP Paper/samplepaper.tex
intro

--- 2022-01-28 ICLP Paper/samplepaper.tex
@@
 \usepackage{url}
-
+\usepackage{enumitem}
 \usepackage{comment}
@@
 Hence, more efficient approaches to approximate good-quality schedules 
-instead of striving for optimal solutions have attracted wide research interest.
+instead of striving for optimal solutions have attracted broad research interest.
 On the one hand, respective methods include greedy and local search techniques such as
@@
 
-In this paper, we greatly extend our preliminary study \citep{elkgeb20a} 
+In this paper, we significantly extend our preliminary study \citep{elkgeb20a} 
 on problem decomposition into time windows and successive schedule optimization
-by means of multi-shot ASP modulo DL solving with \clingodl.
+through multi-shot ASP modulo DL solving with \clingodl.
 The goal of the decomposition is to split highly complex scheduling tasks into
 balanced portions for which partial schedules of good quality can be reliably
-found within tight runtime limits, and then be merged into a global solution of
-significantly better quality than obtainable in similar runtime 
-with single-shot optimization on the full problem.
-In this process, problem decomposition must respect the precedence of operations
-within their encompassing jobs to guarantee the feasibility of schedules, and as
-a secondary objective, the order by time windows of operations sharing a machine should come close to the sequences of operations in (unknown) optimal schedules.
+found within tight runtime limits.
+Then, the partial schedules are merged into a global solution of significantly better quality than obtainable in similar runtime with single-shot optimization on the entire problem.
+In this process, problem decomposition must satisfy two criteria: 
+\begin{enumerate*}[label=\emph{(\roman*)}]
+  \item the precedence of operations within their encompassing jobs to guarantee the feasibility of partial schedules, and 
+  \item the ordering by time windows of operations sharing a machine should come close to the sequences of operations in (unknown) optimal schedules.
+\end{enumerate*}
 We address computational efficiency as well as solution quality by devising and investigating decomposition strategies regarding the size of
@@
 
-The contributions of our work going beyond the study in \citep{elkgeb20a} are:
+This work goes beyond the study of \cite{elkgeb20a} and makes the following contributions:
 \begin{itemize}
@@
       ASP programs (without DL constraints).
-\item Considering that a decomposition into time windows may be incompatible with
+\item Since a decomposition into time windows may be incompatible with
       the optimal sequences of operations sharing a machine, we incorporate
-      time window overlapping into the iterative scheduling process to
-      offer a chance for revising "decomposition mistakes".
-      Moreover, as the makespan objective we apply for the optimization of
-      (partial) schedules tolerates unnecessary idle times of machines as long as
-      they do not yield a greater scheduling horizon,
-      we devise a stratified ASP encoding to postprocess and compress partial
-      schedules by reassigning operations to earlier idle slots available on their machines.
+      overlapping time windows into the iterative scheduling process to
+      offer a chance for revising ``decomposition mistakes''.
+      Moreover, the makespan objective, which we apply to optimize (partial) schedules, tolerates unnecessary idle times of machines as long as they do not yield a greater scheduling horizon. 
+      Therefore, we devise a stratified ASP encoding to postprocess and compress partial schedules by reassigning operations to earlier idle slots available on their machines.
 \item We experimentally evaluate decomposition strategies varying the number of
-      operations per time window as well as multi-shot ASP solving
+      operations per time window, as well as multi-shot ASP, solving
       augmented with overlapping and compression techniques
@@
 
-The rest of this 
-paper is organized as follows.
+The paper is organized as follows:
 Section~\ref{sec:preliminaries} briefly introduces ASP along with the relevant
@@
 In Section~\ref{sec:problem}, we present our successive optimization approach
-and detail the ASP programs encoding problem decomposition or iterative scheduling by
-time windows, respectively.
-Section~\ref{sec:experiments} provides experimental results on JSP benchmark sets, % of several sizes,
-assessing different decomposition strategies as well as the impact of overlapping and
-compression techniques.
+and detail the ASP programs encoding problem decomposition or iterative scheduling by time windows, respectively.
+Section~\ref{sec:experiments} provides experimental results on JSP benchmark sets, assessing different decomposition strategies as well as the impact of overlapping and compression techniques.
 Conclusions and future work are discussed in Section~\ref{sec:conclusions}.  
@@
 the grounding and search phases of a stateful ASP system.
-For referring to a collection of rules to instantiate,
+For identifying a collection of rules to instantiate,
 the input language of \clingo supports \lstinline{#program name(c).}
-directives, where \lstinline{name} denotes a \emph{subprogram} comprising
-the rules below such a directive and the parameter~\lstinline{c} is a
-placeholder for some value, e.g., the current time step in case of a
-planning problem, supplied upon instantiating the subprogram.
-Moreover, \lstinline{#external h : b}$_1$\lstinline{, }$\dots$\lstinline{, b}$_n$\lstinline{.}
-statements are formed similar to rules yet declare an atom~\lstinline{h} as
-\emph{external} when the body is satisfied:
-such an external atom can be freely set to true or false by means of the
-Python interface of \clingo,
-so that rules including it in the body can be selectively (de)activated
-to direct the search.
-%
-% Multi-shot solving ASP allows solving logic programs dynamically in sequential order.
+directives, where the constant~\lstinline{c} is a placeholder for some value, e.g.,
+the current time step in case of a planning problem,
+supplied upon instantiating the rules below such a directive.
+
+Multi-shot solving ASP allows solving logic programs dynamically in sequential order.
 This can be manipulated using APIs implementation via an imperative programming language.
@@
 
-\paragraph{ASP modulo DL}
-integrates DL constraints \citep{cotmal06a}, i.e., expressions written as
-\lstinline|&diff{t|$_1$\lstinline| - t|$_2$\lstinline|} <= t|$_3$, in the head of rules.
-With the exception of the constant~\lstinline{0}, which denotes the number zero,
-the terms \lstinline|t|$_1$ and \lstinline|t|$_2$ represent DL variables that can be
-assigned any integer value.
-However, the difference \lstinline|t|$_1$\lstinline| - t|$_2$ must not exceed the
-integer constant~\lstinline|t|$_3$ if the body of a rule with the DL constraint in the
-head is satisfied.
-That is, the DL constraints asserted by rules whose body is satisfied restrict the
-feasible values for DL variables,
-and the \clingodl system extends \clingo by checking the consistency of DL constraints
-imposed by an answer set.
-If these DL constraints are satisfiable, a canonical assignment of smallest feasible
-integer values to DL variables can be determined in polynomial time and is output
-together with the answer set.
-%
-% \emph{clingo}[DL] is an extension of the input language \emph{clingo} by theory atoms representing difference constraints \cite{gebser2016theory,janhunen2017clingo}.
+\paragraph{ASP Difference Logic:}
+\emph{clingo}[DL] is an extension of the input language \emph{clingo} by theory atoms representing difference constraints \cite{gebser2016theory,janhunen2017clingo}.
 Difference constraints are defined by specific constraint atoms of the form $\text{\lstinline{&diff}}\{u-v\} \leq k$ where $u$ and $v$ are terms which are interpreted as integer variables and $k$ is an integer constant.
 \emph{clingo}[DL] provides the following extension of the normal form \eqref{rule1}:
-% \begin{align*}
-% 	\text{\lstinline{&diff}}\{u-v\} \leq k \gets a_1, \ldots, a_m, \sim a_{m+1}, \ldots, \sim a_n.
-% \end{align*}
-%  This rule shows that whenever the body is $True$ the inequality constraints must be satisfied.
+\begin{align*}
+	\text{\lstinline{&diff}}\{u-v\} \leq k \gets a_1, \ldots, a_m, \sim a_{m+1}, \ldots, \sim a_n.
+\end{align*}
+ This rule shows that whenever the body is $True$ the inequality constraints must be satisfied.
 

### 2022-01-29 ICLP Paper/samplepaper.tex

--- 2022-01-29 ICLP Paper/samplepaper.tex
@@
 
+\usepackage{enumitem}
 \usepackage{multirow}

### 2022-01-29 ICLP Paper/samplepaper.tex
text on Table 2 revised

--- 2022-01-29 ICLP Paper/samplepaper.tex
@@
 \end{table}
+%
+In Table~\ref{tab:Table02}, we compare other problem decomposition strategies
+to the static Machine-based EST approach, providing the baseline abbreviated by M-EST.
+Further decomposition strategies are denoted by M-MTWR for the static Machine-based MTWR
+variant, corresponding dynamic versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$ that
+determine EST and MTWR values w.r.t.\ partial schedules from previous time windows, as well as
+static Job-based counterparts J-EST and J-MTWR in the first two rows.
+For each decomposition strategy and JSP instance size, we present results for the number
+of time windows,
+which amounts to the next greater integer of the average number of interrupted optimization
+runs in columns headed by Ints, leading to the shortest average makespan.
+The considerably increased average makespans in the first two rows clearly indicate that
+time windows determined with Job-based decomposition strategies are less adequate than
+those investigating bottleneck machines in the first place, and then picking their operations
+based on the smallest EST or greatest MTWR value as secondary criterion.
+While Machine-based decomposition turns out to be advantageous,
+there is no clear gap between the static M-EST and M-MTWR variants as well as their dynamic
+versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$.
+That is, the heuristic to schedule operations from highly loaded machines
+as soon as possible works well for decomposing the considered JSP instances,
+yet further solution quality differences due to minor deviations in the
+decomposition strategy or optimization performance of \clingodl are negligible.
 
@@
 
-Table \ref{tab:Table02} demonstrates the results of the four static decomposition strategies and the machine-based MTWR with the dynamic assignment.
+% Table \ref{tab:Table02} demonstrates the results of the four static decomposition strategies and the machine-based MTWR with the dynamic assignment.
 Every three sub-columns have illustrated the results of a set of the benchmark instances and each row refers to a decomposition strategy with static (S) or dynamic (D) assignment.

### 2022-01-29 ICLP Paper/samplepaper.tex
experiments section completely revised

@@
 \end{table}
+%
+Focusing again on the static M-EST decomposition strategy,
+Table~\ref{tab:Table03} assesses different percentages of overlapping
+operations from previous time windows as well as the impact of compressing
+(optimized) partial schedules in a postprocessing phase.
+The baseline approach as above is included in the first row with 0\% of the
+operations per time window overlap and without the use of compression.
+The next three rows exclude compression as well but introduce 10\%, 20\%, or
+30\% of overlapping operations that can be rescheduled together with operations
+of the next time window.
+In view of the considered number of time windows, picked according to the shortest
+average makespan achieved with 0\% overlap, between $18$ and $35$ operations constitute
+an overlap of 10\%, so that 30\% range up to more than $100$ operations.
+Such non-negligible amounts can make the optimization of subproblems for time windows
+harder, while the moderate overlaps of 10\% or 20\% manage to consistently
+improve the solution quality.
+Even more significant advantages are obtained by applying compression in the last
+four rows,
+where the average makespan gets already shorter with 0\% overlap.
+Further improvements of roughly 2\% in solution quality are achieved with 20\% overlap,
+which yield consistently better schedules than the smaller overlap of 10\% and
+avoid deteriorations as observed with 30\% overlap on the largest instances
+with $100\times20$ jobs and machines.
+In summary, a Machine-based problem 
+decomposition strategy along with
+postprocessing to compress partial schedules as well as a modest overlap of about
+20\% of the time window size turn out to be most successful for the JSP
+benchmark sets under consideration.
 
@@
 %The best average makespan of $50\times15$ and $100\times20$ are  $3060.3$, and $5948.6$.
-We applied the overlap between the windows to repair the mistakes generated by the decomposition; in addition, we compressed each optimized window to ensure that all the operations had been started as early as possible.
+% We applied the overlap between the windows to repair the mistakes generated by the decomposition; in addition, we compressed each optimized window to ensure that all the operations had been started as early as possible.
 Table \ref{tab:Table03} summarizes the results with/without the compression and the overlap.

### 2022-01-29 ICLP Paper/samplepaper.tex
Section 3.2

--- 2022-01-29 ICLP Paper/samplepaper.tex
@@
 
-This work goes beyond the study of \cite{elkgeb20a} and makes the following contributions:
+The contributions of our work going beyond the study in \citep{elkgeb20a} are:
 \begin{itemize}
@@
 followed by a condition
-\lstinline|: c|$_1$\lstinline{, }$\dots$\lstinline{, c}$_l$, 
+\lstinline|: c|$_1$\lstinline{,}$\dots$\lstinline{,c}$_l$, 
 a built-in comparison \lstinline{t}$_1\circ{}$\lstinline{t}$_2$
@@
 \lstinline|t|$_0$\lstinline| = #count{t|$_1$\lstinline{,}$\dots$\lstinline{,t}$_m$%
-\lstinline| : c|$_1$\lstinline{, }$\dots$\lstinline{, c}$_l$\lstinline|}|.
+\lstinline| : c|$_1$\lstinline{,}$\dots$\lstinline{,c}$_l$\lstinline|}|.
 Each \lstinline|t|$_j$ denotes a term, i.e., a constant, variable, tuple,
@@
 planning problem, supplied upon instantiating the subprogram.
-Moreover, \lstinline{#external h : b}$_1$\lstinline{, }$\dots$\lstinline{, b}$_n$\lstinline{.}
+Moreover, \lstinline{#external h : b}$_1$\lstinline{,}$\dots$\lstinline{,b}$_n$\lstinline{.}
 statements are formed similar to rules yet declare an atom~\lstinline{h} as
@@
 For instance, \emph{operation(1,1)} is processed by machine 2.
-\lstinputlisting[float=t,label=prg:facts,caption={Example JSP instance},linerange={1-3},numbers=none,xleftmargin=3pt]{listing/facts.lp}
-
-Figure~\ref{fig:schedule} depicts an optimal schedule in terms of makespan,
-i.e., the latest completion time of any job/operation,
-for the JSP instance from Listing~\ref{prg:facts}.
+\lstinputlisting[float=b,label=prg:facts,caption={Example JSP instance},linerange={1-3},numbers=none,xleftmargin=3pt]{listing/facts.lp}
+
+Figure~\ref{fig:schedule} depicts a schedule with the optimal makespan, i.e., the latest completion time of any job/operation, for the JSP instance from Listing~\ref{prg:facts}.
 The \textbf{J}-\textbf{S} pairs in horizontal bars indicated for the
@@
 intersect, so that the operations are performed in sequential order.
-Moreover, operations belonging to the same job are scheduled one after the other.
+Moreover, operations belonging to the same job are scheduled one after another.
 For example,
@@
 enable a successive extension of good-quality partial schedules to a global solution.
-In the following, we first detail problem decomposition based on earliest starting
-times of operations, and then outline further strategies that can be encoded
-by stratified ASP programs as well.
-
-\lstinputlisting[float=t,label=prg:decomposition,caption={Job-based EST decomposition encoding}]{listing/decomposition.lp}
-%
-Our encoding for \emph{Job-based Earliest Starting Time (EST)} decomposition
+In the following, we first detail problem decomposition based on the earliest starting times of operations, and then outline further strategies that can be encoded by stratified ASP programs as well.
+
+\lstinputlisting[float=t,label=prg:decomposition,caption={J-EST decomposition encoding}]{listing/decomposition.lp}
+% Only 
+Our encoding for \emph{Job-based Earliest Starting Time (J-EST)} decomposition
 in Listing~\ref{prg:decomposition} takes a JSP instance specified by facts over
@@
 The rules in lines~\ref{prg:assignment_process:est:begin} and~\ref{prg:assignment_process:est:end}
-encode the EST calculation per operation of a job,
+encode the J-EST calculation per operation of a job,
 given by the sum of processing times for predecessor operations belonging to the same job.
@@
 (ii) the same earliest starting time and shorter processing time, or
-(iii) a smaller job identifier as tie breaker in case of similar
+(iii) a smaller job identifier as a tie-breaker in case of similar
 earliest starting and processing times.
@@
 indexes increase according to the precedence of operations within their jobs,
-given that earliest starting times grow along the sequence of operations in a job.
+given that the earliest starting times grow along the sequence of operations in a job.
 
@@
 
-In addition to Job-based EST decomposition,
-we have devised a similar ASP program for
-\emph{Job-based Most Total Work Remaining (MTWR)} decomposition,
+In addition to J-EST decomposition, we have devised a similar ASP program for
+\emph{Job-based Most Total Work Remaining (J-MTWR)} decomposition,
 where the total order of operations is decreasing by the sum of processing
@@
 matching the time for executing all three operations of each job.
-Hence, the first operation of job~\lstinline{3} is considered most
-important and associated with the index~\lstinline{0}, and the other
-operations follow in the total order taken for partitioning into time windows.
-As MTWR values are decreasing along the sequence of operations in a job, the obtained time windows
-also respect the precedence of operations.
-
-Instead of assessing EST and MTWR values in a purely job-based fashion,
-we have encoded \emph{Machine-based} decomposition in which an operation
-from a bottleneck machine with greatest sum of processing times for yet
-unordered operations is considered next.
+Hence, the first operation of job~\lstinline{3} is considered as the most
+important and is associated with the index~\lstinline{0}. The other
+operations follow the total order taken for partitioning into time windows.
+As J-MTWR values are decreasing along the sequence of operations in a job, the obtained time windows also respect the precedence of operations.
+
+In addition to assessing decompositions in a purely Job-based fashion,
+we have encoded \emph{Machine-based} decompositions \emph{M-EST} and \emph{M-MTRW} in which an operation from a bottleneck machine with the greatest sum of processing times for yet unordered operations is considered next.
 For our example instance, the sum of processing times for operations 
@@
 and already scheduled operations from previous time windows may increase the
-EST and MTWR values used to order the yet unscheduled operations.
+EST and MTWR values are used to order the yet unscheduled operations.
 However, the principle of applying a stratified ASP program to determine
@@
 \end{table}
-%
-In Table~\ref{tab:Table02}, we compare other problem decomposition strategies
-to the static Machine-based EST approach, providing the baseline abbreviated by M-EST.
-Further decomposition strategies are denoted by M-MTWR for the static Machine-based MTWR
-variant, corresponding dynamic versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$ that
-determine EST and MTWR values w.r.t.\ partial schedules from previous time windows, as well as
-static Job-based counterparts J-EST and J-MTWR in the first two rows.
-For each decomposition strategy and JSP instance size, we present results for the number
-of time windows,
-which amounts to the next greater integer of the average number of interrupted optimization
-runs in columns headed by Ints, leading to the shortest average makespan.
-The considerably increased average makespans in the first two rows clearly indicate that
-time windows determined with Job-based decomposition strategies are less adequate than
-those investigating bottleneck machines in the first place, and then picking their operations
-based on the smallest EST or greatest MTWR value as secondary criterion.
-While Machine-based decomposition turns out to be advantageous,
-there is no clear gap between the static M-EST and M-MTWR variants as well as their dynamic
-versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$.
-That is, the heuristic to schedule operations from highly loaded machines
-as soon as possible works well for decomposing the considered JSP instances,
-yet further solution quality differences due to minor deviations in the
-decomposition strategy or optimization performance of \clingodl are negligible.
 
@@
 
-% Table \ref{tab:Table02} demonstrates the results of the four static decomposition strategies and the machine-based MTWR with the dynamic assignment.
+Table \ref{tab:Table02} demonstrates the results of the four static decomposition strategies and the machine-based MTWR with the dynamic assignment.
 Every three sub-columns have illustrated the results of a set of the benchmark instances and each row refers to a decomposition strategy with static (S) or dynamic (D) assignment.

### 2022-01-29 ICLP Paper/samplepaper.tex
conclusions updated

--- 2022-01-29 ICLP Paper/samplepaper.tex
@@
 \section{Conclusions}\label{sec:conclusions}
-In this work, different decomposition strategies have been proposed for solving the Job-shop Scheduling Problem (JSP) using a multi-shot Answer Set Programming with Difference Logic constraints to handle the decomposition approach and the problem constraints.
+%
+Our work develops multi-shot ASP modulo DL methods for JSP solving,
+based on problem decomposition into balanced time windows that respect the operation
+precedence within jobs and successive makespan optimization while extending
+partial schedules to a global solution.
+We demonstrate that splitting highly complex JSP instances into
+balanced portions, for which partial schedules of good quality can be reliably
+found within tight runtime limits, leads to significantly better global solutions
+than obtainable with single-shot ASP modulo DL optimization in similar runtime.
+Related works taking advantage of single- \citep{abjoossctowa21a}
+and multi-shot solving \citep{francescutto2021solving} by
+\clingodl to tackle real-world scheduling problems evaluate threshold values
+on DL variables and perform optimization by means of common 
+\lstinline{#minimize} statements of \clingo rather than minimizing the
+value of a DL variable like \lstinline{makespan} in Listing~\ref{prg:encoding}.
+Applying a similar approach can be advantageous in our future work as it may
+potentially improve computational efficiency as well as solution quality further
+and is even a necessity to switch from the makespan as optimization objective to tardiness
+w.r.t.\ deadlines, which are crucial in many application areas.
+Moreover, we aim to generalize our methods further, e.g.,
+by considering flexible machine allocations and varying processing times,
+and scale them up to large-scale instances with an order of magnitude more
+operations than in synthetic benchmarks.
+%
+% In this work, different decomposition strategies have been proposed for solving the Job-shop Scheduling Problem (JSP) using a multi-shot Answer Set Programming with Difference Logic constraints to handle the decomposition approach and the problem constraints.
 The proposed model split the problem into sub-problems using different decomposition strategies and then optimized each separately while minimizing the Makespan.

### 2022-01-30 ICLP Paper/samplepaper.tex
minor reformatting
section 3.2

@@
 % \end{lstlisting}
-\lstinline{h :- b}$_1$\lstinline{, }$\dots$\lstinline{, b}$_n$\lstinline{.},
+\lstinline{h :- b}$_1$\lstinline{,}$\dots$\lstinline{,b}$_n$\lstinline{.},
 in which the head \lstinline{h} is an atom
@@
 times for an operation and its successors in a job.
-For example, we obtain the MTWR values~\lstinline{7}, \lstinline{12}, and~\lstinline{20}
+For example, we obtain the J-MTWR values~\lstinline{7}, \lstinline{12}, and~\lstinline{20}
 for the first operation of job~\lstinline{1}, \lstinline{2}, or~\lstinline{3},
@@
 Hence, the operation to be inserted into the total order first is picked from
-machine~\lstinline{2}, where the smallest EST or greatest MTWR value is
+machine~\lstinline{2}, where the smallest M-EST or greatest M-MTWR value is
 used for choosing a particular job/operation.
-Unlike Job-based decomposition, this may lead to the choice of an operation
+Unlike Job-based decompositions, this may lead to the choice of an operation
 such that predecessor operations processed by other machines are yet unordered.
 In this case, they are inserted into the total order directly before the
-operation from a bottleneck machine,
-and all machine loads are reduced by processing times
-before determining the next machine to consider.
-
-Moreover, we have devised dynamic versions of the Job- and Machine-based 
+operation from a bottleneck machine, and all machine loads are reduced by processing times before determining the next machine to consider.
+
+Moreover, we have devised \emph{dynamic versions} of the Job- and Machine-based 
 decomposition strategies in which partial schedules are taken into account,
-and already scheduled operations from previous time windows may increase the
-EST and MTWR values are used to order the yet unscheduled operations.
+and already scheduled operations from previous time windows may change the EST and MTWR values used to order the yet unscheduled operations.
 However, the principle of applying a stratified ASP program to determine
@@
 those investigating bottleneck machines in the first place, and then picking their operations
-based on the smallest EST or greatest MTWR value as secondary criterion.
+based on the smallest EST or greatest MTWR value as a secondary criterion.
 While Machine-based decomposition turns out to be advantageous,

### 2022-01-30 ICLP Paper/samplepaper.tex
Section 4

@@
 \newcommand{\figspace}{\vspace{-1ex}}
-
+\newcommand{\ksverify}[1]{\textcolor{red}{#1}}
 \makeatletter
@@
 The \lstinline{step(w)} subprogram until line~\ref{prg:diff_log:oper_limit:end} 
-constitutes the central part of our multi-shot ASP modulo DL encoding in Listing~\ref{prg:encoding}, whose parameter~\lstinline{w} stands for consecutive integers from~\lstinline{1}
+in Listing~\ref{prg:encoding} constitutes the central part of our multi-shot ASP modulo DL encoding, whose parameter~\lstinline{w} stands for consecutive integers from~\lstinline{1}
 identifying time windows to schedule.
@@
 
+%
+Our first experiments, shown in Table~\ref{tab:Table01}, concern instances with $50\times15$, $50\times20$, and $100\times20$ jobs and machines with $750$, $1000$, or $2000$ operations, respectively.
+These operations were split into time windows with the M-EST decomposition strategy as a baseline because it leads to more robust solution quality than Job-based problem decompositions.
+% KS removed, since it was already explained in Section 3.2
+%The latter augments the encoding given in Listing~\ref{prg:decomposition} with bottleneck machines for inserting their operations into time windows, 
+For each of the three problem sizes, we report the average makespan over ten instances, where smaller values indicate better schedules, the average runtime of \clingodl, and the average number of interrupted optimization runs on subproblems.
+Each run was executed with a total timeout of $1000$ seconds, where each of $n$ subproblems got at most $1000/n$ seconds computation time.
+
+For every set of instances, we gradually increase the number of time windows from~$1$ to~$6$ and additionally include results for~$10$ time windows to outline the trend of degrading solution quality when the partition into time windows becomes too fragmented for merging (optimized) partial schedules into proficient global solutions.
+In fact, the shortest makespans, highlighted in boldface, were obtained with problem decomposition into $3$, $4$, or $6$ time windows, respectively, for the instances of size $50\times15$, $50\times20$, and $100\times20$.
+Compared to the results for $1$ time window, 
+which represent global optimization on the full problem,
+we observe substantial makespan improvements, amounting to almost one order
+of magnitude for the largest instances with $100\times20$ jobs and machines.
+These advantages are not surprising, considering that the JSP instances are highly combinatorial \citep{shysha18a} and each global optimization run times out with a more or less optimized solution, where a good-quality schedule is the more challenging to find, the greater the number of operations gets.
+Except for about 12\% completed optimization runs in multi-shot solving with $4$ time windows on the $50\times20$ instances, the best schedules were obtained with optimization runs timing out on subproblems.
+Nevertheless, they resulted in good-quality partial schedules that can be merged into better global solutions than achievable with more time windows, which become smaller and easier to solve but also cut off good-quality schedules.
+
 \begin{table}[t]
@@
 %       & {}        & \multicolumn{6}{c}{Group of Instances}  \\
-%    &  &  \multicolumn{7}{c}{Number of Time Windows} \\
+%  &  &  \multicolumn{7}{c}{Number of Time Windows} \\ [2pt] \cline{3-9} 
   Size & Average & 1 & 2 & 3 & 4 & 5 & 6 & 10  % \\
@@
   & Time             & $1000.0$	& $1000.0$   & $1000.0$	 & $934.4$ & $615.5$  & $243.8$  & $18.5$  \\
-    & Interrupts    & $1.0$	    & $2.0$      & $3.0$	& $3.6$  & $2.7$  & $1.0$  & $0.1$       \\[1.5mm]
+  & Interrupts    & $1.0$	    & $2.0$      & $3.0$	& $3.6$  & $2.7$  & $1.0$  & $0.1$       \\ [1.5mm]
 
@@
 \end{table}
-%
-Our first series of experiments shown in Table~\ref{tab:Table01} concerns the number
-of time windows into which the $750$, $1000$, or $2000$ operations, respectively,
-of the instances with $50\times15$, $50\times20$, and $100\times20$ jobs and machines
-are split.
-We use the static Machine-based EST decomposition strategy, which augments the encoding
-given in Listing~\ref{prg:decomposition} with bottleneck machines for inserting 
-their operations into time windows, as baseline because it leads to more robust
-solution quality than Job-based problem decomposition.
-For each of the three problem sizes,
-we report the average makespan over ten instances,
-where smaller values indicate better schedules,
-the average runtime of \clingodl, and the
-average number of interrupted optimization runs on subproblems.
-We gradually increase the number of time windows from~$1$ to~$6$ and additionally
-include results for~$10$ time windows to outline the trend of degrading
-solution quality when the partition into time windows becomes too fragmented
-for merging (optimized) partial schedules into proficient global solutions.
-In fact, the shortest makespans, highlighted in boldface,
-are obtained with problem decomposition into
-$3$, $4$, or $6$ time windows, respectively, for the instances of size 
-$50\times15$, $50\times20$, and $100\times20$.
-Compared to the results for $1$ time window,
-which represent global optimization on the full problem up to the
-runtime limit of $1000$ seconds,
-we observe substantial makespan improvements, amounting to almost one order
-of magnitude for the largest instances with $100\times20$ jobs and machines.
-These advantages do not come as a surprise considering that the JSP instances
-are highly combinatorial \citep{shysha18a} and each global optimization run
-times out with a more or less optimized solution, 
-where a good-quality schedule is the more difficult to find the greater the
-number of operations gets.
-Except for about 12\% completed optimization runs in multi-shot solving with
-$4$ time windows on the $50\times20$ instances,
-the best schedules are obtained with optimization runs timing out on subproblems,
-yet with good-quality partial schedules that can be merged into better global
-solutions than achievable with more time windows, which
-become smaller and easier to solve but also cut off good-quality
-schedules.
 
@@
 
-\begin{table}[t]
+\begin{table}[b]
     \caption{Experiments comparing Job- and Machine-based problem decomposition strategies\label{tab:Table02}}%
@@
 %
-In Table~\ref{tab:Table02}, we compare other problem decomposition strategies
-to the static Machine-based EST approach, providing the baseline abbreviated by M-EST.
-Further decomposition strategies are denoted by M-MTWR for the static Machine-based MTWR
-variant, corresponding dynamic versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$ that
-determine EST and MTWR values w.r.t.\ partial schedules from previous time windows, as well as
-static Job-based counterparts J-EST and J-MTWR in the first two rows.
+In Table~\ref{tab:Table02}, we provide a comparison of the Job- and Machine-based EST and MTWR approaches with %, providing the baseline abbreviated by M-EST.
+%Further decomposition strategies are denoted by M-MTWR for the static Machine-based MTWR variant, 
+dynamic versions of M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$ that
+determine M-EST and M-MTWR values w.r.t.\ partial schedules from previous time windows.
+\ksverify{We excluded the dynamic variants of Job-based strategies, since they performed worse than their static originals.}
+%, as well as static Job-based counterparts J-EST and J-MTWR in the first two rows.
 For each decomposition strategy and JSP instance size, we present results for the number

### 2022-01-30 ICLP Paper/samplepaper.tex
changing the place where to mention dynamic job-based strategies in experiments sections

@@
 
-In addition to assessing decompositions in a purely Job-based fashion,
+Beyond partitioning operations in a purely Job-based fashion,
 we have encoded \emph{Machine-based} decompositions \emph{M-EST} and \emph{M-MTRW} in which an operation from a bottleneck machine with the greatest sum of processing times for yet unordered operations is considered next.
@@
 determine M-EST and M-MTWR values w.r.t.\ partial schedules from previous time windows.
-\ksverify{We excluded the dynamic variants of Job-based strategies, since they performed worse than their static originals.}
+% \ksverify{We excluded the dynamic variants of Job-based strategies, since they performed worse than their static originals.}
 %, as well as static Job-based counterparts J-EST and J-MTWR in the first two rows.
@@
 there is no clear gap between the static M-EST and M-MTWR variants as well as their dynamic
-versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$.
+versions M-EST$^\text{\,d}$ and M-MTWR$^\text{\,d}$\ksverify{,
+and the same indifference applies to Job-based decomposition strategies whose dynamic counterparts are thus omitted}.
 That is, the heuristic to schedule operations from highly loaded machines

### 2022-02-01 ICLP Paper/Implementation/dlO.py
update python api

@@ -39,7 +39,7 @@ class Application:
         to_join = []
         for name, value in assignment:
             if str(name) != "makespan":
-                facts_format = "startTime({}, {}, {}). ".format(name, value, i)
+                facts_format = "start({}, {}, {}). ".format(name, value, i)
                 to_join.append(facts_format)
             else:
                 makespan = int(value)
@@ -51,7 +51,7 @@ class Application:
     def step_to_ground(self, prg, step, total_facts):
         parts = []
         if step > 0:
-            parts.append(("subproblem", [step]))
+            parts.append(("step", [step]))
             if step > 1:
                 parts.append(("solutionTimeWindow", []))
                 prg.add("solutionTimeWindow", [], total_facts)
@@ -63,8 +63,8 @@ class Application:
     # add a new constraint to get lower value of makespan (Optimization Part)
     def add_new_constraint(self, prg, makespan):
         prg.cleanup()
-        prg.ground([("opt", [makespan-1])])
-        prg.assign_external(Function("makespan", [makespan-1]), True)
+        prg.ground([("optimize", [makespan-1])])
+        prg.assign_external(Function("horizon", [makespan-1]), True)
     # ********************************************************************
 
     def main(self, prg, files):
@@ -91,7 +91,7 @@ class Application:
             #adjust = self.__theory.lookup_symbol(clingo.Number(0))
             makespan = 0
             while True:
-                prg.assign_external(Function("makespan", [lastmakespan-1]), False)
+                prg.assign_external(Function("horizon", [lastmakespan-1]), False)
                 lastmakespan = makespan
                 tic = time.time()
                 if time_used >= timeout_for_window:
@@ -105,7 +105,7 @@ class Application:
                     for model in handle:
                         a = self.__theory.assignment(model.thread_id)
                         total_facts, makespan = self.get_total_facts(a, i)
-                        
+                        #print(a)
                         break
                     else:
                         non_interrupted_calls += 1
@@ -119,7 +119,7 @@ class Application:
             if i != 0:
                 makespan_time_window.append(lastmakespan)
             i = i + 1      # Go to the next Time Window
-            print("Makespan {} : Assignment {}".format(lastmakespan, total_facts))
+            print("Makespan {} : Operation assignment is --> {}".format(lastmakespan, total_facts))
             print("**************************************************************")
         for x in range(NUM_OF_TIME_WINDOWS):
             print("Completion Time for Window {} : {} ".format(x+1, makespan_time_window[x]))

### 2022-02-03 ICLP Paper/samplepaper.tex
taking out red text color
set page numbers to 14
changed placement of decomposition encoding to bottom of page
update 2nd Table
added own related work in conclusions

@@
 \newcommand{\figspace}{\vspace{-1ex}}
-\newcommand{\ksverify}[1]{\textcolor{red}{#1}}
+\newcommand{\ksverify}[1]{#1}%{\textcolor{red}{#1}}
 \makeatletter
@@
 
-\jnlPage{1}{8}
+\jnlPage{1}{14}
 \jnlDoiYr{2021}
@@
 
-\lstinputlisting[float=t,label=prg:decomposition,caption={J-EST decomposition encoding}]{listing/decomposition.lp}
+\lstinputlisting[float=b,label=prg:decomposition,caption={J-EST decomposition encoding}]{listing/decomposition.lp}
 % Only 
@@
     \midline
-    0\%              & $3013.3$  & $572.3$   & $3.7$	   & $3186.8$      & $540.8$ & $3.3$ & $5713.0$ & $589.2$ & $8.6$\\
+    0\%              & $3013.3$  & $972.3$   & $3.7$	   & $3186.8$      & $940.8$ & $3.3$ & $5713.0$ & $989.2$ & $8.6$\\
     % [1.5mm]
 
-    10\%             & $2970.2$  & $600.0$   & $4.0$	   & $3149.8$      & $600.0$ & $4.0$ & $5656.0$ & $600.0$ & $9.0$\\
-
-    20\%             & $2952.2$  & $600.0$   & $4.0$	   & \boldmath{$3147.7$}      & $600.0$ & $4.0$ & \boldmath{$5619.2$} & $600.0$ & $9.0$\\
-
-    30\%             & \boldmath{$2943.1$}  & $600.0$   & $4.0$	   & $3162.2$      & $600.0$ & $4.0$ & $5857.0$ & $600.0$ & $9.0$%\\ 
+    10\%             & $2970.2$  & $1000.0$   & $4.0$	   & $3149.8$      & $1000.0$ & $4.0$ & $5656.0$ & $1000.0$ & $9.0$\\
+
+    20\%             & $2952.2$  & $1000.0$   & $4.0$	   & \boldmath{$3147.7$}      & $1000.0$ & $4.0$ & \boldmath{$5619.2$} & $1000.0$ & $9.0$\\
+
+    30\%             & \boldmath{$2943.1$}  & $1000.0$   & $4.0$	   & $3162.2$      & $1000.0$ & $4.0$ & $5857.0$ & $1000.0$ & $9.0$%\\ 
                      % [1.5mm]
@@
 w.r.t.\ deadlines, which are crucial in many application areas.
+While Machine-based decomposition turned out to be helpful
+for finding better schedules, the strategies presented in Section~\ref{subsec:decomposition}
+can be regarded as naive heuristics based on straightforward instance properties,
+and incorporating machine-learning approaches \citep{elscge22a,tagesc21a}
+for more adept partitioning is another promising direction for future work.
 Moreover, we aim to generalize our methods further, e.g.,
 by considering flexible machine allocations and varying processing times,
-and scale them up to large-scale instances with an order of magnitude more
+and scale them up to large real-world scheduling problems with an order of magnitude more
 operations than in synthetic benchmarks.

### 2022-03-17 ICLP Paper/samplepaper.tex
add word instead about --minimize-variable

@@
 For the successive optimization by time windows, where the scheduling
-horizon gradually increases, we require a dedicated treatment of DL constraints
+horizon gradually increases, we instead require a dedicated treatment of DL constraints
 limiting the value for \lstinline{makespan}.

### 2022-04-04 ICLP Paper/samplepaper.tex
updating based-on reviews

--- 2022-02-03 ICLP Paper/samplepaper.tex
@@
 \usepackage{comment}
+\usepackage{algorithm} 
+\usepackage{algpseudocode}
+\usepackage{color,soul}
+
 % Used for displaying a sample figure. If possible, figure files should
@@
 % \renewcommand\UrlFont{\color{blue}\rmfamily}
+
+\usepackage[most]{tcolorbox}
+\usepackage{caption}
 
@@
 machine~\lstinline{3} when compressing the decomposed schedule displayed in Figure~\ref{fig:decomposed}. 
+\hl{Figure~{\ref{fig:enum}} summarizes the main steps of the Multi-shot solving controlled by a python script.}
+
+\begin{tcolorbox}[title={The main steps of the Multi-shot solving},breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
+  \vspace{0.1\baselineskip}
+  \begin{enumerate}
+    \item[\textbf{1-}] Ground the base program.
+    \item[\textbf{2-}] Ground the next shot.
+    \item[\textbf{3-}] Optimize the current window till reaching the optima or the timeout.
+    \item[\textbf{4-}] Apply the compression of a window (postprocessing phase).
+    \item[\textbf{5-}] Keep the operations assignment to be passed to the next window.
+    \item[\textbf{6-}] Go to \textbf{2} until all windows are optimized.
+  \end{enumerate}
+\end{tcolorbox}
+\captionof{figure}{Main Steps of \textit{Multi-shot}\label{fig:enum}}
 
@@
 20\% of the time window size turn out to be most successful for the JSP
-benchmark sets under consideration.
+benchmark sets under consideration.\\
 
@@
 Therefore, the overlap and compression have improved the makespan results by $4.5\%, 3.4\%$ and $6.8\%$ and the running time by $40\%$.
+\hl{We conducted other experiments to assess the quality of the solutions obtained by our proposed method by comparing ASP with single-shot and OR tools.
+Table~{\ref{tab:Table04}} summarizes the results of our experiments.
+The values in the table represent the average distance to the optimal solutions for each set of instances.
+The first column represents the size of the instances with the decomposition parameters.
+The second and third columns show the average distance of the results obtained by ASP single and multi-shot, respectively.
+The last column shows the average distance obtained by OR tools.
+The results show that the single-shot ASP is far from the optima, especially with the largest dataset.
+In addition, it is shown that the developed model improves the results compared to the pure ASP.
+However, the performance of OR tools is better than ours, especially for the first two sets of instances.
+For $100 \times 20$, our proposed model is a bit comparable to OR tools, which means that if the instance size increases, it could be possible to obtain better results with our approach.}
+
+\begin{table}[t]
+  \caption{Experiments comparing the Single, Multi-shot ASP and OR tools\label{tab:Table04}}%
+%    \mbox{~}\hfill
+%    \resizebox{0.8\textwidth}{!}{%
+%    \begin{minipage}{\textwidth}
+  \centering
+  \begin{tabular}{l l r r }
+
+  \topline%\hline
+  %Timeout & \multicolumn{2}{c}{$1000$ sec} & \multicolumn{2}{c}{$500$ sec} \\
+   & \multicolumn{2}{c}{ASP} & OR tools \\
+  Instance set & Single-shot & Multi-shot &
+  \midline
+
+  $50\times15$  & $768.3$ & $169.3$ & $5.0$ \\
+
+  %$50\times15$  & $169.3$ & $5.0$ & $163.3$ & $26.0$ \\
+  ($4$w, M-EST, 30\%) & & & \\
+
+  $50\times20$  & $1035.5$ & $303.8$ & $70.4$ \\
+  %$50\times20$  & $303.8$ & $70.4$ & $304.4$ & $86.2$ \\
+  ($4$w, M-EST, 20\%) & & & \\
+
+  $100\times20$  & $41420.4$ & $253.5$ & $193.6$ \\
+  %$100\times20$  & $253.5$ & $193.6$ & $286.3$ & $284.1$ \\
+  ($9$w, M-EST, 20\%) & & &
+
+  \botline %\hline
+  \end{tabular}
+\end{table}
+%
 
@@
 w.r.t.\ deadlines, which are crucial in many application areas.
-While Machine-based decomposition turned out to be helpful
-for finding better schedules, the strategies presented in Section~\ref{subsec:decomposition}
-can be regarded as naive heuristics based on straightforward instance properties,
-and incorporating machine-learning approaches \citep{elscge22a,tagesc21a}
-for more adept partitioning is another promising direction for future work.
 Moreover, we aim to generalize our methods further, e.g.,
 by considering flexible machine allocations and varying processing times,
-and scale them up to large real-world scheduling problems with an order of magnitude more
+and scale them up to large-scale instances with an order of magnitude more
 operations than in synthetic benchmarks.

### 2022-04-04 ICLP Paper/samplepaper.tex
add some points

@@
 bottleneck operations \citep{zhang2010hybrid,zhai2014decomposition} 
-have been proposed to partition large-scale instances into better manageable subproblems,
+have been proposed to partition large-scale instances into better manageable subproblems, \hl{while minimizing the total tardiness}
 where no single strategy strictly dominates \citep{ovacik2012decomposition}.
+\hl{Most of the previous work has applied metaheuristic techniques or local search algorithms for solving the problem without sharing their implementation.}
 
@@
 To this end, we consider strategies for partitioning the operations of JSP instances
-into balanced time windows, which respect the operation precedence within jobs and
+into balanced time windows \hl{(the number of operations assigned to each time window is equal)}, which respect the operation precedence within jobs and
 enable a successive extension of good-quality partial schedules to a global solution.
@@
 $50\times15$, $50\times20$, or $100\times20$ jobs and machines.
+\hl{We have selected the largest set of instances generated and excluded the small instances because they could be solved by the existing methods and applying decomposition leads to unnecessary overheads.}
 The instances are generated such that each job includes one operation per machine, e.g.,
@@
 20\% of the time window size turn out to be most successful for the JSP
-benchmark sets under consideration.\\
+benchmark sets under consideration.
 
@@
 w.r.t.\ deadlines, which are crucial in many application areas.
+\hl{Furthermore, it would be beneficial to apply the same approach to Flow-shop and open-shop scheduling problems.
+However, the proposed model must be changed to solve open-shop since the precedence constraint is not applied}
 Moreover, we aim to generalize our methods further, e.g.,

### 2022-04-11 ICLP Paper/samplepaper.tex
update a table
complete updating table 3 & 4
added sentence reacting to Reviewer 2

@@
 \lstinline{order((}$j_1$\lstinline{,}$s_1$\lstinline{),(}$j_2$\lstinline{,}$s_2$\lstinline{),}$p_1$\lstinline{,}$w$\lstinline{)}
-to signal that \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} needs to be completed
+to express that \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} needs to be completed
 before performing \lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}, i.e.,
@@
 the rule in lines \ref{prg:base:order:begin}-\ref{prg:base:order:end}
-derives an atom signaling the inverse, given that the operations must not intersect and some sequence has to be determined.%
-%
+derives an atom expressing the inverse, given that the operations must not intersect and some sequence has to be determined.
+\hl{Note that applications of the rules in lines} \ref{prg:base:order}-\ref{prg:base:order:end}
+\hl{for choosing an execution order require that the
+operations}
+\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} \hl{and}
+\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}
+\hl{are part of the same time window~$w$ and belong to distinct jobs, so that}
+\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} \hl{and}
+\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}
+\hl{are out of scope of the rules
+imposing a fixed order in lines} \ref{prg:base:ordered}-\ref{prg:base:same_job:end}.%
 \lstinputlisting[float=t,label=prg:encoding,caption={Multi-shot ASP modulo DL encoding}]{listing/encoding.lp}
@@
     \midline
-    0\%              & $3013.3$  & $972.3$   & $3.7$	   & $3186.8$      & $940.8$ & $3.3$ & $5713.0$ & $989.2$ & $8.6$\\
+    0\%              & $3010.3$  & $907.5$   & $3.4$	   & $3174.5$      & $839.6$ & $3.0$ & $5671.9$ & $1000.0$ & $7.0$\\
     % [1.5mm]
 
-    10\%             & $2970.2$  & $1000.0$   & $4.0$	   & $3149.8$      & $1000.0$ & $4.0$ & $5656.0$ & $1000.0$ & $9.0$\\
-
-    20\%             & $2952.2$  & $1000.0$   & $4.0$	   & \boldmath{$3147.7$}      & $1000.0$ & $4.0$ & \boldmath{$5619.2$} & $1000.0$ & $9.0$\\
-
-    30\%             & \boldmath{$2943.1$}  & $1000.0$   & $4.0$	   & $3162.2$      & $1000.0$ & $4.0$ & $5857.0$ & $1000.0$ & $9.0$%\\ 
+    10\%             & $2960.6$  & $989.4$   & $3.9$	   & $3154.6$      & $956.4$ & $3.9$ & \boldmath{$5637.5$} & $1000.0$ & $7.0$\\
+
+    20\%             & $2946.9$  & $1000.0$   & $4.0$	   & \boldmath{$3120.3$}      & $1000.0$ & $4.0$ & $5640.7$ & $1000.0$ & $7.0$\\
+
+    30\%             & \boldmath{$2932.9$}  & $1000.0$   & $4.0$	   & $3169.0$      & $1000.0$ & $4.0$ & $5722.7$ & $1000.0$ & $7.0$%\\ 
                      % [1.5mm]
@@
 
+\begin{comment}
   \begin{table}[t]
   \caption{Experiments comparing the Single, Multi-shot ASP and OR tools\label{tab:Table04}}%
-%    \mbox{~}\hfill
-%    \resizebox{0.8\textwidth}{!}{%
-%    \begin{minipage}{\textwidth}
   \centering
@@
 %
+\end{comment}
+
+\begin{table}[t]
+  \caption{Experiments comparing the Single, Multi-shot ASP and OR tools\label{tab:Table04}}%
+  \centering
+  \begin{tabular}{l l r r r r r r}
+
+  \topline%\hline
+  \multirow{3}{*}{Instances} & \multicolumn{4}{c}{Makespan} & \multicolumn{3}{c}{Distance measures} \\
+  & Optima & OR-tools & \multicolumn{2}{c}{ASP} & OR-tools & \multicolumn{2}{c}{ASP} \\
+
+  & & & Single-shot & Multi-shot &  & Single-shot & Multi-shot \\
+
+  $50\times15$ & $2733.8$ & $2778.8$ & $3542.1$ & $2932.9$ & $5.0$ & $768.3$ & $159.1$ \\
+  $50\times20$ & $2843.9$ & $2914.3$ & $3879.4$ & $3120.3$ & $70.4$ & $1035.5$ & $276.4$ \\
+  $100\times20$ & $5365.7$ & $5559.3$ & $46786.1$ & $5637.5$ & $193.6$ & $41420.4$ & $271.8$
+
+  \botline %\hline
+  \end{tabular}
+\end{table}
 

### 2022-04-12 ICLP Paper/samplepaper.tex
working on control loop

@@
 fixed to the value~$t$ in an (optimized) partial schedule for time window $w-1$, 
-as supplied by reified facts \lstinline{start(}$o${,}$t$\lstinline{,}$w-1$\lstinline{).}
+as supplied by reified facts \lstinline{start(}$o$\lstinline{,}$t$\lstinline{,}$w-1$\lstinline{).}
 In line~\ref{prg:diff_log:non_nega:begin},
@@
 via the command-line option \lstinline{--minimize-variable=makespan}.
+\hl{This option, however, is implemented by means of a fixed control loop that
+cannot be combined with (other) multi-shot solving methods.}
 For the successive optimization by time windows, where the scheduling
-horizon gradually increases, we require a dedicated treatment of DL constraints
+horizon gradually increases, we thus require a dedicated treatment of DL constraints
 limiting the value for \lstinline{makespan}.
@@
 and thus make sure that the iterative scheduling progresses.
+\hl{Figure~{\ref{fig:enum}} summarizes the main steps of the Multi-shot solving controlled by a python script.}
+
+\begin{figure}[tb]
+\begin{tcolorbox}[breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
+%  \vspace{0.1\baselineskip}
+Starting with $H:=\emptyset$, for $w:= 1$ to number $n$ of time windows, do:
+\begin{enumerate}[label=\textbf{\arabic*.}]
+  \item % \begin{tabbing}
+        If $w= 1$, % \= 
+        then $P:=\{\text{\lstinline{base}}\}$;
+        % \\
+        % \>
+
+        % \hfill
+        else $P := \{\text{\lstinline{start((}$j$\lstinline{,}$s$\lstinline{),}$t$\lstinline{,}$w-1$\lstinline{).}}
+                    \mid
+                    \{\text{\lstinline{window(}$j$\lstinline{,}$s$\lstinline{,}$w-1$\lstinline{)}},
+                    \text{\lstinline{(}$j$\lstinline{,}$s$\lstinline{)}}=t\}\subseteq X\}$.
+        % \end{tabbing}
+  \item Ground $P\cup\{\text{\lstinline{step(}$w$\lstinline{)}}\}$
+  \item While the time limit per time window is not reached, do:
+  \begin{enumerate}[label=\textbf{\alph*.}]
+  \item Find some answer set $X$
+  \end{enumerate}
+  \item Apply the compression of a window (postprocessing phase).
+  \item Keep the operations assignment to be passed to the next window.
+  \item Go to \textbf{2} until all windows are optimized.
+\end{enumerate}
+  \begin{enumerate}[label=\textbf{\arabic*.}]
+    \item Ground the base program.
+    \item Ground the next shot.
+    \item Optimize the current window till reaching the optima or the timeout.
+    \item Apply the compression of a window (postprocessing phase).
+    \item Keep the operations assignment to be passed to the next window.
+    \item Go to \textbf{2} until all windows are optimized.
+  \end{enumerate}
+\end{tcolorbox}
+\caption{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
+\end{figure}
+% \captionof{figure}{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
 
@@
 machine~\lstinline{3} when compressing the decomposed schedule displayed in Figure~\ref{fig:decomposed}.
-\hl{Figure~{\ref{fig:enum}} summarizes the main steps of the Multi-shot solving controlled by a python script.}
-
-\begin{tcolorbox}[title={The main steps of the Multi-shot solving},breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
-  \vspace{0.1\baselineskip}
-  \begin{enumerate}
-    \item[\textbf{1-}] Ground the base program.
-    \item[\textbf{2-}] Ground the next shot.
-    \item[\textbf{3-}] Optimize the current window till reaching the optima or the timeout.
-    \item[\textbf{4-}] Apply the compression of a window (postprocessing phase).
-    \item[\textbf{5-}] Keep the operations assignment to be passed to the next window.
-    \item[\textbf{6-}] Go to \textbf{2} until all windows are optimized.
-  \end{enumerate}
-\end{tcolorbox}
-\captionof{figure}{Main Steps of \textit{Multi-shot}\label{fig:enum}}
 

### 2022-04-12 ICLP Paper/samplepaper.tex
control algorithm updated

@@
 
+\begin{figure}[tb]
+  \begin{tcolorbox}[breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
+  %  \vspace{0.1\baselineskip}
+  Starting with $H:=\emptyset$, for $w:= 1$ to number $n$ of time windows:
+  \begin{enumerate}[label=\textbf{\arabic*.}]
+    \item For each $h\in H$, set \lstinline{horizon(}$h$\lstinline{)} to false
+    \item % \begin{tabbing}
+          If $w= 1$, % \= 
+          then $P:=\{\text{\lstinline{base}}\}$;
+          % \\
+          % \>
+
+          % \hfill
+          else $P := \{\text{\lstinline{start((}$j$\lstinline{,}$s$\lstinline{),}$t$\lstinline{,}$w-1$\lstinline{).}}
+                      \mid
+                      \{\text{\lstinline{window(}$j$\lstinline{,}$s$\lstinline{,}$w-1$\lstinline{)}},
+                      \text{\lstinline{(}$j$\lstinline{,}$s$\lstinline{)}}=t\}\subseteq X\}$
+          % \end{tabbing}
+    \item Ground $P\cup\{\text{\lstinline{step(}$w$\lstinline{)}}\}$
+    \item While there is some answer set $X$ and the time limit per time window is not reached:
+      \begin{enumerate}[label=\textbf{\alph*.}]
+      \item $h:=t-1$, where $($\lstinline{makespan}${}=t)\in X$
+      \item If $h\notin H$, then $H := H\cup\{h\}$ and ground $\{$\lstinline{optimize(}$h$\lstinline{)}$\}$
+      \item Set \lstinline{horizon(}$h$\lstinline{)} to true
+      \end{enumerate}
+    \end{enumerate}
+  \end{tcolorbox}
+  \caption{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
+  \end{figure}
+  % \captionof{figure}{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
+%
 The task of optimizing the horizon of a (partial) schedule means choosing
@@
 
-\begin{figure}[tb]
-\begin{tcolorbox}[breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
-%  \vspace{0.1\baselineskip}
-Starting with $H:=\emptyset$, for $w:= 1$ to number $n$ of time windows, do:
-\begin{enumerate}[label=\textbf{\arabic*.}]
-  \item % \begin{tabbing}
-        If $w= 1$, % \= 
-        then $P:=\{\text{\lstinline{base}}\}$;
-        % \\
-        % \>
-
-        % \hfill
-        else $P := \{\text{\lstinline{start((}$j$\lstinline{,}$s$\lstinline{),}$t$\lstinline{,}$w-1$\lstinline{).}}
-                    \mid
-                    \{\text{\lstinline{window(}$j$\lstinline{,}$s$\lstinline{,}$w-1$\lstinline{)}},
-                    \text{\lstinline{(}$j$\lstinline{,}$s$\lstinline{)}}=t\}\subseteq X\}$.
-        % \end{tabbing}
-  \item Ground $P\cup\{\text{\lstinline{step(}$w$\lstinline{)}}\}$
-  \item While the time limit per time window is not reached, do:
-  \begin{enumerate}[label=\textbf{\alph*.}]
-  \item Find some answer set $X$
-  \end{enumerate}
-  \item Apply the compression of a window (postprocessing phase).
-  \item Keep the operations assignment to be passed to the next window.
-  \item Go to \textbf{2} until all windows are optimized.
-\end{enumerate}
-  \begin{enumerate}[label=\textbf{\arabic*.}]
-    \item Ground the base program.
-    \item Ground the next shot.
-    \item Optimize the current window till reaching the optima or the timeout.
-    \item Apply the compression of a window (postprocessing phase).
-    \item Keep the operations assignment to be passed to the next window.
-    \item Go to \textbf{2} until all windows are optimized.
-  \end{enumerate}
-\end{tcolorbox}
-\caption{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
-\end{figure}
-% \captionof{figure}{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
-
 % This section will show how to solve the problem dynamically, where the output of each iteration is an input to the next and merging them to obtain the solution of the whole problem.

### 2022-04-12 ICLP Paper/samplepaper.tex
additions on problem encoding part

@@
 \newcommand{\clingodl}{\emph{clingo}[DL]\xspace}
-\newcommand{\figspace}{\vspace{-1ex}}
+\newcommand{\figspace}{}%{\vspace{-1ex}}
 \newcommand{\ksverify}[1]{#1}%{\textcolor{red}{#1}}
@@
 present our ASP encoding with DL constraints for optimizing the makespan of partial schedules, and
-finally outline the incorporation of time window overlapping and compression
-into the iterative scheduling process.   
+finally outline the iterative scheduling process 
+along with the incorporation of time window overlapping and compression techniques.   
 % The Job-shop Scheduling Problem (JSP) has been a complex and combinatorial optimization problem since the 1950s, and it was shown to be NP-hard.
@@
 machines are displayed in Listing~\ref{prg:facts}.
+%
+% In order to encode the problem using ASP, we should define a set of predicates that represents the problem instances.
+For instance, the operations are encoded with the predicate \lstinline{operation/2} where the first variable denotes the job number and the second is the operation number, see lines \ref{prg:facts:ops:begin}-\ref{prg:facts:ops:end} in Listing \ref{prg:facts}.
+The time needed to finish an operation is represented by the predicate \lstinline{pro/3} in which the operation is determined by the first two terms and the third represents the processing time in lines \ref{prg:facts:pro:begin}-\ref{prg:facts:pro:end}.
+For the machine assignment, the predicate \lstinline{assign/3} shows that a particular operation is executed by a machine in lines \ref{prg:facts:mach:begin}-\ref{prg:facts:mach:end}.
+For instance, \emph{operation(1,1)} is processed by machine 2.
+\lstinputlisting[float=b,label=prg:facts,caption={Example JSP instance},linerange={1-3},numbers=none,xleftmargin=3pt]{listing/facts.lp}
+%
 An atom of the form
@@
 and likewise for operations sharing machine~\lstinline{2} or~\lstinline{3}, respectively.
-%
-% In order to encode the problem using ASP, we should define a set of predicates that represents the problem instances.
-For instance, the operations are encoded with the predicate \lstinline{operation/2} where the first variable denotes the job number and the second is the operation number, see lines \ref{prg:facts:ops:begin}-\ref{prg:facts:ops:end} in Listing \ref{prg:facts}.
-The time needed to finish an operation is represented by the predicate \lstinline{pro/3} in which the operation is determined by the first two terms and the third represents the processing time in lines \ref{prg:facts:pro:begin}-\ref{prg:facts:pro:end}.
-For the machine assignment, the predicate \lstinline{assign/3} shows that a particular operation is executed by a machine in lines \ref{prg:facts:mach:begin}-\ref{prg:facts:mach:end}.
-For instance, \emph{operation(1,1)} is processed by machine 2.
-\lstinputlisting[float=b,label=prg:facts,caption={Example JSP instance},linerange={1-3},numbers=none,xleftmargin=3pt]{listing/facts.lp}
 
@@
 scalability of complete optimization methods necessitates problem decomposition.
-To this end, we consider strategies for partitioning the operations of JSP instances
-into balanced time windows \hl{(the number of operations assigned to each time window is equal)}, which respect the operation precedence within jobs and
-enable a successive extension of good-quality partial schedules to a global solution.
+% To this end, 
+In order to enable a successive extension of good-quality partial schedules to a global solution,
+we consider strategies for partitioning the operations of JSP instances
+into balanced time windows, \hl{each comprising an equal number of operations such that their precedence
+within jobs is respected}. % , which respect the operation precedence within jobs and
+% in order to enable a successive extension of good-quality partial schedules to a global solution.
 In the following, we first detail problem decomposition based on the earliest starting times of operations, and then outline further strategies that can be encoded by stratified ASP programs as well.
@@
 %
-\lstinputlisting[float=t,label=prg:tw,caption={Example time windows},numbers=none,xleftmargin=3pt]{listing/TW.lp}
+\lstinputlisting[float=b,label=prg:tw,caption={Example time windows},numbers=none,xleftmargin=3pt]{listing/TW.lp}
 
@@
 are in the lines~\ref{prg:diff_log:freez1:begin} and~\ref{prg:diff_log:freez2:begin}
-fixed to the value~$t$ in an (optimized) partial schedule for time window $w-1$, 
+fixed \hl{by restricting them from above or below, respectively,}
+to the value~$t$ in an (optimized) partial schedule for time window $w-1$, 
 as supplied by reified facts \lstinline{start(}$o$\lstinline{,}$t$\lstinline{,}$w-1$\lstinline{).}
@@
 
+The task of optimizing the horizon of a (partial) schedule means choosing
+the execution order of operations of the latest time window sharing a machine
+such that the value for the \lstinline{makespan} variable is minimized.
+In single-shot ASP modulo DL solving with \clingodl, this can be accomplished
+via the command-line option \lstinline{--minimize-variable=makespan}.
+\hl{This option, however, is implemented by means of a fixed control loop that
+cannot be combined with (other) multi-shot solving methods.}
+For the successive optimization by time windows, where the scheduling
+horizon gradually increases, we thus require a dedicated treatment of DL constraints
+limiting the value for \lstinline{makespan}.
+To this end, the \lstinline{optimize(m)} subprogram below line~\ref{prg:diff_log:oper_min}
+declares an external atom \lstinline{horizon(m)} for controlling whether a DL
+constraint asserted in line~\ref{prg:diff_log:oper_min:end} is active and limits
+the \lstinline{makespan} value to an integer supplied for the parameter~\lstinline{m}.
+
+\subsection{\hl{Iterative Scheduling}}\label{subsec:scheduling}
+
+\hl{The main steps of our control loop for successive schedule optimization
+by multi-shot ASP modulo DL solving, implemented by means of the Python interface of \clingodl,
+are displayed in Figure~{\ref{fig:enum}} and further detailed in the following.}
+When launching the optimization process for a new time window, any instance of the
+\lstinline{horizon(m)} atom introduced before is set to false in step~1 for making sure that
+some (partial) schedule is feasible.
+\hl{No such atoms have been introduced yet for the first time window $w=1$,
+where the static (default) subprogram called} \lstinline{base}\hl{,
+supplying a JSP instance along with its decomposition in terms of facts over} \lstinline{window/3}\hl{,
+and the} \lstinline{step(}$1$\lstinline{)} \hl{subprogram for operations of the first time window
+are instantiated in steps~2 and~3.}
+%
 \begin{figure}[tb]
@@
   \caption{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
-  \end{figure}
+  \end{figure}%
   % \captionof{figure}{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
 %
-The task of optimizing the horizon of a (partial) schedule means choosing
-the execution order of operations of the latest time window sharing a machine
-such that the value for the \lstinline{makespan} variable is minimized.
-In single-shot ASP modulo DL solving with \clingodl, this can be accomplished
-via the command-line option \lstinline{--minimize-variable=makespan}.
-\hl{This option, however, is implemented by means of a fixed control loop that
-cannot be combined with (other) multi-shot solving methods.}
-For the successive optimization by time windows, where the scheduling
-horizon gradually increases, we thus require a dedicated treatment of DL constraints
-limiting the value for \lstinline{makespan}.
-To this end, the \lstinline{optimize(m)} subprogram below line~\ref{prg:diff_log:oper_min}
-declares an external atom \lstinline{horizon(m)} for controlling whether a DL
-constraint asserted in line~\ref{prg:diff_log:oper_min:end} is active and limits
-the \lstinline{makespan} value to an integer supplied for the parameter~\lstinline{m}.
-When launching the optimization process for a new time window, any instance of the
-\lstinline{horizon(m)} atom introduced before is set to false for making sure that
-some (partial) schedule is feasible.
-Once some schedule with a horizon $h+1$ is found,
-the subprogram \lstinline{optimize(}$h$\lstinline{)} is instantiated on demand, i.e.,
+Once some schedule with a horizon $h+1$ is found in step~4a,
+the step~4b consists of instantiating
+the subprogram \lstinline{optimize(}$h$\lstinline{)} % is instantiated
+on demand, i.e.,
 in case $h$ has not been passed as a value for~\lstinline{m} before, and
-the corresponding \lstinline{horizon(}$h$\lstinline{)} atom is set to true for activating the DL
+the corresponding \lstinline{horizon(}$h$\lstinline{)} atom is set to true in step~4c for activating the DL
 constraint reducing the admitted scheduling horizon to~$h$.
-The process of successively reducing the horizon~$h$ in order to find better partial schedules
-is repeated until the imposed \lstinline{makespan} value turns out to be infeasible,
+The step 4 of successively reducing the horizon~$h$ in order to find better partial schedules
+% is repeated until 
+stops when the imposed \lstinline{makespan} value turns out to be infeasible,
 meaning that an optimal partial schedule has been found.
 As already mentioned above, the introduced instances of the external \lstinline{horizon(m)} atom
-are then set to false and the successive optimization proceeds by instantiating the
-\lstinline{step(}$w+1$\lstinline{)} subprogram for the next time window $w+1$ (if any), and
+are then set to false in step~1, and the successive optimization proceeds by in steps~2 and~3 instantiating the
+\lstinline{step(}$w+\nolinebreak1$\lstinline{)} subprogram for the next time window $w+1$ (if any) and
 also supplying the determined starting times of operations from time window~$w$
@@
 The described control loop for successively extending good-quality partial schedules to
-a global solution is implemented by means of the Python interface of \clingodl,
-where a time limit can be configured to restrict the optimization efforts per time window
+a global solution 
+% is implemented by means of the Python interface of \clingodl,
+% where a time limit 
+can be configured with a time limit,
+included as secondary stopping condition in step~4,
+to restrict the optimization efforts per time window
 and thus make sure that the iterative scheduling progresses.
-\hl{Figure~{\ref{fig:enum}} summarizes the main steps of the Multi-shot solving controlled by a python script.}
 
@@
 with different number and size of time windows as well as the impact of
-time window overlapping and compression techniques described in Section~\ref{subsec:encoding}.
+time window overlapping and compression techniques described in Section~\ref{subsec:scheduling}.
 For the comparability of results between runs with a different number of time windows

### 2022-04-12 ICLP Paper/samplepaper.tex
checked experiments section

@@
 In order to counteract limitations of window-wise successive optimization
-due to "decomposition mistakes" and unnecessary idle times that do not directly
+due to ``decomposition mistakes'' and unnecessary idle times that do not directly
 affect the makespan to some extent,
@@
 $50\times15$, $50\times20$, or $100\times20$ jobs and machines.
-\hl{We have selected the largest set of instances generated and excluded the small instances because they could be solved by the existing methods and applying decomposition leads to unnecessary overheads.}
+\hl{These are the largest instances provided by common JSP benchmark sets,
+so that complete optimization methods do not terminate within tight runtime limits
+and the amount of operations makes problem decomposition worthwhile.}
+% and excluded the small instances because they could be solved by the existing methods and applying decomposition leads to unnecessary overheads.}
 The instances are generated such that each job includes one operation per machine, e.g.,
@@
 
+\begin{table}[b]
+  \caption{Comparison of single- and multi-shot ASP modulo DL solving approaches to OR-tools\label{tab:Table04}}%
+  \centering
+  \begin{tabular}{r r r r r r r r}
+
+  \topline%\hline
+%  \multirow{3}{*}{Size} & \multicolumn{4}{c}{Makespan} & \multicolumn{3}{c}{Distance measures} \\
+%  & Optima & OR-tools & \multicolumn{2}{c}{ASP} & OR-tools & \multicolumn{2}{c}{ASP} \\
+
+ % & & & Single-shot & Multi-shot &  & Single-shot & Multi-shot 
+ & & \multicolumn{2}{c}{OR-tools} & \multicolumn{2}{c}{Multi-shot} & \multicolumn{2}{c}{Single-shot} \\
+ Size & Optima & Makespan & Distance & Makespan & Distance & Makespan & Distance \midline
+
+  $50\times15$ & $2773.8$ & $2778.8$ & $5.0$ & $2932.9$ & $159.1$ & $3542.1$ & $768.3$ \\
+  $50\times20$ & $2843.9$ & $2914.3$ & $70.4$ & $3120.3$ & $276.4$ & $3879.4$ & $1035.5$ \\
+  $100\times20$ & $5365.7$ & $5559.3$ & $193.6$ & $5637.5$ & $271.8$ & $46786.1$ & $41420.4$
+
+  \botline %\hline
+  \end{tabular}
+\end{table}
 % Table \ref{tab:Table02} demonstrates the results of the four static decomposition strategies and the machine-based MTWR with the dynamic assignment.
@@
 Therefore, the overlap and compression have improved the makespan results by $4.5\%, 3.4\%$ and $6.8\%$ and the running time by $40\%$.
-\hl{We conducted other experiments to assess the quality of the solutions obtained by our proposed method by comparing ASP with single-shot and OR tools.
-Table~{\ref{tab:Table04}} summarizes the results of our experiments.
+\hl{For putting our results, obtained with single-shot ASP modulo DL solving ($1$ time window in Table}~\ref{tab:Table01}\hl{) and the best-performing multi-shot solving approach
+indicated in the lower half of Table}~\ref{tab:Table03}\hl{, in relation,
+we also ran the state-of-the-art CP system OR-tools (version 9.2)} \citep{ortools}\hl{
+with a timeout of $1000$ seconds,
+using the JSP encoding provided by} \cite{tagesc21a}\hl{.
+The respective average makespans and distances to the known optima reported by}
+\cite{taillard1993benchmarks}\hl{ are shown in Table}~\ref{tab:Table04}\hl{.
+Notably, the optimal schedules from the literature have been determined by greedy or
+local search techniques, which can in general not compute (provably) optimal solutions,
+yet the optimality of their schedules is guaranteed by theoretical lower bounds that were
+not supplied as background knowledge to} \clingodl \hl{and OR-tools.
+As a consequence, all runs of the complete OR-tools and} \clingodl \hl{solvers are
+interrupted at the runtime limit and do thus not yield optimality proofs.
+However, we observe that the schedules of OR-tools are of high quality and
+come close to the optima for the instances of size $50\times15$.
+The larger the instances get, the closer is our decomposition and multi-shot solving approach,
+while it remains behind OR-tools.
+On the one hand, this can be due to ``decomposition mistakes'' of our method,
+deteriorating the quality of the resulting global solutions.
+On the other hand, the CP encoding taken for OR-tools features dedicated interval variables
+and global constraints, which are tailored to scheduling problems and give it an advantage
+over the use of DL constraints.
+Nevertheless, the improvements due to problem decomposition and multi-shot solving are apparent
+relative to the single-shot solving approach, which is significantly behind for each of the three problem sizes.}
+% \hl{We conducted other experiments to assess the quality of the solutions obtained by our proposed method by comparing ASP with single-shot and OR-tools (version 9.2)} \citep{ortools}\hl{.
+Table}~\ref{tab:Table04} \hl{summarizes the results of our experiments.
 The values in the table represent the average distance to the optimal solutions for each set of instances.
@@
 \end{comment}
-
-\begin{table}[t]
-  \caption{Experiments comparing the Single, Multi-shot ASP and OR tools\label{tab:Table04}}%
-  \centering
-  \begin{tabular}{l l r r r r r r}
-
-  \topline%\hline
-  \multirow{3}{*}{Instances} & \multicolumn{4}{c}{Makespan} & \multicolumn{3}{c}{Distance measures} \\
-  & Optima & OR-tools & \multicolumn{2}{c}{ASP} & OR-tools & \multicolumn{2}{c}{ASP} \\
-
-  & & & Single-shot & Multi-shot &  & Single-shot & Multi-shot \\
-
-  $50\times15$ & $2733.8$ & $2778.8$ & $3542.1$ & $2932.9$ & $5.0$ & $768.3$ & $159.1$ \\
-  $50\times20$ & $2843.9$ & $2914.3$ & $3879.4$ & $3120.3$ & $70.4$ & $1035.5$ & $276.4$ \\
-  $100\times20$ & $5365.7$ & $5559.3$ & $46786.1$ & $5637.5$ & $193.6$ & $41420.4$ & $271.8$
-
-  \botline %\hline
-  \end{tabular}
-\end{table}
 
@@
 w.r.t.\ deadlines, which are crucial in many application areas.
+While Machine-based decomposition turned out to be helpful
+for finding better schedules, the strategies presented in Section~\ref{subsec:decomposition}
+can be regarded as naive heuristics based on straightforward instance properties,
+and incorporating machine-learning approaches \citep{elscge22a,tagesc21a}
+for more adept partitioning is another promising direction for future work.
 \hl{Furthermore, it would be beneficial to apply the same approach to Flow-shop and open-shop scheduling problems.
@@
 by considering flexible machine allocations and varying processing times,
-and scale them up to large-scale instances with an order of magnitude more
+and scale them up to large real-world scheduling problems with an order of magnitude more
 operations than in synthetic benchmarks.

### 2022-04-13 ICLP Paper/samplepaper.tex
ready with paper

@@
 
-\usepackage{comment}
-\usepackage{algorithm} 
-\usepackage{algpseudocode}
-\usepackage{color,soul}
+%\usepackage{comment}
+%\usepackage{algorithm} 
+%\usepackage{algpseudocode}
+\usepackage{soul}%{color,soul}
 
@@
 bottleneck operations \citep{zhang2010hybrid,zhai2014decomposition} 
-have been proposed to partition large-scale instances into better manageable subproblems, \hl{while minimizing the total tardiness}
-where no single strategy strictly dominates \citep{ovacik2012decomposition}.
-\hl{Most of the previous work has applied metaheuristic techniques or local search algorithms for solving the problem without sharing their implementation.}
+have been proposed to partition large-scale instances into better manageable subproblems,
+where no single strategy strictly dominates \hl{in minimizing the tardiness} \citep{ovacik2012decomposition}.
+% \hl{Most of the previous work has applied metaheuristic techniques or local search algorithms for solving the problem without sharing their implementation.}
 
@@
 for more adept partitioning is another promising direction for future work.
-\hl{Furthermore, it would be beneficial to apply the same approach to Flow-shop and open-shop scheduling problems.
+\hl{Provided the adjustment of problem decomposition strategies, 
+our multi-shot solving methods can also be applied to the 
+Flow-shop and Open-shop Scheduling Problems} \citep{taillard1993benchmarks}.
+% \hl{Furthermore, it would be beneficial to apply the same approach to Flow-shop and open-shop scheduling problems.
 However, the proposed model must be changed to solve open-shop since the precedence constraint is not applied}
@@
 
+\paragraph{Acknowledgments.}{This work was partially funded by
+KWF project 28472,
+cms electronics GmbH,
+FunderMax GmbH,
+Hirsch Armbänder GmbH,
+incubed IT GmbH,
+Infineon Technologies Austria AG,
+Isovolta AG,
+Kostwein Holding GmbH, and
+Privatstiftung Kärntner Sparkasse.
+We thank the anonymous reviewers for their valuable suggestions and comments.}
+
 %

### 2022-04-13 ICLP Paper/samplepaper.tex
small change

@@
 yet highly combinatorial JSP instances \citep{elkgeb20a}, for some of which optimal solutions are so far unknown \citep{shysha18a}.
-Successful application areas beyond JSP include 
+Successful \hl{applications in} areas beyond JSP include 
 industrial printing \citep{balduccini11a},
@@
 w.r.t.\ deadlines, which are crucial in many application areas.
-While Machine-based decomposition turned out to be helpful
-for finding better schedules, the strategies presented in Section~\ref{subsec:decomposition}
-can be regarded as naive heuristics based on straightforward instance properties,
-and incorporating machine-learning approaches \citep{elscge22a,tagesc21a}
-for more adept partitioning is another promising direction for future work.
-\hl{Provided the adjustment of problem decomposition strategies, 
+\hl{While Machine-based decomposition turned out to be helpful
+for finding better schedules, the strategies presented in Section}~\ref{subsec:decomposition}
+\hl{can be regarded as naive heuristics based on straightforward instance properties,
+and incorporating machine-learning approaches} \citep{elscge22a,tagesc21a}
+\hl{for more adept partitioning is another promising direction for future work.
+Provided the adjustment of problem decomposition strategies, 
 our multi-shot solving methods can also be applied to the 

### 2022-04-13 ICLP Paper/Responses.tex
drafted response letter

\documentclass[11pt]{article}
\usepackage{amsmath, amssymb, amsthm, graphicx, epsfig, fancyhdr}
\usepackage[a4paper, total={7.5in, 10in}]{geometry}
\usepackage{enumitem}
\usepackage{url}
\setlength{\headheight}{10pt}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[R]{M. El-Kholany, M. Gebser, K. Schekotihin}
%\fancyfoot[C]{\thepage}

\begin{document}
We thank the reviewers for their valuable feedback helping us to improve the paper. In the following, we reply to the comments/questions. % Some of the questions are grouped when corresponding answers are similar or related.

\subsection*{Review 1}
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item \textbf{The effectiveness of the proposed approach is still not clear to me.
      For this, in my opinion, the authors need to show some comparison
	  results with previous best known bounds obtained by other different
	  approaches.}
\item[] We did so by adding a comparison to known optima and OR-tools at the end of Section 4.
        It turns out that OR-tools with dedicated interval variables and global constraints
		finds better schedules than our decomposition and multi-shot solving
		approach, yet the gap isn't too large and substantially reduced in comparison to
		single-shot ASP modulo DL solving.
		As implementations of greedy and local search methods applied to obtain the known
		optima for the instances we tested are, to our knowledge, not publically available,
		we couldn't run them but still report their results as optimal values.
\item \textbf{Could you explain the reason why the authors selected a subset of
       Taillard benchmark instances in [1,2] and didn't use the other
       instances in OR-library [1]?}
\item[] We used the largest families of instances included in common JSP benchmark sets,
        and now explain our choice in the first paragraph of Section 4.
\item \textbf{Can the proposed approach be effective to other shop scheduling
problems like flow-shop and open-shop?}
\item[] Yes, our problem encoding can be readily applied to these problem variants,
        yet the decomposition should be adjusted to them.
		We now mention this in Section 5.
\end{enumerate}

\subsection*{Review 2}
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item \textbf{%
	Some more intuitive description of the main encoding part dealing with the difference logic constraints can be beneficial to the reader. For instance, what is the reason to include two rules (instead of one of these) in lines 20 and 21 of Listing 4 to constrain starting times of initial operations in a schedule? Regarding optimization based on DL variables, is the --minimize-variable option still used in the whole algorithm or is it not applicable due to the dynamically added optimize(m) program parts?}
\item[] We use the two rules in lines 20 and 21 of Listing 4 to restrict starting times from
        above and below. Maybe this is somewhat overshooting, but it makes sure that the starting
		times of operations from earlier time windows are fixed.
		Moreover, the \texttt{--minimize-variable} option is internally implemented by multi-shot solving in \textit{clingo}[DL].
		When control is handed over to a user-defined script, the option cannot be used anymore,
		and we take care of the optimization ourselves.
		The two issues are now addressed by additions to the last two paragraphs of Section 3.3.
\item \textbf{%
    The pseudo-code of the main control loop that implements the overall algorithm can be given so the reader can better comprehend the overall solution technique, especially with some of the additional features like window overlapping and compression.}
\item[] We now provide the pseudocode in Figure 2 and extended corresponding descriptions
        in the first paragraph of (the newly introduced) Section 3.4. 
\item \textbf{%		
    There might be a problem with the main encoding in Listing 4. The paper mentions when the X argument of share predicate is 1 (denoting the respective J1 operation is coming from the previous window) the ordering of respective operations is fixed via Line 12. However, line 4 will also generate use(M,W,w) instance and there will also be another copy of the share predicate instance only different in the X argument, this time 0. So, Line 16 will also try to choose an ordering for the same respective operations. Does this cause a problem in general or in the overlapping variant specifically where new ordering needs to be chosen for some of the already scheduled operations?}
\item[] When the operation order can be chosen by the rules in lines 16-18, the cases for a fixed
        ordering by the rules in lines 12-14 cannot apply. 
		This is now commented at the end of the second paragraph of Section 3.3.
\end{enumerate}

\subsection*{Review 3}
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item \textbf{%
      On the negative side, the novelty of the presented approach and the significance of the results are not clear to me. The authors cite in the introduction some past problem decomposition strategies based on a rolling horizon (Singer 2001; Liu et al. 2008) or bottleneck operations (Zhang and Wu 2010; Zhai et al. 2014), but there is no discussion on how they differ from those presented in the paper and why there is a need for new strategies. Indeed, the paper lacks a related work section. The experimental study does not include results from some standard methods to solve JSP nor compares the presented approach to related work. It is therefore not clear how the presented work advances state-of-the art.}
\item[] The problem here is that implementations of these previous methods are, to our knowledge,
        not publicly available. This seems to be generally the case for greedy and local search
		methods, so that we opted for OR-tools for a comparison with the state of the art added at the end of Section 4.
		It turns out that OR-tools with interval variables and global constraints in the encoding
		is ahead, but the gap isn't too large and much smaller for decomposition and multi-shot
		solving than with single-shot ASP modulo DL solving.
\item \textbf{%
      You do not define what you mean by partitioning the operations into balanced time windows. It would also be nice to give the main idea first before presenting the decomposition strategies. For instance, before start explaining Listing 2, you could explain what the J-EST is aiming to do. Similarly for the others.}
\item[] We tried to explain better what ``balanced'' is supposed to say in the first paragraph of Section 3.2.
        Otherwise, the decomposition strategies merely use different metrics to generate a total order of operations.
		In Section 5, we now comment on this and write that machine-learning approaches may be
		used for better problem decomposition.
\item \textbf{The last sentence of the third paragraph in page 2 ("Successful application areas beyond JSP include...") does not follow the previous sentence, you may want to reformulate it.}
\item[] We incorporated a (small) reformulation that hopefully helps here.
\item \textbf{The example for idle slots (Figure 2) is indeed negligible, could you present a better example to make your point? }
\item[] That's difficult. We wanted to have a small example such that the decomposition matches
        the J-EST approach. We would need to make the example substantially larger for having idle slots that propagate on to the makespan.
\item \textbf{You could structure Section 4 using subsections to improve readability.}
\item[] Here we usually take one paragraph per table with results, so that introducing subsections
        may be overshooting. 
\end{enumerate}
We highlighted the reformulations and additions relevant to the above comments/replies in yellow in our revised version, which is provided together with this reply.
\end{document}

### 2022-04-13 ICLP Paper/samplepaper.tex
abstract

@@
 \begin{abstract}
-Scheduling methods are important for effective production and logistics management, where tasks need to be allocated and performed by limited resources.
-In particular, the Job-shop Scheduling Problem (JSP) is a well-known and challenging
-combinatorial optimization problem in which tasks sharing a machine are to be arranged
-in sequence such that encompassing jobs can be completed as early as possible.
-Given that already moderately sized JSP instances can turn out as highly combinatorial,
-so that neither optimal schedules nor the runtime to termination of complete optimization methods
-is known, efficient approaches to approximate good-quality schedules are of interest.
-In this paper, we propose problem decomposition into time windows whose operations
-can be successively scheduled and optimized by means of multi-shot ASP solving.
-From a computational perspective, decomposition aims to split highly complex scheduling tasks into better manageable subproblems with a balanced number of 
-operations, so that good-quality or even optimal partial solutions can be
+Scheduling methods are important for effective production and logistics management, where tasks need to be allocated and performed with limited resources.
+In particular, the Job-shop Scheduling Problem (JSP) is a well-known and challenging combinatorial optimization problem in which tasks sharing a machine are to be arranged in a sequence such that encompassing jobs can be completed as early as possible.
+% remove double 'so that'
+Given that already moderately sized JSP instances can be highly combinatorial, and neither optimal schedules nor the runtime to termination of complete optimization methods is known, efficient approaches to approximate good-quality schedules are of interest.
+In this paper, we propose problem decomposition into time windows whose operations can be successively scheduled and optimized by means of multi-shot ASP solving.
+From a computational perspective, decomposition aims to split highly complex scheduling tasks into better manageable subproblems with a balanced number of  operations so that good-quality or even optimal partial solutions can be
 reliably found in a small fraction of runtime.
 Regarding the feasibility and quality of solutions, problem decomposition must
-respect the precedence of operations within their jobs, and partial schedules
+respect the precedence of operations within their jobs and partial schedules
 optimized by time windows should yield better global solutions than obtainable
-in similar runtime on the full problem.
+in similar runtime on the entire instance.
 We devise and investigate a variety of decomposition strategies in terms of the number and size of time windows as well as heuristics for choosing their operations.
 Moreover, we incorporate time window overlapping and compression techniques
-into the iterative scheduling process in order to counteract limitations of window-wise optimization restricted to partial schedules.
+into the iterative scheduling process to counteract window-wise optimization limitations restricted to partial schedules.
 Our experiments on JSP benchmark sets of several sizes show that
@@
 In general, scheduling operations is one of the most critical issues in the planning and managing of the manufacturing processes\cite{uzsoy2000performance}.
-Effective scheduling methods are essential for complex manufacturing and logistics systems,
-where allocating and performing diverse tasks within resource capacity limits 
-is one of the most critical challenges for production management \citep{uzsoy2000performance}.
+Effective scheduling methods are essential for complex manufacturing and transportation systems, where allocating and performing diverse tasks within resource capacity limits is one of the most critical challenges for production management \citep{uzsoy2000performance}.
 The Job-shop Scheduling Problem (JSP) \citep{baker1974introduction,taillard1993benchmarks}
 constitutes a well-known mathematical abstraction of industrial production scheduling in
-which sequences of operations need to be processed by machines such that a given objective
-like the makespan for completing all jobs or their tardiness w.r.t.\ deadlines is minimized.
-Finding optimal JSP solutions, determined by a sequence of operations for each machine,
-is an NP-hard combinatorial problem
-\citep{garey1976complexity,lenstra1977complexity,liu2008prediction},
-and optimal schedules as well as termination guarantees can be extremely challenging
-or beyond reach 
-of complete optimization methods already for moderately sized instances.
-For example, it took about $20$ years to develop a search procedure able to find a
-(provably) optimal solution for an instance called FT10 with $10$ jobs
-\citep{adams1988shifting,zhang2010hybrid},
-each consisting of a sequence of $10$ operations to be processed by $10$ machines.
+which sequences of operations need to be processed by machines such that a given objective,
+like the makespan for completing all jobs or their tardiness w.r.t.\ deadlines, is minimized.
+Finding optimal JSP solutions, determined by a sequence of operations for each machine, is an NP-hard combinatorial problem \citep{garey1976complexity,lenstra1977complexity,liu2008prediction}.
+Therefore, optimal schedules and termination guarantees can be extremely challenging or even unreachable for complete optimization methods already for moderately sized instances.
+For example, it took about $20$ years to develop a search procedure able to find a (provably) optimal solution for an instance called FT10 with $10$ jobs
+\citep{adams1988shifting,zhang2010hybrid}, each consisting of a sequence of $10$ operations to be processed by $10$ machines.
 

### 2022-04-13 ICLP Paper/samplepaper.tex
small changes

@@
 such that the value for the \lstinline{makespan} variable is minimized.
-In single-shot ASP modulo DL solving with \clingodl, this can be accomplished
-via the command-line option \lstinline{--minimize-variable=makespan}.
+In single-shot ASP modulo DL solving with \clingodl, the minimization can be accomplished via the command-line option \lstinline{--minimize-variable=makespan}.
 \hl{This option, however, is implemented by means of a fixed control loop that
@@
 \lstinline{horizon(m)} atom introduced before is set to false in step~1 for making sure that
-some (partial) schedule is feasible.
+some (partial) schedule $X$ is feasible.
 \hl{No such atoms have been introduced yet for the first time window $w=1$,
@@
 %
-Once some schedule with a horizon $h+1$ is found in step~4a,
+Once some schedule $X$ with a horizon $h+1$ is found in step~4a,
 the step~4b consists of instantiating

### 2022-04-17 Doctoral consortium ICLP 2022/example.tex
uploading DC ICLP 2022

\documentclass[submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{ICLP DC 2022} % Name of the event you are submitting to
\usepackage{breakurl}             % Not needed if you use pdflatex only.
\usepackage{underscore}           % Only needed if you use pdflatex.
\usepackage{pgfplots, pgfplotstable}

\title{Decomposition strategies for solving Scheduling Problems in Industrial Applications}
\author{Mohammed M. S. El-Kholany
\institute{University of Klagenfurt \\ Klagenfurt, Austria}
\institute{Cairo University \\ Cairo, Egypt}
\email{mohammed.el-kholany@aau.at}
\and
Martin Gebser
\institute{University of Klagenfurt \thanks{} \\ Klagenfurt, Austria}
\institute{Graz University of Technology \\ Graz, Austria}
\email{martin.gebser@aau.at}
\and
Konstantin Schekotihin
\institute{University of Klagenfurt \\ Klagenfurt, Austria}
\email{konstantin.schekotihin@aau.at}
}
\def\titlerunning{Decomposition strategies for solving Scheduling Problems in Industrial Applications}
\def\authorrunning{M. El-Kholany, M. Gebser \& K. Schekotihin}
\begin{document}
\maketitle

\begin{abstract}
  This article presents an overview of a research study of a crucial optimization problem in the Computer Science/Operations research field: The job-shop Scheduling Problem (JSP). The JSP is a challenging task in which a set of operations must be processed using a set of scarce machines to optimize a particular objective. The main purpose of the JSP is to determine the execution order of the processes assigned to each machine to optimize an objective. Our main interest in this study is to investigate developing decomposition strategies using logic programming to solve the JSP. We split our goal into two main phases. The first phase is to apply the decomposition approach and evaluate the proposed model by solving a set of known benchmark instances. The second phase is to apply the successful decomposition methods obtained from the first phase to solve a scheduling problem in the real-life application. In the current state, we finished the first phase and started the second one aiming to have a model that can provide a schedule of a factory for a short-time period.
\end{abstract}

\section{Introduction}

Real-life scheduling problems are very complex because of the combinatorial nature and constraints dictated by different production environments. The classical Job-shop Scheduling Problem (JSP) has been known as a complex and combinatorial optimization problem since the 1950s and it was shown to be NP-hard in \cite{lenstra1977complexity}. Therefore, it is not easy to obtain the optimal solution. Until now, the optimal solution of some benchmark instances with medium size is unknown. In reality, the scheduling problems are much more complicated than the classical ones because many jobs should be scheduled and it may be up to $10000$ in some mechanical workshops \cite{zhang2010hybrid}. In addition, there are a lot of integrated factors that make the problem more complex. The general overview of the scheduling is to find the sequence of the tasks assigned to a scarce resource while optimizing an objective. Our study focuses on the Job-shop Scheduling Problem (JSP). The JSP is defined as a set of $n$ jobs $J_{1}, ..., J_{n}$ have to be executed by a set of $m$ machines $M_{1}, ... M_{m}$ while finishing all of these jobs as early as possible. Each job has a series of $m$ operations with a predefined order and must visit each machine one time. Each machine can process an operation at a time. A machine cannot be interrupted once it starts processing an operation. The processing time of each operation is fixed. The naive objective of the JSP is to complete all jobs in the shortest time, which is called makespan. The main goal of solving such a problem is to find the sequence of operations assigned to each machine that completes all the jobs in the shortest possible time while satisfying the precedence constraint. Since the problem is complex and developing a model that provides an efficient schedule is challenging, many researchers have studied the JSP and proposed several techniques to solve it. The following section summarizes the previous work that tackled the JSP.

\section{Literature Review}
This section summarizes the work studied on solving scheduling problems, particularly the JSP. The scheduling problem has been addressed using exact and/or heuristics techniques. Several researchers have developed and proposed different dispatching rules to order the operations stated in front of each machine to obtain an efficient schedule in a reasonable time. Many dispatching rules have been proposed in \cite{holthaus1997efficient,kaban2012comparison,paul2015investigation} for solving different scheduling problems. These studies focused on minimizing either the total tardiness or makespan. However, the main issue with the dispatching rule is that there is no optimization process behind it. It aims to define combined rules based on the information of the problem and evaluate the obtained schedule w.r.t the performance measures. Other researchers have applied the data mining techniques that study the sequence pattern of the operations assigned to each machine and develop some rules to order the operations in front of each machine \cite{shahzad2010discovering,koonce2000using}. Another group of researchers has presented exact methods such as Mixed-integer Programming models or constraint programming for solving small benchmark instances of the JSP while minimizing the tardiness \cite{ku2016mixed,meng2020mixed}. The drawback of the exact methods is that it fails to obtain a good solution in a reasonable time for large instances. This leads many researchers to follow another decomposition approach that aims to partition the problem into subparts, optimize each separately, and merge all sub-solutions to solve the main problem. Different decomposition strategies have been proposed during the last two decades. Most of the previous work decomposed the problem based on different criteria and applied heuristics/metaheuristics to solve the problem\cite{singer2001decomposition, zhai2014decomposition,byeon1998decomposition}. This study aims to introduce new decomposition strategies to split the problem efficiently into parts (time windows) and solve them to obtain good solutions reasonably. The decomposition optimization processes are applied using a Logic Programming language called Answer Set Programming.

\section{Research Goal}
The main goal of our research is to build a model that can provide an efficient schedule for manufacturing industrial systems for a short period using Answer Set Programming (ASP). ASP is a programming methodology rooted in artificial intelligence and computational logic research \cite{lifschitz2002answer}. The main reason for selecting ASP as a solver is because it showed its effectiveness for solving different combinatorial optimization problems in several fields \cite{abseher2016shift,ricca2012team,abels2019train}. Since the scheduling problem is very complex and increasing the number of operations to a schedule makes the problem much harder. We aim to use the decomposition approach to obtain a good schedule in a reasonable time. We can split our main goal into two main steps. The first is to develop different decomposition strategies and investigate their effectiveness. Therefore, we can determine the most critical factors that significantly impact the solution quality. Secondly is to build the scheduling model using ASP, which optimizes each time window separately by applying a multi-shot ASP solving technique\cite{gebser2019multi}.

\subsection{Accomplished work}
In order to accomplish our goal, we decided to build a scheduling model using ASP and assess its performance by solving a set of known benchmark instances. Firstly, we have built a basic model of ASP to solve the classical JSP and we found that it cannot find good solutions for the instances with more than 500 operations. So, we investigated applying the decomposition approach in which the problem is cut into sub-parts called (time windows) and the ASP optimizer solves each dynamically using the multi-shot solving technique \cite{gebser2019multi} and get the solution of the original problem. The main contribution of this approach is to find a smart way to split the problem into time windows. More specifically, determining which operations are assigned in which time window with considering the precedence constraint and obtaining a near-optimal solution in a reasonable time. We have developed/proposed different decomposition methods based on different techniques and tested them on a set of benchmark instances generated in \cite{taillard1993benchmarks,demirkol1998benchmarks}. Firstly, we applied a constrained k-Means clustering algorithm for the decomposition process. This work depends on collecting information from the problem itself and from the solutions obtained by heuristics like FIFO/MTWR to define the similar operations and assign them to the same cluster where a cluster is a time window and applied multi-shot ASP for the optimization. This work has been published in PADL 2022 \cite{el2022decomposition}. 
Secondly, We have developed other decomposition strategies that depend on ranking the operations. Our two main decomposition ideas in that work are time/machine-based methods. The time-based approach ranks the operations based on the Earliest possible Starting Time (EST) or the Most Total Work Remaining (MTWR) and the operation with a smaller rank will be assigned to an earlier time window. However, the machine-based method ranks the operations based on the time-based method and determines the bottleneck machines. The operation with a smaller rank and assigned to the bottleneck machine will be in an earlier time window. This work is currently under the second review of ICLP 2022. In addition, we applied different techniques to improve the solution quality. Firstly, we applied the overlapping between the time windows. This process aims to overcome the mistakes due to the decomposition. Secondly, we performed a post-processing phase (compression) after optimizing each time window. In this phase, we check the possibility of executing an operation earlier without violating the precedence constraint.

\section{Results}
This section shows a summary of the obtained results with a set of benchmark instances created in \cite{taillard1993benchmarks}. We selected the sets with the largest size which are $50\times15$, $50\times20$ and $100\times20$. We will focus only in this article on the size $50\time15$ to avoid presenting too many results of our experiments. We performed the experiments on our selected instances with different time windows ($2 - 10$ ) to determine the number of time windows that provide a good makespan with a timeout of $1000$ seconds. Figure ~\ref{fig:figure01} shows the average makespan of the instances set $50\times15$ with different time windows while applying EST Time-based and Machine-based decomposition methods. The results show that the Machine-based decomposition method is more effective than the Time-based method throughtout all the time windows. The best number of time windows for such a set is $3$. After determining the best number of time windows for the instances, we applied the overlapping technique with different percentages.
The overlapping percentage represents the number of operations assigned to the previous time window and will be rescheduled with the current time window. Increasing the percentage could provide better results; however, the computational time will increase. So, we run the experiments with overlapping percentages to determine the best percentage for the selected set, which provides a good result in a reasonable time. Figure ~\ref{fig:figure02} shows the makespan of each instance when the EST Time-based is applied with $03$ time windows with/without overlapping. The horizontal line shows the selected instances and the vertical shows the gap to the optimality. Each line represents a particular overlapping percentage where $0\%$ means no overlapping. We can see from the figure that increasing the overlapping percentage improves the solution quality. When we set the overlapping percentage to $40\%$ and $50\%$, the results were worse than the obtained. The last step we applied to improve the solution quality is compression. The main goal of this phase is to check after optimizing the current window if it is possible to execute some operations earlier without violating the precedence constraint. Figure ~\ref{fig:figure03} shows the average makespan throughtout the time windows when we applied the compression and compares the results without compression. As shown in the figure, the compression improves the solution quality for all time windows. We can conclude from the obtained results in this section that considering the bottleneck machines (has the highest processing time to finish) during the decomposition process is beneficial. In addition, the overlapping plays a significant role in improving the solution quality because it overcomes the decomposition mistakes. Also, applying the compression phase after optimizing each time window has a positive impact.

\pgfplotstableread[row sep=\\,col sep=&]{
    interval & EST Time-based & EST Machine-based \\
    TW 02     & 3525  & 3150   \\
    TW 03     & 3622 & 3084   \\
    TW 04     & 3870 & 3111  \\
    TW 05     & 3908 & 3214  \\
    TW 06     & 4044  & 3225  \\
    TW 07     & 4064  & 3351 \\
    TW 08     & 4269  & 3379  \\
    TW 09     & 4146  & 3495  \\
    TW 10     & 4127  & 3525  \\
    }\mydata

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}[
                ybar,
                symbolic x coords={TW 02,TW 03,TW 04,TW 05,TW 06,TW 07,TW 08,
                TW 09,TW 10},
                width=\textwidth,
                ymin=0,
                ymax=4400,
                height=8cm,
                ylabel={Average makespan},
                xtick=data,
            ]
            \addplot table[x=interval,y=EST Time-based]{\mydata};
            \addplot table[x=interval,y=EST Machine-based]{\mydata};
            \legend{EST Time-based, EST Machine-based};
    \end{axis}
  \end{tikzpicture}
  \caption{Comparison between (Time and Machine)-based decomposition strategy}  
  \label{fig:figure01}
\end{figure}

\pgfplotstableread[row sep=\\,col sep=&]{
instance  &  0    & 10    & 20    & 30 \\
ta51      & 24.71 & 21.30 & 20.69 & 21.78 \\
ta52      & 36.07 & 34.87 & 27.43 & 26.71 \\
ta53      & 28.71 & 25.80 & 20.39 & 18.00 \\
ta54      & 22.90 & 19.65 & 18.99 & 16.94 \\
ta55      & 28.37 & 25.23 & 26.35 & 25.12 \\
ta56      & 24.81 & 23.52 & 23.34 & 21.32 \\
ta57      & 28.41 & 26.33 & 24.97 & 21.34 \\
ta58      & 30.92 & 30.50 & 25.89 & 22.77 \\
ta59      & 27.50 & 28.66 & 25.73 & 22.75 \\
ta60      & 18.22 & 18.47 & 16.05 & 18.66 \\
}\overlapping
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}[
       xtick=data,
       xticklabels from table={\overlapping}{instance},      
       bar width=7mm, y=4mm,
        nodes near coords align={vertical},
        ymin = 15,
        ymax = 37,
        height = 4cm,
        width=\textwidth,
        ylabel={Gap to optimality (\%)},
      ]
    \addplot table [ybar, y=0, x expr=\coordindex,] {\overlapping};  %Works not
    \addplot table [y=10, x expr=\coordindex] {\overlapping}; %Works
    \addplot table [y=20, x expr=\coordindex] {\overlapping}; %Works
    \addplot table [y=30, x expr=\coordindex] {\overlapping}; %Works
    \legend{0\% , 10\%, 20\%, 30\%}
    \end{axis}
  \end{tikzpicture}  
  \caption{EST Time-based with overlapping (03 TW)} 
  \label{fig:figure02}
\end{figure}

\pgfplotstableread[row sep=\\,col sep=&]{
    interval & No-compression & Compression \\
    TW 02     & 3525  & 3426   \\
    TW 03     & 3622 & 3474   \\
    TW 04     & 3870 & 3502  \\
    TW 05     & 3908 & 3479  \\
    TW 06     & 4044  & 3445  \\
    TW 07     & 4064  & 3422 \\
    TW 08     & 4269  & 3463  \\
    TW 09     & 4146  & 3454  \\
    TW 10     & 4127  & 3441  \\
}\compression

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}[
                ybar,
                symbolic x coords={TW 02,TW 03,TW 04,TW 05,TW 06,TW 07,TW 08,
                TW 09,TW 10},
                width=\textwidth,
                ymin=0,
                ymax=4400,
                height=8cm,
                ylabel={Average makespan},
                xtick=data,
            ]
            \addplot table[x=interval,y=No-compression]{\compression};
            \addplot table[x=interval,y=Compression]{\compression};
            \legend{No-compression, Compression};
    \end{axis}
  \end{tikzpicture}
  \caption{Makespan EST Time-based with/without compression}
  \label{fig:figure03}
\end{figure}

\section{Expected Achievement}
The classical Job-shop Scheduling Problem (JSP) is limited and has many assumptions which make the problem unrealistic. We currently aim to apply the successful (Machine-based) decomposition strategies, overlapping and compression, to semiconductor manufacturing systems. In order to accomplish this, we are currently working on a dataset generated from a simulator called SMT2020 \cite{kopp2020smt2020}. The advantage of this dataset is that it links between the academic and industrial perspectives. Such a dataset considers many factors such as machine maintenance, cascading, and batching. We aim to build a scheduling model-based ASP by considering the mentioned factors and testing it on a small-scale dataset. After that, we aim to link our model with the simulator SMT2020 and investigate how our proposed model behaves with such a problem compared to other dispatching rules developed by colleagues in our research group.

\nocite{*}
\bibliographystyle{eptcs}
\bibliography{generic}
\end{document}

### 2022-05-11 ICLP Paper/samplepaper.tex
extended abstract and introduction in view of 2nd round comments by Reviewer 3

@@
 \usepackage{soul}%{color,soul}
+\usepackage{pagecolor}
+\sethlcolor{\thepagecolor}
 
@@
 Given that already moderately sized JSP instances can be highly combinatorial, and neither optimal schedules nor the runtime to termination of complete optimization methods is known, efficient approaches to approximate good-quality schedules are of interest.
-In this paper, we propose problem decomposition into time windows whose operations can be successively scheduled and optimized by means of multi-shot ASP solving.
+In this paper, we propose problem decomposition into time windows whose operations can be successively scheduled and optimized by means of multi-shot Answer Set Programming (ASP) solving.
 From a computational perspective, decomposition aims to split highly complex scheduling tasks into better manageable subproblems with a balanced number of  operations so that good-quality or even optimal partial solutions can be
@@
 where the gap increases with the number of operations to schedule.
+While the obtained solution quality still remains behind a state-of-the-art Constraint
+Programming system, our multi-shot solving approach comes closer the larger the instance
+size, demonstrating good scalability by problem decomposition.
 %
@@
       where the gap increases with the number of operations to schedule.
+      While the state-of-the-art CP system OR-tools \citep{ortools} is still ahead regarding the solution quality, our multi-shot solving approach % based on problem decomposition
+      comes the closer the larger the instance size gets.
 \end{itemize}
@@
 indicated in the lower half of Table}~\ref{tab:Table03}\hl{, in relation,
-we also ran the state-of-the-art CP system OR-tools (version 9.2)} \citep{ortools}\hl{
+we also ran the state-of-the-art CP system OR-tools (version 9.2)}\hl{
 with a timeout of $1000$ seconds,

### 2022-06-01 ICLP_DC/ICLP_DC.tex
updating_ICLP_DC

--- 2022-04-17 Doctoral consortium ICLP 2022/example.tex
@@
 \usepackage{pgfplots, pgfplotstable} 
+\usepackage{mathtools}
+\RequirePackage{doi}
 
@@
 \email{mohammed.el-kholany@aau.at}
-\and
-Martin Gebser
-\institute{University of Klagenfurt \thanks{} \\ Klagenfurt, Austria}
-\institute{Graz University of Technology \\ Graz, Austria}
-\email{martin.gebser@aau.at}
-\and
-Konstantin Schekotihin
-\institute{University of Klagenfurt \\ Klagenfurt, Austria}
-\email{konstantin.schekotihin@aau.at}
 }
@@
 \begin{abstract}
-  This article presents an overview of a research study of a crucial optimization problem in the Computer Science/Operations research field: The job-shop Scheduling Problem (JSP).
+  This article presents an overview of a research study of a crucial optimization problem in the Computer Science/Operations research field: The Job-shop Scheduling Problem (JSP).
 The JSP is a challenging task in which a set of operations must be processed using a set of scarce machines to optimize a particular objective.
@@
 Real-life scheduling problems are very complex because of the combinatorial nature and constraints dictated by different production environments.
-The classical Job-shop Scheduling Problem (JSP) has been known as a complex and combinatorial optimization problem since the 1950s and it was shown to be NP-hard in \cite{lenstra1977complexity}.
+The classical JSP has been known as a complex and combinatorial optimization problem since the 1950s and it was shown to be NP-hard in \cite{lenstra1977complexity}.
 Therefore, it is not easy to obtain the optimal solution.
@@
 Our study focuses on the Job-shop Scheduling Problem (JSP).
-The JSP is defined as a set of $n$ jobs $J_{1}, ..., J_{n}$ have to be executed by a set of $m$ machines $M_{1}, ...
+The JSP is defined as a set of $n$ jobs $J_{1}, ..., J_{n}$ that have to be executed by a set of $m$ machines $M_{1}, ...
 M_{m}$ while finishing all of these jobs as early as possible.
@@
 It aims to define combined rules based on the information of the problem and evaluate the obtained schedule w.r.t the performance measures.
-Other researchers have applied the data mining techniques that study the sequence pattern of the operations assigned to each machine and develop some rules to order the operations in front of each machine \cite{shahzad2010discovering,koonce2000using}.
+Other researchers have applied data mining techniques that study the sequence pattern of the operations assigned to each machine and develop some rules to order the operations in front of each machine \cite{shahzad2010discovering,koonce2000using}.
 Another group of researchers has presented exact methods such as Mixed-integer Programming models or constraint programming for solving small benchmark instances of the JSP while minimizing the tardiness \cite{ku2016mixed,meng2020mixed}.
-The drawback of the exact methods is that it fails to obtain a good solution in a reasonable time for large instances.
+The drawback of the exact methods is that they fail to obtain a good solution in a reasonable time for large instances.
 This leads many researchers to follow another decomposition approach that aims to partition the problem into subparts, optimize each separately, and merge all sub-solutions to solve the main problem.
 Different decomposition strategies have been proposed during the last two decades.
-Most of the previous work decomposed the problem based on different criteria and applied heuristics/metaheuristics to solve the problem\cite{singer2001decomposition, zhai2014decomposition,byeon1998decomposition}.
+Most of the previous work decomposed the problem based on different criteria and applied heuristics/metaheuristics to solve the problem~\cite{singer2001decomposition, zhai2014decomposition,byeon1998decomposition}.
 This study aims to introduce new decomposition strategies to split the problem efficiently into parts (time windows) and solve them to obtain good solutions reasonably.
@@
 Firstly, we applied a constrained k-Means clustering algorithm for the decomposition process.
-This work depends on collecting information from the problem itself and from the solutions obtained by heuristics like FIFO/MTWR to define the similar operations and assign them to the same cluster where a cluster is a time window and applied multi-shot ASP for the optimization.
+This work depends on collecting information from the problem itself and the solutions obtained by heuristics like FIFO/MTWR to define the similar operations and assign them to the same cluster where a cluster is a time window and apply multi-shot ASP for the optimization.
 This work has been published in PADL 2022 \cite{el2022decomposition}.
@@
 The time-based approach ranks the operations based on the Earliest possible Starting Time (EST) or the Most Total Work Remaining (MTWR) and the operation with a smaller rank will be assigned to an earlier time window.
-However, the machine-based method ranks the operations based on the time-based method and determines the bottleneck machines.
-The operation with a smaller rank and assigned to the bottleneck machine will be in an earlier time window.
-This work is currently under the second review of ICLP 2022.
+For simplicity, we will describe only the EST strategy.
+In the EST (time-based), the EST of first operations of all jobs is $0$ because they have no predecessors.
+The EST of the successors is the sum of processing times for the predecessor operations belonging to the same job.
+The operation with smaller EST will be assigned to an earlier time window.
+If the EST of two or more operations is similar, the operation with a shorter processing time will be assigned to an earlier time window.
+However, in the EST (machine-based), the first step is to rank the operations based on EST.
+Secondly, we determine the bottleneck machines with the longest processing time to execute.
+Thirdly, we select the operation assigned to a bottleneck machine with the smallest EST; if the predecessor of that operation is not assigned to a time window, it will be assigned to that time window to satisfy the precedence constraint.
+This work has been accepted and will be presented in ICLP 2022.
 In addition, we applied different techniques to improve the solution quality.
@@
 We selected the sets with the largest size which are $50\times15$, $50\times20$ and $100\times20$.
-We will focus only in this article on the size $50\time15$ to avoid presenting too many results of our experiments.
-We performed the experiments on our selected instances with different time windows ($2 - 10$ ) to determine the number of time windows that provide a good makespan with a timeout of $1000$ seconds.
+We will focus only in this article on the size $50\times15$ to avoid presenting too many results of our experiments.
+We performed the experiments on our selected instances with different time windows ($2 - 10$) to determine the number of time windows that provide a good makespan with a timeout of $1000$ seconds.
 Figure ~\ref{fig:figure01} shows the average makespan of the instances set $50\times15$ with different time windows while applying EST Time-based and Machine-based decomposition methods.
@@
 We can see from the figure that increasing the overlapping percentage improves the solution quality.
-When we set the overlapping percentage to $40\%$ and $50\%$, the results were worse than the obtained.
+When we set the overlapping percentage to $40\%$ and $50\%$, the results were worse than the obtained with smaller percentages.
 The last step we applied to improve the solution quality is compression.
@@
 \section{Expected Achievement}
-The classical Job-shop Scheduling Problem (JSP) is limited and has many assumptions which make the problem unrealistic.
+The classical JSP is limited and has many assumptions which make the problem unrealistic.
 We currently aim to apply the successful (Machine-based) decomposition strategies, overlapping and compression, to semiconductor manufacturing systems.
@@
 The advantage of this dataset is that it links between the academic and industrial perspectives.
-Such a dataset considers many factors such as machine maintenance, cascading, and batching.
-We aim to build a scheduling model-based ASP by considering the mentioned factors and testing it on a small-scale dataset.
-After that, we aim to link our model with the simulator SMT2020 and investigate how our proposed model behaves with such a problem compared to other dispatching rules developed by colleagues in our research group.
+Since the SMT2020 data set is quite big, it is not possible to provide a schedule with a simulator that simulates the system for $2$ years, as has been presented in \cite{kopp2020smt2020} using ASP.
+We aim to develop a scheduler model that could provide a schedule in the manufacturing system for a short period, like one week.
+The given dataset is completely different from the classical JSP, where there is a set of lots instead of jobs.
+Each lot belongs to a particular product, consists of steps (operations in the JSP), and is released at a particular time.
+Each step can be executed by different machines \textit{Flexible scheduling problem} (FSB); however, in the JSP, each operation is executed by only one machine.
+Since the problem is FSB, it is required to find a criterion to assign each step to a machine before starting to schedule.
+In addition, the optimization criterion is to minimize the total tardiness of the lots while considering other factors such as machine maintenance, cascading, and batching.
+Given that information, our plan is:
+\begin{enumerate}
+  \item Firstly, we aim to build a basic ASP model that assigns each step to a machine and then schedules the lots considering only the machine maintenance.
+  \item Secondly, develop decomposition strategies \textit{(machine-based)} and apply the multi-shot solving ASP.
+  \item Thirdly, we will consider the other mentioned factors.
+\end{enumerate}
+
+In order to assess our model, we will compare our obtained results with the results of colleagues (in our research group) who are working on the same dataset and applying Deep Learning and dispatching rules.
 
